### 2019-10-08 template.tex
add latex template/edit readme

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejihpe, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: Place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: Describe briefly the main methods or treatments applied; (3) Results: Summarize the article's main findings; and (4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact latex@mdpi.com.
%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}
The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}. Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\begin{quote}
This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection}
\unskip
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}[leftmargin=*,labelsep=5.8mm]
\item	First bullet
\item	Second bullet
\item	Third bullet
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item	First item 
\item	Second item
\item	Third item
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=2 cm]{Definitions/logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

Text

Text

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

Text

Text

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\textwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\textwidth}{1pt}
%\end{listing}

\subsection{Formatting of Mathematical Components}

This is an example of an equation:

\begin{equation}
a + b = c
\end{equation}
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}
This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendix
\section{}
\unskip
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book}
Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2019-10-10 hmda_sample_for_paper.py
add first version of the HMDA data and the python code to create it

@@ -1,0 +1,1742 @@
+from typing import Union
+import pandas as pd
+import numpy as np
+import sqlalchemy
+
+
+def optimize_data_types(
+        df: Union[pd.Series, pd.DataFrame],
+        verbose: int = 0,
+        shrink_ints=1,
+        shrink_floats=1,
+        shrink_strings=1
+):
+    """
+    Reviews all variables in a dataframe or values in a series to determine the
+     smallest datatype that can be used while not losing information.
+    :param df - Pandas dataframe or series for use in the datatype review
+    :param verbose - int that determines whether information is printed
+    value = 0 means no printing; higher values lead to more detail
+    being printed.
+    """
+    if isinstance(df, pd.DataFrame):
+        df = df.copy()
+        orig_mem = df.memory_usage(deep=True).sum() / 1024 ** 2
+        if shrink_ints:
+            df_int = df.select_dtypes(include=['int'])
+            converted_int = df_int.apply(pd.to_numeric, downcast='integer')
+            df[converted_int.columns] = converted_int
+            int_reduce = df.memory_usage(deep=True).sum() / 1024 ** 2
+        else:
+            int_reduce = df.memory_usage(deep=True).sum() / 1024 ** 2
+
+        if shrink_floats:
+            df_float = df.select_dtypes(include=['float'])
+            converted_float = df_float.apply(pd.to_numeric, downcast='float')
+            df[converted_float.columns] = converted_float
+            float_reduce = df.memory_usage(deep=True).sum() / 1024 ** 2
+        else:
+            float_reduce = df.memory_usage(deep=True).sum() / 1024 ** 2
+
+        if shrink_strings:
+            df_object = df.select_dtypes(include=['object'])
+            for col in df_object:
+                unique = len(df_object[col].unique())
+                total = len(df_object[col])
+                if unique / total < 0.5:
+                    df[col] = df_object[col].astype('category')
+            category_reduce = df.memory_usage(deep=True).sum() / 1024 ** 2
+        else:
+            category_reduce = df.memory_usage(deep=True).sum() / 1024 ** 2
+        if verbose > 0:
+            df.info(memory_usage='deep')
+            print("{:03.2f} MB Originally".format(orig_mem))
+            print(
+                "{0:03.2f} MB {1:5.2f}% After Int Reduction".format(
+                    int_reduce, (int_reduce - orig_mem) / orig_mem * 100.0
+                )
+            )
[…]

### 2019-10-10 hmda_sample_for_paper.py
add first version of the HMDA data and the python code to create it

@@ -1736,7 +1736,7 @@ if __name__ == '__main__':
         'tract_median_age_of_housing_units',
     ]
     lar = lar[output_columns]
-    lar_sample = lar.sample(n=40000, random_state=5419)
+    lar_sample = lar.sample(n=40000)
     lar_sample.to_csv(
         '/mnt/ldrive/census/hmda lar 2018-static/hmda_lar_2018_orig_mtg_sample.csv'
     )

### 2019-11-01 article-information-2019.tex
basic LaTeX structure

--- 2019-10-08 template.tex
@@
 %=================================================================
-\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[information,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 % Full title of the paper (Capitalized)
-\Title{Title}
+\Title{Responsible Machine Learning\\\large{Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}}
 
@@
 % Authors, for the paper (add full first names)
-\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
+\Author{Navdeep Gill $^{1, \ddagger}$, Patrick Hall $^{1,\ddagger,*}$, Kim Montgomery $^{1,\ddagger}$, and Nicholas Schmidt $^{2, \ddagger}$}
 
 % Authors, for metadata in PDF
-\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+\AuthorNames{Navdeep Gill, Patrick Hall, Kim Montgomery, and Nicholas Schmidt}
 
@@
 \address{%
-$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
-$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
+$^{1}$ \quad H2O.ai\\
+$^{2}$ \quad BLDS, LLC}
 
 % Contact information of the corresponding author
-\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
+\corres{Correspondence: phall@h2o.ai; nschmidt@bldsllc.com}
 
 % Current address and/or shared authorship
-\firstnote{Current address: Affiliation 3} 
+%\firstnote{Current address: Affiliation 3} 
 \secondnote{These authors contributed equally to this work.}
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{A single paragraph of about 200 words maximum.
-For research articles, abstracts should give a pertinent overview of the work.
-We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: Place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: Describe briefly the main methods or treatments applied; (3) Results: Summarize the article's main findings; and (4) Conclusion: Indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
+\abstract{This text outlines a viable approach for training and evaluating complex machine learning systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (M-GBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated datasets with known feature importance and sociological bias characteristics and on realistic, publicly available example datasets.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of individual conditional expectation (ICE) and global and local gradient-based or Shapley feature importance.
+The constrained model predictions are also tested for disparate impact and other types of sociological bias using straightforward group fairness measures.
+By combining innovations in interpretable models, post-hoc explanation, and bias testing with accessible software tools, this text aims to provide a template workflow for important machine learning applications that require high accuracy and interpretability and low disparate impact.}
 
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}
+\keyword{Machine Learning; Neural Network; Gradient Boosting Machine; Interpretable; Explanation; Fairness; Disparate Impact; Python}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{section}{-1} %% Remove this when starting to work on the template.
-\section{How to Use this Template}
-The template details the sections that can be used in a manuscript.
+\setcounter{section}{-1} 
+%% Remove this when starting to work on the template.
+%\section{How to Use this Template}
+%The template details the sections that can be used in a manuscript.
 Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section).
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-The introduction should briefly place the study in a broad context and highlight why it is important.
-It should define the purpose of the work and its significance.
-The current state of the research field should be reviewed carefully and key publications cited.
-Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-And now citing a book reference \cite{ref-book}.
-Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Materials and Methods}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Data Description}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Model Description}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Software Resources}
+% Description of code
 
@@
 
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\begin{quote}
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\end{quote}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simulated Data Results}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection}
-\unskip
-\subsubsection{Subsubsection}
+\subsection{Loan Data Results}
 
-Bulleted lists look like this:
-\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-\item	First bullet
-\item	Second bullet
-\item	Third bullet
-\end{itemize}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
 
-Numbered lists can be added as follows:
-\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item	First item 
-\item	Second item
-\item	Third item
-\end{enumerate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+% Model debugging
+% Model management and monitoring
+% Appeal
 
-The text continues here.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
 
-\subsection{Figures, Tables and Schemes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{, N.G.; , P.H.; , K.M.; , N.S.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work reported.}
 
-All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\funding{This research received no external funding.}
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=2 cm]{Definitions/logo-mdpi}
-\caption{This is a figure, Schemes follow the same formatting.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\acknowledgments{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conflictsofinterest{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+\abbreviations{The following abbreviations are used in this manuscript:\\}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\reftitle{References}
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\externalbibliography{yes}
+\bibliography{article-information-2019}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%This section may be divided by subheadings.
+It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
+%\begin{quote}
+%This section may be divided by subheadings.
+It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
+%\end{quote}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Subsection}
+%\unskip
+%\subsubsection{Subsubsection}
+
+%Bulleted lists look like this:
+%\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+%\item	First bullet
+%\item	Second bullet
+%\item	Third bullet
+%\end{itemize}
+
+%Numbered lists can be added as follows:
+%\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
+%\item	First item 
+%\item	Second item
+%\item	Third item
+%\end{enumerate}
+
+%The text continues here.
+
+%\subsection{Figures, Tables and Schemes}
+
+%All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
+
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=2 cm]{Definitions/logo-mdpi}
+%\caption{This is a figure, Schemes follow the same formatting.
 If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
@@
 A caption on a single line should be centered.}
-\end{figure}
+%\end{figure}
 
-Text
+%Text
 
-Text
+%Text
 
-\begin{table}[H]
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
-\centering
+%\begin{table}[H]
+%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
+%\centering
 %% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-entry 1		& data			& data\\
-entry 2		& data			& data\\
-\bottomrule
-\end{tabular}
-\end{table}
+%\begin{tabular}{ccc}
+%\toprule
+%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+%\midrule
+%entry 1		& data			& data\\
+%entry 2		& data			& data\\
+%\bottomrule
+%\end{tabular}
+%\end{table}
 
-Text
+%Text
 
-Text
+%Text
 
@@
 
-\subsection{Formatting of Mathematical Components}
+%\subsection{Formatting of Mathematical Components}
 
-This is an example of an equation:
+%This is an example of an equation:
 
-\begin{equation}
-a + b = c
-\end{equation}
+%\begin{equation}
+%a + b = c
+%\end{equation}
 %% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
@@
 
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
+%Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
 %% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
+%\begin{Theorem}
+%Example text of a theorem.
+%\end{Theorem}
 
-The text continues here. Proofs must be formatted as follows:
+%The text continues here. Proofs must be formatted as follows:
 
 %% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
+%\begin{proof}[Proof of Theorem 1]
+%Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
+%\end{proof}
+%The text continues here.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
+%\section{Discussion}
 
-Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
+%Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
 The findings and their implications should be discussed in the broadest context possible.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
+%\section{Materials and Methods}
 
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
+%Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
 Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
@@
 
-Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
+%Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
 If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
@@
 
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
+%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+%\section{Conclusions}
 
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
+%This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
+%\section{Patents}
+%This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work reported.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}
+%\noindent 
+%\begin{tabular}{@{}ll}
+%MDPI & Multidisciplinary Digital Publishing Institute\\
+%DOAJ & Directory of open access journals\\
+%TLA & Three letter acronym\\
+%LD & linear dichroism
+%\end{tabular}}
 
@@
 %% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
-
-\noindent 
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & linear dichroism
-\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendix
-\section{}
-\unskip
-\subsection{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
+%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+%\appendix
+%\section{}
+%\unskip
+%\subsection{}
+%The appendix is an optional section that can contain details and data supplemental to the main text.
 For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
@@
 
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\reftitle{References}
+%\section{}
+%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
 
@@
 according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: external bibliography
-%=====================================
-%\externalbibliography{yes}
-%\bibliography{your_external_BibTeX_file}
+% Citations an
 
@@
 %=====================================
-\begin{thebibliography}{999}
+%\begin{thebibliography}{999}
 % Reference 1
-\bibitem[Author1(year)]{ref-journal}
-Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+%\bibitem[Author1(year)]{ref-journal}
+%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
 % Reference 2
-\bibitem[Author2(year)]{ref-book}
-Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-\end{thebibliography}
+%\bibitem[Author2(year)]{ref-book}
+%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; %Publishing House: City, Country, 2007; pp. 32--58.
+%\end{thebibliography}
 
@@
 %% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
+%\sampleavailability{Samples of the compounds ...... are available from the authors.}
 
@@
 %}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}

### 2019-11-03 data/scripts/information_paper_initial_processing.py
Revise HMDA train and test datasets.

@@ -9,10 +9,10 @@ Subset the original data and create a training and test set.
 random.seed(42)
 
 # Load the data
-DATA = pd.read_csv('./hmda_lar_2018_orig_mtg_sample.csv')
+DATA = pd.read_csv('data/hmda_lar_2018_orig_mtg_sample.csv')
 
 
-
+'''
 # Keep only certain subsets
 DATA = DATA[DATA["derived_dwelling_category"]=="Single Family (1-4 Units):Site-Built"]
 # Subset to "action_taken"=="Loan originated"
@@ -39,14 +39,14 @@ keep_columns += ['applicant_credit_score_type', 'co_applicant_credit_score_type'
 
 # Keep only the interesting columns
 DATA = DATA[keep_columns]
-
+'''
 
 
 # Shuffle the dataset then create a training and test set
 fraction_train = 0.8
 DATA = DATA.sample(frac=1)
-TRAIN = DATA.iloc[0:(int(fraction_train*len(DATA))),:].copy()
-TEST= DATA.iloc[(int(fraction_train*len(DATA))):,:].copy()
+TRAIN = DATA.iloc[0:(int(fraction_train*len(DATA))), :].copy()
+TEST = DATA.iloc[(int(fraction_train*len(DATA))):, :].copy()
 
 
 # Create a cv column in the training set
@@ -54,8 +54,8 @@ TRAIN['cv_fold'] = random.choices(list(range(0, 5)), k=len(TRAIN))
 
 
 # Save the training and test set
-TRAIN.to_csv('output/TRAIN.csv', index=False)
-TEST.to_csv('output/TEST.csv', index=False)
+TRAIN.to_csv('data/output/hmda_train.csv', index=False)
+TEST.to_csv('data/output/hmda_test.csv', index=False)
 
 
 

### 2019-11-04 article-information-2019.tex
adding some notation

@@
 % Full title of the paper (Capitalized)
-\Title{Responsible Machine Learning\\\large{Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}}
+\Title{Responsible Machine Learning\\\large{with Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Notation}
+
+To facilitate descriptions of data, modeling, explanatory, and social bias techniques, notation for input and output spaces, datasets, and models is defined.
+
+\subsubsection{Spaces}
+
+	\begin{itemize}
+		\item Input features come from the set $\mathcal{X}$ contained in a \textit{P}-dimensional input space, $\mathcal{X} \subset \mathbb{R}^P$.
+An arbitrary, potentially unobserved, or future instance of $\mathcal{X}$ is denoted $\mathbf{x}$, $\mathbf{x} \in \mathcal{X}$.
+		\item Labels corresponding to instances of $\mathcal{X}$ come from the set $\mathcal{Y}$.
+		\item Learned output responses come from the set $\mathcal{\hat{Y}}$.
+% For regression models, the set $\mathcal{\hat{Y}}_r$ is also contained in a $C$-dimensional output space, $\mathcal{\hat{Y}}_r \subset \mathbb{R}^{C_r}$.
+For classification models, the set $\mathcal{\hat{Y}}_c$ typically contains a column vector for each unique class in $\mathcal{Y}$.
+Hence, $\mathcal{\hat{Y}}_c$ is contained in a $C'$-dimensional output space,  $\mathcal{\hat{Y}}_c \subset \mathbb{R}^{C'_c}$.
+	\end{itemize}
+
+\subsubsection{Datasets}
+
+	\begin{itemize}
+		\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
+		\item Each $i$-th observation of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
+		$[x_0^{(i)}, x_1^{(i)}, \dots, x_{\textit{P}-1}^{(i)}]$, with corresponding $i$-th labels in $\mathbf{Y}, \mathbf{y}^{(i)}$, and corresponding predictions in $\mathbf{\hat{Y}}, \mathbf{\hat{y}}^{(i)}$.
+% = [y_0^{(i)}, y_1^{(i)}, \dots, y_{\textit{C}-1}^{(i)}]$.
+		\item $\mathbf{X}$ and $\mathbf{Y}$ consist of $N$ tuples of observations: $[(\mathbf{x}^{(0)},\mathbf{y}^{(0)}), (\mathbf{x}^{(1)},\mathbf{y}^{(1)}), \dots,(\mathbf{x}^{(N-1)},\mathbf{y}^{(N-1)})]$.
+%\\ $\mathbf{x}^{(i)} \in \mathcal{X}$, $\mathbf{y}^{(i)} \in \mathcal{Y}$.
+		\item Each $j$-th input column vector of $\mathbf{X}$ is denoted as $X_j = [x_{j}^{(0)}, x_{j}^{(1)}, \dots, x_{j}^{(N-1)}]^T$.
+	\end{itemize}
+
+\subsubsection{Models}
+
+	\begin{itemize}
+		\item A type of machine learning model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$:
+		$ \mathbf{X}, \mathbf{Y} \xrightarrow{\mathcal{A}} g$, such that $g \approx f$.
+		\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
+		\item The model to be explained is denoted as $g$.
+	\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Data Description}
@@
 \subsection{Model Description}
+
+\subsubsection{Explainable Neural Network}
+
+\subsubsection{Monotonically Constrained Gradient Boosting Machine}
+
+Monotonic gradient boosting machines (M-GBMs) constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints.
+The M-GBM remains an additive combination of many trees, $T_b$, but with a set of splitting rules that respect the monotonicity constraints,  $\Theta^\text{mono}_b$.
+
+\begin{equation}
+\begin{aligned}\label{eq:rf}
+g^{\text{mono}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
+\end{aligned}
+\end{equation}
+
+\noindent As in unconstrained GBM, $\Theta_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the M-GBM, $g^{\text{mono}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+For each $i$-th observation  and the $b$-th iteration the loss function, $\mathcal{L}_{i,b}$, can be defined as:
+
+\begin{equation}
+\begin{aligned}
+\mathcal{L}_{i,b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{mono}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
+\end{aligned}
+\end{equation}
+
+\noindent In addition to $\mathcal{L}_{i,b}$, $\Theta^\text{mono}_b$ is constrained by applying additional splitting rules for each binary split rule, $\theta_{b,j,k} \in \Theta_b$.
+Each $\theta_{b,j,k}$ is associated with a feature, $X_j$, and can be the $k$-th such split associated with $X_j$ in $T_b$.
+Each $\theta_{b,j,k}$ also results in left and right child nodes with a numeric weight, $\{w_{b,j,k,L}, w_{b,j,kR}\}$.
+For terminal nodes, each $w_{b,j,k}$ is essentially the model prediction.
+For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$, $\Theta_b$ is said to be positive monotonic if:
+
+\begin{enumerate}
+\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ causing the left child weight to be greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$ is not considered.
+\item For any subsequent left child node involving $X_j$, any $\theta_{b,j, 1+}$ causing $T(x_j; \theta_{b,j,1+}) = \{w_{b,j,1+,L}, w_{b,j,1+R}\}$ where $w_{b,j,1+,L} > w_{b,j,1+,R}$ is not considered.
+\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k+}) = \{w_{b,j,k,L}, w_{b,j,k,R}\}$ for $k > 0$, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ are bound by the preceding set of $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$ such that $ \{w_{b,j,k,L}, w_{b,j,k,R}\} \le \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $\{w_{b,j,k,L}, w_{b,j,k,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+
+\end{enumerate}
+
+\noindent Note that for any $X_j$ in any $g_{\text{mono}}~T_b$ left subtrees will alway produce lower predictions than right subtrees, and that any $g_{\text{mono}}(\mathbf{x})$ is a sequential addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classifications.
+Also note that each tree's root node will always obey monotonicity constraints, as $T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0})$, ensuring $T(x^{\alpha}_j; \theta_{b,j,0}) = w_{b,j,0,L} \le T(x^{\beta}_j; \theta_{b,j,0}) = w_{b,j,0,R}$.
+For negative monotonic constraints left and right splitting rules are switched, and tree pruning strategies can be applied.
+
+% Tree pruning?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Explanatory Method Description}
+
+\subsubsection{Partial Dependence and Individual Conditional Expectation}
+
+Partial dependence (PD) plots are a widely-used method for describing the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
+Individual conditional expectation (ICE) plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$.
+Partial dependence and ICE can be combined in the same plot to compensate for known weaknesses of partial dependence, to identify interactions modeled by $g$, and to create a holistic portrait of the predictions of a complex model for some $X_j$  \cite{ice_plots}.
+
+Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{(-j)} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{(-j)} = \mathbf{X}$) is considered.
+$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{(-j)}$ is left unchanged.
+$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{(-j)}$ are left untouched.
+Partial dependence and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$.
+
+\subsubsection{Shapley Values}
+
+Shapley explanations, including Tree SHAP (SHapley Additive exPlanations) , are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}.
+Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ and that Shapley explanation values can never decrease for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}.
+
+For some observation $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
+
+\begin{equation}
+\label{eq:shap_additive}
+\begin{aligned}
+g(\mathbf{x}) = \phi_0 + \sum_{j=0}^{j=\mathcal{P} - 1} \phi_j \mathbf{z}_j
+\end{aligned}
+\end{equation}
+
+\noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
+Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
+
+Shapley values can be estimated in different ways.
+Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal tree structures to estimate the impact of each $x_j$ for some $g(\mathbf{x})$ of interest \cite{tree_shap}.
+
+\begin{equation}
+\label{eq:shap_contrib}
+\begin{aligned}
+\phi_{j} = \sum_{S \subseteq \mathcal{P} \setminus \{j\}}\frac{|S|!(\mathcal{P} -|S| -1)!}{\mathcal{P}!}[g_x(S \cup \{j\}) - g_x(S)]
+\end{aligned}
+\end{equation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Social Bias Test Description}
 
@@
 \section{Discussion}
+% Bivariate monotonicity constraints likely dampen impact of interactions
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{}
+\acknowledgments{Wen Phan for work in formalizing our notation.}
 

### 2019-11-05 article-information-2019.tex
add description of XNN

@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{This text outlines a viable approach for training and evaluating complex machine learning systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (M-GBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated datasets with known feature importance and sociological bias characteristics and on realistic, publicly available example datasets.
-For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of individual conditional expectation (ICE) and global and local gradient-based or Shapley feature importance.
+\abstract{This text outlines a viable approach for training and evaluating machine learning systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (M-GBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and sociological bias characteristics and on realistic, publicly available lending data.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence and individual conditional expectation (ICE) and global and local gradient-based or Shapley feature importance.
 The constrained model predictions are also tested for disparate impact and other types of sociological bias using straightforward group fairness measures.
@@
 		\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
-		\item The model to be explained is denoted as $g$.
+		\item The model to be explained and tested for unwanted social bias is denoted as $g$.
 	\end{itemize}
@@
 \subsection{Model Description}
+% both additive
+% XNN smooth
 
 \subsubsection{Explainable Neural Network}
+
+Explainable neural networks (XNNs) are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
+XNNs also bare a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e.
+GAMs which consider main effects and a small number of 2-way interactions and incorporate boosting in their training \cite{esl}, \cite{ga2m}.
+Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
+XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum^J_{j=0}\beta_{k, j}x_j$.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:xnn}
+g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=1}^K\gamma_k n_k(\sum^J_{j=1}\beta_{k, j}x_j)
+\end{aligned}
+\end{equation}
+
+\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:
+
+\begin{enumerate}
+\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, is known as the \textit{projection layer} and is fully connected to each input feature, $X_j$.
+\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
+Each $n_k$ is a neural network, which can be parameterized to suite a given modeling task.
+To facilitate direct visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
+\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
+Again, subnetwork output is restricted to 1-dimension for visualization purposes.
+\end{enumerate}
+
+\noindent $g^{\text{XNN}}$ is typically trained by mini-batch stochastic gradient descent (SGD).
+$L_1$ regularization is often applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
 
@@
 Monotonic gradient boosting machines (M-GBMs) constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints.
-The M-GBM remains an additive combination of many trees, $T_b$, but with a set of splitting rules that respect the monotonicity constraints,  $\Theta^\text{mono}_b$.
+The M-GBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, but each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
 
@@
 
-\noindent As in unconstrained GBM, $\Theta_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the M-GBM, $g^{\text{mono}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
-For each $i$-th observation  and the $b$-th iteration the loss function, $\mathcal{L}_{i,b}$, can be defined as:
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the M-GBM, $g^{\text{mono}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
 
@@
 \begin{aligned}
-\mathcal{L}_{i,b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{mono}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
+\mathcal{L}_{b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{mono}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
 \end{aligned}
@@
 
-\noindent In addition to $\mathcal{L}_{i,b}$, $\Theta^\text{mono}_b$ is constrained by applying additional splitting rules for each binary split rule, $\theta_{b,j,k} \in \Theta_b$.
-Each $\theta_{b,j,k}$ is associated with a feature, $X_j$, and can be the $k$-th such split associated with $X_j$ in $T_b$.
-Each $\theta_{b,j,k}$ also results in left and right child nodes with a numeric weight, $\{w_{b,j,k,L}, w_{b,j,kR}\}$.
-For terminal nodes, each $w_{b,j,k}$ is essentially the model prediction.
-For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$, $\Theta_b$ is said to be positive monotonic if:
+\noindent In addition to $\mathcal{L}_{b}$, $g^\text{mono}$ training is characterized by additional splitting rules and constraints on tree node weights.
+Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,kR}\}$.
+For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,kR}\}$ can be direct numeric components of some $g^\text{mono}$ prediction.
+For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$, $\Theta_b$ is restricted to be positive monotonic w.r.t.
+$X_j$ by the following rules and constraints.
 
 \begin{enumerate}
-\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ causing the left child weight to be greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$ is not considered.
-\item For any subsequent left child node involving $X_j$, any $\theta_{b,j, 1+}$ causing $T(x_j; \theta_{b,j,1+}) = \{w_{b,j,1+,L}, w_{b,j,1+R}\}$ where $w_{b,j,1+,L} > w_{b,j,1+,R}$ is not considered.
-\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k+}) = \{w_{b,j,k,L}, w_{b,j,k,R}\}$ for $k > 0$, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ are bound by the preceding set of $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$ such that $ \{w_{b,j,k,L}, w_{b,j,k,R}\} \le \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
-\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $\{w_{b,j,k,L}, w_{b,j,k,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
-
+\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in the left child weight being greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
+\item For any subsequent left child node involving $X_j$, any $\theta_{b,j, k\ge1}$ resulting in $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ where $w_{b,j,k\ge1,L} > w_{b,j,k\ge1,R}$, is not considered.
+\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$, $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ are bound by the parent set of node weights, $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$, such that $ \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \le \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
 \end{enumerate}
 
-\noindent Note that for any $X_j$ in any $g_{\text{mono}}~T_b$ left subtrees will alway produce lower predictions than right subtrees, and that any $g_{\text{mono}}(\mathbf{x})$ is a sequential addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classifications.
-Also note that each tree's root node will always obey monotonicity constraints, as $T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0})$, ensuring $T(x^{\alpha}_j; \theta_{b,j,0}) = w_{b,j,0,L} \le T(x^{\beta}_j; \theta_{b,j,0}) = w_{b,j,0,R}$.
-For negative monotonic constraints left and right splitting rules are switched, and tree pruning strategies can be applied.
-
-% Tree pruning?
+\noindent Note that for any one $X_j$ and $T_b \in g^{\text{mono}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{mono}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
+Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,j,0}) = w_{b,0}$.
+Together these additional splitting rules and node weight constraints ensure that $g^{\text{mono}}(x^{\alpha}_j)  \le g^{\text{mono}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
+For a negative monotonic constraint, i.e.
+$g^{\text{mono}}(x^{\alpha}_j)  \ge g^{\text{mono}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
 
@@
 \section{Discussion}
+
+% richness of possible interpretable model architecture
+
+%scalability, easy computation of derivatives
+% XNN is quite sensitive to initialization, and ...
+% variable selection
+
 % Bivariate monotonicity constraints likely dampen impact of interactions
+% Tree pruning?
 

### 2019-11-06 article-information-2019.tex
wireframe notebooks, lower-case d

@@
 %=================================================================
-\documentclass[information,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[information,article,submit,moreauthors,pdftex]{definitions/mdpi}
 
@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
 %\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name
@@
 % Authors, for the paper (add full first names)
-\Author{Navdeep Gill $^{1, \ddagger}$, Patrick Hall $^{1,\ddagger,*}$, Kim Montgomery $^{1,\ddagger}$, and Nicholas Schmidt $^{2, \ddagger}$}
+\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger}$}
 
@@
 % Current address and/or shared authorship
-%\firstnote{Current address: Affiliation 3} 
-\secondnote{These authors contributed equally to this work.}
+\firstnote{All authors contributed equally to this work.} 
+%\secondnote{}
 % The commands \thirdnote{} till \eighthnote{} are available for further notes
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{This text outlines a viable approach for training and evaluating machine learning systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (M-GBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and sociological bias characteristics and on realistic, publicly available lending data.
-For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence and individual conditional expectation (ICE) and global and local gradient-based or Shapley feature importance.
-The constrained model predictions are also tested for disparate impact and other types of sociological bias using straightforward group fairness measures.
-By combining innovations in interpretable models, post-hoc explanation, and bias testing with accessible software tools, this text aims to provide a template workflow for important machine learning applications that require high accuracy and interpretability and low disparate impact.}
+\abstract{This text outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and sociological bias characteristics and on realistic, publicly available lending data.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local gradient-based or Shapley feature importance.
+The constrained model predictions are also tested for disparate impact (DI) and other types of sociological bias using straightforward group fairness measures.
+By combining innovations in interpretable models, post-hoc explanation, and bias testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and low disparate impact.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
+ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{See: ``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate social biases \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
+ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
+The authors make no claim that the interdependent issues of opaqueness, unwanted sociological bias, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
+the fairwashing of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{shokri2019privacy}.
+However, this text does present an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI testing implemented in widely available, free, and open source Python tools.
+Section \ref{sec:m_and_m} describes methods and materials herein, including simulated and collected training datasets, interpretable and constrained architectures used to train highly transparent models, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other unwanted sociological bias, and public and open source software resources.
+In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and DI testing results are also presented.
+Section \ref{sec:disc} then discusses some nuances of the presented modeling, explanation, and DI testing methods and results.
+Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML and also touches on the authors' future work.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Notation}
+\section{Materials and Methods}\label{sec:m_and_m}
 
-To facilitate descriptions of data, modeling, explanatory, and social bias techniques, notation for input and output spaces, datasets, and models is defined.
+Section \ref{sec:m_and_m} presents descriptions of notation, training data, ML models, post-hoc explanation techniques, DI testing methods, and software resources as follows:\\
+
+\vspace{-10pt}
+\textsection \ref{ssec:not} \textbf{Notation}: for spaces, datasets, and models.\\
+\indent \textsection \ref{ssec:data} \textbf{Training data}: simulated data with known feature importance and social bias characteristics and collected home mortgage data.\\
+\indent \textsection \ref{ssec:models} \textbf{ML models}: interpretable XNN and constrained MGBM models.\\
+\indent \textsection \ref{ssec:expl} \textbf{Post-hoc explanation techniques}: PD, ICE, and Shapley values.\\
+\indent \textsection \ref{ssec:di}  \textbf{DI testing methods}: .\\
+\indent \textsection \ref{ssec:soft} \textbf{Software resources}: GitHub repository for the results presented herein and utilized and useful Python packages.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Notation}\label{ssec:not}
+
+To facilitate descriptions of data and modeling, explanatory, and DI testing techniques, notation for input and output spaces, datasets, and models is defined here.
 
@@
 	\begin{itemize}
-		\item A type of machine learning model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$:
+		\item A type of ML model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$:
 		$ \mathbf{X}, \mathbf{Y} \xrightarrow{\mathcal{A}} g$, such that $g \approx f$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Data Description}
+\subsection{Data Description}\label{ssec:data}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Model Description}
+\subsection{Model Description}\label{ssec:models}
 % both additive
@@
 
-Explainable neural networks (XNNs) are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
+XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
 XNNs also bare a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e.
@@
 
-Monotonic gradient boosting machines (M-GBMs) constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints.
-The M-GBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, but each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
+MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints.
+The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, but each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
 
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the M-GBM, $g^{\text{mono}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the MGBM, $g^{\text{mono}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
 For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Explanatory Method Description}
+\subsection{Explanatory Method Description}\label{ssec:expl}
 
@@
 
-Partial dependence (PD) plots are a widely-used method for describing the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
-Individual conditional expectation (ICE) plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$.
-Partial dependence and ICE can be combined in the same plot to compensate for known weaknesses of partial dependence, to identify interactions modeled by $g$, and to create a holistic portrait of the predictions of a complex model for some $X_j$  \cite{ice_plots}.
+PD plots are a widely-used method for describing the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
+ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$.
+PD and ICE can be combined in the same plot to compensate for known weaknesses of PD, to identify interactions modeled by $g$, and to create a holistic portrait of the predictions of a complex model for some $X_j$  \cite{ice_plots}.
 
@@
 $\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{(-j)}$ are left untouched.
-Partial dependence and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$.
+PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Social Bias Test Description}
+\subsection{Social Bias Test Description}\label{ssec:di}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Resources}
+\subsection{Software Resources}\label{ssec:soft}
 % Description of code
 
+% github repo, main packges
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
+\section{Results}\label{sec:res}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-% richness of possible interpretable model architecture
+\section{Discussion}\label{sec:disc}
 
@@
 
+% somewhat aligned with United States regulation-mandated adverse action notices?
+
+%Interpretable models are also tested for disparate impact and other unwanted sociological bias and viable bias mitigation strategies are discussed.
+While the focus of this paper is not ML security, best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+\section{Conclusion}\label{sec:con}
 % Model debugging
@@
 % Appeal
+% local bias
+
+% Moreover, intersectional dangers at the nexus of sociological bias, opaque, and insecure decision-making systems: privacy dangers of explanations, fairwashing, local bias, proxies
+% accuracy, fairness, security are not static
 
@@
 %% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\}
+\abbreviations{The following abbreviations are used in this manuscript: ML -- machine learning, MGBM -- monotonic gradient boosting machine, XNN -- explainable neural network, PD -- partial dependence, ICE -- individual conditional expectation, DI -- disparate impact, SGD -- stochastic gradient descent.}
 

### 2019-11-06 article-information-2019.tex
fix all the typos I made when rushing before meeting

@@
 % Authors, for the paper (add full first names)
-\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger}$}
+\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,3,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger}$}
 
@@
 $^{1}$ \quad H2O.ai\\
-$^{2}$ \quad BLDS, LLC}
+$^{2}$ \quad BLDS, LLC\\
+$^{3}$ \quad George Washington University}
 

### 2019-11-06 article-information-2019.tex
minor change

@@
 \indent \textsection \ref{ssec:di}  \textbf{DI testing methods}: .\\
-\indent \textsection \ref{ssec:soft} \textbf{Software resources}: GitHub repository for the results presented herein and utilized and useful Python packages.
+\indent \textsection \ref{ssec:soft} \textbf{Software resources}: GitHub repository for the presented results; utilized and potentially useful Python packages.
 

### 2019-11-07 notebooks/scripts/xnn_fin2_info_short_run.py
initial checkin

@@ -1,0 +1,633 @@
+
+import numpy as np
+import pandas as pd
+import shap
+
+import keras
+from keras import backend
+from keras.layers import Activation, Add, Dense, Dropout, Input, Lambda, Concatenate
+from keras.models import Model
+
+#import plotly.plotly as py
+#import plotly.tools as tls
+#import matplotlib.pyplot as plt
+
+
+import subprocess
+import sys
+
+def install(library):
+    subprocess.call([sys.executable, "-m", "pip", "install", library])
+
+try:
+    import pydot
+except ImportError:
+    install('pydot')
+    import pydot
+    
+
+from timeit import default_timer as timer
+import tensorflow as tf
+from keras import backend as K
+
+lll="quick_info_first_"
+""" Changing MLP init """
+""" The previous init changes were lousy"""
+""" Let's try again """
+""" First two mlp's """
+
+seed = 12345
+seed = 4422
+np.random.seed(seed)
+
+my_init = keras.initializers.RandomUniform(seed=seed)
+
+
+def projection_initializer(shape, dtype=None):
+    print(shape)
+    inps = shape[0]
+    subs = shape[1]
+    if subs > pow(inps, 2) - 1:
+        raise ValueError("Currently we support only up to 2^features - 1 number of subnetworks.")
+    
+    weights = []
+    # TODO impl when subs > inps
+    for i in range(subs):
+        w = [0] * inps
+        w[i] = 1
+        weights.append(w)
+    return weights
+
+class XNN:
+    # define base model
+    def __init__(self, features, ridge_functions=3, arch=[20,12], bg_samples=100, seed=None, is_categorical=False):
+        self.seed = seed
+        self.bg_samples = bg_samples
+        self.is_categorical = is_categorical
+        
+        #
+        # Prepare model arch
+        #
+        
+        # Input to the network, our observation containing all the features
+        input = Input(shape=(features,), name='main_input')
+
+        init = keras.initializers.RandomUniform(minval=-5, maxval=5, seed=None)
+        # Input to ridge function number i is the dot product of our original input vector times coefficients
+        ridge_input = Dense(ridge_functions,
+                            name="projection_layer",
[…]

### 2019-11-07 article-information-2019.tex
several updates

@@
 The authors make no claim that the interdependent issues of opaqueness, unwanted sociological bias, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
-the fairwashing of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{shokri2019privacy}.
+the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.
 However, this text does present an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI testing implemented in widely available, free, and open source Python tools.
@@
 \subsection{Data Description}\label{ssec:data}
+
+Results are presented for simulated data with known feature importance and social bias characteristics and for more realistic lending data sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{\href{https://www.consumerfinance.gov/data-research/hmda/}{See: Mortgage data (HMDA)}.}
+
+\subsubsection{Simulated Data}
+
+\subsubsection{Lending Data}
+
+Lending data is sampled from the public HMDA database to create a real-world dataset on which to assess interpretable models, post-hoc explanations, and DI tests.
+The sampled lending data contains recent and anonymized consumer loans issued by many major United States (US) financial institutions.
+Consumers and loans in the dataset are characterized by Bayesian improved surname geocoding (BISG) demographic features, consumer credit scores and various consumer financial features, and loan purpose, amount, and term features.\footnote{\href{https://files.consumerfinance.gov/f/201409_cfpb_report_proxy-methodology.pdf}{See: Using publicly available information to proxy for unidentified race and ethnicity}.} The lending data is divided randomly into training and test partitions.
+The training data contains 33 total features and 144,000 rows, each representing a unique loan, and a fold identifier to ensure consistent 5-fold cross-validation accuracy and error measurements across different types of models.
+Consumer financials and loan descriptors are used for training.
+Demographic features are not used in model training.
+The lending test data contains 36,000 loans.
 
@@
 % Description of code
-
 % github repo, main packges
+Python code to reproduce the results presented in this text are available at: \url{https://github.com/h2oai/article-information-2019}.
+The authors primarily make use of the \href{https://github.com/h2oai/datatable}{datatable}, \href{https://github.com/h2oai/h2o-3}{h2o}, \href{https://matplotlib.org/}{matplotlib}, \href{https://pandas.pydata.org/}{pandas}, \href{https://scikit-learn.org/stable/}{scikit-learn}, \href{https://seaborn.pydata.org/}{seaborn}, and \href{https://github.com/slundberg/shap}{shap} packages for data manipulation, modeling, and reporting results.
 

### 2019-11-09 data/scripts/augmented_friedman1_creator.py
Rewrite simulated program; add simulated data.

@@ -3,55 +3,41 @@ import numpy as np
 import pandas as pd
 from sklearn import datasets as ds
 from scipy import stats
-import os
-import datetime
 
 
 if __name__ == "__main__":
 
-    nCorrNormal = 20
-    nNormal = 20
-    nObs = 100000
+    n_corr_normal = 20
+    n_normal = 20
+    n_obs = 100000
     intercept = -13.8955411
-    getsOfferPct = 0.5
-    pctFlippedByNoise = 0.20
-
-    todaysDate = datetime.date.today().strftime("%Y%m%d")
 
     # Each call of a random number generator will call a value from the list
     # of random seeds.
-    rnd = np.random.RandomState(61803)
-    randSeed = list(np.unique(rnd.randint(0, 1000000, 100)))
-    randIter = 0
+    np.random.seed(61803)
+    rand_seed_list = list(np.unique(np.random.randint(0, 1000000, 100)))
 
-    dataFolder = '/mnt/rdrive/Artificial Intelligence/article-information-2019/'
-    os.chdir(dataFolder)
-    
     # Creates matrix of uncorrelated normal variables
-    rnd = np.random.RandomState(randSeed[randIter])
-    varsNormal = pd.DataFrame(rnd.randn(nObs, nNormal), columns=['norm ' + str(x + 1) for x in range(nNormal)])
+    features_normal = pd.DataFrame(np.random.randn(n_obs, n_normal),
+                                   columns=['norm' + str(x + 1) for x in range(n_normal)])
 
     # Creates matrix of correlated normal variables, first by creating the
     # positive definite matrix, which serves as the variance-covariance matrix.
-    randIter = + 1
-    rnd = np.random.RandomState(randSeed[randIter])
-    corrNormalVarCovar = ds.make_spd_matrix(n_dim=nCorrNormal, random_state=randSeed[randIter])
+    var_covar_cor_normal = ds.make_spd_matrix(n_dim=n_corr_normal, random_state=rand_seed_list.pop())
+    np.random.seed(rand_seed_list.pop())
+    features_cor_normal = np.random.multivariate_normal(mean=[0] * n_corr_normal,
+                                                        cov=var_covar_cor_normal,
+                                                        size=n_obs)
 
[…]

### 2019-11-10 data/scripts/di_testing.py
Add DI Testing program (w/AIR)

@@ -1,0 +1,36 @@
+
+import h2o
+import pandas as pd
+import numpy as np
+from typing import Union
+
+
+class DisparateImpactTesting(object):
+    def __init__(self, lower_score_better: bool,
+                 pg_names: Union[list, tuple] = ('black', 'hispanic' 'asian', 'female', 'older'),
+                 cg_names: Union[list, tuple] = ('white', 'white', 'white', 'male', 'younger'),
+                 pgcg_names: Union[list, tuple] = ( 'black', 'hispanic', 'asian', 'white',
+                                                    'female', 'male', 'older', 'younger'),
+                 ):
+        self.pg_names = pg_names
+        self.cg_names = cg_names
+        self.lower_score_better = lower_score_better
+        self.pgcg_names = pgcg_names
+
+    def adverse_impact_ratio(self,
+                             data: Union[pd.DataFrame, h2o.H2OFrame],
+                             label: str,
+                             ):
+        di_cols = ["class", "ctrl_class", "total", "gets_offer", "pct", "air"]
+        di_table = pd.DataFrame(np.full(shape=(len(self.pgcg_names), len(di_cols)), fill_value=np.nan),
+                                index=self.pgcg_names, columns=di_cols)
+        di_table["class"] = self.pgcg_names
+        di_table.loc[di_table["class"].isin(self.pg_names), "ctrl_class"] = self.cg_names
+        di_table["gets_offer"] = (np.sum(data[label].to_numpy().reshape(-1, 1) * data[self.pgcg_names]))
+        di_table["total"] = data[self.pgcg_names].sum(axis=0)
+        di_table["pct"] = di_table["gets_offer"] / di_table["total"]
+        for cgi, pgi in zip(self.pg_names, self.cg_names):
+            di_table.loc[di_table["class"] == cgi, "air"] = di_table.loc[di_table["class"] == cgi, "pct"].to_numpy() / \
+                                                            di_table.loc[di_table["class"] == pgi, "pct"].to_numpy()
+        di_table.index = range(len(di_table))
+        return di_table

### 2019-11-10 data/scripts/augmented_friedman1_creator.py
Add DI Testing program (w/AIR), update simu data

@@ -3,12 +3,12 @@ import numpy as np
 import pandas as pd
 from sklearn import datasets as ds
 from scipy import stats
-
+from data.scripts.di_testing import DisparateImpactTesting
 
 if __name__ == "__main__":
 
-    n_corr_normal = 20
-    n_normal = 20
+    n_corr_normal = 20  # do not change; variable naming is specific to 20
+    n_normal = 20  # do not change; variable naming is specific to 20
     n_obs = 100000
     intercept = -13.8955411
 
@@ -71,7 +71,7 @@ if __name__ == "__main__":
     for vci in ["binary1", "binary2", "cat1"]:
         print(X[vci].value_counts(dropna=False, normalize=True))
 
-    dum = pd.get_dummies(X["cat1"], drop_first=False, prefix="cat1")
+    dum = pd.get_dummies(X["cat1"], drop_first=True, prefix="cat1")
     X = pd.concat([X, dum], sort=False, axis=1)
 
     # Friedman + binaries & categoricals, no noise:
@@ -136,8 +136,15 @@ if __name__ == "__main__":
                 or str.startswith(x, "cat1_")
                 or str.startswith(x, "other_uncor_norm")]
     X.drop(inplace=True, columns=droppers + ["other_cor_uni5", "other_cor_uni6"])
-
     simulated = pd.merge(label_set, X, left_index=True, right_index=True, how='inner')
+
+    pd.crosstab(simulated["binary1"], simulated["label_with_noise"], normalize=True, margins=True)
+    pd.crosstab(simulated["other_binary1"], simulated["label_with_noise"], normalize=True, margins=True)
+
+    simulated.rename(inplace=True, columns={'binary1': 'ctrl_class1',
+                                            'other_binary1': 'ctrl_class2'})
+    simulated["prot_class1"] = 1 - simulated["ctrl_class1"]
+    simulated["prot_class2"] = 1 - simulated["ctrl_class2"]
 
     simu_train = simulated[:int(round(n_obs) * (4 / 5))].copy()
     np.random.seed(rand_seed_list.pop())
@@ -146,3 +153,10 @@ if __name__ == "__main__":
 
     simu_train.to_csv("./data/output/simu_train.csv")
     simu_test.to_csv("./data/output/simu_test.csv")
+
+    di = DisparateImpactTesting(lower_value_favorable=True,
+                                pgcg_names=["prot_class1", "ctrl_class1", "prot_class2", "ctrl_class2"],
[…]

### 2019-11-10 data/scripts/di_testing.py

@@ -6,31 +6,74 @@ from typing import Union
 
 
 class DisparateImpactTesting(object):
-    def __init__(self, lower_score_better: bool,
+    def __init__(self, lower_value_favorable: bool,
                  pg_names: Union[list, tuple] = ('black', 'hispanic' 'asian', 'female', 'older'),
                  cg_names: Union[list, tuple] = ('white', 'white', 'white', 'male', 'younger'),
-                 pgcg_names: Union[list, tuple] = ( 'black', 'hispanic', 'asian', 'white',
-                                                    'female', 'male', 'older', 'younger'),
+                 pgcg_names: Union[list, tuple] = ('black', 'hispanic', 'asian', 'white',
+                                                   'female', 'male', 'older', 'younger'),
                  ):
+        """
+        This class creates metrics used to test disparate impact.  See each method for detail on the
+        particular calculations.
+
+        :param lower_value_favorable: Boolean specifying whether a lower outcome value being tested
+        is considered favorable.  For example, a marketing offer would generally be associated with a score
+        where a higher value is favorable.  For credit, generally a lower score is favorable, as the score
+        is typically a measure of the probability of default.
+
+        :param pg_names: list of names of the variables that indicate whether an individual is a member of a
+        protected group.
+
+        :param cg_names: list of names of the variables that indicate whether an individual is a member of a
+        control groups. These must correspond to the protected group values in "pg_names."  If a control group is
+        used for more than one comparison (e.g., Non-Hispanic Whites as compared to Blacks, Hispanics, or Asians),
+        then that control group variable should be repeated for each instance of an associated protected group in
+        "pg_names".  For example:
+            pg_names = ("black", "hispanic", "asian)
+            cg_names = ("white", "white", "white")
+
+        :param pgcg_names: This is a unique list of the values contained in pg_names and cg_names.  It is only used
+        as a way to make the output more readable: it specifies the order in which each group is displayed.
+        """
         self.pg_names = pg_names
         self.cg_names = cg_names
-        self.lower_score_better = lower_score_better
+        self.lower_value_favorable = lower_value_favorable
         self.pgcg_names = pgcg_names
 
     def adverse_impact_ratio(self,
[…]

### 2019-12-01 data/scripts/augmented_friedman1_creator.py
Fixing simulation dataset (not done).

@@ -1,4 +1,3 @@
-
 """
 License
 Copyright 2019 Navdeep Gill, Patrick Hall, Kim Montgomery, Nick Schmidt
@@ -20,11 +19,11 @@ from scipy import stats
 from data.scripts import di_testing
 from sklearn.metrics import roc_auc_score, classification_report
 from sklearn.ensemble import RandomForestClassifier
+import matplotlib.pyplot as plt
 
 pd.set_option('display.max_rows', 500)
 pd.set_option('display.max_columns', 500)
 pd.set_option('display.width', 1000)
-
 
 if __name__ == "__main__":
 
@@ -43,7 +42,7 @@ if __name__ == "__main__":
     # Test uncorrelated version
     # var_covar_cor_normal = np.identity(n=n_corr_normal)
     #####################################################
-    
+
     np.random.seed(rand_seed_list.pop())
     features_cor_normal = pd.DataFrame(np.random.multivariate_normal(mean=[0] * n_corr_normal,
                                                                      cov=var_covar_cor_normal,
@@ -81,6 +80,7 @@ if __name__ == "__main__":
     friedman_weights = [10., 20., 10., 5.]
     binary_categorical_weights = [1.0, -1.5, 2.0, -2.5, 1.25, -3.0]
     coefficients = pd.Series(friedman_weights + binary_categorical_weights, index=names_weights)
+    coefficients = coefficients.divide(10)
 
     # Friedman-Only Variables:
     contrib = pd.DataFrame(np.full(shape=(len(X), len(names_weights)), fill_value=np.nan),
@@ -112,20 +112,33 @@ if __name__ == "__main__":
     X["intercept"] = intercept
 
     np.random.seed(rand_seed_list.pop())
-    X["logistic_noise"] = np.random.logistic(scale=24, size=len(X))
-    X["logistic_noise"] = X["logistic_noise"] - X["logistic_noise"].mean()
-    print(f'Logistic Random Noise Distribution Characteristics:\n{X["logistic_noise"].agg(["mean", "std"])}')
-    X['latent_with_noise'] = X["latent_no_noise"] + X["logistic_noise"]
-    X["latent_no_noise"].hist()
-    X["latent_with_noise"].hist()
-    X["logistic_noise"].hist()
-    X["outcome"] = pd.DataFrame(np.where(X["latent_with_noise"] > 0, 1, 0))
[…]

### 2019-12-01 data/scripts/augmented_friedman1_creator.py
Fixing simulation dataset (not done).

@@ -80,7 +80,7 @@ if __name__ == "__main__":
     friedman_weights = [10., 20., 10., 5.]
     binary_categorical_weights = [1.0, -1.5, 2.0, -2.5, 1.25, -3.0]
     coefficients = pd.Series(friedman_weights + binary_categorical_weights, index=names_weights)
-    coefficients = coefficients.divide(10)
+    coefficients = coefficients.divide(5)
 
     # Friedman-Only Variables:
     contrib = pd.DataFrame(np.full(shape=(len(X), len(names_weights)), fill_value=np.nan),
@@ -167,7 +167,7 @@ if __name__ == "__main__":
     X = X[['outcome', "outcome_no_noise"] +
           ["fried" + str(x) for x in range(1, 6)] +
           ["binary1", "binary2", "cat1", 'intercept', "prot_class1", "ctrl_class1", "prot_class2", "ctrl_class2",
-           "latent_no_noise", "logistic_noise", "latent_with_noise"]]
+           "latent_no_noise", "logistic_noise", "latent_with_noise", "random_flip_outcome"]]
     print("\n", X.head())
 
     X_train = X[:int(round(n_obs) * (4 / 5))].copy()

### 2019-12-01 data/scripts/augmented_friedman1_creator.py
Fixing simulation data.

@@ -20,6 +20,7 @@ from data.scripts import di_testing
 from sklearn.metrics import roc_auc_score, classification_report
 from sklearn.ensemble import RandomForestClassifier
 import matplotlib.pyplot as plt
+from sklearn.model_selection import train_test_split
 
 pd.set_option('display.max_rows', 500)
 pd.set_option('display.max_columns', 500)
@@ -112,22 +113,25 @@ if __name__ == "__main__":
     X["intercept"] = intercept
 
     np.random.seed(rand_seed_list.pop())
-    # X["logistic_noise"] = np.random.logistic(scale=0.5, size=len(X))
-    # X["logistic_noise"] = (X["logistic_noise"] - X["logistic_noise"].mean())  # / X["logistic_noise"].std()
-    # print(f'\nLogistic Random Noise Distribution Characteristics:\n{X["logistic_noise"].agg(["mean", "std"])}')
-    # X['latent_with_noise'] = X["latent_no_noise"] + X["logistic_noise"]
-    # X["latent_with_noise"].describe()
-    # X["latent_with_noise"].hist(alpha=0.5)
-    # X["logistic_noise"].hist(alpha=0.5)
+    X["logistic_noise"] = np.random.logistic(scale=0.5, size=len(X))
+    X["logistic_noise"] = (X["logistic_noise"] - X["logistic_noise"].mean())  # / X["logistic_noise"].std()
+    print(f'\nLogistic Random Noise Distribution Characteristics:\n{X["logistic_noise"].agg(["mean", "std"])}')
+    X['latent_with_noise'] = X["latent_no_noise"] + X["logistic_noise"]
+    X["latent_with_noise"].describe()
+    X["latent_with_noise"].hist(alpha=0.5)
+    X["logistic_noise"].hist(alpha=0.5)
+    X["outcome"] = pd.DataFrame(np.where(X["latent_with_noise"] > 0, 1, 0))
 
-    # X["outcome"] = pd.DataFrame(np.where(X["latent_with_noise"] > 0, 1, 0))
     X["outcome_no_noise"] = pd.DataFrame(np.where(X["latent_no_noise"] > 0, 1, 0))
-    np.random.seed(rand_seed_list.pop())
-    X["outcome"] = X["outcome_no_noise"]
-    X["random_flip_outcome"] = np.random.rand(len(X)) >= 0.10
[…]

### 2019-12-01 data/scripts/hmda_sample_for_paper.py

@@ -1,4 +1,3 @@
-
 """
 License
 Copyright 2019 Navdeep Gill, Patrick Hall, Kim Montgomery, Nick Schmidt
@@ -15,1503 +14,1505 @@ DISCLAIMER: This notebook is not legal compliance advice.
 
 import numpy as np
 import pandas as pd
-from sklearn.model_selection import train_test_split
+from scipy import optimize
+import matplotlib.pyplot as plt
+
 
 def get_hmda_descriptions(data):
     data['conforming_loan_limit_desc'] = pd.Series(index=data.index, data=
-        np.select(
-            [
-                data.conforming_loan_limit == 'C',
-                data.conforming_loan_limit == 'NC',
-                data.conforming_loan_limit == 'U',
-                data.conforming_loan_limit == 'NA'
-            ],
-            [
-                'Conforming',
-                'Nonconforming',
-                'Undetermined',
-                'Not applicable'
-            ],
-            ''
-        ),
-        dtype='category'
-    )
+    np.select(
+        [
+            data.conforming_loan_limit == 'C',
+            data.conforming_loan_limit == 'NC',
+            data.conforming_loan_limit == 'U',
+            data.conforming_loan_limit == 'NA'
+        ],
+        [
+            'Conforming',
+            'Nonconforming',
+            'Undetermined',
+            'Not applicable'
+        ],
+        ''
+    ),
+                                                   dtype='category'
+                                                   )
     data['action_taken_desc'] = pd.Series(index=data.index, data=
-        np.select(
-            (
-                data.action_taken == 1, data.action_taken == 2, data.action_taken == 3,
-                data.action_taken == 4, data.action_taken == 5, data.action_taken == 6,
-                data.action_taken == 7, data.action_taken == 8
-            ),
-            (
-                'Loan originated', 'Application approved but not accepted',
-                'Application denied', 'Application withdrawn by applicant',
-                'File closed for incompleteness', 'Purchased loan',
-                'Preapproval request denied',
-                'Preapproval request approved but not accepted'
-            ),
-            default=''
+    np.select(
+        (
+            data.action_taken == 1, data.action_taken == 2, data.action_taken == 3,
+            data.action_taken == 4, data.action_taken == 5, data.action_taken == 6,
+            data.action_taken == 7, data.action_taken == 8
         ),
-        dtype='category'
-    )
+        (
+            'Loan originated', 'Application approved but not accepted',
+            'Application denied', 'Application withdrawn by applicant',
[…]

### 2019-12-01 data/scripts/hmda_sample_for_paper.py
Add test of features and outcome for use in estimating monotonic constraints.

@@ -14,8 +14,6 @@ DISCLAIMER: This notebook is not legal compliance advice.
 
 import numpy as np
 import pandas as pd
-from scipy import optimize
-import matplotlib.pyplot as plt
 
 
 def get_hmda_descriptions(data):
@@ -134,22 +132,6 @@ def get_hmda_descriptions(data):
     ),
                                               dtype='category'
                                               )
-    data['open_end_line_of_credit_desc'] = pd.Series(index=data.index, data=
-    np.select(
-        (
-            data.open_end_line_of_credit == 1,
-            data.open_end_line_of_credit == 2,
-            data.open_end_line_of_credit == 1111
-        ),
-        (
-            'Open-end line of credit',
-            'Not an open-end line of credit',
-            'Exempt'
-        ),
-        default=''
-    ),
-                                                     dtype='category'
-                                                     )
     data['business_or_commercial_purpose_desc'] = pd.Series(index=data.index, data=
     np.select(
         (
@@ -174,6 +156,22 @@ def get_hmda_descriptions(data):
     ),
                                           dtype='category'
                                           )
+    data['open_end_line_of_credit_desc'] = pd.Series(index=data.index, data=
+    np.select(
+        (
+            data.open_end_line_of_credit == 1,
+            data.open_end_line_of_credit == 2,
+            data.open_end_line_of_credit == 1111
+        ),
+        (
+            'Open-end line of credit',
+            'Not an open-end line of credit',
+            'Exempt'
+        ),
+        default=''
+    ),
+                                                     dtype='category'
+                                                     )
     data['negative_amortization_desc'] = pd.Series(index=data.index, data=
     np.select(
         (
@@ -1681,16 +1679,19 @@ if __name__ == '__main__':
     train.to_csv('./data/output/hmda_train.csv')
     test.to_csv('./data/output/hmda_test.csv')
 
-    ##################################
-    ##################################
-    ##################################
+    # Analysis for Implementation of Monotonic Constraints:
     train = pd.read_csv('./data/output/hmda_train.csv')
+    features = ['loan_amount', 'loan_to_value_ratio', 'no_intro_rate_period', 'intro_rate_period',
[…]

### 2019-12-02 article-information-2019.tex
progress on model and explanation intros and explanation text

--- 2019-11-06 article-information-2019.tex
@@
 \abstract{This text outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and sociological bias characteristics and on realistic, publicly available lending data.
-For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local gradient-based or Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of sociological bias using straightforward group fairness measures.
-By combining innovations in interpretable models, post-hoc explanation, and bias testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and low disparate impact.}
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on realistic, publicly available mortgage data.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and \textcolor{red}{global and local gradient-based or Shapley feature importance}.
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using straightforward group fairness measures.
+By combining innovations in interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}
 
@@
 
-ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{See: ``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate social biases \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
+ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
-The authors make no claim that the interdependent issues of opaqueness, unwanted sociological bias, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
-the fairwashing of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{shokri2019privacy}.
-However, this text does present an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI testing implemented in widely available, free, and open source Python tools.
-Section \ref{sec:m_and_m} describes methods and materials herein, including simulated and collected training datasets, interpretable and constrained architectures used to train highly transparent models, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other unwanted sociological bias, and public and open source software resources.
-In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and DI testing results are also presented.
-Section \ref{sec:disc} then discusses some nuances of the presented modeling, explanation, and DI testing methods and results.
+The authors make no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
+the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.
+However, this text is an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools.
+Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other discrimination, and public and open source software resources.
+In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and discrimination testing results are also presented for interpretable models.
+Section \ref{sec:disc} then discusses some nuances of the presented modeling, explanation, and discrimination testing methods and results.
 Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML and also touches on the authors' future work.
@@
 
-Section \ref{sec:m_and_m} presents descriptions of notation, training data, ML models, post-hoc explanation techniques, DI testing methods, and software resources as follows:\\
+To provide a sense of accuracy differences, this text compares the performance of more interpretable constrained ML models and less interpretable unconstrained ML models on simulated data and on collected mortgage data.
+The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1991multivariate}.
+The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a more realistic template workflow for future users of these methods and tools.
+Unconstrained gradient boosting machines (GBMs) (e.g.
+\cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
+\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}) are deemed uninterpretable black-box ML models herein.
 
-\vspace{-10pt}
-\textsection \ref{ssec:not} \textbf{Notation}: for spaces, datasets, and models.\\
-\indent \textsection \ref{ssec:data} \textbf{Training data}: simulated data with known feature importance and social bias characteristics and collected home mortgage data.\\
-\indent \textsection \ref{ssec:models} \textbf{ML models}: interpretable XNN and constrained MGBM models.\\
-\indent \textsection \ref{ssec:expl} \textbf{Post-hoc explanation techniques}: PD, ICE, and Shapley values.\\
-\indent \textsection \ref{ssec:di}  \textbf{DI testing methods}: .\\
-\indent \textsection \ref{ssec:soft} \textbf{Software resources}: GitHub repository for the results presented herein and utilized and useful Python packages.
+Because black-box ML models can be difficult to understand, trust, and appeal, even after the application of post-hoc explanation techniques, explanation analysis and discrimination testing are applied only to the constrained interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
+Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} and XNNs (\cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analysis.
+Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}.
+PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
+\textcolor{red}{Discussed discrimination testing methods and metrics include ...
+}.
+All discussed materials and methods are implemented with open source Python packages and code and are available in the software resources associated with this text.
+Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:\\
+
+\textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}\\
+\indent\textbf{Training data}: simulated data \& collected home mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}\\
+\indent\textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}\\
+\indent\textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}\\
+\indent\textbf{Discrimination testing methods}: -- \textsection\ref{ssec:di}\\
+\indent\textbf{Software resources}: GitHub repository for this text; utilized \& useful Python packages -- \textsection\ref{ssec:soft}
 
@@
 
-To facilitate descriptions of data and modeling, explanatory, and DI testing techniques, notation for input and output spaces, datasets, and models is defined here.
+To facilitate descriptions of data and modeling, explanatory, and discrimination testing techniques, notation for input and output spaces, datasets, and models is defined.
 
@@
 		\item Labels corresponding to instances of $\mathcal{X}$ come from the set $\mathcal{Y}$.
-		\item Learned output responses come from the set $\mathcal{\hat{Y}}$.
+		\item Learned output responses of models are contained in the set $\mathcal{\hat{Y}}$.
 % For regression models, the set $\mathcal{\hat{Y}}_r$ is also contained in a $C$-dimensional output space, $\mathcal{\hat{Y}}_r \subset \mathbb{R}^{C_r}$.
@@
 		\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
-		\item The model to be explained and tested for unwanted social bias is denoted as $g$.
+		\item The model to be explained and tested for discrimination testing is denoted as $g$.
 	\end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Data Description}\label{ssec:data}
+\subsection{Simulated Data}\label{ssec:sim_data}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Model Description}\label{ssec:models}
-% both additive
-% XNN smooth
+\subsection{Mortgage Data}\label{ssec:mort_data}
 
-\subsubsection{Explainable Neural Network}
+The training data contains 33 total features and 144,000 rows, each representing a unique loan, and a fold identifier to ensure consistent 5-fold cross-validation accuracy and error measurements across different types of models.
+Consumer finance and loan descriptors are used for training.
+Demographic features are not used in model training.
+The mortgage test data contains 36,000 loans.
 
-XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
-XNNs also bare a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e.
-GAMs which consider main effects and a small number of 2-way interactions and incorporate boosting in their training \cite{esl}, \cite{ga2m}.
-Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
-XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum^J_{j=0}\beta_{k, j}x_j$.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:xnn}
-g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=1}^K\gamma_k n_k(\sum^J_{j=1}\beta_{k, j}x_j)
-\end{aligned}
-\end{equation}
-
-\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:
-
-\begin{enumerate}
-\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, is known as the \textit{projection layer} and is fully connected to each input feature, $X_j$.
-\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
-Each $n_k$ is a neural network, which can be parameterized to suite a given modeling task.
-To facilitate direct visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
-\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
-Again, subnetwork output is restricted to 1-dimension for visualization purposes.
-\end{enumerate}
-
-\noindent $g^{\text{XNN}}$ is typically trained by mini-batch stochastic gradient descent (SGD).
-$L_1$ regularization is often applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
-
-\subsubsection{Monotonically Constrained Gradient Boosting Machine}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Monotonically Constrained Gradient Boosting Machine}\label{ssec:mgbm}
 
@@
 For a negative monotonic constraint, i.e.
-$g^{\text{mono}}(x^{\alpha}_j)  \ge g^{\text{mono}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
+$g^{\text{mono}}(x^{\alpha}_j)  \ge g^{\text{mono}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.\\
+
+Herein, two $g^\text{mono}$ models are trained.
+One on the simulated data and one on the mortgage data.
+In both cases, positive and negative monotonic constraints for each $X_j$ are selected using the sign of the Pearson correlation between each $X_j$ and $y$, random grid search is used to determine other hyperparameters, and five-fold cross validation and test partitions are used for model assessment.
+For exact parameterization of the MGBM models, see the software resources referenced in Subsection \ref{ssec:soft}.
+
+\subsection{Explainable Neural Network}\label{ssec:xnn}
+
+XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
+XNNs also bare a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e.
+GAMs which consider main effects and a small number of 2-way interactions and incorporate boosting in their training \cite{esl}, \cite{ga2m}.
+Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
+XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum^J_{j=0}\beta_{k, j}x_j$.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:xnn}
+g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=1}^K\gamma_k n_k(\sum^J_{j=1}\beta_{k, j}x_j)
+\end{aligned}
+\end{equation}
+
+\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:
+
+\begin{enumerate}
+\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, is known as the \textit{projection layer} and is fully connected to each input feature, $X_j$.
+\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
+Each $n_k$ is a neural network, which can be parameterized to suite a given modeling task.
+To facilitate direct visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
+\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
+Again, subnetwork output is restricted to 1-dimension for visualization purposes.
+\end{enumerate}
+
+Here, each $g^\text{XNN}$ is trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
+Each $g^\text{XNN}$ is assessed in five training folds and in a test data partition.
+$L_1$ regularization is applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
+The $g^\text{XNN}$ models are highly sensitive to weight initialization and manual feature selection is informed by Shapley feature importance from the $g^\text{mono}$.
+For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Explanatory Method Description}\label{ssec:expl}
+\subsection{Partial Dependence and Individual Conditional Expectation}\label{ssec:pd_ice}
 
-\subsubsection{Partial Dependence and Individual Conditional Expectation}
-
-PD plots are a widely-used method for describing the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
-ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$.
-PD and ICE can be combined in the same plot to compensate for known weaknesses of PD, to identify interactions modeled by $g$, and to create a holistic portrait of the predictions of a complex model for some $X_j$  \cite{ice_plots}.
+PD plots are a widely-used method for describing and plotting the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
+ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$ \cite{ice_plots}.
+PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g.
+inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions of a complex model for some $X_j$ \cite{ice_plots}.
 
@@
 $\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{(-j)}$ are left untouched.
-PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$.
+PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
 
-\subsubsection{Shapley Values}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Shapley Values}\label{ssec:shap}
 
-Shapley explanations, including Tree SHAP (SHapley Additive exPlanations) , are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}.
-Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ and that Shapley explanation values can never decrease for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}.
+Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}, \cite{shapley1988shapley}.
+Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values can never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
+Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing the same model with augmented training data \cite{molnar}.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific and personalized adverse action notices for automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.
 
-For some observation $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
+For some instance $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
 
@@
 Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
-
-Shapley values can be estimated in different ways.
-Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal tree structures to estimate the impact of each $x_j$ for some $g(\mathbf{x})$ of interest \cite{tree_shap}.
+Local, per-instance explanations using Shapley values tend to involve presenting a ranking of $x_j$ by $\phi_j$ values or simply presenting a set of the $x_j$ associated with the largest $\phi_j$ for some $\mathbf{x}$.
+Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across the observations in some partition of data $\mathbf{X}$.
+Each $\phi_j$ is a weighted combination of model scores on the $g_x(\mathbf{x})$ with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$.
 
@@
 
+\noindent Where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
+
+Shapley values can be estimated in different ways, many of which are intractable for datasets with large $\mathcal{P}$.
+Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
+Tree SHAP has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this text.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Social Bias Test Description}\label{ssec:di}
+\subsection{Discrimination Metrics and Test Description}\label{ssec:di}
 
@@
 % Description of code
+% github repo, main packges
+Python code to reproduce the results presented in this text are available at: \url{https://github.com/h2oai/article-information-2019}.
+The authors primarily make use of the \href{https://github.com/h2oai/datatable}{datatable}, \href{https://github.com/h2oai/h2o-3}{h2o}, \href{https://matplotlib.org/}{matplotlib}, \href{https://pandas.pydata.org/}{pandas}, \href{https://scikit-learn.org/stable/}{scikit-learn}, \href{https://seaborn.pydata.org/}{seaborn}, and \href{https://github.com/slundberg/shap}{shap} packages for data manipulation, modeling, and reporting results.
 
-% github repo, main packges
+% interesting packages
 
@@
 
+\subsubsection{Unconstrained Model Fit Assessment}
+
+\subsubsection{Constrained Model Fit Assessment}
+
+\subsubsection{Interpretability and Post-hoc Explanation Results}
+
+\subsubsection{Discrimination Testing Results}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Loan Data Results}
+\subsection{Mortgage Data Results}
+
+\subsubsection{Unconstrained Model Fit Assessment}
+
+\subsubsection{Constrained Model Fit Assessment}
+
+\subsubsection{Interpretability and Post-hoc Explanation Results}
+
+\subsubsection{Discrimination Testing Results}
 
@@
 
+\subsection{The Burgeoning Ecosystem of Interpretable Models}
+
+% these are just examples of good models
 %scalability, easy computation of derivatives
@@
 
-% Bivariate monotonicity constraints likely dampen impact of interactions
+% both additive
+% XNN smooth
+
+% Bivariate monotonicity constraints likely dampen impact of interactions - same for XNN
 % Tree pruning?
+\subsection{Impact of Discrimination Testing on Model Use and Adoption}
+
+\subsection{Viable Discrimination Remediation Approaches}
 
 % somewhat aligned with United States regulation-mandated adverse action notices?
+
+\subsection{Intersectionality of Interpretability, Fairness, and Security in ML}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{, N.G.; , P.H.; , K.M.; , N.S.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+\authorcontributions{N.G., GBM and MGBM assessment and results; P.H., primary author; K.M., ANN and XNN implementation, assessment, and results; N.S., secondary author, data simulation and collection, and discrimination testing.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
 The following statements should be used ``conceptualization, X.X.
@@
 %% optional
-\abbreviations{The following abbreviations are used in this manuscript: ML -- machine learning, MGBM -- monotonic gradient boosting machine, XNN -- explainable neural network, PD -- partial dependence, ICE -- individual conditional expectation, DI -- disparate impact, SGD -- stochastic gradient descent.}
+\abbreviations{The following abbreviations are used in this text: ANN -- artificial neural network, DI -- disparate impact, GBM -- gradient boosting machine, ICE -- individual conditional expectation, MGBM -- monotonic gradient boosting machine, ML -- machine learning, PD -- partial dependence, SGD -- stochastic gradient descent, US -- United States, XNN -- explainable neural network.}
 

### 2019-12-03 article-information-2019.tex
progress on discussion section

@@
 % Full title of the paper (Capitalized)
-\Title{Responsible Machine Learning\\\large{with Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}}
+\Title{Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}}
 
@@
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
-The authors make no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
+This text makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
 the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.
-However, this text is an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools.
+This text does however propose some workable countermeasures to address these vexing problems in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools.
 Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other discrimination, and public and open source software resources.
 In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and discrimination testing results are also presented for interpretable models.
-Section \ref{sec:disc} then discusses some nuances of the presented modeling, explanation, and discrimination testing methods and results.
-Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML and also touches on the authors' future work.
+Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results.
+Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML.
 
@@
 The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1991multivariate}.
-The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a more realistic template workflow for future users of these methods and tools.
-Unconstrained gradient boosting machines (GBMs) (e.g.
+The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a more realistic template workflow for future users of similar methods and tools.
+
+Because unconstrained ML models, like gradient boosting machines (GBMs) (e.g.
 \cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
-\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}) are deemed uninterpretable black-box ML models herein.
-
-Because black-box ML models can be difficult to understand, trust, and appeal, even after the application of post-hoc explanation techniques, explanation analysis and discrimination testing are applied only to the constrained interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
+\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), can be difficult to understand, trust, and appeal, even after the application of post-hoc explanation techniques, explanation analysis and discrimination testing are applied only to the constrained interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
 Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} and XNNs (\cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analysis.
@@
 }.
-All discussed materials and methods are implemented with open source Python packages and code and are available in the software resources associated with this text.
-Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:\\
+All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated with this text.
+Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:
 
-\textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}\\
-\indent\textbf{Training data}: simulated data \& collected home mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}\\
-\indent\textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}\\
-\indent\textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}\\
-\indent\textbf{Discrimination testing methods}: -- \textsection\ref{ssec:di}\\
-\indent\textbf{Software resources}: GitHub repository for this text; utilized \& useful Python packages -- \textsection\ref{ssec:soft}
+\begin{itemize}
+	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
+	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
+	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
+	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
+	\item \textbf{Discrimination testing methods}: -- \textsection\ref{ssec:di}
+	\item \textbf{Software resources}: GitHub repository for this text; utilized \& useful Python packages -- \textsection\ref{ssec:soft}
+\end{itemize}
 
@@
 \begin{aligned}\label{eq:rf}
-g^{\text{mono}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
+g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
 \end{aligned}
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the MGBM, $g^{\text{mono}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the MGBM, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
 For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
@@
 \begin{aligned}
-\mathcal{L}_{b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{mono}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
+\mathcal{L}_{b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{MGBM}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
 \end{aligned}
@@
 
-\noindent In addition to $\mathcal{L}_{b}$, $g^\text{mono}$ training is characterized by additional splitting rules and constraints on tree node weights.
+\noindent In addition to $\mathcal{L}_{b}$, $g^\text{MGBM}$ training is characterized by additional splitting rules and constraints on tree node weights.
 Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,kR}\}$.
-For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,kR}\}$ can be direct numeric components of some $g^\text{mono}$ prediction.
+For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,kR}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction.
 For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$, $\Theta_b$ is restricted to be positive monotonic w.r.t.
@@
 
-\noindent Note that for any one $X_j$ and $T_b \in g^{\text{mono}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{mono}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
+\noindent Note that for any one $X_j$ and $T_b \in g^{\text{MGBM}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
 Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,j,0}) = w_{b,0}$.
-Together these additional splitting rules and node weight constraints ensure that $g^{\text{mono}}(x^{\alpha}_j)  \le g^{\text{mono}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
+Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
 For a negative monotonic constraint, i.e.
-$g^{\text{mono}}(x^{\alpha}_j)  \ge g^{\text{mono}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.\\
+$g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.\\
 
-Herein, two $g^\text{mono}$ models are trained.
+Herein, two $g^\text{MGBM}$ models are trained.
 One on the simulated data and one on the mortgage data.
@@
 For exact parameterization of the MGBM models, see the software resources referenced in Subsection \ref{ssec:soft}.
+Also consider that MGBM models with one-dimensional monotonicity constraints between some $X_j$ and $\hat{Y}$ likely restrict non-monotonic interactions between multiple $X_j$.
+Moreover, if monotonicity constraints are not applied to all $X_j \in \mathbf{X}$, any strong non-monotonic signal in training data associated with some important $X_j$ maybe forced onto some other arbitrary unconstrained $X_j$ under some $g^{MGBM}$ models, compromising the end goal of interpretability.
 
@@
 $L_1$ regularization is applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
-The $g^\text{XNN}$ models are highly sensitive to weight initialization and manual feature selection is informed by Shapley feature importance from the $g^\text{mono}$.
+The $g^\text{XNN}$ models are highly sensitive to weight initialization and manual feature selection is informed by Shapley feature importance from the $g^\text{MGBM}$.
 For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft}.
+Be aware that $g^{XNN}$ model architectures may restrict important interactions, appear highly sensitive to initialization routines, and require manual and judicious feature selection due to burdensome training times.
 
@@
 PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
+Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
+PD should always be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.
 
@@
 Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values can never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
-Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing the same model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific and personalized adverse action notices for automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.
-
 For some instance $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
@@
 Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
-Local, per-instance explanations using Shapley values tend to involve presenting a ranking of $x_j$ by $\phi_j$ values or simply presenting a set of the $x_j$ associated with the largest $\phi_j$ for some $\mathbf{x}$.
-Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across the observations in some partition of data $\mathbf{X}$.
-Each $\phi_j$ is a weighted combination of model scores on the $g_x(\mathbf{x})$ with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$.
+Local, per-instance explanations using Shapley values tend to involve ranking of $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$.
+Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some partition of data $\mathbf{X}$.
+
+Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
 
@@
 
-\noindent Where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
-
 Shapley values can be estimated in different ways, many of which are intractable for datasets with large $\mathcal{P}$.
@@
 Tree SHAP has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this text.
+Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing the same model with augmented training data \cite{molnar}.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific and personalized adverse action notices for automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.
 
@@
 \subsection{Software Resources}\label{ssec:soft}
-% Description of code
-% github repo, main packges
-Python code to reproduce the results presented in this text are available at: \url{https://github.com/h2oai/article-information-2019}.
-The authors primarily make use of the \href{https://github.com/h2oai/datatable}{datatable}, \href{https://github.com/h2oai/h2o-3}{h2o}, \href{https://matplotlib.org/}{matplotlib}, \href{https://pandas.pydata.org/}{pandas}, \href{https://scikit-learn.org/stable/}{scikit-learn}, \href{https://seaborn.pydata.org/}{seaborn}, and \href{https://github.com/slundberg/shap}{shap} packages for data manipulation, modeling, and reporting results.
+% Description of code?
 
-% interesting packages
+Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed therein are:
+
+\begin{itemize}
+	\item \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation.
+	\item \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing.
+	\item \href{https://matplotlib.org/}{\texttt{matplotlib}} and \href{https://seaborn.pydata.org/}{\texttt{seaborn}} for plotting. 
+\end{itemize}
+
+Additional, relevant Python packages include: \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}} for discrimination testing, \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} for discrimination testing and remediation, \href{https://github.com/interpretml/interpret}{\texttt{interpret}} for interpretable models and post-hoc explanation, and \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}} for post-hoc explanation.\footnote{See: \texttt{\url{https://github.com/jphall663/awesome-machine-learning-interpretability}} for a longer, curated list of related software packages and resources.}
 
@@
 \section{Results}\label{sec:res}
+
+Results are laid out for the simulated and mortgage datasets.
+Accuracy is compared for unconstrained, less interpretable $g^{\text{GBM}}$ and $g^{\text{ANN}}$ models and constrained, more interpretable $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models.
+Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretabilily, post-hoc explanation, and discrimination testing results are presented.
 
@@
 
-\subsection{The Burgeoning Ecosystem of Interpretable Models}
+\subsection{The Burgeoning Ecosystem of Interpretable Machine Learning Models}
 
-% these are just examples of good models
-%scalability, easy computation of derivatives
-% XNN is quite sensitive to initialization, and ...
-% variable selection
+MGBM and XNN architectures were selected for this text because they are straightforward variants of popular unconstrained ML models.
+If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
+The same can be said for the discussed post-hoc explanation and discrimination testing techniques.
+Practitioners should be able to augment many interpretable modeling workflows with Shapley and derivative-based post-hoc explanatory techniques and \textcolor{red}{...} discrimination testing methods.
+These approaches are promising responses to the opaqueness and discrimination problems in ML.
+However, they are imperfect and just a small part of a larger and burgeoning responsible ML ecosystem.
 
-% both additive
-% XNN smooth
+% GA2M, Rudin Group
 
-% Bivariate monotonicity constraints likely dampen impact of interactions - same for XNN
-% Tree pruning?
 \subsection{Impact of Discrimination Testing on Model Use and Adoption}
@@
 \subsection{Viable Discrimination Remediation Approaches}
-
-% somewhat aligned with United States regulation-mandated adverse action notices?
 
@@
 While the focus of this paper is not ML security, best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+
+%appeal
+
+% somewhat aligned with United States regulation-mandated adverse action notices?
+
+% security
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{N.G., GBM and MGBM assessment and results; P.H., primary author; K.M., ANN and XNN implementation, assessment, and results; N.S., secondary author, data simulation and collection, and discrimination testing.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+\authorcontributions{NG, data cleaning, GBM and MGBM assessment and results; PH, primary author; KM, ANN and XNN implementation, assessment, and results; NS, secondary author, data simulation and collection, and discrimination testing.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
 The following statements should be used ``conceptualization, X.X.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{Wen Phan for work in formalizing our notation.}
+\acknowledgments{Wen Phan for work in formalizing notation.}
 

### 2019-12-04 article-information-2019.tex
progress on figures, tables, discussion, and intro

@@
 
+Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e.
+ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}} Even if responsible AI feels like a somewhat broad and amorphous notion, at least one aspect should be crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
 ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
-This text makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g.
-the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.
-This text does however propose some workable countermeasures to address these vexing problems in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools.
+Though this text makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), this text does propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems.
+
 Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other discrimination, and public and open source software resources.
@@
 Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results.
-Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML.
+Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML.
 
@@
 \section{Materials and Methods}\label{sec:m_and_m}
+
+Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:
+
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
+	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
+	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
+	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
+	\item \textbf{Discrimination testing methods}: -- \textsection\ref{ssec:di}
+	\item \textbf{Software resources}: GitHub repository for this text; utilized \& useful Python packages -- \textsection\ref{ssec:soft}
+\end{itemize}
 
@@
 All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated with this text.
-Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:
-
-\begin{itemize}
-	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
-	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
-	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
-	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
-	\item \textbf{Discrimination testing methods}: -- \textsection\ref{ssec:di}
-	\item \textbf{Software resources}: GitHub repository for this text; utilized \& useful Python packages -- \textsection\ref{ssec:soft}
-\end{itemize}
 
@@
 
-	\begin{itemize}
+	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
 		\item Input features come from the set $\mathcal{X}$ contained in a \textit{P}-dimensional input space, $\mathcal{X} \subset \mathbb{R}^P$.
@@
 
-	\begin{itemize}
+	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
 		\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
@@
 
-	\begin{itemize}
+	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
 		\item A type of ML model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$:
@@
 
-\begin{enumerate}
+\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
 \item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in the left child weight being greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
@@
 
-\begin{enumerate}
+\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
 \item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, is known as the \textit{projection layer} and is fully connected to each input feature, $X_j$.
@@
 Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
-Tree SHAP has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this text.
+Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this text.
 Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing the same model with augmented training data \cite{molnar}.
@@
 
-Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed therein are:
-
-\begin{itemize}
-	\item \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation.
-	\item \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing.
-	\item \href{https://matplotlib.org/}{\texttt{matplotlib}} and \href{https://seaborn.pydata.org/}{\texttt{seaborn}} for plotting. 
-\end{itemize}
-
-Additional, relevant Python packages include: \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}} for discrimination testing, \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} for discrimination testing and remediation, \href{https://github.com/interpretml/interpret}{\texttt{interpret}} for interpretable models and post-hoc explanation, and \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}} for post-hoc explanation.\footnote{See: \texttt{\url{https://github.com/jphall663/awesome-machine-learning-interpretability}} for a longer, curated list of related software packages and resources.}
+Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}.
+The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and \href{https://matplotlib.org/}{\texttt{matplotlib}} and \href{https://seaborn.pydata.org/}{\texttt{seaborn}} for plotting.
 
@@
 
-\subsubsection{Unconstrained Model Fit Assessment}
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}
 
-\subsubsection{Constrained Model Fit Assessment}
+\begin{table}[H]
+\label{tab:sim_acc}
+\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on simulated data.}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{ccccc}
+\toprule
+\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
+\midrule
+$g^\text{GBM}$		& 				& 			& 				& \\
+$g^\text{MGBM}$	& 				& 			& 				& \\
+$g^\text{ANN}$		& 				& 			& 				& \\
+$g^\text{XNN}$		& 				& 			& 				& \\
+\bottomrule
+\end{tabular}
+\end{table}
 
 \subsubsection{Interpretability and Post-hoc Explanation Results}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:sim_mgbm_glob}
+\caption{PD and ICE across deciles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the simulated data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:sim_mgbm_loc}
+\caption{Mean Tree SHAP values across quintiles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the simulated data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:sim_xnn_glob}
+\caption{Ridge functions \textcolor{red}{and ICE across deciles} for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the simulated data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:fig:sim_xnn_glob}
+\caption{Mean \textcolor{red}{local feature importance} values across quintiles for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the simulated data.}
+\end{figure}
 
@@
 
-\subsubsection{Unconstrained Model Fit Assessment}
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}
 
-\subsubsection{Constrained Model Fit Assessment}
+\begin{table}[H]
+\label{tab:mort_acc}
+\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on mortgage data.}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{ccccc}
+\toprule
+\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
+\midrule
+$g^\text{GBM}$		& 				& 			& 				& \\
+$g^\text{MGBM}$	& 				& 			& 				& \\
+$g^\text{ANN}$		& 				& 			& 				& \\
+$g^\text{XNN}$		& 				& 			& 				& \\
+\bottomrule
+\end{tabular}
+\end{table}
 
 \subsubsection{Interpretability and Post-hoc Explanation Results}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:mort_mgbm_glob}
+\caption{PD and ICE across deciles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the mortgage data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:mort_mgbm_loc}
+\caption{Mean Tree SHAP values across quintiles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the mortgage data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:mort_xnn_glob}
+\caption{Ridge functions \textcolor{red}{and ICE across deciles} for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the mortgage data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:fig:mort_xnn_glob}
+\caption{Mean \textcolor{red}{local feature importance} values across quintiles for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the mortgage data.}
+\end{figure}
 
@@
 
-\subsection{The Burgeoning Ecosystem of Interpretable Machine Learning Models}
+\subsection{The Burgeoning Python Ecosystem for Responsible Machine Learning}
 
-MGBM and XNN architectures were selected for this text because they are straightforward variants of popular unconstrained ML models.
+MGBM and XNN interpretable model architectures were selected for this text because they are straightforward variants of popular unconstrained ML models.
 If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
-The same can be said for the discussed post-hoc explanation and discrimination testing techniques.
-Practitioners should be able to augment many interpretable modeling workflows with Shapley and derivative-based post-hoc explanatory techniques and \textcolor{red}{...} discrimination testing methods.
-These approaches are promising responses to the opaqueness and discrimination problems in ML.
-However, they are imperfect and just a small part of a larger and burgeoning responsible ML ecosystem.
-
-% GA2M, Rudin Group
+These approaches are promising responses to the black-box problem in ML.
+However, they are just a small part of a burgeoning ecosystem of Python tools for interpretable models, post-hoc explanation, and discrimination testing and remediation.
+Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers Microsoft and Cornell and led by Rich Caruana.
+In keeping with a major theme of this text, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
+Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and ``this looks like that'' interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{Optimal sparse decision trees: \url{https://github.com/xiyanghu/OSDT}.}\textsuperscript{,}\footnote{``This looks like that'' interpretable deep learing: \url{https://github.com/cfchen-duke/ProtoPNet}.} Additional, relevant Python packages include: \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}} for discrimination testing, \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} for discrimination testing and remediation, and \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}} for post-hoc explanation.\footnote{See: \texttt{\url{https://github.com/jphall663/awesome-machine-learning-interpretability}} for a longer, curated list of related software packages and resources.}
 
@@
 
-%appeal
+%appeal: interpretable models + explanations
 
@@
 % security
+
+% e.g. the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy})
 
@@
 % accuracy, fairness, security are not static
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=8cm]{img/blueprint.png}
+\label{fig:blueprint}
+\caption{A diagram of a proposed holistic ML workflow in which explanations (highlighted in red) are used along with interpretable models, DI analysis and remediation techniques, and other review and appeal mechanisms to create an understandable and trustworthy ML system.}
+\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{This research received no external funding.}
+\funding{This work received no external funding.}
 
@@
 %% optional
-\abbreviations{The following abbreviations are used in this text: ANN -- artificial neural network, DI -- disparate impact, GBM -- gradient boosting machine, ICE -- individual conditional expectation, MGBM -- monotonic gradient boosting machine, ML -- machine learning, PD -- partial dependence, SGD -- stochastic gradient descent, US -- United States, XNN -- explainable neural network.}
+\abbreviations{The following abbreviations are used in this text: AI -- artificial intelligence, ANN -- artificial neural network, DI -- disparate impact, EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e.
+variants GAMs that consider two-way interactions and incorporate boosting into training, GAM -- generalized additive model, GBM -- gradient boosting machine, ICE -- individual conditional expectation, MGBM -- monotonic gradient boosting machine, ML -- machine learning, PD -- partial dependence, SGD -- stochastic gradient descent, US -- United States, XNN -- explainable neural network.}
 

### 2019-12-05 article-information-2019.tex
more progress on the text

@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{This text outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
+\abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
 The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on realistic, publicly available mortgage data.
@@
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
-Though this text makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), this text does propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems.
+Though this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), this text does propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems.
 
@@
 Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results.
-Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML.
+Section \ref{sec:con} closes this manuscript with a brief outline of proposed additional steps to increase human trust and understanding in ML.
 
@@
 
-To provide a sense of accuracy differences, this text compares the performance of more interpretable constrained ML models and less interpretable unconstrained ML models on simulated data and on collected mortgage data.
+To provide a sense of accuracy differences, this manuscript compares the performance of more interpretable constrained ML models and less interpretable unconstrained ML models on simulated data and on collected mortgage data.
 The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1991multivariate}.
@@
 Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
-Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this text.
+Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this manuscript.
 Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing the same model with augmented training data \cite{molnar}.
@@
 If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
-These approaches are promising responses to the black-box problem in ML.
-However, they are just a small part of a burgeoning ecosystem of Python tools for interpretable models, post-hoc explanation, and discrimination testing and remediation.
+The same can be said of Shapley values and \textcolor{red}{discrimination tests}.
+Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines.
+While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML.
+Figure \ref{fig:blueprint} is a workflow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy machine learning system.\footnote{See: \texttt{\url{https://github.com/jphall663/hc_ml}} for details regarding the workflow in Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure ~\ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted here as further reading.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=8cm]{img/blueprint.png}
+\caption{A diagram of a proposed holistic ML workflow in which explanations (highlighted in red) are used along with interpretable models, DI analysis and remediation techniques, and other review and appeal mechanisms to create an understandable and trustworthy ML system.}
+\label{fig:blueprint}
+\end{figure}
+
 Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers Microsoft and Cornell and led by Rich Caruana.
-In keeping with a major theme of this text, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
-Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and ``this looks like that'' interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{Optimal sparse decision trees: \url{https://github.com/xiyanghu/OSDT}.}\textsuperscript{,}\footnote{``This looks like that'' interpretable deep learing: \url{https://github.com/cfchen-duke/ProtoPNet}.} Additional, relevant Python packages include: \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}} for discrimination testing, \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} for discrimination testing and remediation, and \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}} for post-hoc explanation.\footnote{See: \texttt{\url{https://github.com/jphall663/awesome-machine-learning-interpretability}} for a longer, curated list of related software packages and resources.}
+In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
+Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and ``this looks like that'' interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{Optimal sparse decision trees: \url{https://github.com/xiyanghu/OSDT}.}\textsuperscript{,}\footnote{``This looks like that'' interpretable deep learning: \url{https://github.com/cfchen-duke/ProtoPNet}.} Additional, relevant references and Python functionality includes:
+
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+\item \textbf{Exploratory data visualization (EDA)}: \texttt{H2OAggregatorEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}{\texttt{h2o}} \cite{wilkinson2018visualizing}.
+\item \textbf{Sparse feature extraction}: \texttt{H2OGeneralizedLowRankEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogeneralizedlowrankestimator}{\texttt{h2o}} \cite{udell2016generalized}.
+\item \textbf{Privacy preprocessing and private models}: differential privacy for ML in \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}} \cite{diffprivlib}, \cite{ji2014differential}; private models in \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{pate}, \cite{abadi2016deep}.
+\item \textbf{Post-hoc explanation}: %\href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}} gradients
+\item \textbf{Discrimination testing}: in \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.
+\item \textbf{Discrimination remediation}: Reweighing, adversarial de-biasing, learning fair representations, and reject option classification in \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} \cite{kamiran2012data}, \cite{zhang2018mitigating}, \cite{lfr}, \cite{kamiran2012decision}.
+\item \textbf{Model debugging}: %tensorflow stuff cleverhans, foolbox, SALib what-if
+\item \textbf{Model documentation}: Models cards \cite{model_cards}, e.g. \texttt{\url{https://github.com/openai/gpt-2}}, \texttt{\url{https://modelcards.withgoogle.com/object-detection}}. 
+\end{itemize}
+
+\noindent See: \texttt{\url{https://github.com/jphall663/awesome-machine-learning-interpretability}} for a longer, curated list of related software packages and resources.
 
@@
 
-%appeal: interpretable models + explanations
-
 % somewhat aligned with United States regulation-mandated adverse action notices?
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}\label{sec:con}
-% Model debugging
-% Model management and monitoring
-% Appeal
-% local bias
-
 % Moreover, intersectional dangers at the nexus of sociological bias, opaque, and insecure decision-making systems: privacy dangers of explanations, fairwashing, local bias, proxies
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=8cm]{img/blueprint.png}
-\label{fig:blueprint}
-\caption{A diagram of a proposed holistic ML workflow in which explanations (highlighted in red) are used along with interpretable models, DI analysis and remediation techniques, and other review and appeal mechanisms to create an understandable and trustworthy ML system.}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}\label{sec:con}
+
+This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
+These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible machine learning workflow.
+The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects.
+Of course there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and this manuscript also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
+While the messy, complex, and human problems of racism, sexism, privacy violations, and cybercrime can probably not be solved by technology alone, this work and many others illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{}
+\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank. Wells Fargo is a customer and investor of H2O.ai and a customer of BLDS, LLC.}
 

### 2019-12-07 data/scripts/di_testing.py
Changes to DI tests

@@ -55,16 +55,17 @@ class DisparateImpactTesting(object):
 
     def adverse_impact_ratio(self,
                              data: pd.DataFrame,
-                             label: str,
+                             outcome: str,
                              ) -> pd.DataFrame:
         """
         This calculates the Adverse Impact Ratio (AIR), a standard measure of disparate impact used in
         U.S. legal and regulatory settings.  It is defined as:
             AIR = (% Protected Class Chosen) / (% Control Group Chosen)
 
-        :param data: Data containing the label and protected and control group information for each person
+        :param data: Data containing the outcome being measured
+         and protected and control group information for each person
 
-        :param label: The string name of the label
+        :param outcome: The string name of the label
 
         :return: A table containing the AIR and other summary statistics for each group.
         """
@@ -76,9 +77,9 @@ class DisparateImpactTesting(object):
         di_table.loc[di_table["class"].isin(self.pg_names), "Control Class"] = self.cg_names
 
         if self.lower_value_favorable:
-            outcome = (1 - data[label]).to_numpy().reshape(-1, 1)
+            outcome = (1 - data[outcome]).to_numpy().reshape(-1, 1)
         else:
-            outcome = data[label].to_numpy().reshape(-1, 1)
+            outcome = data[outcome].to_numpy().reshape(-1, 1)
 
         di_table["Total"] = data[self.pgcg_names].sum(axis=0)
         di_table["Total Favorable"] = np.sum(outcome * data[self.pgcg_names])

### 2019-12-07 article-information-2019.tex
progress on intersectionality

@@
 
-% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
+% If you would like to post an early version of this manuscript  a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
 The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
@@
 The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on realistic, publicly available mortgage data.
-For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and \textcolor{red}{global and local gradient-based or Shapley feature importance}.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
 The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using straightforward group fairness measures.
@@
 ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}} Even if responsible AI feels like a somewhat broad and amorphous notion, at least one aspect should be crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
-ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
+ML models can be inaccurate, noncompliant, and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
-Though this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), this text does propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems.
+Though this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), this text does propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the U.S., interpretable models, explanations, disparate impact testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (E.U.) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
 
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-\textcolor{red}{Discussed discrimination testing methods and metrics include ...
-}.
+Discussed discrimination testing methods include measures of DI with marginal effects, the adverse impact ratio (AIR), and the standardized mean difference (SMD) \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment\cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated with this text.
@@
 \label{tab:sim_acc}
-\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on simulated data.}
+\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on simulated test data.}
 \centering
@@
 \label{fig:sim_mgbm_glob}
-\caption{PD and ICE across deciles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the simulated data.}
+\caption{Global Tree SHAP feature importance for $g^\text{MGBM}$ on the test partition of the simulated data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:sim_mgbm_glob_pdp_ice}
+\caption{PD and ICE across deciles for the three most important input features for $g^\text{MGBM}$ on the test partition of the simulated data.}
 \end{figure}
@@
 \label{fig:sim_mgbm_loc}
-\caption{Mean Tree SHAP values across quintiles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the simulated data.}
+\caption{Mean Tree SHAP values across quintiles the three most important input features for $g^\text{MGBM}$ on the test partition of the simulated data.}
 \end{figure}
@@
 \label{fig:sim_xnn_glob}
-\caption{Ridge functions \textcolor{red}{and ICE across deciles} for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the simulated data.}
+\caption{Global Deep SHAP feature importance for $g^\text{XNN}$ on the test partition of the simulated data.}
 \end{figure}
@@
 %\includegraphics[width=8cm]{img/}
-\label{fig:fig:sim_xnn_glob}
-\caption{Mean \textcolor{red}{local feature importance} values across quintiles for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the simulated data.}
+\label{fig:sim_xnn_glob_pdp_ice}
+\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the test partition of the simulated data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:sim_xnn_loc}
+\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the test partition of the simulated data.}
 \end{figure}
@@
 \label{tab:mort_acc}
-\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on mortgage data.}
+\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on mortgage test data.}
 \centering
@@
 \label{fig:mort_mgbm_glob}
-\caption{PD and ICE across deciles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the mortgage data.}
+\caption{Global Tree SHAP feature importance for $g^\text{MGBM}$ on the test partition of the simulated data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:mort_mgbm_glob_pdp_ice}
+\caption{PD and ICE across deciles for the three most important input features for $g^\text{MGBM}$ on the test partition of the mortgage data.}
 \end{figure}
@@
 \label{fig:mort_mgbm_loc}
-\caption{Mean Tree SHAP values across quintiles for the five most important input features for $g^\text{MGBM}(\mathbf{X})$ on the test partition of the mortgage data.}
+\caption{Mean Tree SHAP values across quintiles the three most important input features for $g^\text{MGBM}$ on the test partition of the mortgage data.}
 \end{figure}
@@
 \label{fig:mort_xnn_glob}
-\caption{Ridge functions \textcolor{red}{and ICE across deciles} for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the mortgage data.}
+\caption{Global Deep SHAP feature importance for $g^\text{XNN}$ on the test partition of the mortgage data.}
 \end{figure}
@@
 %\includegraphics[width=8cm]{img/}
-\label{fig:fig:mort_xnn_glob}
-\caption{Mean \textcolor{red}{local feature importance} values across quintiles for the five most important input features for $g^\text{XNN}(\mathbf{X})$ on the test partition of the mortgage data.}
+\label{fig:mort_xnn_glob_pdp_ice}
+\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the test partition of the mortgage data.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=8cm]{img/}
+\label{fig:mort_xnn_loc}
+\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the test partition of the mortgage data.}
 \end{figure}
@@
 If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
-The same can be said of Shapley values and \textcolor{red}{discrimination tests}.
+The same can be said of the presented explanation methods and discrimination tests.
 Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines.
@@
 \begin{itemize}[leftmargin=*,labelsep=5.8mm]
+
 \item \textbf{Exploratory data visualization (EDA)}: \texttt{H2OAggregatorEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}{\texttt{h2o}} \cite{wilkinson2018visualizing}.
+
 \item \textbf{Sparse feature extraction}: \texttt{H2OGeneralizedLowRankEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogeneralizedlowrankestimator}{\texttt{h2o}} \cite{udell2016generalized}.
+
 \item \textbf{Privacy preprocessing and private models}: differential privacy for ML in \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}} \cite{diffprivlib}, \cite{ji2014differential}; private models in \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{pate}, \cite{abadi2016deep}.
-\item \textbf{Post-hoc explanation}: %\href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}} gradients
-\item \textbf{Discrimination testing}: in \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.
-\item \textbf{Discrimination remediation}: Reweighing, adversarial de-biasing, learning fair representations, and reject option classification in \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} \cite{kamiran2012data}, \cite{zhang2018mitigating}, \cite{lfr}, \cite{kamiran2012decision}.
-\item \textbf{Model debugging}: %tensorflow stuff cleverhans, foolbox, SALib what-if
-\item \textbf{Model documentation}: Models cards \cite{model_cards}, e.g. \texttt{\url{https://github.com/openai/gpt-2}}, \texttt{\url{https://modelcards.withgoogle.com/object-detection}}. 
+
+\item \textbf{Post-hoc explanation}: structured data explanations with \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}}; image classification explanations with \href{https://github.com/marcoancona/DeepExplain}{\texttt{DeepExplainer}}; natural language explanations with \href{https://github.com/allenai/allennlp}{\texttt{allennlp}} \cite{wachter2017counterfactual}, \cite{grad_attr}, \cite{wallace2019allennlp}.
+
+\item \textbf{Discrimination testing}: with \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.
+
+\item \textbf{Discrimination remediation}: Reweighing, adversarial de-biasing, learning fair representations, and reject option classification with \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} \cite{kamiran2012data}, \cite{zhang2018mitigating}, \cite{lfr}, \cite{kamiran2012decision}.
+
+\item \textbf{Model debugging}: with \href{https://github.com/bethgelab/foolbox}{\texttt{foolbox}}, \href{https://github.com/SALib/SALib}{\texttt{SALib}}, \href{https://github.com/tensorflow/cleverhans}{\texttt{tensorflow/cleverhans}}, and \href{https://github.com/tensorflow/model-analysis}{\texttt{tensorflow/model-analysis}} \cite{rauber2017foolbox}, \cite{papernot2018cleverhans}, \cite{modeltracker}, \cite{papernot2018marauder}.
+
+\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g. \texttt{\url{https://github.com/openai/gpt-2}}, \texttt{\url{https://modelcards.withgoogle.com/object-detection}}.
+
 \end{itemize}
@@
 
-%Interpretable models are also tested for disparate impact and other unwanted sociological bias and viable bias mitigation strategies are discussed.
-While the focus of this paper is not ML security, best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
-
-% somewhat aligned with United States regulation-mandated adverse action notices?
-
-% security
-
-% e.g. the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy})
-
-% Moreover, intersectional dangers at the nexus of sociological bias, opaque, and insecure decision-making systems: privacy dangers of explanations, fairwashing, local bias, proxies
-% accuracy, fairness, security are not static
+The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own.
+However, evidence is mounting that these harms can also manifest as complex intersectional challenges e.g.
+the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.\footnote{See: \href{https://www.telegraph.co.uk/technology/2016/03/24/microsofts-teen-girl-ai-turns-into-a-hitler-loving-sex-robot-wit/}{\textit{Microsoft deletes 'teen girl' AI after it became a Hitler-loving sex robot within 24 hours}}.} (While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\footnote{See: \href{https://www.theatlantic.com/technology/archive/2012/03/the-philosopher-whose-fingerprints-are-all-over-the-ftcs-new-approach-to-privacy/254365/}{\textit{The Philosopher Whose Fingerprints Are All Over the FTC's New Approach to Privacy}}.}\textsuperscript{,}\footnote{See: \textcolor{red}{Andrew Burt HBR article - upcoming.}}) Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+Instead, it aims to highlight these issues as both singular entities and intersectional phenomena.
+Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as remedies to the black-box, and discrimination issues in ML.
+However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks and that explanations can mask ML discrimination.
+Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory.
+Of course, accuracy, discrimination, and security characteristics of a system can change over time, and simply testing for these problems at training time, as presented herein, is not adequate for high-stakes, human-centered, or regulated ML systems.
+Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
 
@@
 This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
-These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible machine learning workflow.
+These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible machine learning workflow high-stakes, human-centered, or regulated ML applications.
 The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects.
 Of course there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and this manuscript also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
-While the messy, complex, and human problems of racism, sexism, privacy violations, and cybercrime can probably not be solved by technology alone, this work and many others illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
+While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work and many others illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{Wen Phan for work in formalizing notation.}
+\acknowledgments{Wen Phan for work in formalizing notation. \textcolor{red}{BLDS editor} for editing.}
 

### 2019-12-11 article-information-2019.tex
edits

@@
 % Full title of the paper (Capitalized)
-\Title{Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}}
+\Title{Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
 
@@
 \abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on realistic, publicly available mortgage data.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on publicly available mortgage data.
 For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using straightforward group fairness measures.
-By combining innovations in interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using adverse impact ratio (AIR), standardized mean difference (SMD), and additional straightforward group fairness measures.
+By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}
 
@@
 Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e.
-ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}} Even if responsible AI feels like a somewhat broad and amorphous notion, at least one aspect should be crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
-ML models can be inaccurate, noncompliant, and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
+ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
+ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
-Though this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), this text does propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the U.S., interpretable models, explanations, disparate impact testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (E.U.) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+While this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
 
-Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other discrimination, and public and open source software resources.
+Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an appealable decision-making framework, tests for DI and other social discrimination, and public and open source software resources.
 In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and discrimination testing results are also presented for interpretable models.
 Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results.
-Section \ref{sec:con} closes this manuscript with a brief outline of proposed additional steps to increase human trust and understanding in ML.
+Section \ref{sec:con} closes this manuscript with a brief summary of the outlined methods, materials, results, and discussion.
 
@@
 	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
-	\item \textbf{Discrimination testing methods}: -- \textsection\ref{ssec:di}
-	\item \textbf{Software resources}: GitHub repository for this text; utilized \& useful Python packages -- \textsection\ref{ssec:soft}
+	\item \textbf{Discrimination testing methods}: AIR, SMD, and other group fairness metrics -- \textsection\ref{ssec:di}
+	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
 \end{itemize}
 
-To provide a sense of accuracy differences, this manuscript compares the performance of more interpretable constrained ML models and less interpretable unconstrained ML models on simulated data and on collected mortgage data.
+To provide a sense of accuracy differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
 The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1991multivariate}.
-The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a more realistic template workflow for future users of similar methods and tools.
-
-Because unconstrained ML models, like gradient boosting machines (GBMs) (e.g.
+The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Because unconstrained ML models, like gradient boosting machines (GBMs) (e.g.
 \cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
@@
 Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} and XNNs (\cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analysis.
+
+Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a more realistic template workflow for future users of similar methods and tools.
 Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}.
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods include measures of DI with marginal effects, the adverse impact ratio (AIR), and the standardized mean difference (SMD) \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment\cite{zafar2017fairness}.
-All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated with this text.
+Discussed discrimination testing methods include measures of DI with marginal effects, AIR and SMD \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
+All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.
 
@@
 		\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
-		\item The model to be explained and tested for discrimination testing is denoted as $g$.
+		\item The model to be explained and tested for discrimination is denoted as $g$.
 	\end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monotonically Constrained Gradient Boosting Machine}\label{ssec:mgbm}
+\subsection{Monotonic Gradient Boosting Machine}\label{ssec:mgbm}
 
-MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints.
-The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, but each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
+MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints, with respect to each $X_j$ and $\mathbf{y}$ independently.
+The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
 
@@
 \noindent In addition to $\mathcal{L}_{b}$, $g^\text{MGBM}$ training is characterized by additional splitting rules and constraints on tree node weights.
-Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,kR}\}$.
-For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,kR}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction.
-For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$, $\Theta_b$ is restricted to be positive monotonic w.r.t.
-$X_j$ by the following rules and constraints.
+Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$.
+For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction.
+For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, $g^{\text{MGBM}}$ is positive monotonic with respect to some $X_j$ if $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
+The following rules and constraints ensure positive monotonicity in $\Theta_b$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$.
 
 \begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in the left child weight being greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
+\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in the left child weight being greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{b,j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
 \item For any subsequent left child node involving $X_j$, any $\theta_{b,j, k\ge1}$ resulting in $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ where $w_{b,j,k\ge1,L} > w_{b,j,k\ge1,R}$, is not considered.
-\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$, $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ are bound by the parent set of node weights, $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$, such that $ \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \le \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
-\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$, $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ are bound by the associated $\theta_{b,j,k-1}$ set of node weights, $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$, such that $ \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} < \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $ w_{b,j,k,L} \le w_{b,j,k,R}$ and $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
 \end{enumerate}
@@
 \noindent Note that for any one $X_j$ and $T_b \in g^{\text{MGBM}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
-Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,j,0}) = w_{b,0}$.
+Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0}) = w_{b,0}$.
 Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
 For a negative monotonic constraint, i.e.
-$g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.\\
+$g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
+Also consider that MGBM models with independent monotonicity constraints between some $X_j$ and $\mathbf{y}$ likely restrict non-monotonic interactions between multiple $X_j$.
+Moreover, if monotonicity constraints are not applied to all $X_j \in \mathbf{X}$, any strong non-monotonic signal in training data associated with some important $X_j$ maybe forced onto some other arbitrary unconstrained $X_j$ under some $g^{MGBM}$ models, compromising the end goal of interpretability.
 
@@
 One on the simulated data and one on the mortgage data.
-In both cases, positive and negative monotonic constraints for each $X_j$ are selected using the sign of the Pearson correlation between each $X_j$ and $y$, random grid search is used to determine other hyperparameters, and five-fold cross validation and test partitions are used for model assessment.
-For exact parameterization of the MGBM models, see the software resources referenced in Subsection \ref{ssec:soft}.
-Also consider that MGBM models with one-dimensional monotonicity constraints between some $X_j$ and $\hat{Y}$ likely restrict non-monotonic interactions between multiple $X_j$.
-Moreover, if monotonicity constraints are not applied to all $X_j \in \mathbf{X}$, any strong non-monotonic signal in training data associated with some important $X_j$ maybe forced onto some other arbitrary unconstrained $X_j$ under some $g^{MGBM}$ models, compromising the end goal of interpretability.
+In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and five-fold cross validation and test partitions are used for model assessment.
+For exact parameterization of the two $g^\text{MGBM}$ models, see the software resources referenced in Subsection \ref{ssec:soft}.
 
@@
 XNNs also bare a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e.
-GAMs which consider main effects and a small number of 2-way interactions and incorporate boosting in their training \cite{esl}, \cite{ga2m}.
+GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
 Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
-XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum^J_{j=0}\beta_{k, j}x_j$.
+XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum_j\beta_{k,j}x_j$.
 
@@
 \label{eq:xnn}
-g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=1}^K\gamma_k n_k(\sum^J_{j=1}\beta_{k, j}x_j)
+g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=0}^{K-1}\gamma_k n_k(\sum^{J=\mathcal{P}-1}_{j=0}\beta_{k, j}x_j)
 \end{aligned}
@@
 \item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, is known as the \textit{projection layer} and is fully connected to each input feature, $X_j$.
+Each hidden unit in the projection layer may optionally include a bias term.
 \item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
-Each $n_k$ is a neural network, which can be parameterized to suite a given modeling task.
-To facilitate direct visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
+Each $n_k$ is a neural network, which can be parameterized to suit a given modeling task.
+To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
+Each $n_k$ can contain several bias terms.
 \item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
-Again, subnetwork output is restricted to 1-dimension for visualization purposes.
+Again, subnetwork output is restricted to 1-dimension for interpretation and visualization purposes.
 \end{enumerate}
@@
 $L_1$ regularization is applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
-The $g^\text{XNN}$ models are highly sensitive to weight initialization and manual feature selection is informed by Shapley feature importance from the $g^\text{MGBM}$.
+The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size.
+Be aware that $g^{XNN}$ model architectures may require manual and judicious feature selection due to long training times.
 For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft}.
-Be aware that $g^{XNN}$ model architectures may restrict important interactions, appear highly sensitive to initialization routines, and require manual and judicious feature selection due to burdensome training times.
 
@@
 PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g.
-inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions of a complex model for some $X_j$ \cite{ice_plots}.
+inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
 
-Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{(-j)} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{(-j)} = \mathbf{X}$) is considered.
-$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{(-j)}$ is left unchanged.
-$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{(-j)}$ are left untouched.
+Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P\setminus{j}}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P\setminus{j}}} = \mathbf{X}$) is considered.
+$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P\setminus{j}}}$ is left unchanged.
+$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P\setminus{j}}}$ are left untouched.
 PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
@@
 Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}, \cite{shapley1988shapley}.
-Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values can never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
+Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values should never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
 For some instance $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
@@
 Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
-Local, per-instance explanations using Shapley values tend to involve ranking of $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$.
-Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some partition of data $\mathbf{X}$.
-
 Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
@@
 
+\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
+Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some set $\mathbf{X}$.
+
 Shapley values can be estimated in different ways, many of which are intractable for datasets with large $\mathcal{P}$.
 Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
-Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in this manuscript.
-Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing the same model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific and personalized adverse action notices for automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.
+Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}.
+Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Sections \ref{sec:m_and_m} and \ref{sec:res}.
 
@@
 \subsection{Software Resources}\label{ssec:soft}
-% Description of code?
 
@@
 \label{tab:sim_acc}
-\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on simulated test data.}
+\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
 \centering
@@
 \label{fig:sim_mgbm_glob}
-\caption{Global Tree SHAP feature importance for $g^\text{MGBM}$ on the test partition of the simulated data.}
+\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.}
 \end{figure}
@@
 \label{fig:sim_mgbm_glob_pdp_ice}
-\caption{PD and ICE across deciles for the three most important input features for $g^\text{MGBM}$ on the test partition of the simulated data.}
+\caption{PD, ICE across deciles, and histograms for the three most important input features for $g^\text{MGBM}$ on the simulated test data.}
 \end{figure}
+
+% highlight monotonicity
 
@@
 \label{fig:sim_mgbm_loc}
-\caption{Mean Tree SHAP values across quintiles the three most important input features for $g^\text{MGBM}$ on the test partition of the simulated data.}
+\caption{Mean Tree SHAP values across quintiles for the three most important input features for $g^\text{MGBM}$ on the simulated test data.}
 \end{figure}
@@
 \label{fig:sim_xnn_glob}
-\caption{Global Deep SHAP feature importance for $g^\text{XNN}$ on the test partition of the simulated data.}
+\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the simulated test data.}
 \end{figure}
@@
 \label{fig:sim_xnn_glob_pdp_ice}
-\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the test partition of the simulated data.}
+\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the simulated test data.}
 \end{figure}
@@
 \label{fig:sim_xnn_loc}
-\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the test partition of the simulated data.}
+\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the simulated test data.}
 \end{figure}
@@
 \label{tab:mort_acc}
-\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on mortgage test data.}
+\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data.}
 \centering
@@
 \label{fig:mort_mgbm_glob}
-\caption{Global Tree SHAP feature importance for $g^\text{MGBM}$ on the test partition of the simulated data.}
+\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the mortgage test data.}
 \end{figure}
@@
 \label{fig:mort_mgbm_glob_pdp_ice}
-\caption{PD and ICE across deciles for the three most important input features for $g^\text{MGBM}$ on the test partition of the mortgage data.}
+\caption{PD, ICE across deciles, and histograms for the three most important input features for $g^\text{MGBM}$ on the mortgage test data.}
 \end{figure}
+
+% highlight monotonicity
 
@@
 \label{fig:mort_mgbm_loc}
-\caption{Mean Tree SHAP values across quintiles the three most important input features for $g^\text{MGBM}$ on the test partition of the mortgage data.}
+\caption{Mean Tree SHAP values across quintiles for the three most important input features for $g^\text{MGBM}$ on the mortgage test data.}
 \end{figure}
@@
 \label{fig:mort_xnn_glob}
-\caption{Global Deep SHAP feature importance for $g^\text{XNN}$ on the test partition of the mortgage data.}
+\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the mortgage test data.}
 \end{figure}
@@
 \label{fig:mort_xnn_glob_pdp_ice}
-\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the test partition of the mortgage data.}
+\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the mortgage test data.}
 \end{figure}
@@
 \label{fig:mort_xnn_loc}
-\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the test partition of the mortgage data.}
+\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the mortgage test data.}
 \end{figure}
@@
 The same can be said of the presented explanation methods and discrimination tests.
-Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines.
+Due to their post-hoc nature, they can often be shoe-horned into existing ML work flows and pipelines.
 While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML.
-Figure \ref{fig:blueprint} is a workflow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy machine learning system.\footnote{See: \texttt{\url{https://github.com/jphall663/hc_ml}} for details regarding the workflow in Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure ~\ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted here as further reading.
+Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.\footnote{See: \texttt{\url{https://github.com/jphall663/hc_ml}} for details regarding the work flow in Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.
 
@@
 \includegraphics[width=8cm]{img/blueprint.png}
-\caption{A diagram of a proposed holistic ML workflow in which explanations (highlighted in red) are used along with interpretable models, DI analysis and remediation techniques, and other review and appeal mechanisms to create an understandable and trustworthy ML system.}
+\caption{A diagram of a proposed holistic ML work flow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, and other review and appeal mechanisms can create an understandable and trustworthy ML system.}
 \label{fig:blueprint}
@@
 
-Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers Microsoft and Cornell and led by Rich Caruana.
+Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers at Microsoft and Cornell.
 In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
-Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and ``this looks like that'' interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{Optimal sparse decision trees: \url{https://github.com/xiyanghu/OSDT}.}\textsuperscript{,}\footnote{``This looks like that'' interpretable deep learning: \url{https://github.com/cfchen-duke/ProtoPNet}.} Additional, relevant references and Python functionality includes:
+Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{Optimal sparse decision trees: \url{https://github.com/xiyanghu/OSDT}.}\textsuperscript{,}\footnote{\textit{This looks like that} interpretable deep learning: \url{https://github.com/cfchen-duke/ProtoPNet}.} Additional, relevant references and Python functionality include:
 
@@
 
-\item \textbf{Exploratory data visualization (EDA)}: \texttt{H2OAggregatorEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}{\texttt{h2o}} \cite{wilkinson2018visualizing}.
+\item \textbf{Exploratory data analysis (EDA)}: \texttt{H2OAggregatorEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}{\texttt{h2o}} \cite{wilkinson2018visualizing}.
 
@@
 
-\item \textbf{Privacy preprocessing and private models}: differential privacy for ML in \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}} \cite{diffprivlib}, \cite{ji2014differential}; private models in \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{pate}, \cite{abadi2016deep}.
+\item \textbf{Privacy preprocessing and private models}: differential privacy and private models in \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}} and \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{diffprivlib}, \cite{ji2014differential}, \cite{pate}, \cite{abadi2016deep}.
 
-\item \textbf{Post-hoc explanation}: structured data explanations with \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}}; image classification explanations with \href{https://github.com/marcoancona/DeepExplain}{\texttt{DeepExplainer}}; natural language explanations with \href{https://github.com/allenai/allennlp}{\texttt{allennlp}} \cite{wachter2017counterfactual}, \cite{grad_attr}, \cite{wallace2019allennlp}.
+\item \textbf{Post-hoc explanation}: structured data explanations with \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}}, image classification explanations with \href{https://github.com/marcoancona/DeepExplain}{\texttt{DeepExplain}}, and natural language explanations with \href{https://github.com/allenai/allennlp}{\texttt{allennlp}} \cite{wachter2017counterfactual}, \cite{grad_attr}, \cite{wallace2019allennlp}.
 
@@
 
+\subsection{Model Transparency, Appeal, and Compliance}
+
+Interpretable model architectures and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions.
+As seen in Sections \ref{sec:m_and_m} and \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain.
+Post-hoc explanations generally help summarize model mechanisms and decisions.
+Together they allow human learning from ML, certain types of regulatory compliance, and crucially, logical human appeal or override of automated model decisions \cite{art_and_sci}.
+Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be require by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention.
+ML model are often wrong.\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to appeal or override inevitable wrong decisions is not only a frequent prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
+
 \subsection{Impact of Discrimination Testing on Model Use and Adoption}
@@
 However, evidence is mounting that these harms can also manifest as complex intersectional challenges e.g.
-the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.\footnote{See: \href{https://www.telegraph.co.uk/technology/2016/03/24/microsofts-teen-girl-ai-turns-into-a-hitler-loving-sex-robot-wit/}{\textit{Microsoft deletes 'teen girl' AI after it became a Hitler-loving sex robot within 24 hours}}.} (While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.
+\footnote{See: \href{https://www.telegraph.co.uk/technology/2016/03/24/microsofts-teen-girl-ai-turns-into-a-hitler-loving-sex-robot-wit/}{\textit{Microsoft deletes 'teen girl' AI after it became a Hitler-loving sex robot within 24 hours}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\footnote{See: \href{https://www.theatlantic.com/technology/archive/2012/03/the-philosopher-whose-fingerprints-are-all-over-the-ftcs-new-approach-to-privacy/254365/}{\textit{The Philosopher Whose Fingerprints Are All Over the FTC's New Approach to Privacy}}.}\textsuperscript{,}\footnote{See: \textcolor{red}{Andrew Burt HBR article - upcoming.}}) Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
-Instead, it aims to highlight these issues as both singular entities and intersectional phenomena.
-Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as remedies to the black-box, and discrimination issues in ML.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \textcolor{red}{Andrew Burt HBR article - upcoming.}} Again, Sections \ref{sec:m_and_m} and \ref{sec:res} make no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+Instead, this text aims to highlight these issues as both singular entities and intersectional phenomena.
+Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box, and discrimination issues in ML.
 However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks and that explanations can mask ML discrimination.
 Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory.
-Of course, accuracy, discrimination, and security characteristics of a system can change over time, and simply testing for these problems at training time, as presented herein, is not adequate for high-stakes, human-centered, or regulated ML systems.
+Of course, accuracy, discrimination, and security characteristics of a system can change over time too, and simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems.
 Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
@@
 This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
-These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible machine learning workflow high-stakes, human-centered, or regulated ML applications.
+These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible machine learning work flow high-stakes, human-centered, or regulated ML applications.
 The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects.
-Of course there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and this manuscript also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
+Of course there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
 While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work and many others illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{Wen Phan for work in formalizing notation. \textcolor{red}{BLDS editor} for editing.}
+\acknowledgments{Wen Phan for work in formalizing notation. \textcolor{red}{BLDS editor} for editing.} Andrew Burt for ideas around ML transparency, model debugging, and the transparency paradox.
 
@@
 \abbreviations{The following abbreviations are used in this text: AI -- artificial intelligence, ANN -- artificial neural network, DI -- disparate impact, EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e.
-variants GAMs that consider two-way interactions and incorporate boosting into training, GAM -- generalized additive model, GBM -- gradient boosting machine, ICE -- individual conditional expectation, MGBM -- monotonic gradient boosting machine, ML -- machine learning, PD -- partial dependence, SGD -- stochastic gradient descent, US -- United States, XNN -- explainable neural network.}
+variants GAMs that consider two-way interactions and incorporate boosting into training, GAM -- generalized additive model, GBM -- gradient boosting machine, ICE -- individual conditional expectation, MGBM -- monotonic gradient boosting machine, ML -- machine learning, PD -- partial dependence, SGD -- stochastic gradient descent, SR -- Supervision and Regulation, US -- United States, XNN -- explainable neural network.}
 

### 2019-12-11 article-information-2019.tex
ore edits, some results

@@
 ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
-ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}.
+ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
 While this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
@@
 \midrule
-$g^\text{GBM}$		& 				& 			& 				& \\
-$g^\text{MGBM}$		& 				& 			& 				& \\
+$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
+$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
 $g^\text{ANN}$		& 		& 			& 				& \\
@@
 \midrule	
-$g^\text{GBM}$		& 				& 			& 				& \\
-$g^\text{MGBM}$		& 				& 			& 				& \\
+$g^\text{GBM}$		& 0.795	& 0.828	& 0.252	& 0.276 \\
+$g^\text{MGBM}$	& 0.765	& 0.814	& 0.259	& 0.278 \\
 $g^\text{ANN}$		& 		& 		& 		& \\
-$g^\text{XNN}$		& 				& 			& 				& \\
+$g^\text{XNN}$		& 		& 0.868	& 0.233	& 0.263 \\
 \bottomrule
@@
 
-\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g. \texttt{\url{https://github.com/openai/gpt-2}}, \texttt{\url{https://modelcards.withgoogle.com/object-detection}}.
+\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.
+\texttt{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}, \texttt{\url{https://modelcards.withgoogle.com/object-detection}}.
 
@@
 
-\subsection{Model Transparency, Appeal, and Compliance}
+\subsection{Interpretability, Explainability, Appeal, and Compliance}
 
@@
 As seen in Sections \ref{sec:m_and_m} and \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain.
-Post-hoc explanations generally help summarize model mechanisms and decisions.
-Together they allow human learning from ML, certain types of regulatory compliance, and crucially, logical human appeal or override of automated model decisions \cite{art_and_sci}.
+Post-hoc explanations generally help describe and summarize mechanisms and decisions, potentially yielding an even clearer understanding of ML models.
+Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}.
 Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be require by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention.
-ML model are often wrong.\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to appeal or override inevitable wrong decisions is not only a frequent prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention.
+ML models are often wrong.\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
 
@@
 
-\subsection{Intersectionality of Interpretability, Fairness, and Security in ML}
+\subsection{Intersectionality of Interpretability, Explainability, Discrimination, and Security in ML}
 
 The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own.
-However, evidence is mounting that these harms can also manifest as complex intersectional challenges e.g.
-the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}.
-\footnote{See: \href{https://www.telegraph.co.uk/technology/2016/03/24/microsofts-teen-girl-ai-turns-into-a-hitler-loving-sex-robot-wit/}{\textit{Microsoft deletes 'teen girl' AI after it became a Hitler-loving sex robot within 24 hours}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g.
+the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \textcolor{red}{Andrew Burt HBR article - upcoming.}} Again, Sections \ref{sec:m_and_m} and \ref{sec:res} make no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
-Instead, this text aims to highlight these issues as both singular entities and intersectional phenomena.
-Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box, and discrimination issues in ML.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \textcolor{red}{Andrew Burt HBR article - upcoming.}} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena.
+Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
 However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks and that explanations can mask ML discrimination.
 Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory.
-Of course, accuracy, discrimination, and security characteristics of a system can change over time too, and simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems.
+Accuracy, discrimination, and security characteristics of a system can change over time as well.
+Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems.
 Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
@@
 This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
-These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible machine learning work flow high-stakes, human-centered, or regulated ML applications.
+These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible ML work flow for high-stakes, human-centered, or regulated ML applications.
 The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects.
 Of course there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
-While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work and many others illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
+While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work (and many, many others) illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{Wen Phan for work in formalizing notation. \textcolor{red}{BLDS editor} for editing.} Andrew Burt for ideas around ML transparency, model debugging, and the transparency paradox.
+\acknowledgments{Wen Phan for work in formalizing notation. \textcolor{red}{BLDS editor} for editing. Andrew Burt for ideas around ML transparency, model debugging, and the transparency paradox.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank. Wells Fargo is a customer and investor of H2O.ai and a customer of BLDS, LLC.}
+\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank. Wells Fargo is a customer of, and investor in, H2O.ai and a customer of BLDS, LLC.}
 
@@
 %% optional
-\abbreviations{The following abbreviations are used in this text: AI -- artificial intelligence, ANN -- artificial neural network, DI -- disparate impact, EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e.
-variants GAMs that consider two-way interactions and incorporate boosting into training, GAM -- generalized additive model, GBM -- gradient boosting machine, ICE -- individual conditional expectation, MGBM -- monotonic gradient boosting machine, ML -- machine learning, PD -- partial dependence, SGD -- stochastic gradient descent, SR -- Supervision and Regulation, US -- United States, XNN -- explainable neural network.}
+\abbreviations{The following abbreviations are used in this text: 
+AI -- artificial intelligence, 
+AIR - adverse impact ratio,
+ALE - accumulated local effect,
+ANN -- artificial neural network, 
+DI -- disparate impact, 
+EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e. variants GAMs that consider two-way interactions and may incorporate boosting into training, 
+ECOA - Equal Credit Opportunity Act,
+EDA -- exploratory data analysis,
+EU -- European Union,
+FCRA -- Fair Credit Reporting Act,
+GAM -- generalized additive model, 
+GBM -- gradient boosting machine, 
+GDPR - General Data Protection Regulation,
+HMDA -- Home Mortgage Disclosure Act
+ICE -- individual conditional expectation, 
+MGBM -- monotonic gradient boosting machine, 
+ML -- machine learning, 
+PD -- partial dependence, 
+SGD -- stochastic gradient descent, 
+SHAP -- Shapley additive explanation,
+SMD - standardized mean difference,
+SR -- supervision and regulation, 
+US -- United States, 
+XNN -- explainable neural network.}
 

### 2019-12-13 article-information-2019.tex
some progress on figures

@@
 \label{tab:sim_acc}
-\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
+\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Accuracy is measured at the best F1 threshold for each model.}
 \centering
-%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
 \begin{tabular}{ccccc}
@@
 $g^\text{ANN}$		& 		& 		& 		& \\
-$g^\text{XNN}$		& 		& 			& 				& \\
+$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
 \bottomrule
@@
 \label{tab:mort_acc}
-\caption{Accuracy metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data.}
+\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Accuracy is measured at the best F1 threshold for each model.}
 \centering
@@
 $g^\text{ANN}$		& 		& 		& 		& \\
-$g^\text{XNN}$		& 		& 0.868	& 0.233	& 0.263 \\
+$g^\text{XNN}$		& 0.869	& 0.868	& 0.233	& 0.263 \\
 \bottomrule
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=8cm]{img/mort_mgbm_glob.png}
 \label{fig:mort_mgbm_glob}
-\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the mortgage test data.}
+\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the mortgage test data.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
 \end{figure}
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=15cm]{img/mort_mgbm_glob_pdp_ice.png}
 \label{fig:mort_mgbm_glob_pdp_ice}
-\caption{PD, ICE across deciles, and histograms for the three most important input features for $g^\text{MGBM}$ on the mortgage test data.}
+\caption{PD, ICE across selected percentiles, and histograms for the three most important input features for $g^\text{MGBM}$ on the mortgage test data.}
 \end{figure}
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=15cm]{img/mort_mgbm_loc.png}
 \label{fig:mort_mgbm_loc}
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=8cm]{img/mort_xnn_glob.png}
 \label{fig:mort_xnn_glob}
-\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the mortgage test data.}
+\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}$ on the mortgage test data.
+Deep SHAP values are reported in the probability space, after the application of the logit link function.}
 \end{figure}

### 2019-12-14 article-information-2019.tex
more figures

@@
 $g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
-$g^\text{ANN}$		& 		& 		& 		& \\
+$g^\text{ANN}$		& 0.757	& 0.850	& 0.480 	& 0.398 \\
 $g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
 \label{fig:sim_mgbm_glob}
-\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.}
+\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
 \end{figure}
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
 \label{fig:sim_mgbm_glob_pdp_ice}
-\caption{PD, ICE across deciles, and histograms for the three most important input features for $g^\text{MGBM}$ on the simulated test data.}
+\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
 \end{figure}
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=8cm]{img/sim_xnn_glob.png}
 \label{fig:sim_xnn_glob}
@@
 $g^\text{MGBM}$	& 0.765	& 0.814	& 0.259	& 0.278 \\
-$g^\text{ANN}$		& 		& 		& 		& \\
+$g^\text{ANN}$		& 0.860	& 0.872 	& 0.230 	& 0.262 \\
 $g^\text{XNN}$		& 0.869	& 0.868	& 0.233	& 0.263 \\
@@
 \label{fig:mort_mgbm_glob_pdp_ice}
-\caption{PD, ICE across selected percentiles, and histograms for the three most important input features for $g^\text{MGBM}$ on the mortgage test data.}
+\caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
 \end{figure}
@@
 \label{fig:mort_mgbm_loc}
-\caption{Mean Tree SHAP values across quintiles for the three most important input features for $g^\text{MGBM}$ on the mortgage test data.}
+\caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
 \end{figure}
@@
 \label{fig:mort_xnn_glob_pdp_ice}
-\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the mortgage test data.}
+\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the mortgage test data. }
 \end{figure}

### 2019-12-14 article-information-2019.tex
more work on figures

@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=15cm]{img/sim_mgbm_loc.png}
 \label{fig:sim_mgbm_loc}
-\caption{Mean Tree SHAP values across quintiles for the three most important input features for $g^\text{MGBM}$ on the simulated test data.}
+\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
 \end{figure}
@@
 \label{fig:sim_xnn_loc}
-\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the simulated test data.}
+\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the simulated test data.
+Deep SHAP values are reported in the probability space, after the application of the logit link function.}
 \end{figure}

### 2019-12-14 data/scripts/di_testing.py
Updating disparity testing.

@@ -1,4 +1,3 @@
-
 """
 License
 Copyright 2019 Navdeep Gill, Patrick Hall, Kim Montgomery, Nick Schmidt
@@ -16,15 +15,20 @@ DISCLAIMER: This notebook is not legal compliance advice.
 import pandas as pd
 import numpy as np
 from typing import Union
+from sklearn import metrics
+from scipy import stats
 
 
-class DisparateImpactTesting(object):
-    def __init__(self, lower_value_favorable: bool,
-                 pg_names: Union[list, tuple] = ('black', 'hispanic' 'asian', 'female', 'older'),
-                 cg_names: Union[list, tuple] = ('white', 'white', 'white', 'male', 'younger'),
-                 pgcg_names: Union[list, tuple] = ('black', 'hispanic', 'asian', 'white',
-                                                   'female', 'male', 'older', 'younger'),
-                 ):
+class DisparityTesting(object):
+    def __init__(self,
+                 pg_names: Union[list, tuple],
+                 cg_names: Union[list, tuple],
+                 pgcg_names: Union[list, tuple],
+                 higher_score_favorable: bool = False):
+        self.pg_names = pg_names
+        self.cg_names = cg_names
+        self.pgcg_names = pgcg_names
+        self.higher_score_favorable = higher_score_favorable
         """
         This class creates metrics used to test disparate impact.  See each method for detail on the
         particular calculations.
@@ -48,46 +52,49 @@ class DisparateImpactTesting(object):
         :param pgcg_names: This is a unique list of the values contained in pg_names and cg_names.  It is only used
         as a way to make the output more readable: it specifies the order in which each group is displayed.
         """
-        self.pg_names = pg_names
-        self.cg_names = cg_names
-        self.lower_value_favorable = lower_value_favorable
-        self.pgcg_names = pgcg_names
 
-    def adverse_impact_ratio(self,
-                             data: pd.DataFrame,
-                             outcome: str,
-                             ) -> pd.DataFrame:
-        """
-        This calculates the Adverse Impact Ratio (AIR), a standard measure of disparate impact used in
-        U.S. legal and regulatory settings.  It is defined as:
-            AIR = (% Protected Class Chosen) / (% Control Group Chosen)
+    def categorical_disparity_measures(self, data: pd.DataFrame, label: str, outcome: str):
+        res = pd.DataFrame({'class': self.pgcg_names}, index=self.pgcg_names)
+        for pi, ci in zip(self.pg_names, self.cg_names):
[…]

### 2019-12-14 data/scripts/disp_tests.py

@@ -1,0 +1,39 @@
+"""
+License
+Copyright 2019 Navdeep Gill, Patrick Hall, Kim Montgomery, Nick Schmidt
+
+Licensed under the Apache License, Version 2.0 (the "License"); you may not use this file except in
+compliance with the License. You may obtain a copy of the License at
+http://www.apache.org/licenses/LICENSE-2.0
+Unless required by applicable law or agreed to in writing, software distributed under the License is distributed
+on an "AS IS" BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied. See the License
+for the specific language governing permissions and limitations under the License.
+
+DISCLAIMER: This notebook is not legal compliance advice.
+"""
+
+import pandas as pd
+import numpy as np
+import di_testing.DisparityTesting
+from di_testing import DisparityTesting
+
+pd.set_option('display.max_columns', 30)
+pd.set_option('display.width', 200)
+
+hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
+
+pg_names = ["black", "amind", "hispanic", "female"]
+cg_names = ["white", "white", "non_hispanic", "male"]
+pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
+predicted = "high_priced_mgbm_pred"
+label = "high_priced"
+outcome = "decision"
+pred_prob_for_outcome = 0.20
+higher_score_favorable = False
+
+hmda[outcome] = np.where(hmda[predicted] >= pred_prob_for_outcome, 1, 0)
+hmda[outcome].value_counts(normalize=True)
+
+data = hmda[["Id", predicted, outcome, label] + pgcg_names]
+
+

### 2019-12-14 data/scripts/di_testing.py
Updating disparity testing.

@@ -24,7 +24,7 @@ class DisparityTesting(object):
                  pg_names: Union[list, tuple],
                  cg_names: Union[list, tuple],
                  pgcg_names: Union[list, tuple],
-                 higher_score_favorable: bool = False):
+                 higher_score_favorable: bool):
         self.pg_names = pg_names
         self.cg_names = cg_names
         self.pgcg_names = pgcg_names
@@ -59,7 +59,7 @@ class DisparityTesting(object):
             res.loc[res["class"] == pi, "control"] = ci
 
         for groupi in self.pgcg_names:
-            data_cm = data[~np.isnan(data[groupi])]
+            data_cm = data[data[groupi].notna()]
             cm = metrics.confusion_matrix(y_true=data_cm[label], y_pred=data_cm[outcome], sample_weight=data_cm[groupi])
             tn, fp, fn, tp = cm.ravel()
             res.loc[res["class"] == groupi, "total"] = (fp + tp + fn + fp)
@@ -87,14 +87,42 @@ class DisparityTesting(object):
         res["favorable"] = res["selected"] if self.higher_score_favorable else res["total"] - res["selected"]
         res["percent_selected"] = res["selected"] / res["total"]
         res["percent_favorable"] = res["favorable"] / res["total"]
-        res.loc[res["class"].isin(self.pg_names), "air"] = np.array(res["percent_favorable"][self.pg_names]) / \
-                                                           np.array(res["percent_favorable"][self.cg_names])
+
         res.loc[res["class"].isin(self.pg_names), "control_total"] = np.array(res["total"][self.cg_names])
         res.loc[res["class"].isin(self.pg_names), "control_selected"] = np.array(res["selected"][self.cg_names])
+        res.loc[res["class"].isin(self.pg_names), "control_percent_favorable"] = \
+            np.array(res["percent_favorable"][self.cg_names])
+
+        res["marginal_difference"] = res["control_percent_favorable"] - res["percent_favorable"]
+        res["shortfall"] = res["marginal_difference"] * res["total"]
+        res["air"] = res["percent_favorable"] / res["control_percent_favorable"]
         for fishi in self.pg_names:
             fishers_values = stats.fisher_exact(np.array(
                 res.loc[res["class"] == fishi, ["total", "selected",
                                                 "control_total", "control_selected"]]).reshape(2, 2))
[…]

### 2019-12-14 data/scripts/disp_tests.py

@@ -14,8 +14,7 @@ DISCLAIMER: This notebook is not legal compliance advice.
 
 import pandas as pd
 import numpy as np
-import di_testing.DisparityTesting
-from di_testing import DisparityTesting
+from data.scripts.di_testing import DisparityTesting
 
 pd.set_option('display.max_columns', 30)
 pd.set_option('display.width', 200)
@@ -35,5 +34,12 @@ hmda[outcome] = np.where(hmda[predicted] >= pred_prob_for_outcome, 1, 0)
 hmda[outcome].value_counts(normalize=True)
 
 data = hmda[["Id", predicted, outcome, label] + pgcg_names]
+
+disp_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=pgcg_names,
+                              higher_score_favorable=higher_score_favorable)
+
+cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome=outcome)
+cont_outcomes = disp_tests.continuous_disparity_measures(data=data, predicted=predicted)
+
 
 

### 2019-12-14 data/scripts/di_testing.py
Updating disparity testing.

@@ -30,10 +30,10 @@ class DisparityTesting(object):
         self.pgcg_names = pgcg_names
         self.higher_score_favorable = higher_score_favorable
         """
-        This class creates metrics used to test disparate impact.  See each method for detail on the
+        This class creates metrics used to test various metrics of fairness.  See each method for detail on the
         particular calculations.
 
-        :param lower_value_favorable: Boolean specifying whether a lower outcome value being tested
+        :param higher_score_favorable: Boolean specifying whether a higher outcome value being tested
         is considered favorable.  For example, a marketing offer would generally be associated with a score
         where a higher value is favorable.  For credit, generally a lower score is favorable, as the score
         is typically a measure of the probability of default.
@@ -54,6 +54,16 @@ class DisparityTesting(object):
         """
 
     def categorical_disparity_measures(self, data: pd.DataFrame, label: str, outcome: str):
+        """
+        This method calculates various disparity measures commonly used to assess fairness or discrimination in
+        processes that lead to categorical outcomes.
+
+        @param data: This is the Pandas data frame that contains the protected and control group information for
+        each observation being tested, along with the given categorical outcome and the true outcome
+        @param label: name of the true outcome variable in the data frame
+        @param outcome: name of the given categorical outcome in the data frame
+        @return: Pandas data frame containing a number of disparity measures
+        """
         res = pd.DataFrame({'class': self.pgcg_names}, index=self.pgcg_names)
         for pi, ci in zip(self.pg_names, self.cg_names):
             res.loc[res["class"] == pi, "control"] = ci
@@ -95,19 +105,30 @@ class DisparityTesting(object):
 
         res["marginal_difference"] = res["control_percent_favorable"] - res["percent_favorable"]
         res["shortfall"] = res["marginal_difference"] * res["total"]
-        res["air"] = res["percent_favorable"] / res["control_percent_favorable"]
+        res["adverse_impact_ratio"] = res["percent_favorable"] / res["control_percent_favorable"]
         for fishi in self.pg_names:
             fishers_values = stats.fisher_exact(np.array(
                 res.loc[res["class"] == fishi, ["total", "selected",
                                                 "control_total", "control_selected"]]).reshape(2, 2))
-            res.loc[res["class"] == fishi, "fishers_value"] = fishers_values[0]
-            res.loc[res["class"] == fishi, "fishers_p"] = fishers_values[1]
[…]

### 2019-12-14 data/scripts/disp_tests.py

@@ -27,7 +27,7 @@ pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "
 predicted = "high_priced_mgbm_pred"
 label = "high_priced"
 outcome = "decision"
-pred_prob_for_outcome = 0.20
+pred_prob_for_outcome = 0.30
 higher_score_favorable = False
 
 hmda[outcome] = np.where(hmda[predicted] >= pred_prob_for_outcome, 1, 0)
@@ -39,6 +39,7 @@ disp_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=p
                               higher_score_favorable=higher_score_favorable)
 
 cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome=outcome)
+print(cat_outcomes)
 cont_outcomes = disp_tests.continuous_disparity_measures(data=data, predicted=predicted)
 
 

### 2019-12-14 data/scripts/di_testing.py
Updating disparity testing.

@@ -72,7 +72,7 @@ class DisparityTesting(object):
             data_cm = data[data[groupi].notna()]
             cm = metrics.confusion_matrix(y_true=data_cm[label], y_pred=data_cm[outcome], sample_weight=data_cm[groupi])
             tn, fp, fn, tp = cm.ravel()
-            res.loc[res["class"] == groupi, "total"] = (fp + tp + fn + fp)
+            res.loc[res["class"] == groupi, "total"] = (fp + tp + fn + tn)
             res.loc[res["class"] == groupi, "selected"] = (tp + fn)
             res.loc[res["class"] == groupi, "true_positive"] = tp
             res.loc[res["class"] == groupi, "true_negative"] = tn
@@ -112,7 +112,7 @@ class DisparityTesting(object):
                                                 "control_total", "control_selected"]]).reshape(2, 2))
             res.loc[res["class"] == fishi, "fishers_exact"] = fishers_values[0]
             res.loc[res["class"] == fishi, "p_value"] = fishers_values[1]
-        res.reset_index(inplace=True)
+        res.index = range(len(res))
         return res
 
     def continuous_disparity_measures(self,
@@ -145,5 +145,5 @@ class DisparityTesting(object):
             t_test = stats.ttest_ind(data.loc[data[pgi] == 1, predicted], data.loc[data[cgi] == 1, predicted])
             res.loc[res["class"] == pgi, "t_statistic"] = t_test[0]
             res.loc[res["class"] == pgi, "p_value"] = t_test[1]
-        res.reset_index(inplace=True)
+        res.index = range(len(res))
         return res

### 2019-12-14 data/scripts/disp_tests.py

@@ -21,9 +21,13 @@ pd.set_option('display.width', 200)
 
 hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
 
-pg_names = ["black", "amind", "hispanic", "female"]
-cg_names = ["white", "white", "non_hispanic", "male"]
-pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
+# pg_names = ["black", "amind", "hispanic", "female"]
+# cg_names = ["white", "white", "non_hispanic", "male"]
+# pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
+pg_names = ["black", "female"]
+cg_names = ["white", "male"]
+pgcg_names = ["black", "white", "female", "male"]
+
 predicted = "high_priced_mgbm_pred"
 label = "high_priced"
 outcome = "decision"
@@ -39,8 +43,10 @@ disp_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=p
                               higher_score_favorable=higher_score_favorable)
 
 cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome=outcome)
-print(cat_outcomes)
+
+cat_outcomes = cat_outcomes[['class', 'control', 'total', 'selected', 'false_positive_rate', 'false_negative_rate',
+                             'accuracy', 'percent_favorable', 'control_percent_favorable', 'marginal_difference',
+                             'shortfall', 'adverse_impact_ratio', 'fishers_exact', 'p_value']]
+
 cont_outcomes = disp_tests.continuous_disparity_measures(data=data, predicted=predicted)
-
-
 

### 2019-12-14 data/scripts/di_testing.py
Updating disparity testing.

@@ -111,8 +111,7 @@ class DisparityTesting(object):
                 res.loc[res["class"] == fishi, ["total", "selected",
                                                 "control_total", "control_selected"]]).reshape(2, 2))
             res.loc[res["class"] == fishi, "fishers_exact"] = fishers_values[0]
-            res.loc[res["class"] == fishi, "p_value"] = fishers_values[1]
-        res.index = range(len(res))
+            res.loc[res["class"] == fishi, "fishers_exact_p_value"] = fishers_values[1]
         return res
 
     def continuous_disparity_measures(self,
@@ -144,6 +143,27 @@ class DisparityTesting(object):
         for pgi, cgi in zip(self.pg_names, self.cg_names):
             t_test = stats.ttest_ind(data.loc[data[pgi] == 1, predicted], data.loc[data[cgi] == 1, predicted])
             res.loc[res["class"] == pgi, "t_statistic"] = t_test[0]
-            res.loc[res["class"] == pgi, "p_value"] = t_test[1]
-        res.index = range(len(res))
+            res.loc[res["class"] == pgi, "t_statistic_p_value"] = t_test[1]
         return res
+    
+    @staticmethod
+    def create_combined_output(cat_outcomes: pd.DataFrame,
+                               cont_outcomes: pd.DataFrame,
+                               cat_vars: Union[list, tuple] = ('class', 'control', 'total', 'false_positive_rate',
+                                                               'false_negative_rate', 'accuracy',
+                                                               'marginal_difference', 'shortfall',
+                                                               'adverse_impact_ratio', 'fishers_exact',
+                                                               'fishers_exact_p_value'),
+                               cont_vars: Union[list, tuple] = ('class', 'control', 'standardized_mean_difference', 
+                                                                't_statistic', 't_statistic_p_value'),
+                               ):
+
+        disp_outcomes = pd.merge(cat_outcomes[list(cat_vars)], cont_outcomes[list(cont_vars)], on=["class", "control"])
+        disp_outcomes["class"] = disp_outcomes["class"].str.replace("_", "-").str.title()
+        disp_outcomes["control"] = disp_outcomes["control"].str.replace("_", "-").str.title()
+
+        disp_outcomes.columns = disp_outcomes.columns.str.title()
+        for old, new in zip(["_", "P Value", "T Statistic"], [" ", "P-Value", "T-Statistic"]):
+            disp_outcomes.columns = disp_outcomes.columns.str.replace(old, new)
+        return disp_outcomes
+

### 2019-12-14 data/scripts/disp_tests.py

@@ -21,12 +21,12 @@ pd.set_option('display.width', 200)
 
 hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
 
-# pg_names = ["black", "amind", "hispanic", "female"]
-# cg_names = ["white", "white", "non_hispanic", "male"]
-# pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
-pg_names = ["black", "female"]
-cg_names = ["white", "male"]
-pgcg_names = ["black", "white", "female", "male"]
+pg_names = ["black", "amind", "hispanic", "female"]
+cg_names = ["white", "white", "non_hispanic", "male"]
+pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
+# pg_names = ["black", "female"]
+# cg_names = ["white", "male"]
+# pgcg_names = ["black", "white", "female", "male"]
 
 predicted = "high_priced_mgbm_pred"
 label = "high_priced"
@@ -43,10 +43,7 @@ disp_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=p
                               higher_score_favorable=higher_score_favorable)
 
 cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome=outcome)
-
-cat_outcomes = cat_outcomes[['class', 'control', 'total', 'selected', 'false_positive_rate', 'false_negative_rate',
-                             'accuracy', 'percent_favorable', 'control_percent_favorable', 'marginal_difference',
-                             'shortfall', 'adverse_impact_ratio', 'fishers_exact', 'p_value']]
-
 cont_outcomes = disp_tests.continuous_disparity_measures(data=data, predicted=predicted)
+disparity_measures = DisparityTesting.create_combined_output(cat_outcomes=cat_outcomes, cont_outcomes=cont_outcomes)
+
 

### 2019-12-14 data/scripts/disp_tests.py
Updating disparity testing.

@@ -21,12 +21,12 @@ pd.set_option('display.width', 200)
 
 hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
 
-pg_names = ["black", "amind", "hispanic", "female"]
-cg_names = ["white", "white", "non_hispanic", "male"]
-pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
-# pg_names = ["black", "female"]
-# cg_names = ["white", "male"]
-# pgcg_names = ["black", "white", "female", "male"]
+# pg_names = ["black", "amind", "hispanic", "female"]
+# cg_names = ["white", "white", "non_hispanic", "male"]
+# pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
+pg_names = ["black", "female"]
+cg_names = ["white", "male"]
+pgcg_names = ["black", "white", "female", "male"]
 
 predicted = "high_priced_mgbm_pred"
 label = "high_priced"

### 2019-12-15 data/scripts/disp_tests.py
Updating disparity testing.

@@ -15,35 +15,46 @@ DISCLAIMER: This notebook is not legal compliance advice.
 import pandas as pd
 import numpy as np
 from data.scripts.di_testing import DisparityTesting
+import os
+import matplotlib.pyplot as plt
 
 pd.set_option('display.max_columns', 30)
 pd.set_option('display.width', 200)
 
-hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
+if __name__ == '__main__':
 
-# pg_names = ["black", "amind", "hispanic", "female"]
-# cg_names = ["white", "white", "non_hispanic", "male"]
-# pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
-pg_names = ["black", "female"]
-cg_names = ["white", "male"]
-pgcg_names = ["black", "white", "female", "male"]
+    pg_names = ["black", "native_american", "asian", "hispanic", "female"]
+    cg_names = ["white", "white", "white", "non_hispanic", "male"]
+    pgcg_names = ["black", "native_american", "asian", "white", "hispanic", "non_hispanic", "female", "male"]
+    # pg_names, cg_names, pgcg_names = ["black", "female"], ["white", "male"], ["black", "white", "female", "male"]
 
-predicted = "high_priced_mgbm_pred"
-label = "high_priced"
-outcome = "decision"
-pred_prob_for_outcome = 0.30
-higher_score_favorable = False
+    file_name = 'test_hmda_with_preds.csv'
+    predicted = "high_priced_mgbm_pred"
+    model_name = "HMDA Monotonic GBM"
+    label = "high_priced"
+    outcome = "decision"
+    pred_prob_for_outcome = 0.30
+    higher_score_favorable = False
 
-hmda[outcome] = np.where(hmda[predicted] >= pred_prob_for_outcome, 1, 0)
-hmda[outcome].value_counts(normalize=True)
+    data = pd.read_csv(os.path.join('./data/output/', file_name))
+    if "native_american" in pg_names:
+        data.rename(inplace=True, columns={'amind': 'native_american'})
+    data[outcome] = np.where(data[predicted] >= pred_prob_for_outcome, 1, 0)
+    data[outcome].value_counts(normalize=True)
 
-data = hmda[["Id", predicted, outcome, label] + pgcg_names]
+    disparity_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=pgcg_names,
[…]

### 2019-12-16 article-information-2019.tex
fix up links, compress images, more figures

@@
 ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
-ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
+ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=15cm]{img/sim_xnn_loc.png}
 \label{fig:sim_xnn_loc}
-\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the simulated test data.
+\caption{Deep SHAP values for three observations across selected percentiles of  $g^\text{XNN}$ on the simulated test data.
 Deep SHAP values are reported in the probability space, after the application of the logit link function.}
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=15cm]{img/mort_xnn_loc.png}
 \label{fig:mort_xnn_loc}
-\caption{Mean Deep SHAP values across quintiles for the three most important input features for $g^\text{XNN}$ on the mortgage test data.}
+\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}$ on the mortgage test data.
+Deep SHAP values are reported in the probability space, after the application of the logit link function.}
 \end{figure}
@@
 While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML.
-Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.\footnote{See: \texttt{\url{https://github.com/jphall663/hc_ml}} for details regarding the work flow in Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.
+Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.
 
@@
 In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
-Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{Optimal sparse decision trees: \url{https://github.com/xiyanghu/OSDT}.}\textsuperscript{,}\footnote{\textit{This looks like that} interpretable deep learning: \url{https://github.com/cfchen-duke/ProtoPNet}.} Additional, relevant references and Python functionality include:
+Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{See: \href{https://github.com/xiyanghu/OSDT}{Optimal sparse decision trees}.}\textsuperscript{,}\footnote{See: \href{https://github.com/cfchen-duke/ProtoPNet}{\textit{This looks like that} interpretable deep learning}.} Additional, relevant references and Python functionality include:
 
@@
 \item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.
-\texttt{\url{https://github.com/openai/gpt-2/blob/master/model_card.md}}, \texttt{\url{https://modelcards.withgoogle.com/object-detection}}.
+\href{https://github.com/openai/gpt-2/blob/master/model_card.md}{GPT-2 model card}, \href{https://modelcards.withgoogle.com/object-detection}{Object Detection model card}.
 
@@
 
-\noindent See: \texttt{\url{https://github.com/jphall663/awesome-machine-learning-interpretability}} for a longer, curated list of related software packages and resources.
+\noindent See: \href{https://github.com/jphall663/awesome-machine-learning-interpretability}{Awesome Machine Learning Interpretability} for a longer, community-curated metalist of related software packages and resources.
 
@@
 Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}.
-Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
+Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
 Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention.
@@
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \textcolor{red}{Andrew Burt HBR article - upcoming.}} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
 Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena.

### 2019-12-17 data/scripts/disp_tests.py
Remove old DI testing.

@@ -14,7 +14,7 @@ DISCLAIMER: This notebook is not legal compliance advice.
 
 import pandas as pd
 import numpy as np
-from data.scripts.di_testing import DisparityTesting
+from notebooks.scripts.di_testing import DisparityTesting
 
 pd.set_option('display.max_columns', 30)
 pd.set_option('display.width', 200)
@@ -29,6 +29,7 @@ cg_names = ["white", "male"]
 pgcg_names = ["black", "white", "female", "male"]
 
 predicted = "high_priced_mgbm_pred"
+apredicted = "high_priced_gbm_pred"
 label = "high_priced"
 outcome = "decision"
 pred_prob_for_outcome = 0.30

### 2019-12-17 article-information-2019.tex
ridge function and weight figures
new and improved xnn features, description of HMDA data

@@
 
-The training data contains 33 total features and 144,000 rows, each representing a unique loan, and a fold identifier to ensure consistent 5-fold cross-validation accuracy and error measurements across different types of models.
-Consumer finance and loan descriptors are used for training.
-Demographic features are not used in model training.
-The mortgage test data contains 36,000 loans.
+The US HMDA, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
+This information is first provided to the Consumer Financial Protection Bureau (CFPB), which subsequently releases some of the data to the public.
+Regulators often use HMDA data to, ``...show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory.''\textsuperscript{5} In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
+The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, loan to value (LTV) ratio, debt to income (DTI) ratio, and other important financial descriptors.
+The data also include information on each borrower and co-borrower’s race, ethnicity, gender, and age.
+Because the data includes information on these protected class characteristics, certain metrics that can be indicative of discrimination in lending can be calculated directly using the HDMA data.
+
+The mortgage dataset analyzed here is a random sample of 100,000 consumer-anonymized loans from the HDMA database, separated into a training set and a test set.
+These loans are a subset of all originated mortgage loans in the 2018 HMDA data that were chosen to represent a relatively comparable group of consumer mortgages.
+A selection of features is used to predict whether a loan is \textit{high-priced}, i.e.
+the annual percentage rate (APR) charged was 150 basis points (1.5\%) or more above a survey-based estimate of other similar loans offered around the time of the given loan.
+While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (LTV ratio, DTI ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
+Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in this text.
 
@@
 \centering
-%\includegraphics[width=8cm]{img/}
-\label{fig:sim_xnn_glob_pdp_ice}
-\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the simulated test data.}
+\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
+\label{fig:sim_xnn_glob_ridge}
+\caption{\textbf{A.} Output layer weights, \textbf{B.} corresponding ridge functions, and \textbf{C.} corresponding projection layer weights for $g^\text{XNN}$ on the simulated test data.}
 \end{figure}
@@
 $g^\text{MGBM}$	& 0.765	& 0.814	& 0.259	& 0.278 \\
-$g^\text{ANN}$		& 0.860	& 0.872 	& 0.230 	& 0.262 \\
+$g^\text{ANN}$		& 0.865	& 0.871 	& 0.231 	& 0.262 \\
 $g^\text{XNN}$		& 0.869	& 0.868	& 0.233	& 0.263 \\
@@
 \centering
-%\includegraphics[width=8cm]{img/}
+\includegraphics[width=16cm]{img/mort_xnn_glob_ridge.png}
 \label{fig:mort_xnn_glob_pdp_ice}
-\caption{Ridge functions for the three most important input features for $g^\text{XNN}$ on the mortgage test data. }
+\caption{\textbf{A.} Output layer weights, \textbf{B.} corresponding ridge functions, and \textbf{C.} corresponding projection layer weights for $g^\text{XNN}$ on the mortgage test data.}
 \end{figure}
@@
 ANN -- artificial neural network, 
+APR -- annual percentage rate,
+CFPB -- Consumer Financial Protection Bureau,
 DI -- disparate impact, 
+DTI -- debt to income,
 EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e. variants GAMs that consider two-way interactions and may incorporate boosting into training, 
@@
 ICE -- individual conditional expectation, 
+LTV -- loan to value,
 MGBM -- monotonic gradient boosting machine, 

### 2019-12-17 notebooks/scripts/ann_simulation_code_shapley.py
adding shapley to ann runs

@@ -1,0 +1,106 @@
+import numpy as np
+import pandas as pd
+import shap
+import subprocess
+import sys
+#import pydot
+
+
+import keras    
+from timeit import default_timer as timer
+import tensorflow as tf
+from keras import backend as K
+from keras.layers import Input, Dense
+from keras.models import Model, Sequential
+from keras.layers import Dense
+
+
+seed = 12345
+
+np.random.seed(seed)
+
+my_init = keras.initializers.RandomUniform(seed=seed)
+
+
+
+out_dir = "ann_output4/"
+
+# baseline model
+def ann_model():
+    
+    input = Input(shape=(features,), name='main_input')
+
+    out = Dense(10, input_dim=10, activation='relu')(input)
+    
+    out = Dense(5, input_dim=10, activation='relu')(out)
+    
+    out = Dense(1, activation='sigmoid')(out)
+    
+    model = Model(inputs=input, outputs=out)
+	# Compile model
+    model.compile(loss='binary_crossentropy', optimizer='adam')
+    
+    return model
+
+# Load the dataset
+xnn_data_dir = '~/article-information-2019/data/xnn_output/'
+#xnn_data_dir = ''
+DATA=pd.read_csv(xnn_data_dir + 'train_simulated_transformed.csv')
+print(list(DATA.columns))
+#DATA = DATA.iloc[0:10000,:]
+TEST=pd.read_csv(xnn_data_dir + 'test_simulated_transformed.csv')
+print(list(TEST.columns))
+
+
+selected_vars = ['binary1', 'binary2', 'cat1_0', 'cat1_1', 'cat1_2', 'cat1_3', 'cat1_4', 
+                 'fried1_std', 'fried2_std', 'fried3_std', 'fried4_std', 'fried5_std']
+
+target_var = 'outcome'
+
+X=DATA[selected_vars].values
+Y=DATA[target_var].values
+TEST_X = TEST[selected_vars].values
+TEST_Y = TEST[target_var].values
+features = X.shape[1]
+
+
+inputs = {'main_input': X}
+
+# Fit model
+model = ann_model()
+model.fit(inputs, Y, epochs=2000, batch_size=1024, validation_split=0, verbose=1)
+
+
+# Find the predictions and Shapley values on the test set.
+bg_samples = 1000
+background = DATA[selected_vars].iloc[np.random.choice(DATA[selected_vars].shape[0], bg_samples, replace=False)]
[…]

### 2019-12-17 data/scripts/disp_tests.py
delete old disparity measurement program

@@ -14,19 +14,19 @@ DISCLAIMER: This notebook is not legal compliance advice.
 
 import pandas as pd
 import numpy as np
-from notebooks.scripts.di_testing import DisparityTesting
+from notebooks.scripts.disparity_measurement import DisparityTesting
 
 pd.set_option('display.max_columns', 30)
 pd.set_option('display.width', 200)
 
 hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
 
-# pg_names = ["black", "amind", "hispanic", "female"]
-# cg_names = ["white", "white", "non_hispanic", "male"]
-# pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
-pg_names = ["black", "female"]
-cg_names = ["white", "male"]
-pgcg_names = ["black", "white", "female", "male"]
+pg_names = ["black", "amind", "hispanic", "female"]
+cg_names = ["white", "white", "non_hispanic", "male"]
+pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
+# pg_names = ["black", "female"]
+# cg_names = ["white", "male"]
+# pgcg_names = ["black", "white", "female", "male"]
 
 predicted = "high_priced_mgbm_pred"
 apredicted = "high_priced_gbm_pred"

### 2019-12-17 notebooks/disparity-testing-pre-notebook.py
update of disparity testing program

@@ -19,7 +19,7 @@ from notebooks.scripts.disparity_measurement import DisparityTesting
 pd.set_option('display.max_columns', 30)
 pd.set_option('display.width', 200)
 
-hmda = pd.read_csv('./data/output/test_hmda_with_preds.csv')
+file_name = './data/output/test_hmda_with_preds.csv'
 
 pg_names = ["black", "amind", "hispanic", "female"]
 cg_names = ["white", "white", "non_hispanic", "male"]
@@ -29,16 +29,16 @@ pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "
 # pgcg_names = ["black", "white", "female", "male"]
 
 predicted = "high_priced_mgbm_pred"
-apredicted = "high_priced_gbm_pred"
+predicted = "high_priced_gbm_pred"
 label = "high_priced"
 outcome = "decision"
-pred_prob_for_outcome = 0.30
+probability_for_classification = 0.30
 higher_score_favorable = False
 
-hmda[outcome] = np.where(hmda[predicted] >= pred_prob_for_outcome, 1, 0)
-hmda[outcome].value_counts(normalize=True)
 
-data = hmda[["Id", predicted, outcome, label] + pgcg_names]
+data = pd.read_csv(file_name)
+data[outcome] = np.where(data[predicted] >= probability_for_classification, 1, 0)
+data[outcome].value_counts(normalize=True)
 
 disp_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=pgcg_names,
                               higher_score_favorable=higher_score_favorable)

### 2019-12-17 article-information-2019.tex
some results section text and improved figures

@@
 Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e.
-ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
+ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
 ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
@@
 Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}.
-The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and \href{https://matplotlib.org/}{\texttt{matplotlib}} and \href{https://seaborn.pydata.org/}{\texttt{seaborn}} for plotting.
+The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.
 
@@
 
-\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}
+Results for constrained models on the simulated data are displayed in Subsections \ref{ssec: c_v_uc_sim} -- \ref{ssec:dis_sim}.
+{Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models.
+\textcolor{red}{...}
+
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_sim}
+
+Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+$g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
+Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
+$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
+Accuracy is measured at the best F1 threshold for each model.
 
 \begin{table}[H]
-\label{tab:sim_acc}
-\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Accuracy is measured at the best F1 threshold for each model.}
+\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
 \centering
@@
 \bottomrule
+\label{tab:sim_acc}
 \end{tabular}
@@
 \subsubsection{Interpretability and Post-hoc Explanation Results}
+
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the numeric value of the reported values can interpreted as the absolute mean impact of each $X_j$ on $g^\text{MGBM}$ in the margin space.
 
@@
 
+% highlight monotonicity 
+PD and ICE are always displayed with a histogram herein to highlight any sparse regions in an input feature's domain.
+Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
+Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
+
 \begin{figure}[H]
@@
 \end{figure}
-
-% highlight monotonicity
 
@@
 
+Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$.
+
 \begin{figure}[H]
@@
 \end{figure}
+
+Ridge functions are reminiscent of basis functions ...
 
@@
 
-\subsubsection{Discrimination Testing Results}
+\subsubsection{Discrimination Testing Results}\label{ssec:dis_sim}
 
@@
 
-\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}
+Results for the mortgage data are presented in Subsections \ref{ssec: c_v_uc_mort} -- \ref{ssec:dis_mort}.
+$g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants.
+Assuming that small fit differences do not outweigh the need for intrinsic model interpretability and post-hoc explainability in high-stakes, human-centered, or regulated applications, $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability and post-hoc explainability results are presented along with discrimination testing results.
+For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of monotonic prediction behavior, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
+For $g^\text{XNN}$ intrinsic interpretability manifests as plots of sparse weights in the projection and output layers and plots of subnetwork ridge functions.
+Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
+Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, SMD, and other measures.
+
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_mort}
+
+Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data.
+This is partially due to the need to present direct comparisons between post-hoc explainability results and the data preprocessing required to use neural networks and tensorflow, e.g.
+numerical encoding of categorical features and missing values.
+This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
+$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$.
 
 \begin{table}[H]
-\label{tab:mort_acc}
-\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Accuracy is measured at the best F1 threshold for each model.}
+\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data.}
 \centering
@@
 \bottomrule
+\label{tab:mort_acc}
 \end{tabular}
@@
 
+\noindent Regardless of the fit differences between the two families of hypothesis models, the difference between the fit of constrained and unconstrained variants within the two types of models is small for the GBMs and negligible for ANNs, 3\% and < 1\% worse fit respectively, averaged across the metrics reported in Table \ref{tab:mort_acc}.
+
 \subsubsection{Interpretability and Post-hoc Explanation Results}
+
+Global feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
+$g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
 
@@
 \includegraphics[width=8cm]{img/mort_mgbm_glob.png}
+\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the mortgage test data.}
 \label{fig:mort_mgbm_glob}
-\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the mortgage test data.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
 \end{figure}
+
+\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded flag features may contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
+
+Domain knowledge was used to positively constrain DTI ratio and LTV ratio and to negatively constrain income and the loan term flag under $g^\text{MGBM}$.
+The monotonicity constraints for DTI ratio and LTV ratio are confirmed on average and across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ in Figure \ref{fig:mort_mgbm_glob_pdp_ice}.
+Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles using ICE and on average with PD.
+Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
+Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$, which does allow for the modeling of non-monotonic interactions.
+DTI ratio and LTV ratio also appear to have sparse regions in their univariate distributions.
+The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as the $g^\text{MGBM}$ appears to carry reasonable predictions learned from more populated regions of both features' distributions into the more sparsely populated regions.
 
@@
 \includegraphics[width=15cm]{img/mort_mgbm_glob_pdp_ice.png}
+\caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
 \label{fig:mort_mgbm_glob_pdp_ice}
-\caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
 \end{figure}
 
-% highlight monotonicity
+Figure \ref{fig:mort_mgbm_glob_pdp_ice} also displays PD and ICE for the unconstrained feature property value.
+Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns of ICE curves.
+While PD shows monotonically increasing prediction behavior on average, apparently influenced by the extremely large predictions of high $g^{\text{MGBM}}(\mathbf{X})$ percentile individuals, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$ exhibit different prediction behavior with respect to property value.
+Some individuals at these lower percentiles display monotonically decreasing prediction behavior while others appear to show fluctuating prediction behavior.
+This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}(\mathbf{X})$ \cite{ice_plots}.
+Drivers of this interaction can often be elucidated using surrogate decision trees \cite{art_and_sci}.
+Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}(\mathbf{X})$ can learn.
+For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions.
+However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value.
+Such fluctuations can be caused by local data characteristics, modeled local interactions, or unfortunately, by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
+
+Local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in Figure \ref{fig:mort_mgbm_loc}.
+The selected individuals show the expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
 
@@
 \includegraphics[width=15cm]{img/mort_mgbm_loc.png}
-\label{fig:mort_mgbm_loc}
 \caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.
 Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
+\label{fig:mort_mgbm_loc}
 \end{figure}
+
+\noindent The deeper significance of Figure \ref{fig:mort_mgbm_loc} lies in the fact that any $g^\text{MGBM}(\mathbf{x})$ prediction can be accurately summarized in this manner, which is generally important for enabling appeal of model decisions, and specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to individuals.
+(In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.)
+
+Figure \ref{fig:mort_xnn_glob} displays global feature importance for $g^\text{XNN}$ on the mortgage test data.
+$g^\text{XNN}$ distributes importance more evenly across business drivers and puts strong emphasis on the no introductory rate period flag feature than does $g^\text{MGBM}$.
+Like $g^\text{MGBM}$, $g^\text{XNN}$ puts little emphasis on the other flag features.
 
@@
 \includegraphics[width=8cm]{img/mort_xnn_glob.png}
+\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}$ on the mortgage test data.}
 \label{fig:mort_xnn_glob}
-\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}$ on the mortgage test data.
-Deep SHAP values are reported in the probability space, after the application of the logit link function.}
 \end{figure}
+
+\noindent As compared to $g^\text{MGBM}$, $g^\text{XNN}$ assigns higher importance to property value and loan amount, and lower importance on LTV ratio, and the income related features, DTI ratio and income.
+
+The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
+\ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse weights of the $g^\text{XNN}$ output layer, in which subnetworks 0,1,2,3,5,8, and 9 have large magnitude weights and subnetworks 4,6, and 7 have small to near-zero magnitude weights.
 
@@
 \includegraphics[width=16cm]{img/mort_xnn_glob_ridge.png}
-\label{fig:mort_xnn_glob_pdp_ice}
 \caption{\textbf{A.} Output layer weights, \textbf{B.} corresponding ridge functions, and \textbf{C.} corresponding projection layer weights for $g^\text{XNN}$ on the mortgage test data.}
+\label{fig:mort_xnn_glob_ridge}
 \end{figure}
@@
 \label{fig:mort_xnn_loc}
-\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}$ on the mortgage test data.
-Deep SHAP values are reported in the probability space, after the application of the logit link function.}
+\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}$ on the mortgage test data.}
 \end{figure}
 
-\subsubsection{Discrimination Testing Results}
+\subsubsection{Discrimination Testing Results}\label{ssec:dis_mort}
 
@@
 APR -- annual percentage rate,
+AUC -- area under the curve,
 CFPB -- Consumer Financial Protection Bureau,
@@
 GDPR - General Data Protection Regulation,
-HMDA -- Home Mortgage Disclosure Act
+HMDA -- Home Mortgage Disclosure Act,
 ICE -- individual conditional expectation, 
@@
 PD -- partial dependence, 
+RMSE -- root mean square error,
 SGD -- stochastic gradient descent, 

### 2019-12-18 article-information-2019.tex
more hdma results text

@@
 
-\subsubsection{Interpretability and Post-hoc Explanation Results}
+\subsubsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}
 
@@
 
-Results for the mortgage data are presented in Subsections \ref{ssec: c_v_uc_mort} -- \ref{ssec:dis_mort}.
+Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort}.
 $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants.
@@
 
-\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_mort}
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec:c_v_uc_mort}
 
@@
 
-The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
-\ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse weights of the $g^\text{XNN}$ output layer, in which subnetworks 0,1,2,3,5,8, and 9 have large magnitude weights and subnetworks 4,6, and 7 have small to near-zero magnitude weights.
+The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions, and to do so in an interpretable manner, is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
+\ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse weights of the $g^\text{XNN}$ output layer, in which subnetworks zero, one, two, three, five, eight and nine have large magnitude weights and subnetworks four, six, and seven have small or near-zero magnitude weights.
+Distinctive ridge functions that feed into those weights are highlighted in \ref{fig:mort_xnn_glob_ridge} \textbf{B}.
+As in the subsection \ref{ssec:int_ex_sim}, ridge function plots vary with the output of the corresponding projection layer hidden unit in \ref{fig:mort_xnn_glob_ridge} \textbf{C}, and ridge functions appear to be simple functional forms that the output layer learns to combine to generate accurate predictions.
 
@@
 \end{figure}
+
+\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse weights of the projection layers associated with each subnetwork ridge function.
+For instance, subnetwork three has large weights for LTV ratio, introductory rate period flag, and introductory rate period, whereas subnetwork nine is nearly completely dominated by the weight for income.
+In combination, Figure \ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the weighting of the ridge functions in the $g^\text{XNN}$ output, \ref{fig:mort_xnn_glob_ridge} \textbf{B} displays the ridge functions that feed directly into those output weights, and \ref{fig:mort_xnn_glob_ridge} \textbf{C} describes the weighting of multiple input features into those ridge functions.
 

### 2019-12-18 notebooks/disparity-testing-pre-notebook.py
Fix AIR calculations

@@ -31,13 +31,13 @@ def disparity_tables(file_name,
                      model_name,
                      ):
     data = pd.read_csv(file_name)
-    data["outcome"] = np.where(data[predicted] >= probability_for_classification, 1, 0)
-    data["outcome"].value_counts(normalize=True)
+    data["classification_outcome"] = np.where(data[predicted] >= probability_for_classification, 1, 0)
+    data["classification_outcome"].value_counts(normalize=True)
 
     disp_tests = DisparityTesting(pg_names=pg_names, cg_names=cg_names, pgcg_names=pgcg_names,
                                   higher_score_favorable=higher_score_favorable)
 
-    cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome="outcome")
+    cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome="classification_outcome")
     cont_outcomes = disp_tests.continuous_disparity_measures(data=data, predicted=predicted)
     disparity_measures = DisparityTesting.create_combined_output(cat_outcomes=cat_outcomes, cont_outcomes=cont_outcomes)
     disparity_measures["Model Name"] = model_name
@@ -47,36 +47,51 @@ def disparity_tables(file_name,
 
 if __name__ == '__main__':
 
-    pg_names = ["black", "amind", "hispanic", "female"]
-    cg_names = ["white", "white", "non_hispanic", "male"]
-    pgcg_names = ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"]
-    # pg_names = ["black", "female"]
-    # cg_names = ["white", "male"]
-    # pgcg_names = ["black", "white", "female", "male"]
-    static_params = {"pg_names": ["black", "amind", "hispanic", "female"],
-                     "cg_names": ["white", "white", "non_hispanic", "male"],
-                     "pgcg_names": ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"],
-                     "higher_score_favorable": False,
-                     }
+    hmda_static_params = {"pg_names": ["black", "female"],
+                          "cg_names": ["white", "male"],
+                          "pgcg_names": ["black", "white", "female", "male"],
+                          # "pg_names": ["black", "amind", "hispanic", "female"],
+                          # "cg_names": ["white", "white", "non_hispanic", "male"],
+                          # "pgcg_names": ["black", "amind", "white", "hispanic", "non_hispanic", "female", "male"],
+                          "higher_score_favorable": False,
+                          "probability_for_classification": 0.20,
+                          }
 
[…]

### 2019-12-18 notebooks/scripts/disparity_measurement.py

@@ -73,7 +73,7 @@ class DisparityTesting(object):
             cm = metrics.confusion_matrix(y_true=data_cm[label], y_pred=data_cm[outcome], sample_weight=data_cm[groupi])
             tn, fp, fn, tp = cm.ravel()
             res.loc[res["class"] == groupi, "total"] = (fp + tp + fn + tn)
-            res.loc[res["class"] == groupi, "selected"] = (tp + fn)
+            res.loc[res["class"] == groupi, "selected"] = (tp + fp)
             res.loc[res["class"] == groupi, "true_positive"] = tp
             res.loc[res["class"] == groupi, "true_negative"] = tn
             res.loc[res["class"] == groupi, "false_positive"] = fp

### 2019-12-18 notebooks/disparity-testing-pre-notebook.py
Fix AIR calculations

@@ -85,13 +85,28 @@ if __name__ == '__main__':
                                  )
 
     simu_gbm = disparity_tables(**simu_static_params,
-                                model_name="Standard GBM - Simulated Data",
+                                model_name="XNN - Simulated Data",
                                 file_name='./data/output/test_sim_with_preds.csv',
                                 predicted="outcome_gbm_pred",
                                 label="outcome",
                                 )
 
-    disparity_results = pd.concat([hmda_mgbm, hmda_gbm,
-                                   simu_mgbm, simu_gbm], axis=0)
+    hmda_xnn = disparity_tables(**hmda_static_params,
+                                model_name="XNN - HMDA Data",
+                                file_name="./data/xnn_output/hmda_results/Results_hmda_test_set.csv",
+                                predicted="probability",
+                                label="high_priced",
+                                )
 
-    disparity_results.to_csv('./data/output/gbm_simu_and_hmda_disparity_results.csv')
+    simu_xnn = disparity_tables(**simu_static_params,
+                                model_name="Standard GBM - Simulated Data",
+                                file_name="./data/xnn_output/simulation_results/Results_simulation_test_set.csv",
+                                predicted="probability",
+                                label="outcome",
+                                )
+
+    disparity_results = pd.concat([hmda_gbm, hmda_mgbm, hmda_xnn,
+                                   simu_gbm, simu_mgbm, simu_xnn], axis=0)
+    pc_only = disparity_results.loc[disparity_results["Class"] == "Black"]
+
+    disparity_results.to_csv('./data/output/disparity_results_gbm_xnn_hmda_simu.csv', index=False)

### 2019-12-18 notebooks/scripts/disparity_measurement.py
Revise disparity measures for XNN and GBM models

@@ -103,8 +103,8 @@ class DisparityTesting(object):
         res.loc[res["class"].isin(self.pg_names), "control_percent_favorable"] = \
             np.array(res["percent_favorable"][self.cg_names])
 
-        res["marginal_difference"] = res["control_percent_favorable"] - res["percent_favorable"]
-        res["shortfall"] = res["marginal_difference"] * res["total"]
+        res["marginal_effects"] = res["control_percent_favorable"] - res["percent_favorable"]
+        res["shortfall"] = res["marginal_effects"] * res["total"]
         res["adverse_impact_ratio"] = res["percent_favorable"] / res["control_percent_favorable"]
         for fishi in self.pg_names:
             fishers_values = stats.fisher_exact(np.array(
@@ -151,7 +151,7 @@ class DisparityTesting(object):
                                cont_outcomes: pd.DataFrame,
                                cat_vars: Union[list, tuple] = ('class', 'control', 'total', 'false_positive_rate',
                                                                'false_negative_rate', 'accuracy',
-                                                               'marginal_difference', 'shortfall',
+                                                               'marginal_effects', 'shortfall',
                                                                'adverse_impact_ratio', 'fishers_exact',
                                                                'fishers_exact_p_value'),
                                cont_vars: Union[list, tuple] = ('class', 'control', 'standardized_mean_difference', 

### 2019-12-18 article-information-2019.tex
many edits/improvements

--- 2019-12-17 article-information-2019.tex
@@
 For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using adverse impact ratio (AIR), standardized mean difference (SMD), and additional straightforward group fairness measures.
-By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures.
+By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to describe the art of the possible for important ML applications that require high accuracy and interpretability and minimal discrimination.}
 
@@
 %\setcounter{secnumdepth}{4}
+
+%two part definition
+\newcommand{\twopartdef}[4]
+{
+	\left\{
+		\begin{array}{ll}
+			#1 & \mbox{if } #2 \\
+			#3 & \mbox{if } #4
+		\end{array}
+	\right.
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
-	\item \textbf{Discrimination testing methods}: AIR, SMD, and other group fairness metrics -- \textsection\ref{ssec:di}
+	\item \textbf{Discrimination testing methods}: AIR, ME, and SMD -- \textsection\ref{ssec:di}
 	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
@@
 To provide a sense of accuracy differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
-The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1991multivariate}.
+The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}.
 The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Because unconstrained ML models, like gradient boosting machines (GBMs) (e.g.
@@
 
-Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a more realistic template workflow for future users of similar methods and tools.
+Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a template workflow for future users of similar methods and tools.
 Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}.
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods include measures of DI with marginal effects, AIR and SMD \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
+Discussed discrimination testing methods include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.
@@
 
+Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
+
+\begin{equation}
+\begin{aligned}\label{eq:sim}\small
+f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
+\end{aligned}
+\end{equation}
+
+\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
+In Friedman's texts, a Gaussian noise term was added to create a continuous output variable for testing spline regression methodologies.
+In this manuscript, the signal generating function and input features are modified in several ways.
+Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings.
+%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so.
+For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
+For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
+Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
+The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows.
+The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
+Within the training set, a 5 fold cross validation indicator is used for training and assessing all models.
+For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The mortgage dataset analyzed here is a random sample of 100,000 consumer-anonymized loans from the HDMA database, separated into a training set and a test set.
+The mortgage dataset analyzed here is a random sample of consumer-anonymized loans from the HDMA database.
 These loans are a subset of all originated mortgage loans in the 2018 HMDA data that were chosen to represent a relatively comparable group of consumer mortgages.
@@
 the annual percentage rate (APR) charged was 150 basis points (1.5\%) or more above a survey-based estimate of other similar loans offered around the time of the given loan.
+After data cleaning and preprocessing to encode categorical features and create missing markers, the mortgage data contains ten input features and the binary outcome, \textit{high-priced}.
+The data is split into a training set with 160,338 loans and a marker for 5 fold cross validation and a test set containing 39,662 loans.
 While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (LTV ratio, DTI ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
-Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in this text.
+Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections \ref{sec:m_and_m} and \ref{sec:res}.
 
@@
 \begin{equation}
-\begin{aligned}\label{eq:rf}
+\begin{aligned}\label{eq:gbm}
 g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
@@
 
-Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P\setminus{j}}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P\setminus{j}}} = \mathbf{X}$) is considered.
-$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P\setminus{j}}}$ is left unchanged.
-$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P\setminus{j}}}$ are left untouched.
+Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$) is considered.
+$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
+$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
 PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
 Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
-PD should always be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.
+PD should be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Discrimination Metrics and Test Description}\label{ssec:di}
+\subsection{Discrimination Testing Metrics}\label{ssec:di}
+
+Metrics for testing discrimination in this text are:
+
+\begin{equation}
+\label{eq:air}
+\begin{aligned}
+\text{AIR} \equiv \frac{\text{Pr}(\hat{\mathbf{y}} = 1 | \mathbf{X}_p = 1)}{\text{Pr}(\hat{\mathbf{y}} = 1| \mathbf{X}_c = 1)}
+\end{aligned}
+\end{equation}
+
+\begin{equation}
+\label{eq:me}
+\begin{aligned}
+\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| \mathbf{X}_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | \mathbf{X}_p = 1))
+\end{aligned}
+\end{equation}
+
+\begin{equation}
+\label{eq:smd}
+\begin{aligned}
+\text{SMD} \equiv \frac{\bar{\hat{\mathbf{y}}}_p - \bar{\hat{\mathbf{y}}}_c}{ \sigma_{\hat{\mathbf{y}}}}
+\end{aligned}
+\end{equation}
+
+\noindent where $\mathbf{X}_p$ and $\mathbf{X}_c$ represent binary markers created from some demographic attribute $\mathbf{X}_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and conditional probabilities, $\text{Pr}(\hat{\mathbf{y}}|\mathbf{X}_j = \alpha)$, are evaluated over all $\mathbf{x}^{(i)} \in \{\mathbf{X}_j | x^{(i)}_j = \alpha\}$.
+Accuracy in each demographic is also considered as a measure of validity for a model and a demographic group.
 
@@
 Accuracy is compared for unconstrained, less interpretable $g^{\text{GBM}}$ and $g^{\text{ANN}}$ models and constrained, more interpretable $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models.
-Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretabilily, post-hoc explanation, and discrimination testing results are presented.
+Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretability, post-hoc explanation, and discrimination testing results are presented.
 
@@
 \includegraphics[width=15cm]{img/sim_mgbm_loc.png}
+\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.}
 \label{fig:sim_mgbm_loc}
-\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
-\end{figure}
-
-Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$.
+\end{figure}
 
@@
 
-Ridge functions are reminiscent of basis functions ...
+Ridge functions are reminiscent of basis functions ... distinctive simplistic functions
 
@@
 \label{fig:sim_xnn_glob_ridge}
-\caption{\textbf{A.} Output layer weights, \textbf{B.} corresponding ridge functions, and \textbf{C.} corresponding projection layer weights for $g^\text{XNN}$ on the simulated test data.}
-\end{figure}
+\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the simulated test data.}
+\end{figure}
+
+Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$.
 
@@
 \includegraphics[width=15cm]{img/sim_xnn_loc.png}
+\caption{Deep SHAP values for three observations across selected percentiles of  $g^\text{XNN}$ on the simulated test data.}
 \label{fig:sim_xnn_loc}
-\caption{Deep SHAP values for three observations across selected percentiles of  $g^\text{XNN}$ on the simulated test data.
-Deep SHAP values are reported in the probability space, after the application of the logit link function.}
-\end{figure}
+\end{figure}
+
+Deep SHAP values are reported in the probability space, after the application of the logit link function.
 
@@
 
+\begin{table}[H]
+\caption{Group size and accuracy for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
+\centering
+\begin{tabular}{cccc}
+\toprule
+\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& Accuracy \\
+\midrule
+Protected 1	& 3057.0		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.764\end{tabular} \\	
+\midrule			 
+Control 1		& 16943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.751\end{tabular} \\	
+\midrule
+Protected 2	& 9916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.761\end{tabular} \\	
+\midrule
+Control 2		& 10084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.745\end{tabular} \\	
+\bottomrule
+\label{tab:sim_n_acc}
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+\caption{AIR, ME and SMD for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
+\centering
+\begin{tabular}{cccccc}
+\toprule
+\textbf{Model}		& \textbf{Protected Class}								& \textbf{Control Class}							& \textbf{AIR}							& \textbf{ME}								& \textbf{SMD} \\
+\midrule
+$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0\%}\\ \textcolor{red}{0\%}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular}\\
+\midrule 
+$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0\%}\\ \textcolor{red}{0\%}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular} \\
+\bottomrule
+\label{tab:sim_disc}
+\end{tabular}
+\end{table}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Results for the mortgage data are presented in Subsections \ref{ssec: c_v_uc_mort} -- \ref{ssec:dis_mort}.
+Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort}.
 $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants.
-Assuming that small fit differences do not outweigh the need for intrinsic model interpretability and post-hoc explainability in high-stakes, human-centered, or regulated applications, $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability and post-hoc explainability results are presented along with discrimination testing results.
-For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of monotonic prediction behavior, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
-For $g^\text{XNN}$ intrinsic interpretability manifests as plots of sparse weights in the projection and output layers and plots of subnetwork ridge functions.
+Assuming that small fit differences on static test data do not outweigh the need for intrinsic model interpretability and post-hoc explainability in high-stakes, human-centered, or regulated applications, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
+For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
+For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
 Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
-Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, SMD, and other measures.
-
-\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_mort}
+Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
+
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec:c_v_uc_mort}
 
 Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data.
-This is partially due to the need to present direct comparisons between post-hoc explainability results and the data preprocessing required to use neural networks and tensorflow, e.g.
+This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g.
 numerical encoding of categorical features and missing values.
@@
 
-Global feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
+Global Shapley feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
 $g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
@@
 
-\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded flag features may contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
+\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
 
 Domain knowledge was used to positively constrain DTI ratio and LTV ratio and to negatively constrain income and the loan term flag under $g^\text{MGBM}$.
-The monotonicity constraints for DTI ratio and LTV ratio are confirmed on average and across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ in Figure \ref{fig:mort_mgbm_glob_pdp_ice}.
-Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles using ICE and on average with PD.
+The monotonicity constraints for DTI ratio and LTV ratio are confirmed for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data in Figure \ref{fig:mort_mgbm_glob_pdp_ice}.
+Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles for ICE and on average with PD.
 Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
@@
 DTI ratio and LTV ratio also appear to have sparse regions in their univariate distributions.
-The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as the $g^\text{MGBM}$ appears to carry reasonable predictions learned from more populated regions of both features' distributions into the more sparsely populated regions.
+The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as $g^\text{MGBM}$ appears to carry reasonable predictions learned from populous domains into the sparse domains of both features.
 
@@
 Figure \ref{fig:mort_mgbm_glob_pdp_ice} also displays PD and ICE for the unconstrained feature property value.
-Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns of ICE curves.
-While PD shows monotonically increasing prediction behavior on average, apparently influenced by the extremely large predictions of high $g^{\text{MGBM}}(\mathbf{X})$ percentile individuals, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$ exhibit different prediction behavior with respect to property value.
+Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns established by ICE curves.
+While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$, and lower,  exhibit different prediction behavior with respect to property value.
 Some individuals at these lower percentiles display monotonically decreasing prediction behavior while others appear to show fluctuating prediction behavior.
-This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}(\mathbf{X})$ \cite{ice_plots}.
-Drivers of this interaction can often be elucidated using surrogate decision trees \cite{art_and_sci}.
-Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}(\mathbf{X})$ can learn.
+Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn.
 For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions.
@@
 Such fluctuations can be caused by local data characteristics, modeled local interactions, or unfortunately, by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
-
-Local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in Figure \ref{fig:mort_mgbm_loc}.
-The selected individuals show the expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
+\textcolor{red}{This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}(\mathbf{X})$ \cite{ice_plots}, and drivers of such interactions can often be elucidated using surrogate decision trees \cite{art_and_sci}.
+}
+
+In Figure \ref{fig:mort_mgbm_loc}, local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in the mortgage test data.
+The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
 
@@
 \includegraphics[width=15cm]{img/mort_mgbm_loc.png}
-\caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
+\caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.}
 \label{fig:mort_mgbm_loc}
@@
 
-\noindent The deeper significance of Figure \ref{fig:mort_mgbm_loc} lies in the fact that any $g^\text{MGBM}(\mathbf{x})$ prediction can be accurately summarized in this manner, which is generally important for enabling appeal of model decisions, and specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to individuals.
-(In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.)
+\noindent Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability to accurately summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
+In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.
 
 Figure \ref{fig:mort_xnn_glob} displays global feature importance for $g^\text{XNN}$ on the mortgage test data.
-$g^\text{XNN}$ distributes importance more evenly across business drivers and puts strong emphasis on the no introductory rate period flag feature than does $g^\text{MGBM}$.
+$g^\text{XNN}$ distributes importance more evenly across business drivers and puts stronger emphasis on the no introductory rate period flag feature than does $g^\text{MGBM}$.
 Like $g^\text{MGBM}$, $g^\text{XNN}$ puts little emphasis on the other flag features.
@@
 
-The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
-\ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse weights of the $g^\text{XNN}$ output layer, in which subnetworks 0,1,2,3,5,8, and 9 have large magnitude weights and subnetworks 4,6, and 7 have small to near-zero magnitude weights.
+The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions, and to do so in an interpretable manner, is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
+\ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse $\gamma_k$ weights of the $g^\text{XNN}$ output layer in which the $n_k$ subnetworks with $k \in \{0,1,2,3,5,8,9\}$ have large magnitude weights and $n_k$ subnetworks, $k \in \{4,6,7\}$, have small or near-zero magnitude weights.
+Distinctive ridge functions that feed into those large magnitude $\gamma_k$ weights are highlighted in \ref{fig:mort_xnn_glob_ridge} \textbf{B} and color-coded to pair with their corresponding $\gamma_k$ weight.
+As in the subsection \ref{ssec:int_ex_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge} \textbf{C}.
+In both the simulated and mortgage data, $g^\text{XNN}$ $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, perhaps reminiscent of the visual primitives often learned by low layers of pattern-detecting convolutional neural networks (CNNs) \cite{lecun2015deep}.
 
@@
 \includegraphics[width=16cm]{img/mort_xnn_glob_ridge.png}
-\caption{\textbf{A.} Output layer weights, \textbf{B.} corresponding ridge functions, and \textbf{C.} corresponding projection layer weights for $g^\text{XNN}$ on the mortgage test data.}
+\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the mortgage test data.}
 \label{fig:mort_xnn_glob_ridge}
@@
 
+\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_k$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
+For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
+In combination, Figure \ref{fig:mort_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
+
+To compliment the global interpretability of $g^\text{XNN}$, Figure \ref{fig:mort_xnn_loc} displays local Shapley values for selected individuals, estimated from the projection layer using Deep SHAP in the $g^\text{XNN}$ probability space.
+
 \begin{figure}[H]
@@
 \includegraphics[width=15cm]{img/mort_xnn_loc.png}
+\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}$ on the mortgage test data.}
 \label{fig:mort_xnn_loc}
-\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}$ on the mortgage test data.}
-\end{figure}
+\end{figure}
+
+\noindent While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc} their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions, even for neural network architectures.
 
 \subsubsection{Discrimination Testing Results}\label{ssec:dis_mort}
+
+\begin{table}[H]
+\caption{Group size and accuracy for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+\centering
+\begin{tabular}{cccc}
+\toprule
+\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& Accuracy \\
+\midrule
+Black 		& 2608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} \\	
+\midrule			 
+White 		& 28361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} \\	
+\midrule
+Female 		& 8301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} \\	
+\midrule
+Male 		& 13166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} \\	
+\bottomrule
+\label{tab:mort_n_acc}
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+\caption{AIR, ME and SMD for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+\centering
+\begin{tabular}{cccccc}
+\toprule
+\textbf{Model}		& \textbf{Protected Class}								& \textbf{Control Class}							& \textbf{AIR}							& \textbf{ME}								& \textbf{SMD} \\
+\midrule
+$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular} \\
+\midrule 
+$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular} \\
+
+\bottomrule
+\label{tab:mort_disc}
+\end{tabular}
+\end{table}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{Wen Phan for work in formalizing notation. \textcolor{red}{BLDS editor} for editing. Andrew Burt for ideas around ML transparency, model debugging, and the transparency paradox.}
+\acknowledgments{Wen Phan for work in formalizing notation. Sue Shay for editing. Andrew Burt for ideas around the transparency paradox.}
 
@@
 AUC -- area under the curve,
+CNN -- convolutional neural network,
 CFPB -- Consumer Financial Protection Bureau,
@@
 LTV -- loan to value,
+ME -- marginal effect,
 MGBM -- monotonic gradient boosting machine, 

### 2019-12-19 article-information-2019.tex
finish section 2.2.2, data entry for discrimination tests

@@
 The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures.
-By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to describe the art of the possible for important ML applications that require high accuracy and interpretability and minimal discrimination.}
+By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present the art of the possible for important ML applications that require high accuracy and interpretability and minimal discrimination.}
 
@@
 %\setcounter{secnumdepth}{4}
-
-%two part definition
-\newcommand{\twopartdef}[4]
-{
-	\left\{
-		\begin{array}{ll}
-			#1 & \mbox{if } #2 \\
-			#3 & \mbox{if } #4
-		\end{array}
-	\right.
-}
 
@@
 \midrule
-$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0\%}\\ \textcolor{red}{0\%}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular}\\
-\midrule 
-$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0\%}\\ \textcolor{red}{0\%}\end{tabular}	& \begin{tabular}{@{}c@{}} \textcolor{red}{0} \\ \textcolor{red}{0}\end{tabular} \\
+$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}\\
+\midrule 
+$g^\text{XNN}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.737 \\ 1.04\end{tabular}	& \begin{tabular}{@{}c@{}} 11.1\%\\ -1.7\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.240 \\ 0.551\end{tabular} \\
 \bottomrule
@@
 However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value.
-Such fluctuations can be caused by local data characteristics, modeled local interactions, or unfortunately, by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
-\textcolor{red}{This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}(\mathbf{X})$ \cite{ice_plots}, and drivers of such interactions can often be elucidated using surrogate decision trees \cite{art_and_sci}.
-}
+This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci}.
+However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
 

### 2019-12-19 notebooks/scripts/ann_hdma_code.py
Updates

@@ -1,77 +1,0 @@
-import numpy as np
-import pandas as pd
-import shap
-import subprocess
-import sys
-#import pydot
-
-
-import keras    
-from timeit import default_timer as timer
-import tensorflow as tf
-from keras import backend as K
-from keras.layers import Input, Dense
-from keras.models import Model, Sequential
-from keras.layers import Dense
-
-
-seed = 12345
-
-np.random.seed(seed)
-
-my_init = keras.initializers.RandomUniform(seed=seed)
-
-out_dir = "ann_output/"
-
-# baseline model
-def ann_model():
-	# create model
-
-    input = Input(shape=(features,), name='main_input')
-
-    out = Dense(20, input_dim=10, activation='relu')(input)
-    
-    out = Dense(20, input_dim=20, activation='relu')(out)
-    
-    out = Dense(1, activation='sigmoid')(out)
-    
-    model = Model(inputs=input, outputs=out)
-	# Compile model
-    model.compile(loss='binary_crossentropy', optimizer='adam')
-    
-    return model
-
-# Load the dataset
-xnn_data_dir = '~/article-information-2019/data/xnn_output/'
-#xnn_data_dir = ''
-DATA=pd.read_csv(xnn_data_dir + 'train_transformed.csv')
-#DATA = DATA.iloc[0:10000,:]
-TEST=pd.read_csv(xnn_data_dir + 'test_transformed.csv')
-
-# Select features and split into target and feature sets
-selected_vars = ['term_360', 'conforming']
-selected_vars += ['debt_to_income_ratio_missing','loan_amount_std', 'loan_to_value_ratio_std']
-selected_vars += ['no_intro_rate_period_std', 'intro_rate_period_std']
-selected_vars += ['property_value_std', 'income_std', 'debt_to_income_ratio_std']
-
-
-target_var = "high_priced"
-
-X=DATA[selected_vars].values
-Y=DATA[target_var].values
-TEST_X = TEST[selected_vars].values
-TEST_Y = TEST[target_var].values
-features = X.shape[1]
-
-inputs = {'main_input': X}
-
-
-#Fit model
-model = ann_model()
-model.fit(inputs, Y, epochs=5000, batch_size=1024, validation_split=0, verbose=1)
-
-# Save results
-test_preds = model.predict(TEST_X)
-pd.DataFrame(pd.concat([TEST, pd.DataFrame(test_preds)], axis=1)).to_csv(out_dir + "hmda_ann_results.csv" , index=False)
-
-

### 2019-12-19 notebooks/scripts/ann_hdma_code_shapley.py

@@ -3,8 +3,6 @@ import pandas as pd
 import shap
 import subprocess
 import sys
-#import pydot
-
 
 import keras    
 from timeit import default_timer as timer
@@ -23,9 +21,9 @@ my_init = keras.initializers.RandomUniform(seed=seed)
 
 out_dir = "ann_output5/"
 
-# baseline model
+
 def ann_model():
-	# create model
+	""" Create an artificial neural network model """
 
     input = Input(shape=(features,), name='main_input')
 
@@ -41,6 +39,7 @@ def ann_model():
     
     return model
 
+
 # Load the dataset
 xnn_data_dir = '~/article-information-2019/data/xnn_output/'
 #xnn_data_dir = ''
@@ -48,15 +47,17 @@ DATA=pd.read_csv(xnn_data_dir + 'train_transformed.csv')
 #DATA = DATA.iloc[0:10000,:]
 TEST=pd.read_csv(xnn_data_dir + 'test_transformed.csv')
 
+
 # Select features and split into target and feature sets
 selected_vars = ['term_360', 'conforming']
 selected_vars += ['debt_to_income_ratio_missing','loan_amount_std', 'loan_to_value_ratio_std']
 selected_vars += ['no_intro_rate_period_std', 'intro_rate_period_std']
 selected_vars += ['property_value_std', 'income_std', 'debt_to_income_ratio_std']
 
-
 target_var = "high_priced"
 
+
+# Split the data into feature and target values
 X=DATA[selected_vars].values
 Y=DATA[target_var].values
 TEST_X = TEST[selected_vars].values
@@ -64,7 +65,6 @@ TEST_Y = TEST[target_var].values
 features = X.shape[1]
 
 inputs = {'main_input': X}
-
 
 #Fit model
 model = ann_model()
@@ -80,6 +80,8 @@ preds = model.predict(TEST_X)
 preds = np.concatenate((preds, shap_values[0], preds), axis=1)
 preds[:, -1] = explainer.expected_value
 
+
+# Add the Shapley values and predictions to the dataset
 TEST = pd.DataFrame(pd.concat([TEST, pd.DataFrame(preds)], axis=1))
 
 Feature_names = selected_vars.copy()
@@ -97,6 +99,7 @@ TEST = TEST.rename(columns={0: "probability",
                             10: Feature_names[9]+"_Shapley_score",
                             11: "Intercept_Shapley_score"})
 
+    
 # Save results
[…]

### 2019-12-19 notebooks/scripts/ann_simulation_code.py

@@ -1,79 +1,0 @@
-import numpy as np
-import pandas as pd
-import shap
-import subprocess
-import sys
-#import pydot
-
-
-import keras    
-from timeit import default_timer as timer
-import tensorflow as tf
-from keras import backend as K
-from keras.layers import Input, Dense
-from keras.models import Model, Sequential
-from keras.layers import Dense
-
-
-seed = 12345
-
-np.random.seed(seed)
-
-my_init = keras.initializers.RandomUniform(seed=seed)
-
-#from keras.models import Sequential
-#from keras.layers import Dense
-
-
-out_dir = "ann_output2/"
-
-# baseline model
-def ann_model():
-    
-    input = Input(shape=(features,), name='main_input')
-
-    out = Dense(10, input_dim=10, activation='relu')(input)
-    
-    out = Dense(5, input_dim=10, activation='relu')(out)
-    
-    out = Dense(1, activation='sigmoid')(out)
-    
-    model = Model(inputs=input, outputs=out)
-	# Compile model
-    model.compile(loss='binary_crossentropy', optimizer='adam')
-    
-    return model
-
-# Load the dataset
-xnn_data_dir = '~/article-information-2019/data/xnn_output/'
-#xnn_data_dir = ''
-DATA=pd.read_csv(xnn_data_dir + 'train_simulated_transformed.csv')
-print(list(DATA.columns))
-#DATA = DATA.iloc[0:10000,:]
-TEST=pd.read_csv(xnn_data_dir + 'test_simulated_transformed.csv')
-print(list(TEST.columns))
-
-
-selected_vars = ['binary1', 'binary2', 'cat1_0', 'cat1_1', 'cat1_2', 'cat1_3', 'cat1_4', 
-                 'fried1_std', 'fried2_std', 'fried3_std', 'fried4_std', 'fried5_std']
-
-target_var = 'outcome'
-
-X=DATA[selected_vars].values
-Y=DATA[target_var].values
-TEST_X = TEST[selected_vars].values
-TEST_Y = TEST[target_var].values
-features = X.shape[1]
-
-
-inputs = {'main_input': X}
-
-# Fit model
-model = ann_model()
-model.fit(inputs, Y, epochs=2000, batch_size=1024, validation_split=0, verbose=1)
-
-# Save results
-test_preds = model.predict(TEST_X)
-pd.DataFrame(pd.concat([TEST, pd.DataFrame(test_preds)], axis=1)).to_csv(out_dir + "simulated_ann_results.csv" , index=False)
-
-

### 2019-12-19 notebooks/scripts/ann_simulation_code_shapley.py

@@ -3,9 +3,6 @@ import pandas as pd
 import shap
 import subprocess
 import sys
-#import pydot
-
-
 import keras    
 from timeit import default_timer as timer
 import tensorflow as tf
@@ -21,13 +18,11 @@ np.random.seed(seed)
 
 my_init = keras.initializers.RandomUniform(seed=seed)
 
-
-
 out_dir = "ann_output4/"
 
-# baseline model
+
 def ann_model():
-    
+    """ Create an ann model"""
     input = Input(shape=(features,), name='main_input')
 
     out = Dense(10, input_dim=10, activation='relu')(input)
@@ -51,12 +46,14 @@ print(list(DATA.columns))
 TEST=pd.read_csv(xnn_data_dir + 'test_simulated_transformed.csv')
 print(list(TEST.columns))
 
-
+# Specify the features and target
 selected_vars = ['binary1', 'binary2', 'cat1_0', 'cat1_1', 'cat1_2', 'cat1_3', 'cat1_4', 
                  'fried1_std', 'fried2_std', 'fried3_std', 'fried4_std', 'fried5_std']
 
 target_var = 'outcome'
 
+
+# Split the datasets into feature and target values
 X=DATA[selected_vars].values
 Y=DATA[target_var].values
 TEST_X = TEST[selected_vars].values
@@ -65,6 +62,7 @@ features = X.shape[1]
 
 
 inputs = {'main_input': X}
+
 
 # Fit model
 model = ann_model()
@@ -80,6 +78,8 @@ preds = model.predict(TEST_X)
 preds = np.concatenate((preds, shap_values[0], preds), axis=1)
 preds[:, -1] = explainer.expected_value
 
+
+# Add the predictions and Shapley values to the test set
 TEST = pd.DataFrame(pd.concat([TEST, pd.DataFrame(preds)], axis=1))
 
 Feature_names = selected_vars.copy()

### 2019-12-19 article-information-2019.tex
nicks most recent content

@@
 	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
-	\item \textbf{Discrimination testing methods}: AIR, ME, and SMD -- \textsection\ref{ssec:di}
+	\item \textbf{Discrimination testing methods}: AIR, ME, SMD and confusion matrix metrics -- \textsection\ref{ssec:di}
 	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
+Discussed discrimination testing methods include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.
@@
 
-Metrics for testing discrimination in this text are:
+Because many technical and academic discussions of fairness in ML have been inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in industries like Financial Services.
+The discussed metrics are also representative of fair lending analyses and pair well with the mortgage data.
+Before discussing the techniques, it is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI).
+DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
+With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI, colloquially known as \textit{unintentional discrimination}, occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
+In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.
+However, other factors, such as hyperparameter or algorithm choice, can drive DI.
+Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.
+
+The analyses and metrics herein focus on several measures of disparate impact that are commonly used in US litigation and regulatory settings.
+One is known as marginal effects (ME).
+ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.
+
+\begin{equation}
+\label{eq:me}
+\begin{aligned}
+\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| \mathbf{X}_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | \mathbf{X}_p = 1))
+\end{aligned}
+\end{equation}
+
+\noindent where $\mathbf{X}_p$ and $\mathbf{X}_c$ represent binary markers created from some demographic attribute $\mathbf{X}_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and conditional probabilities, $\text{Pr}(\hat{\mathbf{y}}|\mathbf{X}_j = \alpha)$, are evaluated over all $\mathbf{x}^{(i)} \in \{\mathbf{X}_j | x^{(i)}_j = \alpha\}$.
+ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
 
@@
 
-\begin{equation}
-\label{eq:me}
-\begin{aligned}
-\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| \mathbf{X}_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | \mathbf{X}_p = 1))
-\end{aligned}
-\end{equation}
+\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the percent of the control class that receives a favorable outcome.
+Another long-standing measure of DI is SMD.
+SMD is typically used to assess disparities in continuous variables, such as income differences in employment analyses, or interest rate differences in lending.
+It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}.
+The SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
 
@@
 
-\noindent where $\mathbf{X}_p$ and $\mathbf{X}_c$ represent binary markers created from some demographic attribute $\mathbf{X}_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and conditional probabilities, $\text{Pr}(\hat{\mathbf{y}}|\mathbf{X}_j = \alpha)$, are evaluated over all $\mathbf{x}^{(i)} \in \{\mathbf{X}_j | x^{(i)}_j = \alpha\}$.
-Accuracy in each demographic is also considered as a measure of validity for a model and a demographic group.
+\noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.
+Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of variables, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
+In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.
+Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.
+
+A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute prima facia evidence of illegal discrimination.
+Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not meaningful.
+It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model.
+Moreover, practical significance is also needed model by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production.
+Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
+One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes prima facia evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.
+
+The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.
+The first guidance in U.S.
+courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v.
+Partida}.\footnote{\textit{Castaneda v.
+Partida}, 430 US 482 - Supreme Court (1977)}   Here, the U.S.
+Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v.
+United States.}\footnote{\textit{Hazelwood School Dist.
+v.
+United States}, 433 U.S.
+299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
 
@@
 
+Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
+Several measures of disparities are shown, with the SMD calculated using the probabilities from the models, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIR calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{The decision of what cutoff should be used in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.
+For example, a model developer at a bank might build a credit model where the F1 score is maximized at a delinquency probability cutoff of 0.15.
+For purposes of evaluating the quality of the model, she may review confusion matrix statistics (accuracy, recall, precision, etc.) using cutoffs based on the maximum F1 score.
+But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
+Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assumes that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might want to consider the relative false negative rates as a measure of the class-control disparities.
+Table \ref{tab:sim_disc} shows that protected group 1 has higher relative false negative rates under $g^{\text{XNN}}$ (1.14 vs.
+1.06).
+However, the overall false negative rates were lower for  $g^{\text{XNN}}$ (0.383 vs.
+0.401).
+This illustrates a danger in considering relative class-control metrics in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$.
+This is because FNR accuracy improves for both the protected group 1 and control group 1, but members of control group 1 benefit more than those in protected group 1.
+Of course, the choice of which model is truly fairer is a policy question.
+
 \begin{table}[H]
-\caption{Group size and accuracy for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
-\centering
-\begin{tabular}{cccc}
+\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
+\centering
+\begin{tabular}{cccccc}
 \toprule
-\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& Accuracy \\
-\midrule
-Protected 1	& 3057.0		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.764\end{tabular} \\	
-\midrule			 
-Control 1		& 16943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.751\end{tabular} \\	
-\midrule
-Protected 2	& 9916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.761\end{tabular} \\	
-\midrule
-Control 2		& 10084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.745\end{tabular} \\	
+\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy}  								& \textbf{FPR}									& \textbf{FNR} \\
+\midrule
+Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.764\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.150 \\ 0.167\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.401 \\ 0.383\end{tabular} \\
+\midrule			 
+Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.751\end{tabular}	& \begin{tabular}{@{}c@{}} 0.129 \\ 0.149\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.337\end{tabular} \\
+\midrule
+Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.761\end{tabular}	& \begin{tabular}{@{}c@{}} 0.169 \\ 0.183\end{tabular}	& \begin{tabular}{@{}c@{}} 0.331 \\ 0.306\end{tabular} \\
+\midrule
+Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.745\end{tabular}	& \begin{tabular}{@{}c@{}} 0.091 \\ 0.116\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.370\end{tabular} \\
 \bottomrule
@@
 
+11.1\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.
+Of note is that 11.1\% is not a meaningful difference without context.
+If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 11.1\% could represent an extremely large difference and would require remediation.
+But if they represent substantially different populations, then 11.1\% could represent a reasonable deviation from parity.
+As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 11.1\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population.
+Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.
+It is also highly statistically significant (not shown), which together would indicate that there may be evidence of illegal DI.
+As with ME and other measures, the reasonableness of this disparity is not clear outside of context.
+However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.
+
 \begin{table}[H]
-\caption{AIR, ME and SMD for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
-\centering
-\begin{tabular}{cccccc}
+\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
+\centering
+\begin{tabular}{cccccccc}
 \toprule
-\textbf{Model}		& \textbf{Protected Class}								& \textbf{Control Class}							& \textbf{AIR}							& \textbf{ME}								& \textbf{SMD} \\
-\midrule
-$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}\\
-\midrule 
-$g^\text{XNN}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.737 \\ 1.04\end{tabular}	& \begin{tabular}{@{}c@{}} 11.1\%\\ -1.7\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.240 \\ 0.551\end{tabular} \\
+\textbf{Model}		& \textbf{Protected Class}										& \textbf{Control Class}								& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD}   								& \textbf{FPR Ratio}							& \textbf{FNR Ratio} \\
+\midrule
+$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.16 \\ 1.86\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
+\midrule 
+$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.737 \\ 1.04\end{tabular}	& \begin{tabular}{@{}c@{}} 11.1\%\\ -1.7\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.240 \\ 0.551\end{tabular}	& \begin{tabular}{@{}c@{}} 1.12 \\ 1.59\end{tabular}	& \begin{tabular}{@{}c@{}} 1.14 \\ 0.827\end{tabular} \\
 \bottomrule
@@
 \end{table}
+
+\noindent Even though SMD is often used to measure disparities for non-categorical outcomes, it is calculated here on the models’ probabilities prior to being transformed into classifications.
+This measurement would be particularly relevant if the probabilities are being used in combination with other models to determine an outcome.
+The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.21 versus -0.24), but both are close to Cohen’s small effect threshold of -0.20.
+Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.
 
@@
 \begin{table}[H]
-\caption{Group size and accuracy for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
-\centering
-\begin{tabular}{cccc}
+\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+\centering
+\begin{tabular}{ccccccc}
 \toprule
-\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& Accuracy \\
-\midrule
-Black 		& 2608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} \\	
-\midrule			 
-White 		& 28361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} \\	
-\midrule
-Female 		& 8301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} \\	
-\midrule
-Male 		& 13166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} \\	
+\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy} 									& \textbf{FPR}											& \textbf{FNR} \\
+\midrule
+Black 		& 2,608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} 			& \begin{tabular}{@{}c@{}} 0.457 \\ 0.308\end{tabular} \\
+\midrule			 
+White 		& 28,361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.150 \\ 0.120\end{tabular}			& \begin{tabular}{@{}c@{}} 0.508 \\ 0.360\end{tabular}\\
+\midrule
+Female 		& 8,301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular}		& \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}			& \begin{tabular}{@{}c@{}} 0.402 \\ 0.322\end{tabular}\\
+\midrule
+Male 		& 13,166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.182 \\ 0.131\end{tabular}			& \begin{tabular}{@{}c@{}} 0.497 \\ 0.347\end{tabular}\\
 \bottomrule
@@
 \begin{table}[H]
-\caption{AIR, ME and SMD for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
-\centering
-\begin{tabular}{cccccc}
+\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+\centering
+\begin{tabular}{cccccccc}
 \toprule
-\textbf{Model}		& \textbf{Protected Class}								& \textbf{Control Class}							& \textbf{AIR}							& \textbf{ME}								& \textbf{SMD} \\
-\midrule
-$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular} \\
-\midrule 
-$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular} \\
+\textbf{Model}	& \textbf{Protected Class}	& \textbf{Control Class}	& \textbf{AIR}	& \textbf{ME}	& \textbf{SMD}  & \textbf{FPR Ratio}		& \textbf{FNR Ratio} \\
+\midrule
+$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular}	& \begin{tabular}{@{}c@{}} 2.10 \\ 1.15\end{tabular}	& \begin{tabular}{@{}c@{}} 0.900 \\ 0.810\end{tabular} \\
+\midrule 
+$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular}	& \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular}	& \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\
 
@@
 
-\subsection{Interpretability, Explainability, Appeal, and Compliance}
+\subsection{Appeal and Override of Automated Decisions}
 
@@
 Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention.
-ML models are often wrong.\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
+ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions is difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
 
@@
 
-\subsection{Intersectionality of Interpretability, Explainability, Discrimination, and Security in ML}
+\subsection{Intersectional and Non-static Problems in Machine Learning}
 
@@
 DI -- disparate impact, 
+DT -- disparate treatment,
 DTI -- debt to income,
 EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e. variants GAMs that consider two-way interactions and may incorporate boosting into training, 
+EEOC -- Equal Employment Opportunity Commission,
 ECOA - Equal Credit Opportunity Act,
@@
 FCRA -- Fair Credit Reporting Act,
+FNR -- false negative rate,
+FPR -- false positive rate,
 GAM -- generalized additive model, 

### 2019-12-19 notebooks/disparity-testing-pre-notebook.py
Added relative FPR/FNR and fixed reference to XNN

@@ -39,9 +39,11 @@ def disparity_tables(file_name,
 
     cat_outcomes = disp_tests.categorical_disparity_measures(data=data, label=label, outcome="classification_outcome")
     cont_outcomes = disp_tests.continuous_disparity_measures(data=data, predicted=predicted)
-    disparity_measures = DisparityTesting.create_combined_output(cat_outcomes=cat_outcomes, cont_outcomes=cont_outcomes)
+    disparity_measures = DisparityTesting.create_combined_output(cat_outcomes=cat_outcomes, cont_outcomes=cont_outcomes,
+                                                                 cat_vars=list(cat_outcomes.columns))
     disparity_measures["Model Name"] = model_name
-    disparity_measures = disparity_measures[["Model Name"] + [x for x in disparity_measures.columns if x != "Model Name"]]
+    disparity_measures = disparity_measures[["Model Name"] +
+                                            [x for x in disparity_measures.columns if x != "Model Name"]]
     return disparity_measures
 
 
@@ -64,12 +66,12 @@ if __name__ == '__main__':
                                  label="high_priced",
                                  )
 
-    hmda_gbm = disparity_tables(**hmda_static_params,
-                                model_name="Standard GBM - HMDA Data",
-                                file_name='./data/output/test_hmda_with_preds.csv',
-                                predicted="high_priced_gbm_pred",
-                                label="high_priced",
-                                )
+    # hmda_gbm = disparity_tables(**hmda_static_params,
+    #                             model_name="Standard GBM - HMDA Data",
+    #                             file_name='./data/output/test_hmda_with_preds.csv',
+    #                             predicted="high_priced_gbm_pred",
+    #                             label="high_priced",
+    #                             )
 
     simu_static_params = {"pg_names": ["prot_class1", "prot_class2"],
                           "cg_names": ["ctrl_class1", "ctrl_class2"],
@@ -84,12 +86,12 @@ if __name__ == '__main__':
                                  label="outcome",
                                  )
 
-    simu_gbm = disparity_tables(**simu_static_params,
-                                model_name="XNN - Simulated Data",
-                                file_name='./data/output/test_sim_with_preds.csv',
-                                predicted="outcome_gbm_pred",
-                                label="outcome",
-                                )
+    # simu_gbm = disparity_tables(**simu_static_params,
+    #                             model_name="Standard GBM - Simulated Data",
[…]

### 2019-12-19 notebooks/scripts/disparity_measurement.py

@@ -99,13 +99,19 @@ class DisparityTesting(object):
         res["percent_favorable"] = res["favorable"] / res["total"]
 
         res.loc[res["class"].isin(self.pg_names), "control_total"] = np.array(res["total"][self.cg_names])
+        res.loc[res["class"].isin(self.pg_names), "control_false_positive_rate"] = \
+            np.array(res["false_positive_rate"][self.cg_names])
+        res.loc[res["class"].isin(self.pg_names), "control_false_negative_rate"] = \
+            np.array(res["false_negative_rate"][self.cg_names])
         res.loc[res["class"].isin(self.pg_names), "control_selected"] = np.array(res["selected"][self.cg_names])
         res.loc[res["class"].isin(self.pg_names), "control_percent_favorable"] = \
             np.array(res["percent_favorable"][self.cg_names])
 
         res["marginal_effects"] = res["control_percent_favorable"] - res["percent_favorable"]
-        res["shortfall"] = res["marginal_effects"] * res["total"]
         res["adverse_impact_ratio"] = res["percent_favorable"] / res["control_percent_favorable"]
+        res["relative_false_positive_rate"] = res["false_positive_rate"] / res["control_false_positive_rate"]
+        res["relative_false_negative_rate"] = res["false_negative_rate"] / res["control_false_negative_rate"]
+
         for fishi in self.pg_names:
             fishers_values = stats.fisher_exact(np.array(
                 res.loc[res["class"] == fishi, ["total", "selected",
@@ -150,9 +156,10 @@ class DisparityTesting(object):
     def create_combined_output(cat_outcomes: pd.DataFrame,
                                cont_outcomes: pd.DataFrame,
                                cat_vars: Union[list, tuple] = ('class', 'control', 'total', 'false_positive_rate',
-                                                               'false_negative_rate', 'accuracy',
-                                                               'marginal_effects', 'shortfall',
-                                                               'adverse_impact_ratio', 'fishers_exact',
+                                                               'relative_false_positive_rate',  'false_negative_rate',
+                                                               'relative_false_negative_rate', 'accuracy',
+                                                               'marginal_effects', 'adverse_impact_ratio', 
+                                                               'fishers_exact',
                                                                'fishers_exact_p_value'),
                                cont_vars: Union[list, tuple] = ('class', 'control', 'standardized_mean_difference', 
                                                                 't_statistic', 't_statistic_p_value'),

### 2019-12-20 notebooks/disparity-testing-pre-notebook.py
Added relative FPR/FNR and fixed reference to XNN

@@ -110,5 +110,12 @@ if __name__ == '__main__':
     disparity_results = pd.concat([simu_mgbm, simu_xnn,
                                    hmda_mgbm, hmda_xnn], axis=0)
     # pc_only = disparity_results.loc[disparity_results["Class"].isin(['Prot-Class1', 'Prot-Class2'])]
-
+    disparity_results_for_paper = disparity_results[["Model Name", "Class", "Control", "Total", "Accuracy",
+                                                     "False Positive Rate", "Control False Positive Rate",
+                                                     "Relative False Positive Rate", "False Negative Rate",
+                                                     "Control False Negative Rate", "Relative False Negative Rate",
+                                                     "Marginal Effects", "Adverse Impact Ratio",
+                                                     "Fishers Exact P-Value", "Standardized Mean Difference",
+                                                     "T-Statistic P-Value"]]
     disparity_results.to_csv('./data/output/disparity_results_mgbm_xnn_hmda_simu.csv', index=False)
+    disparity_results_for_paper.to_csv('./data/output/disparity_results_for_paper_mgbm_xnn_hmda_simu.csv', index=False)

### 2019-12-20 notebooks/scripts/disparity_measurement.py

@@ -156,6 +156,7 @@ class DisparityTesting(object):
     def create_combined_output(cat_outcomes: pd.DataFrame,
                                cont_outcomes: pd.DataFrame,
                                cat_vars: Union[list, tuple] = ('class', 'control', 'total', 'false_positive_rate',
+                                                               'control_false_positive', 'control_false_negative',
                                                                'relative_false_positive_rate',  'false_negative_rate',
                                                                'relative_false_negative_rate', 'accuracy',
                                                                'marginal_effects', 'adverse_impact_ratio', 

### 2019-12-20 article-information-2019.tex

@@
 \noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
-In Friedman's texts, a Gaussian noise term was added to create a continuous output variable for testing spline regression methodologies.
+In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
 In this manuscript, the signal generating function and input features are modified in several ways.
@@
 Another long-standing measure of DI is SMD.
-SMD is typically used to assess disparities in continuous variables, such as income differences in employment analyses, or interest rate differences in lending.
+SMD is typically used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
 It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}.
@@
 \noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.
-Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of variables, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
+Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of features, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
 In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.
@@
 
-Results for constrained models on the simulated data are displayed in Subsections \ref{ssec: c_v_uc_sim} -- \ref{ssec:dis_sim}.
-{Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models.
-\textcolor{red}{...}
+Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
 
@@
 $g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
-Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
-Accuracy is measured at the best F1 threshold for each model.
 
@@
 
+\noindent Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
+$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
+Accuracy is measured at the best F1 threshold for each model.
+
 \subsubsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}
 
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the numeric value of the reported values can interpreted as the absolute mean impact of each $X_j$ on $g^\text{MGBM}$ in the margin space.
+Global mean absolute Shapley value feature importance for $g^{\text{MGBM}}(\mathbf{X})$ on the simulated test data is displayed in Figure \ref{fig:sim_mgbm_glob}.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the simulated test data in the $g^\text{MGBM}$ margin space.
 
@@
 \includegraphics[width=8cm]{img/sim_mgbm_glob.png}
+\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.}
 \label{fig:sim_mgbm_glob}
-\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
-\end{figure}
-
-% highlight monotonicity 
+\end{figure}
+
+\noindent As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.
+
+Figure \ref{fig:sim_mgbm_glob_pdp_ice} highlights PD, ICE, and histograms of the most important features from Figure \ref{fig:sim_mgbm_glob}.
+PD gives an indication of the estimated average prediction of $g^\text{MGBM}(\mathbf{X})$ in the test data for some $X_j$.
+ICE serves several purposes.
+It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is itself a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior.
 PD and ICE are always displayed with a histogram herein to highlight any sparse regions in an input feature's domain.
@@
 \includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
+\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
 \label{fig:sim_mgbm_glob_pdp_ice}
-\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
-\end{figure}
-
-Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$.
+\end{figure}
+
+$X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE.
+Also, as the PD curves generally follows the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
+Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is likely not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}.
+The one-dimensional monotonic constraints probably hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data.
+This is an interesting result for responsible ML practitioners.
+In some noisy scenarios, monotonicity constraints can increase model interpretability without causing a drastic drop in model accuracy, even when known interactions and non-monotonic behavior exist in training data.
+
+Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
+These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_glob_pdp_ice} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
+Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
 
@@
 
+\noindent The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result.
+For the lower prediction at the 10\textsuperscript{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative.
+At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive.
+In each case, large local contributions follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.
+
+Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique.
+Deep SHAP values are reported in the probability space, after the application of the logit link function.
+They are also calculated from the projection layer of $g^\text{XNN}$.
+Thus, the Deep SHAP values in Figure \ref{fig:sim_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the simulated test data.
+
 \begin{figure}[H]
@@
 \includegraphics[width=8cm]{img/sim_xnn_glob.png}
+\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the simulated test data.}
 \label{fig:sim_xnn_glob}
-\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the simulated test data.}
-\end{figure}
+\end{figure}
+
+\noindent Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features.
+The consistency between the feature rankings of $g^\text{MGBM}$ and $g^\text{XNN}$ is somewhat striking, given their different hypothesis families and architectures.
+Both $g^\text{MGBM}$ and $g^\text{XNN}$ rank $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ as the most important features, both place $X_{\text{Categorical},2}$ and $X_{\text{Categorical},3}$ above the $X_{\text{Binary},1}$ and $X_{\text{Binary},2}$ features, both rank $X_{\text{Binary},1}$ above $X_{\text{Binary},2}$, and both place the least importance on $X_{\text{Categorical},4}$ and $X_{\text{Categorical},0}$.
 
@@
 \includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
+\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the simulated test data.}
 \label{fig:sim_xnn_glob_ridge}
-\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the simulated test data.}
-\end{figure}
+\end{figure}
+
+In combination, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
 
@@
 \end{figure}
-
-Deep SHAP values are reported in the probability space, after the application of the logit link function.
 
@@
 Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.
-It is also highly statistically significant (not shown), which together would indicate that there may be evidence of illegal DI.
+It is also highly statistically significant (not shown, but available in resources discussed in Section \ref{sec:soft}), which together would indicate that there may be evidence of illegal DI.
 As with ME and other measures, the reasonableness of this disparity is not clear outside of context.
@@
 This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
-$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$.
+$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$ (not shown, but available in resources discussed in Section \ref{sec:soft}).
 
@@
 However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value.
-This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci}.
+This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Section \ref{sec:soft}).
 However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
@@
 For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
-In combination, Figure \ref{fig:mort_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
 
@@
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{\textcolor{red}{24}}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
 Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena.

### 2019-12-20 article-information-2019.tex
finish first draft of 2.1

@@
 
-Ridge functions are reminiscent of basis functions ... distinctive simplistic functions
+Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
+\ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
+The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms.
+Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
 
@@
 
-In combination, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
-
-Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$.
+\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks.
+Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
+In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
+Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features and do display some noticeable sinusoidal characteristics.
+Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
+
+Local Deep SHAP feature importance in Figure \ref{fig:sim_xnn_loc} supplements the global interpretability of $g^\text{XNN}$ displayed in Figures \ref{fig:sim_xnn_glob} and \ref{fig:sim_xnn_glob_ridge}.
+Local Deep SHAP values are extracted from the projection layer of $g^\text{XNN}$ and reported in the probability space.
+Deep SHAP values can be calculated for any arbitrary $g^\text{XNN}(\mathbf{x})$, allowing for detailed, local summarization of individual model predictions.
 
@@
 \end{figure}
+
+\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x}^{(i)})$ predictions.
 
@@
 Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.
-It is also highly statistically significant (not shown, but available in resources discussed in Section \ref{sec:soft}), which together would indicate that there may be evidence of illegal DI.
+It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), which together would indicate that there may be evidence of illegal DI.
 As with ME and other measures, the reasonableness of this disparity is not clear outside of context.
@@
 This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
-$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$ (not shown, but available in resources discussed in Section \ref{sec:soft}).
+$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$ (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
 
@@
 However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value.
-This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Section \ref{sec:soft}).
+This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
 However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.

### 2019-12-20 article-information-2019.tex
proofing

@@
 For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures.
-By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present the art of the possible for important ML applications that require high accuracy and interpretability and minimal discrimination.}
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents: adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures.
+By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}
 
@@
 
-Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e.
-ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
+Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e., ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
 ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
-While this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been solved (even as singular entities, much less as complex intersectional phenomena), Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+While this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been resolved (even as singular entities, much less as complex intersectional phenomena), Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
 
 Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an appealable decision-making framework, tests for DI and other social discrimination, and public and open source software resources.
-In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and discrimination testing results are also presented for interpretable models.
+In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and post-hoc explanation and discrimination testing results are also presented for interpretable models.
 Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results.
@@
 The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
-Within the training set, a 5 fold cross validation indicator is used for training and assessing all models.
+Within the training set, a 5 fold cross validation indicator is used for training all models.
 For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
@@
 
-The US HMDA, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
+The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
 This information is first provided to the Consumer Financial Protection Bureau (CFPB), which subsequently releases some of the data to the public.
@@
 These loans are a subset of all originated mortgage loans in the 2018 HMDA data that were chosen to represent a relatively comparable group of consumer mortgages.
-A selection of features is used to predict whether a loan is \textit{high-priced}, i.e.
-the annual percentage rate (APR) charged was 150 basis points (1.5\%) or more above a survey-based estimate of other similar loans offered around the time of the given loan.
+A selection of features is used to predict whether a loan is \textit{high-priced}, i.e., the annual percentage rate (APR) charged was 150 basis points (1.5\%) or more above a survey-based estimate of other similar loans offered around the time of the given loan.
 After data cleaning and preprocessing to encode categorical features and create missing markers, the mortgage data contains ten input features and the binary outcome, \textit{high-priced}.
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the MGBM, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
 For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
@@
 \begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in the left child weight being greater than the right child weight, $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{b,j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
+\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{b,j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
 \item For any subsequent left child node involving $X_j$, any $\theta_{b,j, k\ge1}$ resulting in $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ where $w_{b,j,k\ge1,L} > w_{b,j,k\ge1,R}$, is not considered.
@@
 Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
-For a negative monotonic constraint, i.e.
-$g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
+For a negative monotonic constraint, i.e., $g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
 Also consider that MGBM models with independent monotonicity constraints between some $X_j$ and $\mathbf{y}$ likely restrict non-monotonic interactions between multiple $X_j$.
@@
 XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
-XNNs also bare a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e.
-GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
+XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e., GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
 Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
@@
 \begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, is known as the \textit{projection layer} and is fully connected to each input feature, $X_j$.
+\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, which should learn higher magnitude weights for important $X_j$, is known as the \textit{projection layer}.
+It is fully connected to each input $X_j$.
 Each hidden unit in the projection layer may optionally include a bias term.
@@
 ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$ \cite{ice_plots}.
-PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g.
-inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
+PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
 
@@
 In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.
-However, other factors, such as hyperparameter or algorithm choice, can drive DI.
+However, other factors, such as hyperparameter or algorithm choices, can drive DI.
 Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.
@@
 \begin{aligned}
-\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| \mathbf{X}_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | \mathbf{X}_p = 1))
+\text{ME} \equiv 100 \cdot (\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)  - \text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1))
 \end{aligned}
@@
 
-\noindent where $\mathbf{X}_p$ and $\mathbf{X}_c$ represent binary markers created from some demographic attribute $\mathbf{X}_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and conditional probabilities, $\text{Pr}(\hat{\mathbf{y}}|\mathbf{X}_j = \alpha)$, are evaluated over all $\mathbf{x}^{(i)} \in \{\mathbf{X}_j | x^{(i)}_j = \alpha\}$.
-ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
+\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute $X_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group,  and $\text{Pr}(\hat{\mathbf{y}}|X_j)$ indicates a conditional probability regarding model outcomes, $\hat{\mathbf{y}}$, and some feature, $X_j$, evaluated over the relevant partition of data.
+ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is adverse impact ratio (AIR), more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
 
@@
 \begin{aligned}
-\text{AIR} \equiv \frac{\text{Pr}(\hat{\mathbf{y}} = 1 | \mathbf{X}_p = 1)}{\text{Pr}(\hat{\mathbf{y}} = 1| \mathbf{X}_c = 1)}
+\text{AIR} \equiv \frac{\text{Pr}(\hat{\mathbf{y}} = 1 | X_p = 1)}{\text{Pr}(\hat{\mathbf{y}} = 1| X_c = 1)}
 \end{aligned}
@@
 
-\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the percent of the control class that receives a favorable outcome.
-Another long-standing measure of DI is SMD.
-SMD is typically used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
+\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the proportion of the control class that receives a favorable outcome.
+Another long-standing measure of DI is standardized mean difference (SMD).
+SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
 It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}.
-The SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
+SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
 
@@
 \noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.
-Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of features, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
+Because of this standardization factor, SMD allows for a comparison across different types of outcomes, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
 In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.
@@
 
-A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute prima facia evidence of illegal discrimination.
-Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not meaningful.
+A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.
+Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
 It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model.
-Moreover, practical significance is also needed model by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production.
+Moreover, practical significance is also needed by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production.
 Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
-One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes prima facia evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
 ``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.
@@
 The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.
-The first guidance in U.S.
-courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v.
+The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v.
 Partida}.\footnote{\textit{Castaneda v.
-Partida}, 430 US 482 - Supreme Court (1977)}   Here, the U.S.
-Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v.
+Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v.
 United States.}\footnote{\textit{Hazelwood School Dist.
 v.
-United States}, 433 U.S.
-299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
+United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
 
@@
 Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
+For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
+For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
+Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
+Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
 
@@
 \includegraphics[width=8cm]{img/sim_mgbm_glob.png}
-\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.}
+\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the simulated test data.}
 \label{fig:sim_mgbm_glob}
@@
 ICE serves several purposes.
-It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is itself a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior.
-PD and ICE are always displayed with a histogram herein to highlight any sparse regions in an input feature's domain.
+It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is also a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior.
+PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain.
 Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
@@
 $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE.
-Also, as the PD curves generally follows the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
-Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is likely not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}.
-The one-dimensional monotonic constraints probably hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data.
+Also, as the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
+Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is probably not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}.
+The one-dimensional monotonic constraints may hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data.
 This is an interesting result for responsible ML practitioners.
@@
 Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
-These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_glob_pdp_ice} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
+These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
 Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
@@
 At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive.
-In each case, large local contributions follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.
+In each case, large local contributions generally follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.
 
@@
 \includegraphics[width=8cm]{img/sim_xnn_glob.png}
-\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the simulated test data.}
+\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
 \label{fig:sim_xnn_glob}
@@
 \includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
-\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the simulated test data.}
+\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
 \label{fig:sim_xnn_glob_ridge}
@@
 In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
-Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features and do display some noticeable sinusoidal characteristics.
-Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
+Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.
+Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
 
@@
 \includegraphics[width=15cm]{img/sim_xnn_loc.png}
-\caption{Deep SHAP values for three observations across selected percentiles of $g^\text{XNN}$ on the simulated test data.}
+\caption{Deep SHAP values for three observations across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
 \label{fig:sim_xnn_loc}
@@
 Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMD calculated using the probabilities from the models, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIR calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{The decision of what cutoff should be used in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{The selection of which cutoff to use in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.
 For example, a model developer at a bank might build a credit model where the F1 score is maximized at a delinquency probability cutoff of 0.15.
@@
 But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
-Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assumes that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might want to consider the relative false negative rates as a measure of the class-control disparities.
-Table \ref{tab:sim_disc} shows that protected group 1 has higher relative false negative rates under $g^{\text{XNN}}$ (1.14 vs.
+Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.}
+Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities.
+Table \ref{tab:sim_disc} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
 1.06).
-However, the overall false negative rates were lower for  $g^{\text{XNN}}$ (0.383 vs.
+However, in Table \ref{tab:sim_n_acc} the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
 0.401).
@@
 \midrule
-Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.764\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.150 \\ 0.167\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.401 \\ 0.383\end{tabular} \\
-\midrule			 
-Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.751\end{tabular}	& \begin{tabular}{@{}c@{}} 0.129 \\ 0.149\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.337\end{tabular} \\
-\midrule
-Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.761\end{tabular}	& \begin{tabular}{@{}c@{}} 0.169 \\ 0.183\end{tabular}	& \begin{tabular}{@{}c@{}} 0.331 \\ 0.306\end{tabular} \\
-\midrule
-Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.745\end{tabular}	& \begin{tabular}{@{}c@{}} 0.091 \\ 0.116\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.370\end{tabular} \\
+Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.150 \\ 0.168\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
+\midrule			 
+Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.129 \\ 0.163\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
+\midrule
+Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular}	& \begin{tabular}{@{}c@{}} 0.169 \\ 0.186\end{tabular}	& \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
+\midrule
+Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.091 \\ 0.138\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\
 \bottomrule
@@
 
-11.1\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.
-Of note is that 11.1\% is not a meaningful difference without context.
-If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 11.1\% could represent an extremely large difference and would require remediation.
-But if they represent substantially different populations, then 11.1\% could represent a reasonable deviation from parity.
-As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 11.1\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population.
-Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.
-It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), which together would indicate that there may be evidence of illegal DI.
+For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.
+Of note is that 12.0\% is not a meaningful difference without context.
+If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 12.0\% could represent an extremely large difference and would require remediation.
+But if they represent substantially different populations, then 12.0\% could represent a reasonable deviation from parity.
+As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 12.0\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population.
+Protected group 1’s AIR under $g^\text{XNN}$ is 0.727, below the EEOC 4/5ths rule threshold.
+It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
+Together these would indicate that there may be evidence of illegal DI.
 As with ME and other measures, the reasonableness of this disparity is not clear outside of context.
 However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.
+Some pertinent remediation strategies for discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
 
@@
 \midrule 
-$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.737 \\ 1.04\end{tabular}	& \begin{tabular}{@{}c@{}} 11.1\%\\ -1.7\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.240 \\ 0.551\end{tabular}	& \begin{tabular}{@{}c@{}} 1.12 \\ 1.59\end{tabular}	& \begin{tabular}{@{}c@{}} 1.14 \\ 0.827\end{tabular} \\
+$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.03 \\ 1.35\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
 \bottomrule
@@
 
-\noindent Even though SMD is often used to measure disparities for non-categorical outcomes, it is calculated here on the models’ probabilities prior to being transformed into classifications.
-This measurement would be particularly relevant if the probabilities are being used in combination with other models to determine an outcome.
-The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.21 versus -0.24), but both are close to Cohen’s small effect threshold of -0.20.
+\noindent SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
+(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.206 vs.
+-0.274), but both are close to Cohen’s small effect threshold of -0.20.
 Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.
@@
 $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants.
-Assuming that small fit differences on static test data do not outweigh the need for intrinsic model interpretability and post-hoc explainability in high-stakes, human-centered, or regulated applications, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
-For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
-For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
-Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
-Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
+Assuming that small fit differences on static test data do not outweigh the need for intrinsic model interpretability and reliable post-hoc explainability in high-stakes, human-centered, or regulated applications, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
 
@@
 Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data.
-This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g.
-numerical encoding of categorical features and missing values.
+This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g., numerical encoding of categorical features and missing values.
 This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
@@
 \includegraphics[width=8cm]{img/mort_mgbm_glob.png}
-\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the mortgage test data.}
+\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
 \label{fig:mort_mgbm_glob}
@@
 Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
-Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$, which does allow for the modeling of non-monotonic interactions.
+Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and perhaps due to less noise in the mortgage data, this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data.
 DTI ratio and LTV ratio also appear to have sparse regions in their univariate distributions.
@@
 For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions.
-However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value.
+However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and appear in a sparse region of property value.
 This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
@@
 
-\noindent Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability to accurately summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
+\noindent Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to \textit{accurately} summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
 In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.
@@
 \includegraphics[width=8cm]{img/mort_xnn_glob.png}
-\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}$ on the mortgage test data.}
+\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
 \label{fig:mort_xnn_glob}
@@
 
-\noindent As compared to $g^\text{MGBM}$, $g^\text{XNN}$ assigns higher importance to property value and loan amount, and lower importance on LTV ratio, and the income related features, DTI ratio and income.
+\noindent As compared to $g^\text{MGBM}$, $g^\text{XNN}$ assigns higher importance to property value, loan amount, and income, and lower importance on LTV ratio and DTI ratio.
 
@@
 As in the subsection \ref{ssec:int_ex_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge} \textbf{C}.
-In both the simulated and mortgage data, $g^\text{XNN}$ $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, perhaps reminiscent of the visual primitives often learned by low layers of pattern-detecting convolutional neural networks (CNNs) \cite{lecun2015deep}.
+In both the simulated and mortgage data, $g^\text{XNN}$ $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
 
@@
 \includegraphics[width=16cm]{img/mort_xnn_glob_ridge.png}
-\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the mortgage test data.}
+\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
 \label{fig:mort_xnn_glob_ridge}
@@
 
-\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_k$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
+\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
 For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
@@
 \includegraphics[width=15cm]{img/mort_xnn_loc.png}
-\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}$ on the mortgage test data.}
+\caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
 \label{fig:mort_xnn_loc}
@@
 \subsubsection{Discrimination Testing Results}\label{ssec:dis_mort}
+
+Tables \ref{tab:mort_n_acc} and \ref{tab:mort_disc} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
+As with the simulated data, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the FPRs, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).
+Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
+Thus, one might consider FPR ratios as a measure of the class-control disparities.
+FPR ratios are higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs.
+2.10) in Table \ref{tab:mort_disc},  but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs.
+0.315) in Table \ref{tab:mort_n_acc}.
+This is the same pattern seen in the simulated data results in Section \ref{ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
+Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
 
@@
 
+Both ME and AIR show higher disparities for blacks under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$.
+Blacks receive high-priced loans 21.4\% more frequently using $g^{\text{XNN}}$ vs.
+18.3\% for $g^{\text{MGBM}}$.
+Both $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ show AIRs that are statistically significantly below parity (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), and which are also below the EEOC’s 0.80 threshold.
+This would typically indicate cause for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
+On the other hand, women improve under $g^{\text{XNN}}$ vs.
+$g^{\text{MGBM}}$ (MEs of 3.6\% vs.
+4.1\%; AIRs of 0.955 vs.
+0.948).
+The AIRs, while statistically significantly below parity, are well above the EEOC’s threshold of 0.80.
+In most situations, the values of these metrics alone would not likely flag a model for further review, though there may be other considerations which would lead to concern.
+
 \begin{table}[H]
@@
 $g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular}	& \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular}	& \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\
-
 \bottomrule
@@
 \end{table}
+
+Black SMDs for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$ are similar: 0.621 and 0.628, respectively.
+These exceed Cohen’s guidelines of 0.5 for a medium effect size and would likely trigger further review.
+Female SMDs are well below Cohen’s definition of small effect size: 0.105 and 0.084 for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$, respectively.
+Similar to results for female AIR, these values alone are unlikely to prompt further review.
 
@@
 In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
-Among others, practitioners can now evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{See: \href{https://github.com/xiyanghu/OSDT}{Optimal sparse decision trees}.}\textsuperscript{,}\footnote{See: \href{https://github.com/cfchen-duke/ProtoPNet}{\textit{This looks like that} interpretable deep learning}.} Additional, relevant references and Python functionality include:
+Among several types of currently available interpretable models, practitioners can now use Python to evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{See: \href{https://github.com/xiyanghu/OSDT}{Optimal sparse decision trees}.}\textsuperscript{,}\footnote{See: \href{https://github.com/cfchen-duke/ProtoPNet}{\textit{This looks like that} interpretable deep learning}.} Additional, relevant references and Python functionality include:
 
@@
 
-\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.
-\href{https://github.com/openai/gpt-2/blob/master/model_card.md}{GPT-2 model card}, \href{https://modelcards.withgoogle.com/object-detection}{Object Detection model card}.
+\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.,  \href{https://github.com/openai/gpt-2/blob/master/model_card.md}{GPT-2 model card}, \href{https://modelcards.withgoogle.com/object-detection}{Object Detection model card}.
 
@@
 Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, providing people affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
 ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions is difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
@@
 
-\subsection{Viable Discrimination Remediation Approaches}
+\subsection{Viable Discrimination Remediation Approaches}\label{ssec:disc_rem}
 
@@
 The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own.
-However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g.
-the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
@@
 Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
-However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks and that explanations can mask ML discrimination.
+However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination.
 Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory.
@@
 The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects.
-Of course there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
+Of course, there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
 While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work (and many, many others) illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank. Wells Fargo is a customer of, and investor in, H2O.ai and a customer of BLDS, LLC.}
+\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank.
+Wells Fargo is a customer of, and investor in, H2O.ai and a customer of BLDS, LLC.
+However, communications regarding XNN between Wells Fargo and the authors have been extremely limited prior to and during the drafting of this manuscript.
+Moreover, Wells Fargo exerted absolutely no editorial control over the text or results herein.}
 

### 2019-12-20 article-information-2019.tex
add last section

@@
 
-\subsection{Impact of Discrimination Testing on Model Use and Adoption}
-
-\subsection{Viable Discrimination Remediation Approaches}\label{ssec:disc_rem}
+\subsection{Discrimination Testing and Remediation in Practice}\label{ssec:disc_rem}
+
+The heightened interest in exploring and fixing algorithmic bias has led to a significant body of work focused in using ML to diminish discrimination in model outcomes \cite{friedler2019comparative}.
+Broadly, these can be defined into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and more recently developed approaches that change the algorithms or input data themselves in order to mitigate disparities.
+Many approaches that have been developed in the last five to seven years focus on altering the algorithm itself, preprocessing the data, or post-processing the predictions in order to diminish class-control correlations or dependencies.
+IBM’s \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} toolkit provides an open source Python package for many of the metrics and techniques that have been developed recently.
+
+Whether these methods are suitable for a particular use case depends on the legal environment in which the model will be used and the use case itself.
+Great care must be taken in order to ensure that the appropriate fairness metrics are chosen, because certain metrics may not be appropriate for some use cases.
+Additionally, the effects of changing the model must be viewed holistically.
+For example, the mortgage data disparity analysis in Section \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
+
+Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
+A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{As discussed in \textit{Bazemore v.
+Friday}, 478 US 385 (1986), a discrimination in pay case, the US Supreme Court wrote that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) usually considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will usually have the same model output.
+Obviously, the inclusion of protected class status will change model output by class.
+With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.
+Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
+
+Of the newer class of fairness enhancing interventions in ML, within-algorithm discrimination mitigation techniques that do not use class information may be more likely to be acceptable in highly regulated settings today.
+These techniques work by incorporating a loss function where more discriminatory paths or weights are penalized and will only be used by the model if the increase in model performance overcomes the penalty (the relative level of performance-to-discrimination penalty is determined via a hyperparameter choice).
+Other mitigation strategies that only alter hyperparameters or algorithm choice are also likely to be acceptable.
+And feature selection techniques that have been used in traditional modeling (ordinary least squares, logistic, and simple decision tree models) are likely to continue to be accepted in regulatory environments.
+For further discussion of potential and utilized techniques that can mitigate disparate impact in financial services, see \citet{schmidt2019introduction}.
+
+Regardless of the methodology chosen to minimize disparities, advances in computing have enhanced the ability to search for less discriminatory models.
+Prior to these advances, only a small number of alternative algorithms could be tested for lower levels of disparity without causing infeasible delays in model implementation.
+Now, tens of thousands of alternative models or more can be quickly tested for lower discrimination and better predictive quality.
+An additional opportunity arises as a result of ML itself: the well-known Rashomon effect, or the multiplicity of good ML models for most datasets.
+Thus, it is now feasible to train more models, find more good models, and test more models for discrimination, and among all those tested models, there are likely to be some with high predictive performance and low discrimination.
 
@@
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{\textcolor{red}{24}}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{25}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
 Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena.

### 2019-12-21 article-information-2019.tex
commit submitted version

@@
 The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents: adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures.
-By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present a template workflow for important ML applications that require high accuracy and interpretability and minimal discrimination.}
+By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
 
@@
 
-\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute $X_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group,  and $\text{Pr}(\hat{\mathbf{y}}|X_j)$ indicates a conditional probability regarding model outcomes, $\hat{\mathbf{y}}$, and some feature, $X_j$, evaluated over the relevant partition of data.
+\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), and $p$ indicates a protected group.
 ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is adverse impact ratio (AIR), more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
@@
 Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
-A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{As discussed in \textit{Bazemore v.
-Friday}, 478 US 385 (1986), a discrimination in pay case, the US Supreme Court wrote that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) usually considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will usually have the same model output.
+A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{In the pay discrimination case, \textit{Bazemore v.
+Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will have the same model output (though there may be some small or random differences in outcomes due to computational issues).
 Obviously, the inclusion of protected class status will change model output by class.
-With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.
-Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
+With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.\footnote{In a reverse discrimination case, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.}  Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
 
@@
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{25}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{26}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
 Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena.
@@
 \conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank.
-Wells Fargo is a customer of, and investor in, H2O.ai and a customer of BLDS, LLC.
+Wells Fargo is a customer of, and investor in, H2O.ai and a client of BLDS, LLC.
 However, communications regarding XNN between Wells Fargo and the authors have been extremely limited prior to and during the drafting of this manuscript.

### 2020-01-27 article-information-2019.tex
initial chop, -4 pages

@@
 
+\usepackage{caption}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Notation}\label{ssec:not}
-
-To facilitate descriptions of data and modeling, explanatory, and discrimination testing techniques, notation for input and output spaces, datasets, and models is defined.
-
-\subsubsection{Spaces}
-
-	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-		\item Input features come from the set $\mathcal{X}$ contained in a \textit{P}-dimensional input space, $\mathcal{X} \subset \mathbb{R}^P$.
-An arbitrary, potentially unobserved, or future instance of $\mathcal{X}$ is denoted $\mathbf{x}$, $\mathbf{x} \in \mathcal{X}$.
-		\item Labels corresponding to instances of $\mathcal{X}$ come from the set $\mathcal{Y}$.
-		\item Learned output responses of models are contained in the set $\mathcal{\hat{Y}}$.
-% For regression models, the set $\mathcal{\hat{Y}}_r$ is also contained in a $C$-dimensional output space, $\mathcal{\hat{Y}}_r \subset \mathbb{R}^{C_r}$.
-For classification models, the set $\mathcal{\hat{Y}}_c$ typically contains a column vector for each unique class in $\mathcal{Y}$.
-Hence, $\mathcal{\hat{Y}}_c$ is contained in a $C'$-dimensional output space,  $\mathcal{\hat{Y}}_c \subset \mathbb{R}^{C'_c}$.
-	\end{itemize}
-
-\subsubsection{Datasets}
-
-	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-		\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
-		\item Each $i$-th observation of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
-		$[x_0^{(i)}, x_1^{(i)}, \dots, x_{\textit{P}-1}^{(i)}]$, with corresponding $i$-th labels in $\mathbf{Y}, \mathbf{y}^{(i)}$, and corresponding predictions in $\mathbf{\hat{Y}}, \mathbf{\hat{y}}^{(i)}$.
-% = [y_0^{(i)}, y_1^{(i)}, \dots, y_{\textit{C}-1}^{(i)}]$.
-		\item $\mathbf{X}$ and $\mathbf{Y}$ consist of $N$ tuples of observations: $[(\mathbf{x}^{(0)},\mathbf{y}^{(0)}), (\mathbf{x}^{(1)},\mathbf{y}^{(1)}), \dots,(\mathbf{x}^{(N-1)},\mathbf{y}^{(N-1)})]$.
-%\\ $\mathbf{x}^{(i)} \in \mathcal{X}$, $\mathbf{y}^{(i)} \in \mathcal{Y}$.
-		\item Each $j$-th input column vector of $\mathbf{X}$ is denoted as $X_j = [x_{j}^{(0)}, x_{j}^{(1)}, \dots, x_{j}^{(N-1)}]^T$.
-	\end{itemize}
-
-\subsubsection{Models}
-
-	\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-		\item A type of ML model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$:
-		$ \mathbf{X}, \mathbf{Y} \xrightarrow{\mathcal{A}} g$, such that $g \approx f$.
-		\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
-		\item The model to be explained and tested for discrimination is denoted as $g$.
-	\end{itemize}
 
@@
 
-The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
-This information is first provided to the Consumer Financial Protection Bureau (CFPB), which subsequently releases some of the data to the public.
-Regulators often use HMDA data to, ``...show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory.''\textsuperscript{5} In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
-The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, loan to value (LTV) ratio, debt to income (DTI) ratio, and other important financial descriptors.
-The data also include information on each borrower and co-borrower’s race, ethnicity, gender, and age.
-Because the data includes information on these protected class characteristics, certain metrics that can be indicative of discrimination in lending can be calculated directly using the HDMA data.
-
 The mortgage dataset analyzed here is a random sample of consumer-anonymized loans from the HDMA database.
@@
 While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (LTV ratio, DTI ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
-Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+See Appendix \ref{a_sec:mort} for general information regarding HMDA data.
 
@@
 \noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
-For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
-
-\begin{equation}
-\begin{aligned}
-\mathcal{L}_{b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{MGBM}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
-\end{aligned}
-\end{equation}
-
-\noindent In addition to $\mathcal{L}_{b}$, $g^\text{MGBM}$ training is characterized by additional splitting rules and constraints on tree node weights.
-Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$.
-For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction.
-For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, $g^{\text{MGBM}}$ is positive monotonic with respect to some $X_j$ if $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
-The following rules and constraints ensure positive monotonicity in $\Theta_b$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$.
-
-\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{b,j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
-\item For any subsequent left child node involving $X_j$, any $\theta_{b,j, k\ge1}$ resulting in $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ where $w_{b,j,k\ge1,L} > w_{b,j,k\ge1,R}$, is not considered.
-\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$, $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ are bound by the associated $\theta_{b,j,k-1}$ set of node weights, $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$, such that $ \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} < \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
-\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $ w_{b,j,k,L} \le w_{b,j,k,R}$ and $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
-\end{enumerate}
-
-\noindent Note that for any one $X_j$ and $T_b \in g^{\text{MGBM}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
-Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0}) = w_{b,0}$.
-Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
-For a negative monotonic constraint, i.e., $g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
-Also consider that MGBM models with independent monotonicity constraints between some $X_j$ and $\mathbf{y}$ likely restrict non-monotonic interactions between multiple $X_j$.
-Moreover, if monotonicity constraints are not applied to all $X_j \in \mathbf{X}$, any strong non-monotonic signal in training data associated with some important $X_j$ maybe forced onto some other arbitrary unconstrained $X_j$ under some $g^{MGBM}$ models, compromising the end goal of interpretability.
 
@@
 In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and five-fold cross validation and test partitions are used for model assessment.
+See Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
 For exact parameterization of the two $g^\text{MGBM}$ models, see the software resources referenced in Subsection \ref{ssec:soft}.
@@
 
-\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:
-
-\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, which should learn higher magnitude weights for important $X_j$, is known as the \textit{projection layer}.
-It is fully connected to each input $X_j$.
-Each hidden unit in the projection layer may optionally include a bias term.
-\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
-Each $n_k$ is a neural network, which can be parameterized to suit a given modeling task.
-To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
-Each $n_k$ can contain several bias terms.
-\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
-Again, subnetwork output is restricted to 1-dimension for interpretation and visualization purposes.
-\end{enumerate}
-
 Here, each $g^\text{XNN}$ is trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
@@
 Be aware that $g^{XNN}$ model architectures may require manual and judicious feature selection due to long training times.
-For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Partial Dependence and Individual Conditional Expectation}\label{ssec:pd_ice}
+For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft} and Appendices  \ref{a_ssec:not} and \ref{a_ssec:xnn}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{One-dimensional Partial Dependence and Individual Conditional Expectation}\label{ssec:pd_ice}
 
@@
 PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
-
-Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$) is considered.
-$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
-$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
-PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
-Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
-PD should be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.
+For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Subsection \ref{ssec:soft} and Appendices \ref{a_ssec:not} and \ref{a_ssec:pd_ice}.
 
@@
 Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values should never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
-For some instance $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
-
-\begin{equation}
-\label{eq:shap_additive}
-\begin{aligned}
-g(\mathbf{x}) = \phi_0 + \sum_{j=0}^{j=\mathcal{P} - 1} \phi_j \mathbf{z}_j
-\end{aligned}
-\end{equation}
-
-\noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
-Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
-Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
-
-\begin{equation}
-\label{eq:shap_contrib}
-\begin{aligned}
-\phi_{j} = \sum_{S \subseteq \mathcal{P} \setminus \{j\}}\frac{|S|!(\mathcal{P} -|S| -1)!}{\mathcal{P}!}[g_x(S \cup \{j\}) - g_x(S)]
-\end{aligned}
-\end{equation}
-
-\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
-Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some set $\mathbf{X}$.
 
@@
 Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.
 
@@
 The discussed metrics are also representative of fair lending analyses and pair well with the mortgage data.
-Before discussing the techniques, it is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI).
-DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
-With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI, colloquially known as \textit{unintentional discrimination}, occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
-In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.
-However, other factors, such as hyperparameter or algorithm choices, can drive DI.
-Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.
-
-The analyses and metrics herein focus on several measures of disparate impact that are commonly used in US litigation and regulatory settings.
-One is known as marginal effects (ME).
+See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination under US laws and Appendix \ref {a_sec:prac_v_stat} for remarks on statistical vs.
+practical significance for discrimination measure values.
+
+One measure is known as marginal effects (ME).
 ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.
@@
 Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.
-
-A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.
-Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
-It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model.
-Moreover, practical significance is also needed by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production.
-Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
-One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
-``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.
-
-The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.
-The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v.
-Partida}.\footnote{\textit{Castaneda v.
-Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v.
-United States.}\footnote{\textit{Hazelwood School Dist.
-v.
-United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
 
@@
 Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{The selection of which cutoff to use in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.
-For example, a model developer at a bank might build a credit model where the F1 score is maximized at a delinquency probability cutoff of 0.15.
-For purposes of evaluating the quality of the model, she may review confusion matrix statistics (accuracy, recall, precision, etc.) using cutoffs based on the maximum F1 score.
-But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
-Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.}
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to cutoff selection and discrimination testing.}
+
 Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities.
@@
 \begin{table}[H]
-\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
-\centering
-\begin{tabular}{cccccc}
+	\captionsetup{width=7cm}
+	\hskip-20pt\parbox{.48\textwidth}{
+	\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
+	\centering
+	\begin{tabular}{ccccc}
 		\toprule
-\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy}  								& \textbf{FPR}									& \textbf{FNR} \\
+		\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} & \textbf{Accuracy} & \textbf{FNR} \\
 		\midrule
-Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.150 \\ 0.168\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
+		Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} 	&  \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
 		\midrule			 
-Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.129 \\ 0.163\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
+		Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
 		\midrule
-Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular}	& \begin{tabular}{@{}c@{}} 0.169 \\ 0.186\end{tabular}	& \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
+		Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular}	&  \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
 		\midrule
-Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.091 \\ 0.138\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\
+		Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\
 		\bottomrule
@@
 	\end{tabular}
+	}\hspace{10pt}
+	\parbox{.48\textwidth}{
+	\vspace{-13pt}
+	\caption{AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
+	\centering
+	\begin{tabular}{ccccccc}
+		\toprule
+		\textbf{Model}		&  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular}					& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD} &  \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}\\
+		\midrule
+		$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}		& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
+		\midrule 
+		$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}		& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
+		\bottomrule
+		\label{tab:sim_disc}
+	\end{tabular}
+	}
 \end{table}
@@
 Some pertinent remediation strategies for discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
-
-\begin{table}[H]
-\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
-\centering
-\begin{tabular}{cccccccc}
-\toprule
-\textbf{Model}		& \textbf{Protected Class}										& \textbf{Control Class}								& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD}   								& \textbf{FPR Ratio}							& \textbf{FNR Ratio} \\
-\midrule
-$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.16 \\ 1.86\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
-\midrule 
-$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.03 \\ 1.35\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
-\bottomrule
-\label{tab:sim_disc}
-\end{tabular}
-\end{table}
 
@@
 \begin{table}[H]
-\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
-\centering
-\begin{tabular}{ccccccc}
+	\captionsetup{width=7cm}
+	\hskip-20pt\parbox{.48\textwidth}{
+		\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.} 
+		\centering
+		\begin{tabular}{ccccc}
 \toprule
-\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy} 									& \textbf{FPR}											& \textbf{FNR} \\
+\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy} 																			& \textbf{FNR} \\
 \midrule
-Black 		& 2,608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} 			& \begin{tabular}{@{}c@{}} 0.457 \\ 0.308\end{tabular} \\
+Black 		& 2,608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} 				& \begin{tabular}{@{}c@{}} 0.457 \\ 0.308\end{tabular} \\
 \midrule			 
-White 		& 28,361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.150 \\ 0.120\end{tabular}			& \begin{tabular}{@{}c@{}} 0.508 \\ 0.360\end{tabular}\\
+White 		& 28,361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular}			& \begin{tabular}{@{}c@{}} 0.508 \\ 0.360\end{tabular}\\
 \midrule
-Female 		& 8,301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular}		& \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}			& \begin{tabular}{@{}c@{}} 0.402 \\ 0.322\end{tabular}\\
+Female 		& 8,301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular}		& \begin{tabular}{@{}c@{}} 0.402 \\ 0.322\end{tabular}\\
 \midrule
-Male 		& 13,166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.182 \\ 0.131\end{tabular}			& \begin{tabular}{@{}c@{}} 0.497 \\ 0.347\end{tabular}\\
+Male 		& 13,166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular}			& \begin{tabular}{@{}c@{}} 0.497 \\ 0.347\end{tabular}\\
 \bottomrule
@@
 		\end{tabular}
+	}\hspace{10pt}
+	\parbox{.48\textwidth}{
+		\vspace{-13pt}
+		\caption{AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+		\centering
+		\begin{tabular}{ccccccc}
+\toprule
+\textbf{Model}		&  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular}					& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD} &  \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}\\
+\midrule
+$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular}	& \begin{tabular}{@{}c@{}} 0.900 \\ 0.810\end{tabular} \\
+\midrule 
+$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular}	& \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\
+\bottomrule
+			\label{tab:mort_disc}
+		\end{tabular}
+	}
 \end{table}
@@
 
-\begin{table}[H]
-\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
-\centering
-\begin{tabular}{cccccccc}
-\toprule
-\textbf{Model}	& \textbf{Protected Class}	& \textbf{Control Class}	& \textbf{AIR}	& \textbf{ME}	& \textbf{SMD}  & \textbf{FPR Ratio}		& \textbf{FNR Ratio} \\
-\midrule
-$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular}	& \begin{tabular}{@{}c@{}} 2.10 \\ 1.15\end{tabular}	& \begin{tabular}{@{}c@{}} 0.900 \\ 0.810\end{tabular} \\
-\midrule 
-$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular}	& \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular}	& \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\
-\bottomrule
-\label{tab:mort_disc}
-\end{tabular}
-\end{table}
-
 Black SMDs for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$ are similar: 0.621 and 0.628, respectively.
@@
 Many approaches that have been developed in the last five to seven years focus on altering the algorithm itself, preprocessing the data, or post-processing the predictions in order to diminish class-control correlations or dependencies.
-IBM’s \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} toolkit provides an open source Python package for many of the metrics and techniques that have been developed recently.
+For comments on why these techniques could result in regulatory non-compliance in certain cases, see Appendix \ref{a_sec:comp}.
 
 Whether these methods are suitable for a particular use case depends on the legal environment in which the model will be used and the use case itself.
-Great care must be taken in order to ensure that the appropriate fairness metrics are chosen, because certain metrics may not be appropriate for some use cases.
-Additionally, the effects of changing the model must be viewed holistically.
-For example, the mortgage data disparity analysis in Section \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
-
-Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
-A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{In the pay discrimination case, \textit{Bazemore v.
-Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will have the same model output (though there may be some small or random differences in outcomes due to computational issues).
-Obviously, the inclusion of protected class status will change model output by class.
-With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.\footnote{In a reverse discrimination case, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.}  Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
-
 Of the newer class of fairness enhancing interventions in ML, within-algorithm discrimination mitigation techniques that do not use class information may be more likely to be acceptable in highly regulated settings today.
@@
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{26}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems.
-Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena.
-Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{26}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
 However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination.
@@
 While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work (and many, many others) illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+\appendixtitles{yes} 
+\appendix
+
+\section{HMDA Mortgage Data Details}\label{a_sec:mort}
+
+The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
+This information is first provided to the Consumer Financial Protection Bureau (CFPB), which subsequently releases some of the data to the public.
+Regulators often use HMDA data to, ``...show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory.''\textsuperscript{5} In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
+The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, loan to value (LTV) ratio, debt to income (DTI) ratio, and other important financial descriptors.
+The data also include information on each borrower and co-borrower’s race, ethnicity, gender, and age.
+Because the data includes information on these protected class characteristics, certain metrics that can be indicative of discrimination in lending can be calculated directly using the HDMA data.
+Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+
+\section{Selected Algorithmic Details}\label{a_sec:algo}
+
+\subsection{Notation}\label{a_ssec:not}
+
+To facilitate descriptions of data and modeling, explanatory, and discrimination testing techniques, notation for input and output spaces, datasets, and models is defined.
+
+\subsubsection{Spaces}
+
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+	\item Input features come from the set $\mathcal{X}$ contained in a \textit{P}-dimensional input space, $\mathcal{X} \subset \mathbb{R}^P$.
+An arbitrary, potentially unobserved, or future instance of $\mathcal{X}$ is denoted $\mathbf{x}$, $\mathbf{x} \in \mathcal{X}$.
+	\item Labels corresponding to instances of $\mathcal{X}$ come from the set $\mathcal{Y}$.
+	\item Learned output responses of models are contained in the set $\mathcal{\hat{Y}}$.
+% For regression models, the set $\mathcal{\hat{Y}}_r$ is also contained in a $C$-dimensional output space, $\mathcal{\hat{Y}}_r \subset \mathbb{R}^{C_r}$.
+For classification models, the set $\mathcal{\hat{Y}}_c$ typically contains a column vector for each unique class in $\mathcal{Y}$.
+Hence, $\mathcal{\hat{Y}}_c$ is contained in a $C'$-dimensional output space,  $\mathcal{\hat{Y}}_c \subset \mathbb{R}^{C'_c}$.
+\end{itemize}
+
+\subsubsection{Datasets}
+
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+	\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
+	\item Each $i$-th observation of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
+	$[x_0^{(i)}, x_1^{(i)}, \dots, x_{\textit{P}-1}^{(i)}]$, with corresponding $i$-th labels in $\mathbf{Y}, \mathbf{y}^{(i)}$, and corresponding predictions in $\mathbf{\hat{Y}}, \mathbf{\hat{y}}^{(i)}$.
+% = [y_0^{(i)}, y_1^{(i)}, \dots, y_{\textit{C}-1}^{(i)}]$.
+	\item $\mathbf{X}$ and $\mathbf{Y}$ consist of $N$ tuples of observations: $[(\mathbf{x}^{(0)},\mathbf{y}^{(0)}), (\mathbf{x}^{(1)},\mathbf{y}^{(1)}), \dots,(\mathbf{x}^{(N-1)},\mathbf{y}^{(N-1)})]$.
+%\\ $\mathbf{x}^{(i)} \in \mathcal{X}$, $\mathbf{y}^{(i)} \in \mathcal{Y}$.
+	\item Each $j$-th input column vector of $\mathbf{X}$ is denoted as $X_j = [x_{j}^{(0)}, x_{j}^{(1)}, \dots, x_{j}^{(N-1)}]^T$.
+\end{itemize}
+
+\subsubsection{Models}
+
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]
+	\item A type of ML model $g$, selected from a hypothesis set $\mathcal{H}$, is trained to represent an unknown signal-generating function $f$ observed as  $\mathbf{X}$ with labels $\mathbf{Y}$ using a training algorithm $\mathcal{A}$:
+	$ \mathbf{X}, \mathbf{Y} \xrightarrow{\mathcal{A}} g$, such that $g \approx f$.
+	\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
+	\item The model to be explained and tested for discrimination is denoted as $g$.
+\end{itemize}
+
+\subsection{Monotonic Gradient Boosting Machine}\label{a_ssec:mgbm}
+
+MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints, with respect to each $X_j$ and $\mathbf{y}$ independently.
+The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
+
+\begin{equation}
+\begin{aligned}\label{eq:gbm}
+g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
+\end{aligned}
+\end{equation}
+
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
+
+\begin{equation}
+\begin{aligned}
+\mathcal{L}_{b} =\sum_{i=0}^{N-1}l(y^{(i)}, g^{\text{MGBM}}_{b-1}(\mathbf{x}^{(i)}), T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)) + \Omega(T_b)\\
+\end{aligned}
+\end{equation}
+
+\noindent In addition to $\mathcal{L}_{b}$, $g^\text{MGBM}$ training is characterized by additional splitting rules and constraints on tree node weights.
+Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$.
+For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction.
+For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, $g^{\text{MGBM}}$ is positive monotonic with respect to some $X_j$ if $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
+The following rules and constraints ensure positive monotonicity in $\Theta_b$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$.
+
+\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
+	\item For the first and highest split in $T_b$ involving $X_j$, any $\theta_{b,j,0}$ resulting in $T(x_j; \theta_{b,j,0}) = \{w_{b,j,0,L}, w_{b,j,0,R}\}$ where $w_{b,j,0,L} > w_{b,j,0,R}$, is not considered.
+	\item For any subsequent left child node involving $X_j$, any $\theta_{b,j, k\ge1}$ resulting in $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ where $w_{b,j,k\ge1,L} > w_{b,j,k\ge1,R}$, is not considered.
+	\item Moreover, for any subsequent left child node involving $X_j$, $T(x_j; \theta_{b,j,k\ge1}) = \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$, $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\}$ are bound by the associated $\theta_{b,j,k-1}$ set of node weights, $\{w_{b,j,k-1,L}, w_{b,j,k-1, R}\}$, such that $ \{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} < \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+	\item (1) and (2) are also applied to all right child nodes, except that for right child nodes $ w_{b,j,k,L} \le w_{b,j,k,R}$ and $\{w_{b,j,k\ge1,L}, w_{b,j,k\ge1,R}\} \ge \frac{w_{b,j,k-1,L} + w_{b,j,k-1,R}}{2}$.
+\end{enumerate}
+
+\noindent Note that for any one $X_j$ and $T_b \in g^{\text{MGBM}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
+Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0}) = w_{b,0}$.
+Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
+For a negative monotonic constraint, i.e., $g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
+Also consider that MGBM models with independent monotonicity constraints between some $X_j$ and $\mathbf{y}$ likely restrict non-monotonic interactions between multiple $X_j$.
+Moreover, if monotonicity constraints are not applied to all $X_j \in \mathbf{X}$, any strong non-monotonic signal in training data associated with some important $X_j$ maybe forced onto some other arbitrary unconstrained $X_j$ under some $g^{MGBM}$ models, compromising the end goal of interpretability.
+
+\subsection{Explainable Neural Network}\label{a_ssec:xnn}
+
+XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
+XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e., GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
+Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
+XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum_j\beta_{k,j}x_j$.
+
+\begin{equation}
+\begin{aligned}
+\label{eq:xnn}
+g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=0}^{K-1}\gamma_k n_k(\sum^{J=\mathcal{P}-1}_{j=0}\beta_{k, j}x_j)
+\end{aligned}
+\end{equation}
+
+\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:
+
+\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
+	\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, which should learn higher magnitude weights for important $X_j$, is known as the \textit{projection layer}.
+It is fully connected to each input $X_j$.
+Each hidden unit in the projection layer may optionally include a bias term.
+	\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
+Each $n_k$ is a neural network, which can be parameterized to suit a given modeling task.
+To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
+Each $n_k$ can contain several bias terms.
+	\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
+Again, subnetwork output is restricted to 1-dimension for interpretation and visualization purposes.
+\end{enumerate}
+
+\subsection{One-dimensional Partial Dependence and Individual Conditional Expectation}\label{a_ssec:pd_ice}
+
+Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$) is considered.
+$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
+$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
+PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
+Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
+PD should be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.
+
+\subsection{Shapley Values}\label{a_ssec:shap}
+
+For some instance $\mathbf{x} \in \mathcal{X}$, Shapley explanations take the form:
+
+\begin{equation}
+\label{eq:shap_additive}
+\begin{aligned}
+g(\mathbf{x}) = \phi_0 + \sum_{j=0}^{j=\mathcal{P} - 1} \phi_j \mathbf{z}_j
+\end{aligned}
+\end{equation}
+
+\noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
+Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
+Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
+
+\begin{equation}
+\label{eq:shap_contrib}
+\begin{aligned}
+\phi_{j} = \sum_{S \subseteq \mathcal{P} \setminus \{j\}}\frac{|S|!(\mathcal{P} -|S| -1)!}{\mathcal{P}!}[g_x(S \cup \{j\}) - g_x(S)]
+\end{aligned}
+\end{equation}
+
+\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
+Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some set $\mathbf{X}$.
+
+\section{Machine Learning Discrimination Under US Law}\label{a_sec:types}
+
+Before discussing the techniques, it is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI).
+DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
+With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI, colloquially known as \textit{unintentional discrimination}, occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
+In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.
+However, other factors, such as hyperparameter or algorithm choices, can drive DI.
+Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.
+The analyses and metrics herein focus on several measures of disparate impact that are commonly used in US litigation and regulatory settings.
+
+\section{Practical vs. Statistical Significance for Discrimination Metrics}\label{a_sec:prac_v_stat}
+
+A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.
+Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
+It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model.
+Moreover, practical significance is also needed by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production.
+Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
+One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.
+
+The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.
+The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v.
+Partida}.\footnote{\textit{Castaneda v.
+Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v.
+United States.}\footnote{\textit{Hazelwood School Dist.
+v.
+United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
+
+\section{Cutoff Selection w.r.t Mitigating Discrimination}\label{a_sec:cut}
+
+The selection of which cutoff to use in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.
+For example, a model developer at a bank might build a credit model where the F1 score is maximized at a delinquency probability cutoff of 0.15.
+For purposes of evaluating the quality of the model, she may review confusion matrix statistics (accuracy, recall, precision, etc.) using cutoffs based on the maximum F1 score.
+But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
+Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.
+
+\section{Recent Fairness Techniques and Compliance}\label{a_sec:comp}
+
+Great care must be taken in order to ensure that the appropriate fairness metrics are chosen, because certain metrics may not be appropriate for some use cases.
+Additionally, the effects of changing the model must be viewed holistically.
+For example, the mortgage data disparity analysis in Section \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
+
+Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
+A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{In the pay discrimination case, \textit{Bazemore v.
+Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will have the same model output (though there may be some small or random differences in outcomes due to computational issues).
+Obviously, the inclusion of protected class status will change model output by class.
+With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.\footnote{In a reverse discrimination case, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.}  Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \reftitle{References}

### 2020-01-29 article-information-2019.tex
simulated results to appendix, update intro

@@
 % Full title of the paper (Capitalized)
-\Title{Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
+\Title{\LARGE{An Example Responsible Machine Learning Workflow}\\\large{With Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
 
@@
 \abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data with known feature importance and discrimination characteristics and on publicly available mortgage data.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on publicly available mortgage data.
 For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
@@
 
+% for subtables
 \usepackage{caption}
+\usepackage{graphicx}
+\usepackage{subcaption}
 
@@
 
-Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds, i.e., ethics, social sciences, and technology.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is crystal clear: ML models, a common application of AI, have problems that responsible practitioners should likely attempt to remediate.
+Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear.
+ML models, a common application of AI, present risks that responsible practitioners should likely attempt to mitigate.
 ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
-While this manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been resolved (even as singular entities, much less as complex intersectional phenomena), Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form of interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools, to address a subset of these vexing problems for high-stakes, human-centered, or regulated ML applications.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
-
-Section \ref{sec:m_and_m} describes methods and materials, including simulated and collected training datasets, interpretable and constrained model architectures, post-hoc explanations used to create an appealable decision-making framework, tests for DI and other social discrimination, and public and open source software resources.
+This manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena.
+However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures to address a subset of these vexing problems: interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+
+Section \ref{sec:m_and_m} describes methods and materials, including collected training datasets, interpretable and constrained models, post-hoc explanations, tests for DI and other social discrimination, and public and open source software resources.
 In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and post-hoc explanation and discrimination testing results are also presented for interpretable models.
-Section \ref{sec:disc} then discusses some nuances of the outlined modeling, explanation, and discrimination testing methods and results.
+Of course, an even wider array of tools and techniques are likely necessary to fully minimize discrimination, inaccuracy, privacy, and security risks associated with ML models.
+Section \ref{sec:disc} addresses the burgeoning Python ecosystem for responsible AI, along with appeal and override of automated decisions, and discrimination testing and remediation in practice.
 Section \ref{sec:con} closes this manuscript with a brief summary of the outlined methods, materials, results, and discussion.
@@
 \begin{itemize}[leftmargin=*,labelsep=5.8mm]
-	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
-	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
+	\item \textbf{Training data}: collected mortgage data -- \textsection\ref{ssec:mort_data}
 	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
@@
 Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} and XNNs (\cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analysis.
+MGBM and XNN interpretable model architectures were selected for this text because they are straightforward variants of popular unconstrained ML models.
+If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
+The same can be said of the presented explanation methods and discrimination tests.
+Due to their post-hoc nature, they can often be shoe-horned into existing ML work flows and pipelines.
+While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML.
+Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simulated Data}\label{ssec:sim_data}
-
-Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
-
-\begin{equation}
-\begin{aligned}\label{eq:sim}\small
-f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
-\end{aligned}
-\end{equation}
-
-\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
-In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
-In this manuscript, the signal generating function and input features are modified in several ways.
-Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings.
-%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so.
-For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
-For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
-Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
-The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows.
-The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
-Within the training set, a 5 fold cross validation indicator is used for training all models.
-For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Mortgage Data}\label{ssec:mort_data}
@@
 Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretability, post-hoc explanation, and discrimination testing results are presented.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Simulated Data Results}
-
-Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
-For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
-For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
-Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
-Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
-
-\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_sim}
-
-Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-$g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
-
-\begin{table}[H]
-\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
-\centering
-\begin{tabular}{ccccc}
-\toprule
-\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
-\midrule
-$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
-$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
-$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
-$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
-\bottomrule
-\label{tab:sim_acc}
-\end{tabular}
-\end{table}
-
-\noindent Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
-Accuracy is measured at the best F1 threshold for each model.
-
-\subsubsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}
-
-Global mean absolute Shapley value feature importance for $g^{\text{MGBM}}(\mathbf{X})$ on the simulated test data is displayed in Figure \ref{fig:sim_mgbm_glob}.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the simulated test data in the $g^\text{MGBM}$ margin space.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
-\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the simulated test data.}
-\label{fig:sim_mgbm_glob}
-\end{figure}
-
-\noindent As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.
-
-Figure \ref{fig:sim_mgbm_glob_pdp_ice} highlights PD, ICE, and histograms of the most important features from Figure \ref{fig:sim_mgbm_glob}.
-PD gives an indication of the estimated average prediction of $g^\text{MGBM}(\mathbf{X})$ in the test data for some $X_j$.
-ICE serves several purposes.
-It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is also a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior.
-PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain.
-Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
-Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
-\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
-\label{fig:sim_mgbm_glob_pdp_ice}
-\end{figure}
-
-$X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE.
-Also, as the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
-Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is probably not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}.
-The one-dimensional monotonic constraints may hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data.
-This is an interesting result for responsible ML practitioners.
-In some noisy scenarios, monotonicity constraints can increase model interpretability without causing a drastic drop in model accuracy, even when known interactions and non-monotonic behavior exist in training data.
-
-Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
-These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
-Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=15cm]{img/sim_mgbm_loc.png}
-\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.}
-\label{fig:sim_mgbm_loc}
-\end{figure}
-
-\noindent The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result.
-For the lower prediction at the 10\textsuperscript{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative.
-At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive.
-In each case, large local contributions generally follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.
-
-Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique.
-Deep SHAP values are reported in the probability space, after the application of the logit link function.
-They are also calculated from the projection layer of $g^\text{XNN}$.
-Thus, the Deep SHAP values in Figure \ref{fig:sim_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the simulated test data.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=8cm]{img/sim_xnn_glob.png}
-\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
-\label{fig:sim_xnn_glob}
-\end{figure}
-
-\noindent Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features.
-The consistency between the feature rankings of $g^\text{MGBM}$ and $g^\text{XNN}$ is somewhat striking, given their different hypothesis families and architectures.
-Both $g^\text{MGBM}$ and $g^\text{XNN}$ rank $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ as the most important features, both place $X_{\text{Categorical},2}$ and $X_{\text{Categorical},3}$ above the $X_{\text{Binary},1}$ and $X_{\text{Binary},2}$ features, both rank $X_{\text{Binary},1}$ above $X_{\text{Binary},2}$, and both place the least importance on $X_{\text{Categorical},4}$ and $X_{\text{Categorical},0}$.
-
-Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
-\ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
-The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms.
-Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
-\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
-\label{fig:sim_xnn_glob_ridge}
-\end{figure}
-
-\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks.
-Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
-In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
-Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.
-Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
-
-Local Deep SHAP feature importance in Figure \ref{fig:sim_xnn_loc} supplements the global interpretability of $g^\text{XNN}$ displayed in Figures \ref{fig:sim_xnn_glob} and \ref{fig:sim_xnn_glob_ridge}.
-Local Deep SHAP values are extracted from the projection layer of $g^\text{XNN}$ and reported in the probability space.
-Deep SHAP values can be calculated for any arbitrary $g^\text{XNN}(\mathbf{x})$, allowing for detailed, local summarization of individual model predictions.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=15cm]{img/sim_xnn_loc.png}
-\caption{Deep SHAP values for three observations across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
-\label{fig:sim_xnn_loc}
-\end{figure}
-
-\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x}^{(i)})$ predictions.
-
-\subsubsection{Discrimination Testing Results}\label{ssec:dis_sim}
-
-Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to cutoff selection and discrimination testing.}
-
-Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities.
-Table \ref{tab:sim_disc} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
-1.06).
-However, in Table \ref{tab:sim_n_acc} the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
-0.401).
-This illustrates a danger in considering relative class-control metrics in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$.
-This is because FNR accuracy improves for both the protected group 1 and control group 1, but members of control group 1 benefit more than those in protected group 1.
-Of course, the choice of which model is truly fairer is a policy question.
-
-\begin{table}[H]
-	\captionsetup{width=7cm}
-	\hskip-20pt\parbox{.48\textwidth}{
-	\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
-	\centering
-	\begin{tabular}{ccccc}
-		\toprule
-		\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} & \textbf{Accuracy} & \textbf{FNR} \\
-		\midrule
-		Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} 	&  \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
-		\midrule			 
-		Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
-		\midrule
-		Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular}	&  \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
-		\midrule
-		Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\
-		\bottomrule
-		\label{tab:sim_n_acc}
-	\end{tabular}
-	}\hspace{10pt}
-	\parbox{.48\textwidth}{
-	\vspace{-13pt}
-	\caption{AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
-	\centering
-	\begin{tabular}{ccccccc}
-		\toprule
-		\textbf{Model}		&  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular}					& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD} &  \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}\\
-		\midrule
-		$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}		& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
-		\midrule 
-		$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}		& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
-		\bottomrule
-		\label{tab:sim_disc}
-	\end{tabular}
-	}
-\end{table}
-
-For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.
-Of note is that 12.0\% is not a meaningful difference without context.
-If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 12.0\% could represent an extremely large difference and would require remediation.
-But if they represent substantially different populations, then 12.0\% could represent a reasonable deviation from parity.
-As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 12.0\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population.
-Protected group 1’s AIR under $g^\text{XNN}$ is 0.727, below the EEOC 4/5ths rule threshold.
-It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
-Together these would indicate that there may be evidence of illegal DI.
-As with ME and other measures, the reasonableness of this disparity is not clear outside of context.
-However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.
-Some pertinent remediation strategies for discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
-
-\noindent SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
-(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.206 vs.
--0.274), but both are close to Cohen’s small effect threshold of -0.20.
-Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.
 
@@
 \centering
-\includegraphics[width=8cm]{img/mort_mgbm_glob.png}
+\includegraphics[width=9cm]{img/mort_mgbm_glob.png}
 \caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
@@
 \centering
-\includegraphics[width=15cm]{img/mort_mgbm_glob_pdp_ice.png}
+\includegraphics[width=16.5cm]{img/mort_mgbm_glob_pdp_ice.png}
 \caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
@@
 \centering
-\includegraphics[width=15cm]{img/mort_mgbm_loc.png}
+\includegraphics[width=16cm]{img/mort_mgbm_loc.png}
 \caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.}
@@
 \centering
-\includegraphics[width=8cm]{img/mort_xnn_glob.png}
+\includegraphics[width=9cm]{img/mort_xnn_glob.png}
 \caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
@@
 \begin{figure}[H]
-\centering
-\includegraphics[width=16cm]{img/mort_xnn_glob_ridge.png}
+\hskip-20pt\includegraphics[width=18cm]{img/mort_xnn_glob_ridge.png}
 \caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
@@
 \centering
-\includegraphics[width=15cm]{img/mort_xnn_loc.png}
+\includegraphics[width=16cm]{img/mort_xnn_loc.png}
 \caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
@@
 \begin{table}[H]
+	\hskip-10pt\begin{subtable}{0.45\textwidth}
 	\captionsetup{width=7cm}
-	\hskip-20pt\parbox{.48\textwidth}{
 	\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.} 
 	\centering
+	\resizebox{\columnwidth}{!}{%
 	\begin{tabular}{ccccc}
@@
 		\label{tab:mort_n_acc}
-		\end{tabular}
-	}\hspace{10pt}
-	\parbox{.48\textwidth}{
-		\vspace{-13pt}
-		\caption{AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
-		\centering
+	\end{tabular}%
+	}
+	\end{subtable}
+	\hspace{10pt}
+	\begin{subtable}{0.45\textwidth}
+	\vspace{-14pt}
+	\captionsetup{width=10cm}
+	\caption{%
+		\leftskip=45pt
+		AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+	\centering
+	\scalebox{0.95}{
 	\begin{tabular}{ccccccc}
@@
 	}
+	\end{subtable}
+\captionsetup{width=10cm}
+\caption{Discrimination measures for the mortgage test data.}
 \end{table}
@@
 
-MGBM and XNN interpretable model architectures were selected for this text because they are straightforward variants of popular unconstrained ML models.
-If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
-The same can be said of the presented explanation methods and discrimination tests.
-Due to their post-hoc nature, they can often be shoe-horned into existing ML work flows and pipelines.
-While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML.
-Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.
+Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, deployment meant to decrease discrimination, inaccuracy, privacy, and security risks for high-stakes, human-centered, or regulated ML applications.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.}
 
@@
 	\centering
-\includegraphics[width=8cm]{img/blueprint.png}
-\caption{A diagram of a proposed holistic ML work flow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, and other review and appeal mechanisms can create an understandable and trustworthy ML system.}
+	\includegraphics[width=11cm]{img/blueprint.png}
+	\caption{An example responsible ML work flow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among other review and appeal mechanisms, can create an understandable and trustworthy ML system.}
 	\label{fig:blueprint}
@@
 
+While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.
 Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers at Microsoft and Cornell.
@@
 
+\section{Simulated Data}\label{a_sec:sim_data}
+
+Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
+
+\begin{equation}
+\begin{aligned}\label{eq:sim}\small
+f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
+\end{aligned}
+\end{equation}
+
+\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
+In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
+In this manuscript, the signal generating function and input features are modified in several ways.
+Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings.
+%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so.
+For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
+For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
+Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
+The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows.
+The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
+Within the training set, a 5 fold cross validation indicator is used for training all models.
+For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
+
 \section{HMDA Mortgage Data Details}\label{a_sec:mort}
@@
 United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
+
+\section{Simulated Data Results}
+
+Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
+For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
+For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
+Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
+Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
+
+\subsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_sim}
+
+Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+$g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
+
+\begin{table}[H]
+	\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
+	\centering
+	\begin{tabular}{ccccc}
+		\toprule
+		\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
+		\midrule
+		$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
+		$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
+		$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
+		$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
+		\bottomrule
+		\label{tab:sim_acc}
+	\end{tabular}
+\end{table}
+
+\noindent Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
+$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
+Accuracy is measured at the best F1 threshold for each model.
+
+\subsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}
+
+Global mean absolute Shapley value feature importance for $g^{\text{MGBM}}(\mathbf{X})$ on the simulated test data is displayed in Figure \ref{fig:sim_mgbm_glob}.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the simulated test data in the $g^\text{MGBM}$ margin space.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
+	\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the simulated test data.}
+	\label{fig:sim_mgbm_glob}
+\end{figure}
+
+\noindent As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.
+
+Figure \ref{fig:sim_mgbm_glob_pdp_ice} highlights PD, ICE, and histograms of the most important features from Figure \ref{fig:sim_mgbm_glob}.
+PD gives an indication of the estimated average prediction of $g^\text{MGBM}(\mathbf{X})$ in the test data for some $X_j$.
+ICE serves several purposes.
+It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is also a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior.
+PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain.
+Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
+Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
+	\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
+	\label{fig:sim_mgbm_glob_pdp_ice}
+\end{figure}
+
+$X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE.
+Also, as the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
+Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is probably not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}.
+The one-dimensional monotonic constraints may hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data.
+This is an interesting result for responsible ML practitioners.
+In some noisy scenarios, monotonicity constraints can increase model interpretability without causing a drastic drop in model accuracy, even when known interactions and non-monotonic behavior exist in training data.
+
+Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
+These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
+Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm]{img/sim_mgbm_loc.png}
+	\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.}
+	\label{fig:sim_mgbm_loc}
+\end{figure}
+
+\noindent The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result.
+For the lower prediction at the 10\textsuperscript{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative.
+At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive.
+In each case, large local contributions generally follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.
+
+Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique.
+Deep SHAP values are reported in the probability space, after the application of the logit link function.
+They are also calculated from the projection layer of $g^\text{XNN}$.
+Thus, the Deep SHAP values in Figure \ref{fig:sim_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the simulated test data.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=8cm]{img/sim_xnn_glob.png}
+	\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
+	\label{fig:sim_xnn_glob}
+\end{figure}
+
+\noindent Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features.
+The consistency between the feature rankings of $g^\text{MGBM}$ and $g^\text{XNN}$ is somewhat striking, given their different hypothesis families and architectures.
+Both $g^\text{MGBM}$ and $g^\text{XNN}$ rank $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ as the most important features, both place $X_{\text{Categorical},2}$ and $X_{\text{Categorical},3}$ above the $X_{\text{Binary},1}$ and $X_{\text{Binary},2}$ features, both rank $X_{\text{Binary},1}$ above $X_{\text{Binary},2}$, and both place the least importance on $X_{\text{Categorical},4}$ and $X_{\text{Categorical},0}$.
+
+Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
+\ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
+The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms.
+Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
+	\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
+	\label{fig:sim_xnn_glob_ridge}
+\end{figure}
+
+\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks.
+Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
+In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
+Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.
+Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
+
+Local Deep SHAP feature importance in Figure \ref{fig:sim_xnn_loc} supplements the global interpretability of $g^\text{XNN}$ displayed in Figures \ref{fig:sim_xnn_glob} and \ref{fig:sim_xnn_glob_ridge}.
+Local Deep SHAP values are extracted from the projection layer of $g^\text{XNN}$ and reported in the probability space.
+Deep SHAP values can be calculated for any arbitrary $g^\text{XNN}(\mathbf{x})$, allowing for detailed, local summarization of individual model predictions.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=15cm]{img/sim_xnn_loc.png}
+	\caption{Deep SHAP values for three observations across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
+	\label{fig:sim_xnn_loc}
+\end{figure}
+
+\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x}^{(i)})$ predictions.
+
+\subsection{Discrimination Testing Results}\label{ssec:dis_sim}
+
+Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to cutoff selection and discrimination testing.}
+
+Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities.
+Table \ref{tab:sim_disc} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
+1.06).
+However, in Table \ref{tab:sim_n_acc} the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
+0.401).
+This illustrates a danger in considering relative class-control metrics in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$.
+This is because FNR accuracy improves for both the protected group 1 and control group 1, but members of control group 1 benefit more than those in protected group 1.
+Of course, the choice of which model is truly fairer is a policy question.
+
+\begin{table}[H]
+	\hskip-10pt
+	\begin{subtable}{0.45\textwidth}
+		\captionsetup{width=7cm}
+		\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
+		\centering
+		\resizebox{\columnwidth}{!}{%
+			\begin{tabular}{ccccc}
+				\toprule
+				\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy} & \textbf{FNR} \\
+				\midrule
+				Protected 1	& 3,057 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} & \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
+				\midrule			 
+				Control 1 & 16,943 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular} & \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
+				\midrule
+				Protected 2	& 9,916 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular} & \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
+				\midrule
+				Control 2 & 10,084 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular} & \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\
+				\bottomrule
+				\label{tab:sim_n_acc}
+			\end{tabular}%
+		}
+	\end{subtable}
+	\hspace{5pt}
+	\begin{subtable}{0.45\textwidth}
+		\vspace{-11pt}
+		\captionsetup{width=10cm}
+		\caption{%
+			\leftskip=45pt
+			AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
+		\centering
+		\scalebox{0.95}{
+			\begin{tabular}{ccccccc}
+				\toprule
+				\textbf{Model} & \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}	& \textbf{ME} & \textbf{SMD} & \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}\\
+				\midrule
+				$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular} & \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
+				\midrule 
+				$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular} & \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
+				\bottomrule
+				\label{tab:sim_disc}
+			\end{tabular}
+		}
+	\end{subtable}
+	\captionsetup{width=10cm}
+	\caption{Discrimination measures for the simulated test data.}
+\end{table}
+
+For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.
+Of note is that 12.0\% is not a meaningful difference without context.
+If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 12.0\% could represent an extremely large difference and would require remediation.
+But if they represent substantially different populations, then 12.0\% could represent a reasonable deviation from parity.
+As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 12.0\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population.
+Protected group 1’s AIR under $g^\text{XNN}$ is 0.727, below the EEOC 4/5ths rule threshold.
+It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
+Together these would indicate that there may be evidence of illegal DI.
+As with ME and other measures, the reasonableness of this disparity is not clear outside of context.
+However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.
+Some pertinent remediation strategies for discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
+
+\noindent SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
+(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.206 vs.
+-0.274), but both are close to Cohen’s small effect threshold of -0.20.
+Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.
 

### 2020-01-29 article-information-2019.tex
good rework of intro section

@@
 % Full title of the paper (Capitalized)
-\Title{\LARGE{An Example Responsible Machine Learning Workflow}\\\large{With Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
+\Title{A Responsible Machine Learning Workflow\\\normalsize{With Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
 
@@
 \abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on publicly available mortgage data.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data and publicly available mortgage data.
 For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents: adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and additional straightforward group fairness measures.
-By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to present a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and also with straightforward group fairness measures.
+By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
 
@@
 Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear.
-ML models, a common application of AI, present risks that responsible practitioners should likely attempt to mitigate.
+ML models, a common application of AI, can present serious risks.
 ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
 ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
-This manuscript makes no claim that the interdependent issues of opaqueness, discrimination, or security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures to address a subset of these vexing problems: interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available, free, and open source Python tools.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
-
-Section \ref{sec:m_and_m} describes methods and materials, including collected training datasets, interpretable and constrained models, post-hoc explanations, tests for DI and other social discrimination, and public and open source software resources.
+This manuscript makes no claim that these interdependent issues of opaqueness, discrimination, privacy harms, and security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena.
+However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+
+Section \ref{sec:m_and_m} describes methods and materials, including collected training datasets, interpretable and constrained models, post-hoc explanations, tests for DI and other social discrimination, and public and open source software resources associated with this text.
 In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and post-hoc explanation and discrimination testing results are also presented for interpretable models.
-Of course, an even wider array of tools and techniques are likely necessary to fully minimize discrimination, inaccuracy, privacy, and security risks associated with ML models.
-Section \ref{sec:disc} addresses the burgeoning Python ecosystem for responsible AI, along with appeal and override of automated decisions, and discrimination testing and remediation in practice.
+Of course, an even wider array of tools and techniques are likely helpful to fully minimize discrimination, inaccuracy, privacy, and security risks associated with ML models.
+Section \ref{sec:disc} puts forward a more holistic responsible ML modeling workflow, and addresses the burgeoning Python ecosystem for responsible AI, along with appeal and override of automated decisions, and discrimination testing and remediation in practice.
 Section \ref{sec:con} closes this manuscript with a brief summary of the outlined methods, materials, results, and discussion.
@@
 
-Detailed descriptions of notation, training data, ML models, post-hoc explanation techniques, discrimination testing methods, and software resources are organized in Section \ref{sec:m_and_m} as follows:
-
-\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-	\item \textbf{Training data}: collected mortgage data -- \textsection\ref{ssec:mort_data}
-	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
-	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
-	\item \textbf{Discrimination testing methods}: AIR, ME, SMD and confusion matrix metrics -- \textsection\ref{ssec:di}
-	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
-\end{itemize}
-
-To provide a sense of accuracy differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
-The simulated data, based on the well-known Friedman datasets and with known feature importance and discrimination characteristics, is used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}.
-The mortgage data is sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} Because unconstrained ML models, like gradient boosting machines (GBMs) (e.g.
+The simulated data (see Subsection \ref{ssec:sim_data}) is based on the well-known Friedman datasets.
+Its known feature importance and discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}.
+The mortgage data (see Subsection \ref{ssec:mort_data}) is sourced from the Home Mortgage Disclosure Act (HMDA) database and is a fairly realistic data source for demonstrating the template workflow.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} To provide a sense of accuracy differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
+Because the unconstrained ML models, gradient boosting machines (GBMs) (e.g.
 \cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
-\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), can be difficult to understand, trust, and appeal, even after the application of post-hoc explanation techniques, explanation analysis and discrimination testing are applied only to the constrained interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
-Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} and XNNs (\cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analysis.
-MGBM and XNN interpretable model architectures were selected for this text because they are straightforward variants of popular unconstrained ML models.
-If practitioners are working with GBM and ANN models, it should be relatively uncomplicated for them to evaluate the constrained versions of these models.
-The same can be said of the presented explanation methods and discrimination tests.
-Due to their post-hoc nature, they can often be shoe-horned into existing ML work flows and pipelines.
-While these approaches are promising responses to the black-box and discrimination problems in ML, they are just a small part of a burgeoning ecosystem of research and Python tools for responsible ML.
-Figure \ref{fig:blueprint} is a work flow blueprint that illustrates some of the additional steps that may be required to build a fully understandable and trustworthy ML system.
-
-Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a template workflow for future users of similar methods and tools.
-Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}.
+\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present previously delineated risks, further explanation and discrimination analysis are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
+Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} (See subsection \ref{ssec:mgbm}) and XNNs (see Subsection \ref{ssec:xnn}, \cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
+MGBM and XNN interpretable model architectures are selected for this text because they are straightforward variants of popular unconstrained ML models.
+If practitioners are working with GBM and ANN models, it should be relatively uncomplicated to also evaluate the constrained versions of these models.
+
+The same can be said of the selected explanation methods and discrimination tests.
+Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines.
+Presented explanation techniques include PD and ICE (see Subsection \ref{ssec:pd_ice}) and Shapley values (see Subsection \ref{ssec:shap}) \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}.
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
-All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.
+Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
+All outlined materials and methods are implemented in open source Python code, and are made available through GitHub (see Subsection \ref{ssec:soft}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Simulated Data}\label{ssec:sim_data}
+
+Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
+
+\begin{equation}
+\begin{aligned}\label{eq:sim}\small
+f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
+\end{aligned}
+\end{equation}
+
+\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
+In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
+In this manuscript, the signal generating function and input features are modified in several ways.
+Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings.
+%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so.
+For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
+For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
+Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
+The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows.
+The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
+Within the training set, a 5 fold cross validation indicator is used for training all models.
+For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
 
@@
 	\includegraphics[width=11cm]{img/blueprint.png}
-	\caption{An example responsible ML work flow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among other review and appeal mechanisms, can create an understandable and trustworthy ML system.}
+	\caption{An example responsible ML workflow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among other review and appeal mechanisms, can create an understandable and trustworthy ML system.}
 	\label{fig:blueprint}
@@
 This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
-These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible ML work flow for high-stakes, human-centered, or regulated ML applications.
+These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible ML workflow for high-stakes, human-centered, or regulated ML applications.
 The discussed methodologies are solid steps toward interpretability, explanation, and minimal discrimination for ML decisions, which should ultimately enable increased fairness and logical appeal processes for ML decision subjects.
@@
 
-\section{Simulated Data}\label{a_sec:sim_data}
-
-Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
-
-\begin{equation}
-\begin{aligned}\label{eq:sim}\small
-f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
-\end{aligned}
-\end{equation}
-
-\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
-In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
-In this manuscript, the signal generating function and input features are modified in several ways.
-Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings.
-%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so.
-For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
-For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
-Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
-The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows.
-The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
-Within the training set, a 5 fold cross validation indicator is used for training all models.
-For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
-
 \section{HMDA Mortgage Data Details}\label{a_sec:mort}

### 2020-01-31 article-information-2019.tex
improvements to XNN section

@@
 \cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
-\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present previously delineated risks, further explanation and discrimination analysis are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
+\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present previously delineated risks, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
 Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} (See subsection \ref{ssec:mgbm}) and XNNs (see Subsection \ref{ssec:xnn}, \cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
-MGBM and XNN interpretable model architectures are selected for this text because they are straightforward variants of popular unconstrained ML models.
+MGBM and XNN interpretable model architectures are selected for the example workflow because they are straightforward variants of popular unconstrained ML models.
 If practitioners are working with GBM and ANN models, it should be relatively uncomplicated to also evaluate the constrained versions of these models.
@@
 Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
-All outlined materials and methods are implemented in open source Python code, and are made available through GitHub (see Subsection \ref{ssec:soft}).
+All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).
 
@@
 
-Simulated data is created based on a function first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
+Simulated data is created based on a signal-generating function, $f$, first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
 
@@
 The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
-Within the training set, a 5 fold cross validation indicator is used for training all models.
+Within the training set, a 5-fold cross-validation indicator is used for training all models.
 For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
@@
 After data cleaning and preprocessing to encode categorical features and create missing markers, the mortgage data contains ten input features and the binary outcome, \textit{high-priced}.
-The data is split into a training set with 160,338 loans and a marker for 5 fold cross validation and a test set containing 39,662 loans.
+The data is split into a training set with 160,338 loans and a marker for 5-fold cross-validation and a test set containing 39,662 loans.
 While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (LTV ratio, DTI ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
@@
 
-MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints, with respect to each $X_j$ and $\mathbf{y}$ independently.
-The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
+MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints, with respect to each input feature, $X_j$, and a target feature, $\mathbf{y}$, independently.
+An MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
+For some row of data, $\mathbf{x}$, a trained MGBM model, $g^{\text{MGBM}}$, takes the form:
 
@@
 \begin{aligned}\label{eq:gbm}
-g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
+g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=0}^{B-1} T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
 \end{aligned}
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), and a regularization term that penalizes complexity in the current tree.
+See Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
 
@@
 One on the simulated data and one on the mortgage data.
-In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and five-fold cross validation and test partitions are used for model assessment.
-See Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
+In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and 5-fold cross-validation and test partitions are used for model assessment.
 For exact parameterization of the two $g^\text{MGBM}$ models, see the software resources referenced in Subsection \ref{ssec:soft}.
@@
 Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
-XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum_j\beta_{k,j}x_j$.
+A trained XNN function, $g^{\text{XNN}}$, applied to some row of data, $\mathbf{x}$, is defined as:
 
@@
 \label{eq:xnn}
-g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=0}^{K-1}\gamma_k n_k(\sum^{J=\mathcal{P}-1}_{j=0}\beta_{k, j}x_j)
+g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=0}^{K-1}\gamma_k n_k(\sum^{J-1}_{j=0}\beta_{k, j}x_j)
 \end{aligned}
@@
 
-Here, each $g^\text{XNN}$ is trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
-Each $g^\text{XNN}$ is assessed in five training folds and in a test data partition.
-$L_1$ regularization is applied to both the projection and combination layers to induce a sparse and interpretable model, where each $n_k$ subnetwork and corresponding combination layer $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
+\noindent where $\mu_0$ is a global bias for $K$ individually specified ANN subnetworks, $n_k$, with scale parameters $\gamma_k$.
+The inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest projection layer.
+
+Here, the two $g^\text{XNN}$ models are trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
+Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition.
+$L_1$ regularization is applied to both the lowest projection layer and to the $n_k$ subnetworks to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
 The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size.
@@
 
-PD plots are a widely-used method for describing and plotting the average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
+PD plots are a widely-used method for describing and plotting the estimated average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
 ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$ \cite{ice_plots}.
@@
 
-\subsection{Intersectional and Non-static Problems in Machine Learning}
+\subsection{Intersectional and Non-static Risks in Machine Learning}
 
@@
 
-MGBMs constrain typical GBM training to consider only tree splits that obey user-defined positive and negative monotonicity constraints, with respect to each $X_j$ and $\mathbf{y}$ independently.
-The MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
-
-\begin{equation}
-\begin{aligned}\label{eq:gbm}
-g^{\text{MGBM}}(\mathbf{x}) &= \sum_{b=1}^B T_b\left(\mathbf{x}; \Theta^\text{mono}_b\right)
-\end{aligned}
-\end{equation}
-
 \noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
@@
 
-XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
-XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e., GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
-Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
-XNNs are composed of a global bias term, $\mu_0$, $K$ individually specified neural networks, $n_k$ with scale parameters $\gamma_k$, and the inputs to each $n_k$ are themselves a linear combination of modeling inputs, $\sum_j\beta_{k,j}x_j$.
-
-\begin{equation}
-\begin{aligned}
-\label{eq:xnn}
-g^{\text{XNN}}(\mathbf{x}) = \mu_0 + \sum_{k=0}^{K-1}\gamma_k n_k(\sum^{J=\mathcal{P}-1}_{j=0}\beta_{k, j}x_j)
-\end{aligned}
-\end{equation}
+% XNN image
 

### 2020-01-31 article-information-2019.tex
131am

@@
 
-Here, the two $g^\text{XNN}$ models are trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
+Two $g^\text{XNN}$ models are trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
 Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition.
@@
 
-PD plots are a widely-used method for describing and plotting the estimated average predictions of a complex model $g$ across some partition of data $\mathbf{X}$ for some interesting input feature $X_j$ \cite{esl}.
-ICE plots are a newer method that describes the local behavior of $g$ for a single instance $\mathbf{x} \in \mathcal{X}$ \cite{ice_plots}.
-PD and ICE can be overlaid in the same plot to compensate for known weaknesses of PD (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}), to identify interactions modeled by $g$, and to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
+PD plots are a widely-used method for describing and plotting the estimated average predictions of a complex model $g$ across some partition of data, $\mathbf{X}$, for some interesting input feature, $X_j$ \cite{esl}.
+ICE plots are a newer method that describes the local behavior of $g$ with regard to a single instance of an input feature, $x_j$, and for it's corresponding single instance $\mathbf{x}$.
+PD and ICE can be overlaid in the same plot to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
+Combining PD and ICE can also be indicative of modeled interactions in $g$ or of known weaknesses in PD estimation (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}) when $\text{PD}(X_j, g)$ and $\text{ICE}(x_j, \mathbf{x}, g)$ diverge.
 For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Subsection \ref{ssec:soft} and Appendices \ref{a_ssec:not} and \ref{a_ssec:pd_ice}.
@@
 
-Shapley values can be estimated in different ways, many of which are intractable for datasets with large $\mathcal{P}$.
+Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features.
 Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
@@
 
-% XNN image
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=8cm]{img/xnn_arch.png}
+	\caption{A diagram of an XNN with three metalayers: the bottom combination layer with  $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, the middle metalayer with with $K$ hidden and separate $n_k$ ridge functions, and the output combination layer that generates $g^{\text{XNN}}$ predictions.
+Figure adapted from \citet{wf_xnn}.}
+	\label{fig:xnn_arch}
+\end{figure}
 

### 2020-01-31 article-information-2019.tex
first pass on XNN sim results

@@
 Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
-The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows.
+The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 instances.
 The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
@@
 An MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
-For some row of data, $\mathbf{x}$, a trained MGBM model, $g^{\text{MGBM}}$, takes the form:
+For some instance, $\mathbf{x}$, a trained MGBM model, $g^{\text{MGBM}}$, takes the form:
 
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), and a regularization term that penalizes complexity in the current tree.
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric loss function (e.g., squared loss, Huber loss), and a regularization term that penalizes complexity in the current tree.
 See Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
@@
 Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
-A trained XNN function, $g^{\text{XNN}}$, applied to some row of data, $\mathbf{x}$, is defined as:
+A trained XNN function, $g^{\text{XNN}}$, applied to some instance , $\mathbf{x}$, is defined as:
 
@@
 Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition.
-$L_1$ regularization is applied to both the lowest projection layer and to the $n_k$ subnetworks to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
+$L_1$ regularization is applied to all network weights to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
 The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size.
@@
 
-PD plots are a widely-used method for describing and plotting the estimated average predictions of a complex model $g$ across some partition of data, $\mathbf{X}$, for some interesting input feature, $X_j$ \cite{esl}.
-ICE plots are a newer method that describes the local behavior of $g$ with regard to a single instance of an input feature, $x_j$, and for it's corresponding single instance $\mathbf{x}$.
+PD plots are a widely-used method for describing and plotting the estimated average prediction of a complex model, $g$, across some partition of data, $\mathbf{X}$, for some interesting input feature, $X_j$ \cite{esl}.
+ICE plots are a newer method that describes the local behavior of $g$ with regard to values of an input feature, $x_j$, in a single instance, $\mathbf{x}$.
 PD and ICE can be overlaid in the same plot to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
-Combining PD and ICE can also be indicative of modeled interactions in $g$ or of known weaknesses in PD estimation (e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}) when $\text{PD}(X_j, g)$ and $\text{ICE}(x_j, \mathbf{x}, g)$ diverge.
+When $\text{PD}(X_j, g)$ and $\text{ICE}(x_j,  g)$ curves diverge, such plots can also be indicative of modeled interactions in $g$ or expose flaws in PD estimation, e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}.
 For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Subsection \ref{ssec:soft} and Appendices \ref{a_ssec:not} and \ref{a_ssec:pd_ice}.
@@
 Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}, \cite{shapley1988shapley}.
-Shapley explanations are the only possible locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for some $\mathbf{x} \in \mathcal{X}$ and that Shapley explanation values should never decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
-
+Shapley explanations are the only locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for any $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
 Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features.
-Tree SHAP is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
-Tree SHAP (SHapley Additive exPlanations) has been implemented efficiently in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+Tree SHAP (SHapley Additive exPlanations) is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
+Tree SHAP has been implemented in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section \ref{sec:res} and Appendix \ref{a_sec:res_sim}.
 Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}.
-Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section \ref{sec:res} and Appendix \ref{a_sec:res_sim}.
 For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\subsection{Discrimination Testing Metrics}\label{ssec:di}
-
-Because many technical and academic discussions of fairness in ML have been inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in industries like Financial Services.
-The discussed metrics are also representative of fair lending analyses and pair well with the mortgage data.
-See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination under US laws and Appendix \ref {a_sec:prac_v_stat} for remarks on statistical vs.
-practical significance for discrimination measure values.
-
-One measure is known as marginal effects (ME).
+\subsection{Discrimination Testing Measures}\label{ssec:di}
+
+Because many technical and academic discussions of fairness in ML have been inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
+The discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
+See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination under US laws and Appendix \ref{a_sec:prac_v_stat} for remarks on statistical vs.
+practical significance for discrimination measures.
+One common measure of disparate impact used in US litigation and regulatory settings is known as marginal effects (ME).
 ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.
@@
 
-\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), and $p$ indicates a protected group.
-ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is adverse impact ratio (AIR), more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
+\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
+ME is a favored DI metric used by the the US Consumer Finance Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is adverse impact ratio (AIR), more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
 
@@
 \noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the proportion of the control class that receives a favorable outcome.
-Another long-standing measure of DI is standardized mean difference (SMD).
+Statistically significant AIR values below 0.8 can be considered \textit{prima facia} evidence of discrimination.
+An additional long-standing measure of DI is standardized mean difference (SMD).
 SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
-It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}.
-SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
+It originates from work on statistical power, and is more formally known as \textit{Cohen’s d}.
+SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
 
@@
 Because of this standardization factor, SMD allows for a comparison across different types of outcomes, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
-In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.
-Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.
+In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail in Appendix \ref{a_sec:prac_v_stat}.
+Finally, confusion matrix measures in demographic groups, such as accuracy, false negative rate (FNR), and their ratios, are also considered as measures of DI in Section \ref{sec:res} and Appendix \ref{a_sec:res_sim}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Simulated Data Results}
+
+As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on debunking the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN results.
+The bulk of the post-hoc explanation and discrimination testing results for the simulated are left to Appendix \ref{a_sec:res_sim}.
+
+\subsubsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec:c_v_uc_sim}
+
+Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
+$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
+Accuracy is measured at the best F1 threshold for each model.
+Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+$g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
+Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
+$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
+Accuracy is measured at the best F1 threshold for each model.
+
+\begin{table}[H]
+	\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
+	\centering
+	\begin{tabular}{ccccc}
+		\toprule
+		\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
+		\midrule
+		$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
+		$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
+		$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
+		$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
+		\bottomrule
+		\label{tab:sim_acc}
+	\end{tabular}
+\end{table}
+
+% add error metrics
+
+\subsubsection{$g^\text{XNN}$ Interpretability Results}
+
+For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
+Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
+\ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
+The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms.
+Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
+
+\begin{figure}[H]
+	\hskip-20pt\includegraphics[width=18cm]{img/sim_xnn_glob_ridge.png}
+	\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
+	\label{fig:sim_xnn_glob_ridge}
+\end{figure}
+
+% update figure
+
+\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks.
+Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
+In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
+Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.
+Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Mortgage Data Results}
 
-Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort}.
+Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort} to showcase the example workflow.
 $g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants.
@@
 $\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
-$\text{ICE}(x_j, \mathbf{x}, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
+$\text{ICE}(x_j, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
 PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
@@
 \noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
-Each $\phi_j$ is the local feature contribution value associated with $x_j$ and $\phi_0$ is the average of $g(\mathbf{X})$.
+Each $\phi_j$ is the local feature contribution value associated with $x_j$, and $\phi_0$ is the average of $g(\mathbf{X})$.
 Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
@@
 
-\section{Machine Learning Discrimination Under US Law}\label{a_sec:types}
-
-Before discussing the techniques, it is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI).
+\section{Machine Learning and Discrimination in US Legal and Regulatory Settings}\label{a_sec:types}
+
+It is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI).
 DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
@@
 Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.
-The analyses and metrics herein focus on several measures of disparate impact that are commonly used in US litigation and regulatory settings.
-
-\section{Practical vs. Statistical Significance for Discrimination Metrics}\label{a_sec:prac_v_stat}
-
-A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.
-Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
+
+\section{Practical vs. Statistical Significance for Discrimination Measures}\label{a_sec:prac_v_stat}
+
+A finding of \textit{practical significance} means that discovered disparity is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.
+Practical significance acknowledges that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
 It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model.
@@
 One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
-``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.
+``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.
 
@@
 
-\section{Simulated Data Results}
-
-Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+\section{Additional Simulated Data Results}\label{a_sec:res_sim}
+
 Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
 For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
-For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
-Post-hoc Shapley explanation techniques are also used to generate global and local feature importance for $g^\text{XNN}$.
+For $g^\text{XNN}$, post-hoc Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Section \ref{}.
 Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
-
-\subsection{Constrained vs. Unconstrained Model Fit Assessment}\label{ssec: c_v_uc_sim}
-
-Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-$g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
-
-\begin{table}[H]
-	\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
-	\centering
-	\begin{tabular}{ccccc}
-		\toprule
-		\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
-		\midrule
-		$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
-		$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
-		$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
-		$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
-		\bottomrule
-		\label{tab:sim_acc}
-	\end{tabular}
-\end{table}
-
-\noindent Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
-Accuracy is measured at the best F1 threshold for each model.
 
@@
 Both $g^\text{MGBM}$ and $g^\text{XNN}$ rank $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ as the most important features, both place $X_{\text{Categorical},2}$ and $X_{\text{Categorical},3}$ above the $X_{\text{Binary},1}$ and $X_{\text{Binary},2}$ features, both rank $X_{\text{Binary},1}$ above $X_{\text{Binary},2}$, and both place the least importance on $X_{\text{Categorical},4}$ and $X_{\text{Categorical},0}$.
-
-Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
-\ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
-The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms.
-Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
-	\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
-	\label{fig:sim_xnn_glob_ridge}
-\end{figure}
-
-\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks.
-Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
-In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
-Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.
-Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
 
@@
 
-\section{Cutoff Selection w.r.t Mitigating Discrimination}\label{a_sec:cut}
+\section{Cutoff Selection and Discrimination Testing}\label{a_sec:cut}
 
@@
 
-\section{Recent Fairness Techniques and Compliance}\label{a_sec:comp}
+\section{Recently Developed Fairness Techniques in US Legal and Regulatory Settings}\label{a_sec:comp}
 

### 2020-02-02 article-information-2019.tex
2/1 revisions to fit tables

@@
 
-As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on debunking the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN results.
+As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on challenging the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN interpretability results.
 The bulk of the post-hoc explanation and discrimination testing results for the simulated are left to Appendix \ref{a_sec:res_sim}.
@@
 $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
-Accuracy is measured at the best F1 threshold for each model.
+Measures that require a probability cutoff are taken at the best F1 threshold for each model.
 
 \begin{table}[H]
-	\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.}
-	\centering
-	\begin{tabular}{ccccc}
+	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Arrows indicate the direction of improvement for each measure.}
+	\centering
+	\begin{tabular}{ccccccccc}
 		\toprule
-		\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
+		\textbf{Model} & 
+		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ &
+		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
 		\midrule
-		$g^\text{GBM}$		& 0.775	& 0.857	& 0.474	& 0.394 \\
-		$g^\text{MGBM}$	& 0.763	& 0.846	& 0.498	& 0.405 \\
-		$g^\text{ANN}$		& 0.757	& 0.850	& 0.480	& 0.398 \\
-		$g^\text{XNN}$		& 0.758	& 0.851	& 0.479	& 0.397 \\
+		$g^\text{GBM}$		& 0.775	& 0.857	& & 0.474 &	& 0.394 & & \\
+		$g^\text{MGBM}$		& 0.763	& 0.846	& & 0.498 &	& 0.405 & & \\
+		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 &	0.525 & 0.398 & 0.858 & \\
+		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 &	0.528 & 0.397 & 0.867 & \\
 		\bottomrule
@@
 \begin{table}[H]
-\caption{Fit metrics for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data.}
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccccc}
+	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Arrows indicate the direction of improvement for each measure.}
+\centering
+\begin{tabular}{ccccccccc}
 	\toprule
-\textbf{Model}		& \textbf{Accuracy}	& \textbf{AUC}	 & \textbf{Logloss}	& \textbf{RMSE}\\
+	\textbf{Model} & 
+	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ &
+	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
 	\midrule	
-$g^\text{GBM}$		& 0.795	& 0.828	& 0.252	& 0.276 \\
-$g^\text{MGBM}$	& 0.765	& 0.814	& 0.259	& 0.278 \\
-$g^\text{ANN}$		& 0.865	& 0.871	& 0.231	& 0.262 \\
-$g^\text{XNN}$		& 0.869	& 0.868	& 0.233	& 0.263 \\
+$g^\text{GBM}$		& 0.795	& 0.828	& & 0.252 &	& 0.276 & & \\
+$g^\text{MGBM}$		& 0.765	& 0.814	& & 0.259 &	& 0.278 & & \\
+$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 &	0.418 & 0.262 & 0.624 & \\
+$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 &	0.409 & 0.263 & 0.593 & \\
 \bottomrule

### 2020-02-02 article-information-2019.tex
a perhaps not broken structure that adheres to the suggested 16-page limit

@@
 This manuscript makes no claim that these interdependent issues of opaqueness, discrimination, privacy harms, and security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
 
@@
 Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition.
-$L_1$ regularization is applied to all network weights to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
+$L_1$ regularization is applied to network weights to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
 The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size.
@@
 Accuracy is compared for unconstrained, less interpretable $g^{\text{GBM}}$ and $g^{\text{ANN}}$ models and constrained, more interpretable $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models.
-Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretability, post-hoc explanation, and discrimination testing results are presented.
+Then, for the  $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ models, intrinsic interpretability, post-hoc explanation, and discrimination testing results are explored.
 
@@
 
+Fit comparisons between unconstrained and constrained models and XNN interpretabililty results are discussed in Subsections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
 As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on challenging the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN interpretability results.
-The bulk of the post-hoc explanation and discrimination testing results for the simulated are left to Appendix \ref{a_sec:res_sim}.
+The bulk of the post-hoc explanation and discrimination testing results for the simulated data are left to Appendix \ref{a_sec:res_sim}.
 
@@
 
-Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+$g^\text{GBM}$ exhibits the best performance, but all models exhibit only a small range fit results.
 Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
-Accuracy is measured at the best F1 threshold for each model.
-Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-$g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.
-Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE).
-Measures that require a probability cutoff are taken at the best F1 threshold for each model.
+$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across all recorded fit metrics.
+Fit measures that require a probability cutoff are taken at the best F1 threshold for each model.
 
@@
 
-% add error metrics
-
-\subsubsection{$g^\text{XNN}$ Interpretability Results}
-
-For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the projection layer.
+% TODO: add error metrics, confirm statements about results
+\subsubsection{$g^\text{XNN}$ Interpretability Results}\label{ssec:xnn_sim}
+
+For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the bottom projection layer.
 Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
@@
 
-% update figure
+% TODO: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
 
@@
 In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
-Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and $n_7$, $n_8$, and $n_9$ do appear to display some noticeable sinusoidal characteristics.
-Conversely, subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the simulated function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.
+Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and these subnetworks, especially $n_7$ and $n_8$, appear to display some noticeable sinusoidal characteristics.
+Subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the signal-generating function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
 
@@
 Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort} to showcase the example workflow.
-$g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec: c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model performance with respect to unconstrained variants.
-Assuming that small fit differences on static test data do not outweigh the need for intrinsic model interpretability and reliable post-hoc explainability in high-stakes, human-centered, or regulated applications, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
+$g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec:c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model fit with respect to unconstrained variants.
+Assuming that in high-stakes applications small fit differences on static test data do not outweigh the need for enhanced model debugging facilitated by high interpretability, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
 
@@
 
-\begin{table}[H]
+\begin{table}[htb]
 	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Arrows indicate the direction of improvement for each measure.}
@@
 
+% TODO: add error metrics, confirm statements about results
+
 \subsubsection{Interpretability and Post-hoc Explanation Results}
 
+For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
 Global Shapley feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
-$g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
-
-\begin{figure}[H]
+$g^\text{MGBM}$ places high importance on loan-to-value (LTV) ratio, perhaps too high, and also weighs debt-to-income (DTI) ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}$ margin space.
+
+\begin{figure}[htb]
 \centering
@@
 
-\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
+\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
 
@@
 The monotonicity constraints for DTI ratio and LTV ratio are confirmed for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data in Figure \ref{fig:mort_mgbm_glob_pdp_ice}.
-Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles for ICE and on average with PD.
+Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for ICE and on average with PD.
 Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
-Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and perhaps due to less noise in the mortgage data, this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data.
-DTI ratio and LTV ratio also appear to have sparse regions in their univariate distributions.
+Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and perhaps due to non-synthetic noise in the mortgage data, this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data not observed in Table \ref{tab:sim_acc} for the simulated data.
+
+PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain.
+Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
+Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
+DTI ratio and LTV ratio do appear to have sparse regions in their univariate distributions.
 The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as $g^\text{MGBM}$ appears to carry reasonable predictions learned from populous domains into the sparse domains of both features.
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=16.5cm]{img/mort_mgbm_glob_pdp_ice.png}
+\begin{figure}[htb]
+\hskip-10pt\includegraphics[width=16.5cm]{img/mort_mgbm_glob_pdp_ice.png}
 \caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
@@
 In Figure \ref{fig:mort_mgbm_loc}, local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in the mortgage test data.
-The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
-
-\begin{figure}[H]
+Each $\phi_j$ value in Figure \ref{fig:mort_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
+Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
+
+\begin{figure}[htb]
 \centering
@@
 
-\noindent Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to \textit{accurately} summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
+\noindent The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
+Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to accurately and consistently summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
 In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.
@@
 Figure \ref{fig:mort_xnn_glob} displays global feature importance for $g^\text{XNN}$ on the mortgage test data.
+Deep SHAP values are reported in the probability space, after the application of the logit link function.
+They are also calculated from the projection layer of $g^\text{XNN}$.
+Thus, the Deep SHAP values in Figure \ref{fig:mort_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the mortgage test data.
 $g^\text{XNN}$ distributes importance more evenly across business drivers and puts stronger emphasis on the no introductory rate period flag feature than does $g^\text{MGBM}$.
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
 \centering
@@
 Distinctive ridge functions that feed into those large magnitude $\gamma_k$ weights are highlighted in \ref{fig:mort_xnn_glob_ridge} \textbf{B} and color-coded to pair with their corresponding $\gamma_k$ weight.
-As in the subsection \ref{ssec:int_ex_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge} \textbf{C}.
-In both the simulated and mortgage data, $g^\text{XNN}$ $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
-
-\begin{figure}[H]
+As in the Subsection \ref{ssec:xnn_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge} \textbf{C}.
+In both the simulated and mortgage data, $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
+\ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
+For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
+
+\begin{figure}[htb]
 \hskip-20pt\includegraphics[width=18cm]{img/mort_xnn_glob_ridge.png}
@@
 
-\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
-For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
+% TODO: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
 
 To compliment the global interpretability of $g^\text{XNN}$, Figure \ref{fig:mort_xnn_loc} displays local Shapley values for selected individuals, estimated from the projection layer using Deep SHAP in the $g^\text{XNN}$ probability space.
-
-\begin{figure}[H]
+Similar to Tree SHAP, local Deep SHAP values should sum to $g^\text{XNN}(\mathbf{x})$.
+While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc}, their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions, even for neural network architectures.
+
+\begin{figure}[htb]
 \centering
@@
 
-\noindent While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc} their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions, even for neural network architectures.
-
 \subsubsection{Discrimination Testing Results}\label{ssec:dis_mort}
 
-Tables \ref{tab:mort_n_acc} and \ref{tab:mort_disc} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
-As with the simulated data, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the FPRs, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).
-Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
+Tables \ref{tab:mort_disc}\textbf{a} and \ref{tab:mort_disc}\textbf{b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
+As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the accuracy, false positive rates (FPRs), and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to cutoff selection and discrimination testing.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
 Thus, one might consider FPR ratios as a measure of the class-control disparities.
 FPR ratios are higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs.
-2.10) in Table \ref{tab:mort_disc},  but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs.
-0.315) in Table \ref{tab:mort_n_acc}.
+2.10) in Table \ref{tab:mort_disc}\textbf{b}, but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs.
+0.315) in Table \ref{tab:mort_disc}\textbf{a}.
 This is the same pattern seen in the simulated data results in Section \ref{ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
@@
 
-\begin{table}[H]
+\begin{table}[htb]
 	\hskip-10pt\begin{subtable}{0.45\textwidth}
@@
 		\toprule
-		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy} & \textbf{FNR} \\
-		\midrule
-		Black & 2,608 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} & \begin{tabular}{@{}c@{}} 0.457 \\ 0.308\end{tabular} \\
-		\midrule			 
-		White & 28,361 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} & \begin{tabular}{@{}c@{}} 0.508 \\ 0.360\end{tabular}\\
-		\midrule
-		Female & 8,301 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} & \begin{tabular}{@{}c@{}} 0.402 \\ 0.322\end{tabular}\\
-		\midrule
-		Male & 13,166 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\ 0.847\end{tabular} & \begin{tabular}{@{}c@{}} 0.497 \\ 0.347\end{tabular}\\
+		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy}$\uparrow$ & \textbf{FPR}$\downarrow$ \\
+		\midrule
+		Black & 2,608 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} & \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} \\
+		\midrule			 
+		White & 28,361 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} & \begin{tabular}{@{}c@{}} 0.150 \\ 0.120\end{tabular}\\
+		\midrule
+		Female & 8,301 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} & \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}\\
+		\midrule
+		Male & 13,166 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\ 0.847\end{tabular} & \begin{tabular}{@{}c@{}} 0.182 \\ 0.131\end{tabular}\\
 		\bottomrule
@@
 		\toprule
-		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR} & \textbf{ME} & \textbf{SMD} & \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}\\
-		\midrule
-		$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}	& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular} & \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular} & \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular} & \begin{tabular}{@{}c@{}} 0.900 \\ 0.810\end{tabular} \\
-		\midrule 
-		$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular} & \begin{tabular}{@{}c@{}} White \\ Male\end{tabular} & \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular} & \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular} & \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular} & \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\
+		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}$\uparrow$ & \textbf{ME}$\downarrow$ & \textbf{SMD}$\downarrow$ & \begin{tabular}[x]{@{}c@{}}\textbf{FPR}\\\textbf{Ratio}\end{tabular}$\downarrow$\\
+		\midrule
+		$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}	& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular} & \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular} & \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular} & \begin{tabular}{@{}c@{}} 2.10 \\ 1.15\end{tabular} \\
+		\midrule 
+		$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular} & \begin{tabular}{@{}c@{}} White \\ Male\end{tabular} & \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular} & \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular} & \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular} & \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular} \\
 		\bottomrule
-		\label{tab:mort_disc}
+		\label{tab:mort_disc_detail}
 	\end{tabular}
@@
 \captionsetup{width=10cm}
-\caption{Discrimination measures for the mortgage test data.}
+\caption{Discrimination measures for the mortgage test data. Arrows indicate the direction of improvement for each measure.}
+\label{tab:mort_disc}
 \end{table}
+
+% TODO: make subfigure reference match XNN charts
 
@@
 Both $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ show AIRs that are statistically significantly below parity (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), and which are also below the EEOC’s 0.80 threshold.
-This would typically indicate cause for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
+This would typically indicate need for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
 On the other hand, women improve under $g^{\text{XNN}}$ vs.
@@
 The AIRs, while statistically significantly below parity, are well above the EEOC’s threshold of 0.80.
-In most situations, the values of these metrics alone would not likely flag a model for further review, though there may be other considerations which would lead to concern.
-
+In most situations, the values of these metrics alone would not likely flag a model for further review.
 Black SMDs for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$ are similar: 0.621 and 0.628, respectively.
@@
 
-Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, deployment meant to decrease discrimination, inaccuracy, privacy, and security risks for high-stakes, human-centered, or regulated ML applications.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.}
+Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, deployment meant to decrease discrimination, inaccuracy, privacy, and security risks for high-stakes, human-centered, or regulated ML applications.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
 
@@
 
-While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading.
-Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM or GA\textsuperscript{2}M models, pioneered by researchers at Microsoft and Cornell.
+Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM, or GA\textsuperscript{2}M models, pioneered by researchers at Microsoft and Cornell.
 In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
@@
 
-\noindent See: \href{https://github.com/jphall663/awesome-machine-learning-interpretability}{Awesome Machine Learning Interpretability} for a longer, community-curated metalist of related software packages and resources.
+\noindent See \href{https://github.com/jphall663/awesome-machine-learning-interpretability}{Awesome Machine Learning Interpretability} for a longer, community-curated metalist of related software packages and resources.
 
@@
 
-Interpretable model architectures and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions.
-As seen in Sections \ref{sec:m_and_m} and \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain.
+Interpretable models and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions.
+As seen in Section \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain.
 Post-hoc explanations generally help describe and summarize mechanisms and decisions, potentially yielding an even clearer understanding of ML models.
@@
 Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, providing people affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
-ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions is difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, providing persons affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
+ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions can be difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for compliance, but also a failsafe procedure for those affected by ML decisions.
 
@@
 
-The heightened interest in exploring and fixing algorithmic bias has led to a significant body of work focused in using ML to diminish discrimination in model outcomes \cite{friedler2019comparative}.
-Broadly, these can be defined into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and more recently developed approaches that change the algorithms or input data themselves in order to mitigate disparities.
-Many approaches that have been developed in the last five to seven years focus on altering the algorithm itself, preprocessing the data, or post-processing the predictions in order to diminish class-control correlations or dependencies.
-For comments on why these techniques could result in regulatory non-compliance in certain cases, see Appendix \ref{a_sec:comp}.
-
-Whether these methods are suitable for a particular use case depends on the legal environment in which the model will be used and the use case itself.
-Of the newer class of fairness enhancing interventions in ML, within-algorithm discrimination mitigation techniques that do not use class information may be more likely to be acceptable in highly regulated settings today.
-These techniques work by incorporating a loss function where more discriminatory paths or weights are penalized and will only be used by the model if the increase in model performance overcomes the penalty (the relative level of performance-to-discrimination penalty is determined via a hyperparameter choice).
-Other mitigation strategies that only alter hyperparameters or algorithm choice are also likely to be acceptable.
-And feature selection techniques that have been used in traditional modeling (ordinary least squares, logistic, and simple decision tree models) are likely to continue to be accepted in regulatory environments.
-For further discussion of potential and utilized techniques that can mitigate disparate impact in financial services, see \citet{schmidt2019introduction}.
+A significant body of research has emerged around exploring and fixing algorithmic discrimination \cite{friedler2019comparative}.
+Broadly, methodologies can be placed into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and more recently developed approaches that change the algorithms or input data themselves in order to mitigate disparities.
+Many approaches that have been developed in the last 5--7 years focus on altering the algorithm itself, preprocessing training data, or post-processing predictions in order to diminish class-control correlations or dependencies.
+Whether these more recent methods are suitable for a particular use case depends on the legal environment in which the model is deployed and on the use case itself.
+For comments on why these techniques could result in regulatory non-compliance in certain scenarios, see Appendix \ref{a_sec:comp}.
+
+Of the newer class of fairness enhancing interventions, within-algorithm discrimination mitigation techniques that do not use class information may be more likely to be acceptable in highly regulated settings today.
+These techniques often work by incorporating a loss function where more discriminatory paths or weights are penalized and will only be used by the model if improvements in model fit overcome the penalty.
+(The relative level of fit-to-discrimination penalty is usually determined via a hyperparameter choice.) Other mitigation strategies that only alter hyperparameters or algorithm choice are also likely to be acceptable.
+And feature selection techniques that have been used in traditional modeling (e.g., linear models and decision trees) are likely to continue to be accepted in regulatory environments.
+For further discussion of potential and utilized techniques that can mitigate disparate impact in US financial services, see \citet{schmidt2019introduction}.
 
@@
 Prior to these advances, only a small number of alternative algorithms could be tested for lower levels of disparity without causing infeasible delays in model implementation.
-Now, tens of thousands of alternative models or more can be quickly tested for lower discrimination and better predictive quality.
+Now, large numbers of models can be quickly tested for lower discrimination and better predictive quality.
 An additional opportunity arises as a result of ML itself: the well-known Rashomon effect, or the multiplicity of good ML models for most datasets.
-Thus, it is now feasible to train more models, find more good models, and test more models for discrimination, and among all those tested models, there are likely to be some with high predictive performance and low discrimination.
+It is now feasible to train more models, find more good models, and test more models for discrimination, and among all those tested models, there are likely to be some with high predictive performance and low discrimination.
 
@@
 
-The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own.
+The often black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the privacy harms and security vulnerabilities inherent in ML are each serious and difficult problems on their own.
 However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{26}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{21}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
 However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination.
@@
 Accuracy, discrimination, and security characteristics of a system can change over time as well.
-Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems.
+Simply testing for these problems at training time, as presented in Section \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems.
 Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
@@
 
-This text puts forward results on simulated data to provide a rough validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
+This text puts forward results on simulated data to provide some validation of constrained ML models, post-hoc explanation techniques, and discrimination testing methods.
 These same modeling, explanation, and discrimination testing approaches are then applied to more realistic mortgage data to provide an example of a responsible ML workflow for high-stakes, human-centered, or regulated ML applications.
@@
 Of course, there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
-While the messy, complex, and human problems of racism, sexism, privacy violations, and cyber crime can probably not be solved by technology alone, this work (and many, many others) illustrate numerous ways for ML practitioners to become part of the solution to these problems, instead of perpetuating and exacerbating them.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+While the complex and messy problems of racism, sexism, privacy violations, and cyber crime can probably never be solved by technology alone, this work (and many others) illustrate numerous ways for ML practitioners to mitigate such risks.
+
+%%%%%%%%
 \authorcontributions{NG, data cleaning, GBM and MGBM assessment and results; PH, primary author; KM, ANN and XNN implementation, assessment, and results; NS, secondary author, data simulation and collection, and discrimination testing.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
@@
 
-Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
-For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
-For $g^\text{XNN}$, post-hoc Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Section \ref{}.
+As seen in Subsection \ref{ssec:c_v_uc_sim}, little or no trade-off is required in terms of model to fit to use the constrained models.
+Hence, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
+Intrinsic interpretability for $g^\text{MGBM}$ is evaluated with PD and ICE, and post-hoc explainability is highlighted via global and local Shapley explanations.
+For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Section \ref{ssec:xnn_sim}.
 Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
@@
 Global mean absolute Shapley value feature importance for $g^{\text{MGBM}}(\mathbf{X})$ on the simulated test data is displayed in Figure \ref{fig:sim_mgbm_glob}.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the simulated test data in the $g^\text{MGBM}$ margin space.
-
-\begin{figure}[H]
+
+\begin{figure}[htb]
 	\centering
@@
 Figure \ref{fig:sim_mgbm_glob_pdp_ice} highlights PD, ICE, and histograms of the most important features from Figure \ref{fig:sim_mgbm_glob}.
-PD gives an indication of the estimated average prediction of $g^\text{MGBM}(\mathbf{X})$ in the test data for some $X_j$.
-ICE serves several purposes.
-It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is also a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior.
-PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain.
-Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
-Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
-
-\begin{figure}[H]
+
+\begin{figure}[htb]
 	\centering
@@
 Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
-These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
-Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
-
-\begin{figure}[H]
+
+\begin{figure}[htb]
 	\centering
@@
 Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique.
-Deep SHAP values are reported in the probability space, after the application of the logit link function.
-They are also calculated from the projection layer of $g^\text{XNN}$.
-Thus, the Deep SHAP values in Figure \ref{fig:sim_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the simulated test data.
-
-\begin{figure}[H]
+
+\begin{figure}[htb]
 	\centering
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
 	\centering
@@
 
-Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to cutoff selection and discrimination testing.}
+Tables \ref{tab:sim_disc}\textbf{a} and \ref{tab:sim_disc}\textbf{b} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\textsuperscript{11}
 
@@
 
-\begin{table}[H]
+\begin{table}[htb]
 	\hskip-10pt
@@
 				\toprule
-				\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy} & \textbf{FNR} \\
+		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy}$\uparrow$ & \textbf{FNR}$\downarrow$ \\
 				\midrule
@@
 				\toprule
-				\textbf{Model} & \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}	& \textbf{ME} & \textbf{SMD} & \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}\\
+		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}$\uparrow$ & \textbf{ME}$\downarrow$ & \textbf{SMD}$\downarrow$ & \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}$\downarrow$\\
 				\midrule
@@
 				\bottomrule
-				\label{tab:sim_disc}
+				\label{tab:sim_disc_detail}
 			\end{tabular}
@@
 	\captionsetup{width=10cm}
-	\caption{Discrimination measures for the simulated test data.}
+	\caption{Discrimination measures for the simulated test data. Arrows indicate the direction of improvement for each measure.}
+	\label{tab:sim_disc}
 \end{table}
@@
 
-\noindent SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
+SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
 (This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.206 vs.
@@
 But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
-Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.
+Because disparity analyses are concerned with how people are affected by the deployed model, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.
 
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-%\begin{quote}
-%This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-%\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Subsection}
-%\unskip
-%\subsubsection{Subsubsection}
-
-%Bulleted lists look like this:
-%\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-%\item	First bullet
-%\item	Second bullet
-%\item	Third bullet
-%\end{itemize}
-
-%Numbered lists can be added as follows:
-%\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-%\item	First item 
-%\item	Second item
-%\item	Third item
-%\end{enumerate}
-
-%The text continues here.
-
-%\subsection{Figures, Tables and Schemes}
-
-%All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
-
-%\begin{figure}[H]
-%\centering
-%\includegraphics[width=2 cm]{Definitions/logo-mdpi}
-%\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
-%\end{figure}
-
-%Text
-
-%Text
-
-%\begin{table}[H]
-%\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
-%\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-%\begin{tabular}{ccc}
-%\toprule
-%\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-%\midrule
-%entry 1		& data			& data\\
-%entry 2		& data			& data\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-
-%Text
-
-%Text
-
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\textwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\textwidth}{1pt}
-%\end{listing}
-
-%\subsection{Formatting of Mathematical Components}
-
-%This is an example of an equation:
-
-%\begin{equation}
-%a + b = c
-%\end{equation}
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-%Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-%\begin{Theorem}
-%Example text of a theorem.
-%\end{Theorem}
-
-%The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-%\begin{proof}[Proof of Theorem 1]
-%Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
-%\end{proof}
-%The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Discussion}
-
-%Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Materials and Methods}
-
-%Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-%Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-%Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Conclusions}
-
-%This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Patents}
-%This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
-%\noindent 
-%\begin{tabular}{@{}ll}
-%MDPI & Multidisciplinary Digital Publishing Institute\\
-%DOAJ & Directory of open access journals\\
-%TLA & Three letter acronym\\
-%LD & linear dichroism
-%\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-%\appendix
-%\section{}
-%\unskip
-%\subsection{}
-%The appendix is an optional section that can contain details and data supplemental to the main text.
-For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-
-%\section{}
-%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations an
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-%\begin{thebibliography}{999}
-% Reference 1
-%\bibitem[Author1(year)]{ref-journal}
-%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% Reference 2
-%\bibitem[Author2(year)]{ref-book}
-%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; %Publishing House: City, Country, 2007; pp. 32--58.
-%\end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\sampleavailability{Samples of the compounds ...... are available from the authors.}
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}

### 2020-02-02 article-information-2019.tex
a perhaps not broken structure that adheres to the suggested 16-page limit

@@
 This is the same pattern seen in the simulated data results in Section \ref{ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
-Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
+Similar results are found for the female-male comparison, buat the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
 

### 2020-02-03 article-information-2019.tex
intermediate commit, checking in-text citations, stoppped at [1], [18], [19]

@@
+% - Structure ------------------------- x 
+% - External links
+% - Internal links
+%   - Sections/subsections/appendices - x
+%   - References
+%   - Tables/figures
+%   - footnotes
+% - Abbreviations --------------------- x
+% - Bibliography 
+%   - Nick/Bryce citation
+% - Notation
+% - Tables/figures
+% - Grammar and spelling
+%   - Hyphenation
+
 %  LaTeX support: latex@mdpi.com 
@@
 This manuscript makes no claim that these interdependent issues of opaqueness, discrimination, privacy harms, and security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) Greater Data Privacy Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
 
@@
 \cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
-\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present previously delineated risks, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
-Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} (See subsection \ref{ssec:mgbm}) and XNNs (see Subsection \ref{ssec:xnn}, \cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
+\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present unmitigated risks, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
+Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} (See Subsection \ref{ssec:mgbm}) and XNNs (see Subsection \ref{ssec:xnn}, \cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
 MGBM and XNN interpretable model architectures are selected for the example workflow because they are straightforward variants of popular unconstrained ML models.
@@
 The data is split into a training set with 160,338 loans and a marker for 5-fold cross-validation and a test set containing 39,662 loans.
-While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (LTV ratio, DTI ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
+While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (loan to value (LTV) ratio, debt to income (DTI) ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
 See Appendix \ref{a_sec:mort} for general information regarding HMDA data.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Monotonic Gradient Boosting Machine}\label{ssec:mgbm}
+\subsection{Monotonic Gradient Boosting Machines}\label{ssec:mgbm}
 
@@
 
-\subsection{Explainable Neural Network}\label{ssec:xnn}
+\subsection{Explainable Neural Networks}\label{ssec:xnn}
 
@@
 Tree SHAP (SHapley Additive exPlanations) is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
-Tree SHAP has been implemented in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section \ref{sec:res} and Appendix \ref{a_sec:res_sim}.
+Tree SHAP has been implemented in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}.
-Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section \ref{sec:res} and Appendix \ref{a_sec:res_sim}.
+Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.
@@
 The discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
-See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination under US laws and Appendix \ref{a_sec:prac_v_stat} for remarks on statistical vs.
-practical significance for discrimination measures.
-One common measure of disparate impact used in US litigation and regulatory settings is known as marginal effects (ME).
+See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination under US laws and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
+statistical significance for discrimination measures.
+One common measure of DI used in US litigation and regulatory settings is ME.
 ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.
@@
 \noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
-ME is a favored DI metric used by the the US Consumer Finance Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is adverse impact ratio (AIR), more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
+ME is a favored DI metric used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
 
@@
 Statistically significant AIR values below 0.8 can be considered \textit{prima facia} evidence of discrimination.
-An additional long-standing measure of DI is standardized mean difference (SMD).
+An additional long-standing measure of DI is SMD.
 SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
@@
 
-\noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.
+\noindent The numerator in the SMD is equivalent to ME but adds the standard deviation divisor as a standardizing factor.
 Because of this standardization factor, SMD allows for a comparison across different types of outcomes, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
 In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail in Appendix \ref{a_sec:prac_v_stat}.
-Finally, confusion matrix measures in demographic groups, such as accuracy, false negative rate (FNR), and their ratios, are also considered as measures of DI in Section \ref{sec:res} and Appendix \ref{a_sec:res_sim}.
+Finally, confusion matrix measures in demographic groups, such as accuracy, false positive rate (FPR), and their ratios, are also considered as measures of DI in Subsection \ref{ssec:dis_mort} and Appendix \ref{a_ssec:dis_sim}.
 
@@
 Fit comparisons between unconstrained and constrained models and XNN interpretabililty results are discussed in Subsections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
-As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on challenging the perceived \textit{accuracy-interpretability tradeoff} for the considered model architectures and on the more novel XNN interpretability results.
+As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly fit measures for the considered data and model architectures and on the more novel $g^\text{XNN}$ interpretability results.
 The bulk of the post-hoc explanation and discrimination testing results for the simulated data are left to Appendix \ref{a_sec:res_sim}.
@@
 % TODO: add error metrics, confirm statements about results
-\subsubsection{$g^\text{XNN}$ Interpretability Results}\label{ssec:xnn_sim}
+\subsubsection{Interpretability Results}\label{ssec:xnn_sim}
 
@@
 This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
-$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$ (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
+$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced area under the curve (AUC) values of $\sim0.81$ (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
 
@@
 
-\subsubsection{Interpretability and Post-hoc Explanation Results}
+\subsubsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_mort}
 
@@
 Global Shapley feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
-$g^\text{MGBM}$ places high importance on loan-to-value (LTV) ratio, perhaps too high, and also weighs debt-to-income (DTI) ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
+$g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
 Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}$ margin space.
@@
 
-PD and ICE are displayed with a histogram in Section \ref{sec:res} to highlight any sparse regions in an input feature's domain.
+PD and ICE are displayed with a histogram to highlight any sparse regions in an input feature's domain.
 Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
@@
 Tables \ref{tab:mort_disc}\textbf{a} and \ref{tab:mort_disc}\textbf{b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
-As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the accuracy, false positive rates (FPRs), and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to cutoff selection and discrimination testing.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
+As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
 Thus, one might consider FPR ratios as a measure of the class-control disparities.
@@
 0.315) in Table \ref{tab:mort_disc}\textbf{a}.
-This is the same pattern seen in the simulated data results in Section \ref{ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
+This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
 Similar results are found for the female-male comparison, buat the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
@@
 	\captionsetup{width=7cm}
-	\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.} 
+	\caption{Group size, accuracy, and FPR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.} 
 	\centering
@@
 		\leftskip=45pt
-		AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+		AIR, ME, SMD, and FPR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
 	\centering
@@
 Both $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ show AIRs that are statistically significantly below parity (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), and which are also below the EEOC’s 0.80 threshold.
-This would typically indicate need for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
+This would typically indicate need for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Subsection \ref{ssec:disc_rem}.
 On the other hand, women improve under $g^{\text{XNN}}$ vs.
@@
 And feature selection techniques that have been used in traditional modeling (e.g., linear models and decision trees) are likely to continue to be accepted in regulatory environments.
-For further discussion of potential and utilized techniques that can mitigate disparate impact in US financial services, see \citet{schmidt2019introduction}.
+For further discussion of potential and utilized techniques that can mitigate DI in US financial services, see \citet{schmidt2019introduction}.
 
@@
 \abbreviations{The following abbreviations are used in this text: 
-AI -- artificial intelligence, 
-AIR - adverse impact ratio,
-ALE - accumulated local effect,
-ANN -- artificial neural network, 
-APR -- annual percentage rate,
-AUC -- area under the curve,
-CNN -- convolutional neural network,
-CFPB -- Consumer Financial Protection Bureau,
-DI -- disparate impact, 
-DT -- disparate treatment,
-DTI -- debt to income,
-EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e. variants GAMs that consider two-way interactions and may incorporate boosting into training, 
-EEOC -- Equal Employment Opportunity Commission,
-ECOA - Equal Credit Opportunity Act,
-EDA -- exploratory data analysis,
-EU -- European Union,
-FCRA -- Fair Credit Reporting Act,
-FNR -- false negative rate,
-FPR -- false positive rate,
-GAM -- generalized additive model, 
-GBM -- gradient boosting machine, 
-GDPR - General Data Protection Regulation,
-HMDA -- Home Mortgage Disclosure Act,
-ICE -- individual conditional expectation, 
-LTV -- loan to value,
-ME -- marginal effect,
-MGBM -- monotonic gradient boosting machine, 
-ML -- machine learning, 
-PD -- partial dependence, 
-RMSE -- root mean square error,
-SGD -- stochastic gradient descent, 
-SHAP -- Shapley additive explanation,
-SMD - standardized mean difference,
-SR -- supervision and regulation, 
-US -- United States, 
-XNN -- explainable neural network.}
+AI -- artificial intelligence, % -------------------------------------------------------------------------- x
+AIR - adverse impact ratio, % ----------------------------------------------------------------------------- x
+ALE - accumulated local effect, % ------------------------------------------------------------------------- x
+ANN -- artificial neural network, % ----------------------------------------------------------------------- x
+APR -- annual percentage rate, % -------------------------------------------------------------------------- x
+AUC -- area under the curve, % ---------------------------------------------------------------------------- x
+CFPB -- Consumer Financial Protection Bureau, % ----------------------------------------------------------- x
+DI -- disparate impact, % --------------------------------------------------------------------------------- x
+DT -- disparate treatment, % ------------------------------------------------------------------------------ x
+DTI -- debt to income, % ---------------------------------------------------------------------------------- x
+EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e.
+variants GAMs that consider two-way interactions and may incorporate boosting into training, % ------------------------------------------------ x
+EEOC -- Equal Employment Opportunity Commission, % -------------------------------------------------------- x
+ECOA - Equal Credit Opportunity Act, % -------------------------------------------------------------------- x
+EDA -- exploratory data analysis, % ----------------------------------------------------------------------- x
+EU -- European Union, % ----------------------------------------------------------------------------------- x
+FCRA -- Fair Credit Reporting Act, % ---------------------------------------------------------------------- x
+FNR -- false negative rate, % ----------------------------------------------------------------------------- x
+FPR -- false positive rate, % ----------------------------------------------------------------------------- x
+GAM -- generalized additive model, % ---------------------------------------------------------------------- x
+GBM -- gradient boosting machine, % ----------------------------------------------------------------------- x
+GDPR - General Data Protection Regulation, % -------------------------------------------------------------- x
+HMDA -- Home Mortgage Disclosure Act, % ------------------------------------------------------------------- x
+ICE -- individual conditional expectation, % -------------------------------------------------------------- x
+LTV -- loan to value, % ----------------------------------------------------------------------------------- x
+MCC -- Matthews correlation coefficient, % ---------------------------------------------------------------- x
+ME -- marginal effect, % ---------------------------------------------------------------------------------- x
+MGBM -- monotonic gradient boosting machine, % ------------------------------------------------------------ x
+ML -- machine learning, % --------------------------------------------------------------------------------- x
+PD -- partial dependence, % ------------------------------------------------------------------------------- x
+RMSE -- root mean square error, % ------------------------------------------------------------------------- x
+SGD -- stochastic gradient descent, % --------------------------------------------------------------------- x
+SHAP -- SHapley Additive exPlanation, % ------------------------------------------------------------------- x
+SMD - standardized mean difference, % --------------------------------------------------------------------- x
+SR -- supervision and regulation, % ----------------------------------------------------------------------- x
+US -- United States, % ------------------------------------------------------------------------------------ x
+XNN -- explainable neural network.} % --------------------------------------------------------------------- x
 
@@
 
-\section{HMDA Mortgage Data Details}\label{a_sec:mort}
+\section{Mortgage Data Details}\label{a_sec:mort}
 
 The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
-This information is first provided to the Consumer Financial Protection Bureau (CFPB), which subsequently releases some of the data to the public.
+This information is first provided to the CFPB, which subsequently releases some of the data to the public.
 Regulators often use HMDA data to, ``...show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory.''\textsuperscript{5} In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
-The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, loan to value (LTV) ratio, debt to income (DTI) ratio, and other important financial descriptors.
+The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, LTV ratio, DTI ratio, and other important financial descriptors.
 The data also include information on each borrower and co-borrower’s race, ethnicity, gender, and age.
@@
 
-\subsubsection{Datasets}
+\subsubsection{Data}
 
@@
 
-\subsection{Monotonic Gradient Boosting Machine}\label{a_ssec:mgbm}
+\subsection{Monotonic Gradient Boosting Machine Details}\label{a_ssec:mgbm}
 
@@
 
-\subsection{Explainable Neural Network}\label{a_ssec:xnn}
+\subsection{Explainable Neural Network Details}\label{a_ssec:xnn}
 
@@
 
-\subsection{One-dimensional Partial Dependence and Individual Conditional Expectation}\label{a_ssec:pd_ice}
+\subsection{One-dimensional Partial Dependence and Individual Conditional Expectation Details}\label{a_ssec:pd_ice}
 
@@
 $\text{ICE}(x_j, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
-PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Section \ref{sec:res}.
+PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
@@
 
-\subsection{Shapley Values}\label{a_ssec:shap}
+\subsection{Shapley Value Details}\label{a_ssec:shap}
 
@@
 
-\section{Machine Learning and Discrimination in US Legal and Regulatory Settings}\label{a_sec:types}
+\section{Types of Machine Learning Discrimination in US Legal and Regulatory Settings}\label{a_sec:types}
 
@@
 However, other factors, such as hyperparameter or algorithm choices, can drive DI.
-Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.
-
-\section{Practical vs. Statistical Significance for Discrimination Measures}\label{a_sec:prac_v_stat}
+Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower DI.
+
+\section{Practical vs. Statistical Significance for Discrimination Testing}\label{a_sec:prac_v_stat}
 
@@
 Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
-One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
 ``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.
@@
 The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.
-The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v.
-Partida}.\footnote{\textit{Castaneda v.
-Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v.
+The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda vs.
+Partida}.\footnote{\textit{Castaneda vs.
+Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts vs.
 United States.}\footnote{\textit{Hazelwood School Dist.
-v.
+vs.
 United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
@@
 As seen in Subsection \ref{ssec:c_v_uc_sim}, little or no trade-off is required in terms of model to fit to use the constrained models.
-Hence, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.
+Hence, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Appendices \ref{a_ssec:int_ex_sim} - \ref{a_ssec:dis_sim}.
 Intrinsic interpretability for $g^\text{MGBM}$ is evaluated with PD and ICE, and post-hoc explainability is highlighted via global and local Shapley explanations.
-For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Section \ref{ssec:xnn_sim}.
+For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Subection \ref{ssec:xnn_sim}.
 Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
 
-\subsection{Interpretability and Post-hoc Explanation Results}\label{ssec:int_ex_sim}
+\subsection{Interpretability and Post-hoc Explanation Results}\label{a_ssec:int_ex_sim}
 
@@
 
-\subsection{Discrimination Testing Results}\label{ssec:dis_sim}
+\subsection{Discrimination Testing Results}\label{a_ssec:dis_sim}
 
@@
 However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.
-Some pertinent remediation strategies for discovered discrimination are discussed in Section \ref{ssec:disc_rem}.
+Some pertinent remediation strategies for discovered discrimination are discussed in Subsection \ref{ssec:disc_rem}.
 
 SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
-(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.206 vs.
+(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less DI than $g^\text{XNN}$ (-0.206 vs.
 -0.274), but both are close to Cohen’s small effect threshold of -0.20.
@@
 
-\section{Cutoff Selection and Discrimination Testing}\label{a_sec:cut}
+\section{Discrimination Testing and Cutoff Selection}\label{a_sec:cut}
 
@@
 
-\section{Recently Developed Fairness Techniques in US Legal and Regulatory Settings}\label{a_sec:comp}
+\section{Recent Fairness Techniques in US Legal and Regulatory Settings}\label{a_sec:comp}
 
@@
 Additionally, the effects of changing the model must be viewed holistically.
-For example, the mortgage data disparity analysis in Section \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
+For example, the mortgage data disparity analysis in Subsection \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
 
 Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
-A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{In the pay discrimination case, \textit{Bazemore v.
+A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{In the pay discrimination case, \textit{Bazemore vs.
 Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will have the same model output (though there may be some small or random differences in outcomes due to computational issues).

### 2020-02-04 article-information-2019.tex
checked overall structure, internal links, and abbreviations

@@
+% TODO:
+% EDITING:
 % - Structure ------------------------- x 
@@
 %   - Sections/subsections/appendices - x
-%   - References
+%   - References ---------------------- x
 %   - Tables/figures
-%   - footnotes
+%   - footnotes ----------------------- x
 % - Abbreviations --------------------- x
@@
 %   - Hyphenation
-
-%  LaTeX support: latex@mdpi.com 
-%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
+%   - "metric" vs. "measure"
+% CONTENT:
+% - sim data: add error metrics, confirm statements about results
+% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
+% - mort data: add error metrics, confirm statements about results
+% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
+% - make subfigure reference match XNN charts
 
@@
 \documentclass[information,article,submit,moreauthors,pdftex]{definitions/mdpi}
-
-% If you would like to post an early version of this manuscript  a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
-The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
-This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
-For preprints.org, the editorial staff will make this change immediately prior to posting.
-
-%--------------------
-% Class Options:
-%--------------------
-%----------
-% journal
-%----------
-% Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejihpe, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
-
-%---------
-% article
-%---------
-% The default type of manuscript is "article", but can be replaced by: 
-% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
-% supfile = supplementary materials
-
-%----------
-% submit
-%----------
-% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
-This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
-Also, line numbering will be removed.
-Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
-
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
-
-%---------
-% pdftex
-%---------
-% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
 
@@
 
-%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
-%\continuouspages{yes}
-
-%------------------------------------------------------------------
-% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
-%\pdfoutput=1
-
-%=================================================================
-% Add packages and commands here.
-The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
-
-%=================================================================
-%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
-%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
-
 %=================================================================
@@
 \Title{A Responsible Machine Learning Workflow\\\normalsize{With Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
-
-% Author Orchid ID: enter ID or remove command
-%\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
-%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name
 
@@
 \firstnote{All authors contributed equally to this work.}
-%\secondnote{}
-% The commands \thirdnote{} till \eighthnote{} are available for further notes
-
-%\simplesumm{} % Simple summary
-
-%\conference{} % An extended version of a conference paper
 
@@
 
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences:
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data:
-%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
-If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
-In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
-
-%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%\setcounter{secnumdepth}{4}
-
 % for subtables
@@
 \setcounter{section}{-1}
-%% Remove this when starting to work on the template.
-%\section{How to Use this Template}
-%The template details the sections that can be used in a manuscript.
-Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section).
-Please check the instructions for authors page of the journal to verify the correct order and names.
-For any questions, please contact the editorial office of the journal or support@mdpi.com.
-For LaTeX related questions please contact latex@mdpi.com.
-%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
 
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
+Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).
@@
 
-% TODO: add error metrics, confirm statements about results
 \subsubsection{Interpretability Results}\label{ssec:xnn_sim}
@@
 \end{figure}
-
-% TODO: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
 
@@
 \noindent Regardless of the fit differences between the two families of hypothesis models, the difference between the fit of constrained and unconstrained variants within the two types of models is small for the GBMs and negligible for ANNs, 3\% and < 1\% worse fit respectively, averaged across the metrics reported in Table \ref{tab:mort_acc}.
-
-% TODO: add error metrics, confirm statements about results
 
@@
 \end{table}
-
-% TODO: make subfigure reference match XNN charts
 
@@
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{21}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
+This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
 However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination.

### 2020-02-04 article-information-2019.tex
fill in accuracy tables
larger, perhaps clearer, XNN charts

@@
 % CONTENT:
-% - sim data: add error metrics, confirm statements about results
-% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
-% - mort data: add error metrics, confirm statements about results
-% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table
-% - make subfigure reference match XNN charts
+% - sim data: add error metrics, confirm statements about results ------------------------------------------------------------------------------------------- x
+% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table --- x
+% - mort data: add error metrics, confirm statements about results ----------------------------------------------------------------------------------------- x
+% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table ------- x
+% - make subfigure reference match XNN charts ---------------------------------------------------------------------------------------------------------------- x
 
@@
 Fit comparisons between unconstrained and constrained models and XNN interpretabililty results are discussed in Subsections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
-As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly fit measures for the considered data and model architectures and on the more novel $g^\text{XNN}$ interpretability results.
+As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on fit measures for the considered data and model architectures and on the more novel $g^\text{XNN}$ interpretability results.
 The bulk of the post-hoc explanation and discrimination testing results for the simulated data are left to Appendix \ref{a_sec:res_sim}.
@@
 
-Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-$g^\text{GBM}$ exhibits the best performance, but all models exhibit only a small range fit results.
+Table \ref{tab:sim_acc} presents a variety of fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
+$g^\text{XNN}$ exhibits the best performance, but the models exhibit only a fairly small range of fit results.
 Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-$g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across all recorded fit metrics.
+For the displayed measures, $g^\text{MGBM}$ performs $\sim$2\% worse then $g^\text{GBM}$.
+$g^\text{XNN}$ performs $\sim$0.5\% better on average than $g^\text{XNN}$, and $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across all fit measures except specificity.
 Fit measures that require a probability cutoff are taken at the best F1 threshold for each model.
@@
 		\midrule
-		$g^\text{GBM}$		& 0.775	& 0.857	& & 0.474 &	& 0.394 & & \\
-		$g^\text{MGBM}$		& 0.763	& 0.846	& & 0.498 &	& 0.405 & & \\
-		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 &	0.525 & 0.398 & 0.858 & \\
-		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 &	0.528 & 0.397 & 0.867 & \\
+		$g^\text{GBM}$		& 0.757 	& 0.847 	& 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & 0.657 \\
+		$g^\text{MGBM}$	& 0.744 	& 0.842 	& 0.771 & 0.502 & 0.504 & 0.407 & 0.864 & 0.625 \\
+		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & 0.657 \\
+		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 & 0.528 & 0.397 & 0.867 & 0.648 \\
 		\bottomrule
@@
 Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
-\ref{fig:sim_xnn_glob_ridge} \textbf{A} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
-The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge} \textbf{B} as simplistic but distinctive functional forms.
-Color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{A} and \ref{fig:sim_xnn_glob_ridge} \textbf{B} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
+\ref{fig:sim_xnn_glob_ridge}\textbf{a} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
+The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge}\textbf{b} as simplistic but distinctive functional forms.
+Color-coding between \ref{fig:sim_xnn_glob_ridge}\textbf{a} and \ref{fig:sim_xnn_glob_ridge}\textbf{b} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
 
@@
 	\hskip-20pt\includegraphics[width=18cm]{img/sim_xnn_glob_ridge.png}
-	\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
+	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
 	\label{fig:sim_xnn_glob_ridge}
@@
 
-\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge} \textbf{B} and \ref{fig:sim_xnn_glob_ridge} \textbf{C} link the $\beta_j$ weights to their $n_k$ subnetworks.
+\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge}\textbf{b} and \ref{fig:sim_xnn_glob_ridge}\textbf{c} link the $\beta_j$ weights to their $n_k$ subnetworks.
 Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
@@
 Subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the signal-generating function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge}\textbf{a}, \textbf{b}, and \textbf{c} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the signal-generating function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
+See Appendix \ref{a_ssec:xnn} for details regarding general XNN architecture.
 
@@
 
-Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data.
+Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data for most of the fit measures.
 This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g., numerical encoding of categorical features and missing values.
@@
 	\midrule	
-$g^\text{GBM}$		& 0.795	& 0.828	& & 0.252 &	& 0.276 & & \\
-$g^\text{MGBM}$		& 0.765	& 0.814	& & 0.259 &	& 0.278 & & \\
-$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 &	0.418 & 0.262 & 0.624 & \\
-$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 &	0.409 & 0.263 & 0.593 & \\
+$g^\text{GBM}$		& 0.795	& 0.828	& 0.376 & 0.252 & 0.314 & 0.276 & 0.634 & 0.813 \\
+$g^\text{MGBM}$	& 0.765	& 0.814	& 0.362 & 0.259 & 0.305 & 0.278 & 0.684 & 0.773 \\
+$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 & 0.418 & 0.262 & 0.624 & 0.891 \\
+$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 & 0.409 & 0.263 & 0.594 & 0.898 \\
 \bottomrule
@@
 
-\noindent Regardless of the fit differences between the two families of hypothesis models, the difference between the fit of constrained and unconstrained variants within the two types of models is small for the GBMs and negligible for ANNs, 3\% and < 1\% worse fit respectively, averaged across the metrics reported in Table \ref{tab:mort_acc}.
+\noindent Regardless of the fit differences between the two families of models, the difference between the constrained and unconstrained variants within the two types of models is small for the GBMs and smaller for the ANNs, $\sim$3.5\% and $\sim$1\% worse fit respectively, averaged across the measures in Table \ref{tab:mort_acc}.
 
@@
 Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}$ margin space.
+The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
 
@@
 \end{figure}
-
-\noindent The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
 
@@
 Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
-Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data, and perhaps due to non-synthetic noise in the mortgage data, this rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data not observed in Table \ref{tab:sim_acc} for the simulated data.
+Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data.
+Perhaps due to the presence of important non-monotonic interactions (e.g., LTV ratio and the no introductory rate period flag, see Figure \ref{fig:mort_xnn_glob_ridge}).
+This rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data not observed for the simulated data, wherein strong interactions appear to be between features with the same monotonicity constraints (e.g., $X_{\text{Friedman},1}$ and $X_{\text{Friedman},2}$, see Figure \ref{fig:sim_xnn_glob_ridge}).
 
 PD and ICE are displayed with a histogram to highlight any sparse regions in an input feature's domain.
-Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
+Because most ML models will always issue a prediction on any instance with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
 Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
@@
 This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
-However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
+However, fluctuations in ICE can also be caused by overfitting or by leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
 
@@
 The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions, and to do so in an interpretable manner, is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
-\ref{fig:mort_xnn_glob_ridge} \textbf{A} presents the sparse $\gamma_k$ weights of the $g^\text{XNN}$ output layer in which the $n_k$ subnetworks with $k \in \{0,1,2,3,5,8,9\}$ have large magnitude weights and $n_k$ subnetworks, $k \in \{4,6,7\}$, have small or near-zero magnitude weights.
-Distinctive ridge functions that feed into those large magnitude $\gamma_k$ weights are highlighted in \ref{fig:mort_xnn_glob_ridge} \textbf{B} and color-coded to pair with their corresponding $\gamma_k$ weight.
-As in the Subsection \ref{ssec:xnn_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge} \textbf{C}.
+\ref{fig:mort_xnn_glob_ridge}\textbf{a} presents the sparse $\gamma_k$ weights of the $g^\text{XNN}$ output layer in which the $n_k$ subnetworks with $k \in \{0,1,2,3,5,8,9\}$ have large magnitude weights and $n_k$ subnetworks, $k \in \{4,6,7\}$, have small or near-zero magnitude weights.
+Distinctive ridge functions that feed into those large magnitude $\gamma_k$ weights are highlighted in \ref{fig:mort_xnn_glob_ridge}\textbf{b} and color-coded to pair with their corresponding $\gamma_k$ weight.
+As in the Subsection \ref{ssec:xnn_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge}\textbf{c}.
 In both the simulated and mortgage data, $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
-\ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
+\ref{fig:mort_xnn_glob_ridge}\textbf{c} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
 For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
+See Appendix \ref{a_ssec:xnn} for details regarding general XNN architecture.
 
@@
 \hskip-20pt\includegraphics[width=18cm]{img/mort_xnn_glob_ridge.png}
-\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
+\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
 \label{fig:mort_xnn_glob_ridge}
@@
 This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
-Similar results are found for the female-male comparison, buat the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
+Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
 

### 2020-02-05 article-information-2019.tex
bigger figures, checked all external links, confirmed figure an table placement, proofed bibliography and tables/figures

@@
 % - Structure ------------------------- x 
-% - External links
-% - Internal links
+% - External links -------------------- x
+% - Internal links --------------------------- x
 %   - Sections/subsections/appendices - x
 %   - References ---------------------- x
-%   - Tables/figures
+%   - Tables/figures ----------------- x
 %   - footnotes ----------------------- x
 % - Abbreviations --------------------- x
-% - Bibliography 
-%   - Nick/Bryce citation
+% - Bibliography  ---------------------- x 
+% - Tables/figures -------------------- x
 % - Notation
-% - Tables/figures
 % - Grammar and spelling
 %   - Hyphenation
-%   - "metric" vs. "measure"
 % CONTENT:
@@
 
-Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/us/en/services/consulting/analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear.
+Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/gx/en/issues/data-and-analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear.
 ML models, a common application of AI, can present serious risks.
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}.
+Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).
@@
 \noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
-ME is a favored DI metric used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
+ME is a favored DI measure used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
 
@@
 It originates from work on statistical power, and is more formally known as \textit{Cohen’s d}.
-SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
+SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this measure to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
 
@@
 Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}.
-The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.
+The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and primarily \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.
 
@@
 \begin{table}[H]
-	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Arrows indicate the direction of improvement for each measure.}
+	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.
+Arrows indicate the direction of improvement for each measure.}
 	\centering
@@
 \begin{figure}[H]
-	\hskip-20pt\includegraphics[width=18cm]{img/sim_xnn_glob_ridge.png}
-	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the simulated test data.}
+	\hskip-35pt\includegraphics[width=19cm]{img/sim_xnn_glob_ridge.png}
+	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ subnetwork ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the simulated data.}
 	\label{fig:sim_xnn_glob_ridge}
@@
 \begin{table}[htb]
-	\caption{Fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the mortgage test data. Arrows indicate the direction of improvement for each measure.}
+	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.
+Arrows indicate the direction of improvement for each measure.}
 \centering
@@
 For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
-Global Shapley feature importance for $g^\text{MGBM}$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
+Global Shapley feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
 $g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
@@
 \centering
-\includegraphics[width=9cm]{img/mort_mgbm_glob.png}
+\captionsetup{width=17cm}
+\includegraphics[width=10cm]{img/mort_mgbm_glob.png}
 \caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
@@
 \centering
-\includegraphics[width=9cm]{img/mort_xnn_glob.png}
+\includegraphics[width=10cm]{img/mort_xnn_glob.png}
 \caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
@@
 \begin{figure}[htb]
-\hskip-20pt\includegraphics[width=18cm]{img/mort_xnn_glob_ridge.png}
-\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_j$ weights for $g^\text{XNN}$ on the mortgage test data.}
+\hskip-35pt\includegraphics[width=19cm]{img/mort_xnn_glob_ridge.png}
+\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the mortgage data.}
 \label{fig:mort_xnn_glob_ridge}
@@
 0.315) in Table \ref{tab:mort_disc}\textbf{a}.
-This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness metric.
+This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness measure.
 Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
@@
 	\captionsetup{width=7cm}
-	\caption{Group size, accuracy, and FPR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.} 
+	\caption{Group size, accuracy, and FPR for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.} 
 	\centering
@@
 		\leftskip=45pt
-		AIR, ME, SMD, and FPR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
+		AIR, ME, SMD, and FPR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
 	\centering
@@
 The AIRs, while statistically significantly below parity, are well above the EEOC’s threshold of 0.80.
-In most situations, the values of these metrics alone would not likely flag a model for further review.
+In most situations, the values of these measures alone would not likely flag a model for further review.
 Black SMDs for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$ are similar: 0.621 and 0.628, respectively.
@@
 	\centering
-	\includegraphics[width=11cm]{img/blueprint.png}
-	\caption{An example responsible ML workflow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among other review and appeal mechanisms, can create an understandable and trustworthy ML system.}
+	\includegraphics[width=12cm]{img/blueprint.png}
+	\caption{An example responsible ML workflow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among several other processes, can create an understandable and trustworthy ML system.}
 	\label{fig:blueprint}
@@
 The data also include information on each borrower and co-borrower’s race, ethnicity, gender, and age.
-Because the data includes information on these protected class characteristics, certain metrics that can be indicative of discrimination in lending can be calculated directly using the HDMA data.
+Because the data includes information on these protected class characteristics, certain measures that can be indicative of discrimination in lending can be calculated directly using the HDMA data.
 Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections \ref{sec:m_and_m} and \ref{sec:res}.
@@
 \begin{figure}[htb]
-	\centering
-	\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
+	\captionsetup{width=16cm}
+	\centering
+	\includegraphics[width=10cm]{img/sim_mgbm_glob.png}
 	\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the simulated test data.}
@@
 
-Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique.
+Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data, using the approximate Deep SHAP technique.
 
@@
 	\centering
-	\includegraphics[width=8cm]{img/sim_xnn_glob.png}
+	\includegraphics[width=10cm]{img/sim_xnn_glob.png}
 	\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
@@
 Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities.
-Table \ref{tab:sim_disc} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
+Table \ref{tab:sim_disc}\textbf{b} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
 1.06).
-However, in Table \ref{tab:sim_n_acc} the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
+However, in Table \ref{tab:sim_disc}\textbf{a}  the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
 0.401).
-This illustrates a danger in considering relative class-control metrics in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$.
+This illustrates a danger in considering relative class-control measures in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$.
 This is because FNR accuracy improves for both the protected group 1 and control group 1, but members of control group 1 benefit more than those in protected group 1.
@@
 		\captionsetup{width=7cm}
-		\caption{Group size, accuracy, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
+		\caption{Group size, accuracy, and FNR for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.} 
 		\centering
@@
 			\leftskip=45pt
-			AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
+			AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
 		\centering
@@
 
-For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.
+For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}\textbf{b}.
 Of note is that 12.0\% is not a meaningful difference without context.
@@
 But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
-Because disparity analyses are concerned with how people are affected by the deployed model, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.
+Because disparity analyses are concerned with how people are affected by the deployed model, it is essential that any confusion matrix-based measures of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.
 
@@
 
-Great care must be taken in order to ensure that the appropriate fairness metrics are chosen, because certain metrics may not be appropriate for some use cases.
+Great care must be taken in order to ensure that the appropriate fairness measures are chosen, because certain measures may not be appropriate for some use cases.
 Additionally, the effects of changing the model must be viewed holistically.

### 2020-02-06 article-information-2019.tex
first edit of main text

@@
 % - External links -------------------- x
-% - Internal links --------------------------- x
+% - Internal links -------------------- x
 %   - Sections/subsections/appendices - x
 %   - References ---------------------- x
-%   - Tables/figures ----------------- x
+%   - Tables/figures ------------------ x
 %   - footnotes ----------------------- x
 % - Abbreviations --------------------- x
-% - Bibliography  ---------------------- x 
+% - Bibliography  --------------------- x 
 % - Tables/figures -------------------- x
@@
 % CONTENT:
-% - sim data: add error metrics, confirm statements about results ------------------------------------------------------------------------------------------- x
-% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table --- x
-% - mort data: add error metrics, confirm statements about results ----------------------------------------------------------------------------------------- x
-% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table ------- x
-% - make subfigure reference match XNN charts ---------------------------------------------------------------------------------------------------------------- x
+% - sim data: add error metrics, confirm statements about results - x
+% - sim data xnn: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table - x
+% - mort data: add error metrics, confirm statements about results - x
+% - mort data: update figure: connections, feed-forward direction, make subfigure reference match discrimination measure table - x
+% - make subfigure reference match XNN charts - x
 
@@
 $^{2}$ \quad BLDS, LLC\\
-$^{3}$ \quad George Washington University}
+$^{3}$ \quad The George Washington University}
 
@@
 \abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBM) and explainable neural networks (XNN), a deep learning architecture well-suited for structured data, are assessed on simulated data and publicly available mortgage data.
-For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio (AIR), marginal effect (ME), standardized mean difference (SMD), and also with straightforward group fairness measures.
-By combining interpretable models, post-hoc explanation, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBMs) and explainable neural networks (XNNs), a deep learning architecture well-suited for structured data, are assessed on simulated data and publicly available mortgage data.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and with global and local Shapley feature importance.
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD), and also with straightforward group fairness measures.
+By combining interpretable models, post-hoc explanations, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
 
 % Keywords
-\keyword{Machine Learning; Neural Network; Gradient Boosting Machine; Interpretable; Explanation; Fairness; Disparate Impact; Python}
+\keyword{Deep Learning; Disparate Impact; Explanation; Fairness; Gradient Boosting Machine; Interpretable; Machine Learning; Neural Network; Python}
 
@@
 ML models, a common application of AI, can present serious risks.
-ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.
-ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
-This manuscript makes no claim that these interdependent issues of opaqueness, discrimination, privacy harms, and security vulnerabilities in ML have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing implemented in widely available Python tools, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve SR 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
-
-Section \ref{sec:m_and_m} describes methods and materials, including collected training datasets, interpretable and constrained models, post-hoc explanations, tests for DI and other social discrimination, and public and open source software resources associated with this text.
+ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}, and ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
+This manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena.
+However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervisory and Regulatory (SR) Letter 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+
+Section \ref{sec:m_and_m} describes methods and materials, including training datasets, interpretable and constrained models, post-hoc explanations, tests for DI and other social discrimination, and public and open source software resources associated with this text.
 In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and post-hoc explanation and discrimination testing results are also presented for interpretable models.
@@
 The simulated data (see Subsection \ref{ssec:sim_data}) is based on the well-known Friedman datasets.
-Its known feature importance and discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}.
-The mortgage data (see Subsection \ref{ssec:mort_data}) is sourced from the Home Mortgage Disclosure Act (HMDA) database and is a fairly realistic data source for demonstrating the template workflow.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} To provide a sense of accuracy differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
-Because the unconstrained ML models, gradient boosting machines (GBMs) (e.g.
-\cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs) (e.g.
-\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present unmitigated risks, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
+Its known feature importance and augmented discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}.
+The mortgage data (see Subsection \ref{ssec:mort_data}) is sourced from the Home Mortgage Disclosure Act (HMDA) database and is a fairly realistic data source for demonstrating the template workflow.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} To provide a sense of fit differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
+Because the unconstrained ML models, gradient boosting machines (GBMs, e.g.
+\cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs, e.g.
+\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present the unmitigated risks discussed above, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
 Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} (See Subsection \ref{ssec:mgbm}) and XNNs (see Subsection \ref{ssec:xnn}, \cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
+Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).
@@
 
-Simulated data is created based on a signal-generating function, $f$, first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
+Simulated data is created based on a signal-generating function, $f$, applied to input data, $\mathbf{X}$, first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
 
@@
 \begin{aligned}\label{eq:sim}\small
-f(\mathbf{X}) = 10\sin(\pi~\mathbf{X}_{\text{Friedman},1}~\mathbf{X}_{\text{Friedman},2}) + 20(\mathbf{X}_{\text{Friedman},3} - 0.5)^2 + 10~\mathbf{X}_{\text{Friedman},4} + 5~\mathbf{X}_{\text{Friedman},5}
+f(\mathbf{X}) = 10\sin(\pi~X_{\text{Friedman},1}~X_{\text{Friedman},2}) + 20(X_{\text{Friedman},3} - 0.5)^2 + 10~X_{\text{Friedman},4} + 5~X_{\text{Friedman},5}
 \end{aligned}
@@
 
-\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$.
+\noindent where each $X_{\text{Friedman},j}$ is a random uniform feature in $[0,1]$.
 In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
@@
 For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
-For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
+For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be $\frac{1}{5}$ of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.
 Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
 The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 instances.
-The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.
+The simulated data is then split into a training and test set, with 80,000 and 20,000 instances, respectively.
 Within the training set, a 5-fold cross-validation indicator is used for training all models.
@@
 While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (loan to value (LTV) ratio, debt to income (DTI) ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
-See Appendix \ref{a_sec:mort} for general information regarding HMDA data.
+See the resources put forward in Section \ref{ssec:soft} and Appendix \ref{a_sec:mort} for more information regarding the HMDA mortgage data.
 
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric loss function (e.g., squared loss, Huber loss), and a regularization term that penalizes complexity in the current tree.
+\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x};\Theta^\text{mono}_b)$, in a numeric loss function (e.g., squared loss, Huber loss), and a regularization term that penalizes complexity in the current tree.
 See Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
@@
 XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e., GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
-Hence, XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
+XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
 A trained XNN function, $g^{\text{XNN}}$, applied to some instance , $\mathbf{x}$, is defined as:
@@
 
-\noindent where $\mu_0$ is a global bias for $K$ individually specified ANN subnetworks, $n_k$, with scale parameters $\gamma_k$.
-The inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest projection layer.
+\noindent where $\mu_0$ is a global bias for $K$ individually specified ANN subnetworks, $n_k$, with weights $\gamma_k$.
+The inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest projection layer of $g^\text{XNN}$.
 
@@
 
-PD plots are a widely-used method for describing and plotting the estimated average prediction of a complex model, $g$, across some partition of data, $\mathbf{X}$, for some interesting input feature, $X_j$ \cite{esl}.
-ICE plots are a newer method that describes the local behavior of $g$ with regard to values of an input feature, $x_j$, in a single instance, $\mathbf{x}$.
+PD plots are a widely-used method for describing and plotting the estimated average prediction of a complex model, $g$, across some partition of data, $\mathbf{X}$, for some interesting input feature, $X_j \in \mathbf{X}$ \cite{esl}.
+ICE plots are a newer method that describes the local behavior of $g$ with regard to values of an input feature in a single instance, $x_j$.
 PD and ICE can be overlaid in the same plot to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
@@
 Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}, \cite{shapley1988shapley}.
-Shapley explanations are the only locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to $g(\mathbf{x})$ for any $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
+Shapley explanations are the only known locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to the model's prediction, $g(\mathbf{x})$, for any instance $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some instance of $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
 Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features.
@@
 
-Because many technical and academic discussions of fairness in ML have been inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
+Because many current technical discussions of fairness in ML appear inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
 The discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
-See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination under US laws and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
+See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
 statistical significance for discrimination measures.
-One common measure of DI used in US litigation and regulatory settings is ME.
+One such common measure of DI used in US litigation and regulatory settings is ME.
 ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.
@@
 
-\noindent where $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
+\noindent where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
 ME is a favored DI measure used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
@@
 
-\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the proportion of the control class that receives a favorable outcome.
+\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome and the proportion of the control class that receives a favorable outcome.
 Statistically significant AIR values below 0.8 can be considered \textit{prima facia} evidence of discrimination.
-An additional long-standing measure of DI is SMD.
+An additional long-standing and pertinent measure of DI is SMD.
 SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
@@
 
-\noindent The numerator in the SMD is equivalent to ME but adds the standard deviation divisor as a standardizing factor.
+\noindent The numerator in the SMD is roughly equivalent to ME but adds the standard deviation divisor as a standardizing factor.
 Because of this standardization factor, SMD allows for a comparison across different types of outcomes, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
 In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail in Appendix \ref{a_sec:prac_v_stat}.
-Finally, confusion matrix measures in demographic groups, such as accuracy, false positive rate (FPR), and their ratios, are also considered as measures of DI in Subsection \ref{ssec:dis_mort} and Appendix \ref{a_ssec:dis_sim}.
+Finally, confusion matrix measures in demographic groups, such as accuracy, false positive rate (FPR), false negative rate (FNR), and their ratios, are also considered as measures of DI in Subsection \ref{ssec:dis_mort} and Appendix \ref{a_ssec:dis_sim}.
 
@@
 Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}.
-The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and primarily \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.
+The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and typically \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.
 
@@
 
-Fit comparisons between unconstrained and constrained models and XNN interpretabililty results are discussed in Subsections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
-As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on fit measures for the considered data and model architectures and on the more novel $g^\text{XNN}$ interpretability results.
+Fit comparisons between unconstrained and constrained models and XNN interpretability results are discussed in Subsections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
+As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on fit measures and the more novel $g^\text{XNN}$ interpretability results.
 The bulk of the post-hoc explanation and discrimination testing results for the simulated data are left to Appendix \ref{a_sec:res_sim}.
@@
 Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-For the displayed measures, $g^\text{MGBM}$ performs $\sim$2\% worse then $g^\text{GBM}$.
+For the displayed measures, $g^\text{MGBM}$ performs $\sim$2\% worse on average than $g^\text{GBM}$.
 $g^\text{XNN}$ performs $\sim$0.5\% better on average than $g^\text{XNN}$, and $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across all fit measures except specificity.
@@
 
-For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_j$ weights in the bottom projection layer.
-Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into $g^\text{XNN}$.
+For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_{j,k}$ weights in the bottom projection layer.
+Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into the structure $g^\text{XNN}$ (also described in Equation \ref{eq:xnn}).
 \ref{fig:sim_xnn_glob_ridge}\textbf{a} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
@@
 
-\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge}\textbf{b} and \ref{fig:sim_xnn_glob_ridge}\textbf{c} link the $\beta_j$ weights to their $n_k$ subnetworks.
+\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge}\textbf{b} and \ref{fig:sim_xnn_glob_ridge}\textbf{c} link the $\beta_{j,k}$ weights to their $n_k$ subnetworks.
 Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
@@
 Subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge}\textbf{a}, \textbf{b}, and \textbf{c} exhibit evidence that $g^\text{XNN}(\mathbf{X})$ has learned about the signal-generating function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
-See Appendix \ref{a_ssec:xnn} for details regarding general XNN architecture.
+As a whole, Figure \ref{fig:sim_xnn_glob_ridge}\textbf{a}, \textbf{b}, and \textbf{c} exhibit evidence that $g^\text{XNN}$ has learned about the signal-generating function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
+See Appendix \ref{a_ssec:xnn} for additional details regarding general XNN architecture.
 
@@
 This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
-$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced area under the curve (AUC) values of $\sim0.81$ (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
+$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced area under the curve (AUC) values of $\sim$0.81 (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
 
@@
 
-For $g^\text{MGBM}$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
+For $g^\text{MGBM}(\mathbf{X})$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
 Global Shapley feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
 $g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
-Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}$ margin space.
-The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}$ as compared to $g^\text{XNN}$.
+Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}(\mathbf{X})$ margin space.
+The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the de-emphasis of the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}(\mathbf{X})$ as compared to $g^\text{XNN}(\mathbf{X})$.
 
@@
 Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
-Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the data.
-Perhaps due to the presence of important non-monotonic interactions (e.g., LTV ratio and the no introductory rate period flag, see Figure \ref{fig:mort_xnn_glob_ridge}).
-This rigidity could also play a role in the performance differences between $g^\text{MGBM}$ and $g^\text{XNN}$ in the mortgage data not observed for the simulated data, wherein strong interactions appear to be between features with the same monotonicity constraints (e.g., $X_{\text{Friedman},1}$ and $X_{\text{Friedman},2}$, see Figure \ref{fig:sim_xnn_glob_ridge}).
+Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the training data (e.g., LTV ratio and the no introductory rate period flag, see Figure \ref{fig:mort_xnn_glob_ridge}).
+This rigidity could also play a role in the performance differences between $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ in the mortgage data not observed for the simulated data, wherein strong interactions appear to be between features with the same monotonicity constraints (e.g., $X_{\text{Friedman},1}$ and $X_{\text{Friedman},2}$, see Figure \ref{fig:sim_xnn_glob_ridge}).
 
 PD and ICE are displayed with a histogram to highlight any sparse regions in an input feature's domain.
-Because most ML models will always issue a prediction on any instance with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction.
+Because most ML models will always issue a prediction on any instance with a correct schema, it's crucial to consider whether a given model learned enough about an instance to make an accurate prediction.
 Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
 DTI ratio and LTV ratio do appear to have sparse regions in their univariate distributions.
-The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as $g^\text{MGBM}$ appears to carry reasonable predictions learned from populous domains into the sparse domains of both features.
+The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as $g^\text{MGBM}(\mathbf{X})$ appears to carry reasonable predictions learned from dense domains into the sparse domains of both features.
 
@@
 This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
-However, fluctuations in ICE can also be caused by overfitting or by leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
+Fluctuations in ICE could also be caused by overfitting or by leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
 
 In Figure \ref{fig:mort_mgbm_loc}, local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in the mortgage test data.
-Each $\phi_j$ value in Figure \ref{fig:mort_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}.
-Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.
+Each Shapley value in Figure \ref{fig:mort_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with this instance of some input feature $x_j$ \cite{molnar}.
+Accordingly, the logit of the sum of the Shapley values and the average of $g^\text{MGBM}(\mathbf{X})$ is $g^\text{MGBM}(\mathbf{x})$, the prediction in the probability space for any $\mathbf{x}$.
 
@@
 Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to accurately and consistently summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
-In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as adverse actions codes.
-
-Figure \ref{fig:mort_xnn_glob} displays global feature importance for $g^\text{XNN}$ on the mortgage test data.
+In the US, applicable regulations are typically ECOA and FCRA, and the consumer-specific reasons are commonly known as \textit{adverse actions codes}.
+
+Figure \ref{fig:mort_xnn_glob} displays global feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.
 Deep SHAP values are reported in the probability space, after the application of the logit link function.
 They are also calculated from the projection layer of $g^\text{XNN}$.
-Thus, the Deep SHAP values in Figure \ref{fig:mort_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the mortgage test data.
+Thus, the Deep SHAP values in Figure \ref{fig:mort_xnn_glob} are the estimated average absolute impact of each input, $X_j$, in the projection layer and probability space of $g^\text{XNN}(\mathbf{X})$ for the mortgage test data.
 $g^\text{XNN}$ distributes importance more evenly across business drivers and puts stronger emphasis on the no introductory rate period flag feature than does $g^\text{MGBM}$.
@@
 In both the simulated and mortgage data, $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
-\ref{fig:mort_xnn_glob_ridge}\textbf{c} displays the sparse $\beta_j$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
+\ref{fig:mort_xnn_glob_ridge}\textbf{c} displays the sparse $\beta_{j,k}$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
 For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
@@
 Similar to Tree SHAP, local Deep SHAP values should sum to $g^\text{XNN}(\mathbf{x})$.
-While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc}, their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions, even for neural network architectures.
+While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc}, their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions.
 
@@
 Tables \ref{tab:mort_disc}\textbf{a} and \ref{tab:mort_disc}\textbf{b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
-As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and $g^{\text{XNN}}$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
+As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and $g^{\text{XNN}}(\mathbf{X})$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
 Thus, one might consider FPR ratios as a measure of the class-control disparities.
@@
 0.315) in Table \ref{tab:mort_disc}\textbf{a}.
-This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, again leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness measure.
-Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}$ and 1.21 for $g^{\text{XNN}}$.
+This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness measure.
+Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}(\mathbf{X})$ and 1.21 for $g^{\text{XNN}}(\mathbf{X})$.
 
@@
 In most situations, the values of these measures alone would not likely flag a model for further review.
-Black SMDs for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$ are similar: 0.621 and 0.628, respectively.
+Black SMDs for $g^{\text{XNN}}(\mathbf{X})$ and $g^{\text{MGBM}}(\mathbf{X})$ are similar: 0.621 and 0.628, respectively.
 These exceed Cohen’s guidelines of 0.5 for a medium effect size and would likely trigger further review.
-Female SMDs are well below Cohen’s definition of small effect size: 0.105 and 0.084 for $g^{\text{XNN}}$ and $g^{\text{MGBM}}$, respectively.
+Female SMDs are well below Cohen’s definition of small effect size: 0.105 and 0.084 for $g^{\text{XNN}}(\mathbf{X})$ and $g^{\text{MGBM}}(\mathbf{X})$, respectively.
 Similar to results for female AIR, these values alone are unlikely to prompt further review.
@@
 
-Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, deployment meant to decrease discrimination, inaccuracy, privacy, and security risks for high-stakes, human-centered, or regulated ML applications.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
+Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, and deployment meant to decrease discrimination, inaccuracy, privacy, and security risks for high-stakes, human-centered, or regulated ML applications.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
 
@@
 	\includegraphics[width=12cm]{img/blueprint.png}
-	\caption{An example responsible ML workflow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among several other processes, can create an understandable and trustworthy ML system.}
+	\caption{An example responsible ML workflow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among several other processes, can create an understandable and trustworthy ML system for high-stakes, human-centered, or regulated applications.}
 	\label{fig:blueprint}
@@
 
-Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBM, or GA\textsuperscript{2}M models, pioneered by researchers at Microsoft and Cornell.
+Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBMs, or GA\textsuperscript{2}Ms, pioneered by researchers at Microsoft and Cornell.
 In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
@@
 
-\item \textbf{Privacy preprocessing and private models}: differential privacy and private models in \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}} and \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{diffprivlib}, \cite{ji2014differential}, \cite{pate}, \cite{abadi2016deep}.
+\item \textbf{Preprocessing \& models for privacy}: \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}}, \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{diffprivlib}, \cite{ji2014differential}, \cite{pate}, \cite{abadi2016deep}.
+
+\item \textbf{Causal inference \& probabilistic programming}: \href{https://github.com/microsoft/dowhy}{dowhy}, \href{https://docs.pymc.io/}{PyMC3} \cite{book_of_why}.
 
@@
 
-\item \textbf{Discrimination testing}: with \href{https://github.com/dssg/aequitas}{\texttt{aequitas}} and \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.
+\item \textbf{Discrimination testing}: \href{https://github.com/dssg/aequitas}{\texttt{aequitas}}, \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.
 
@@
 
-\item \textbf{Model debugging}: with \href{https://github.com/bethgelab/foolbox}{\texttt{foolbox}}, \href{https://github.com/SALib/SALib}{\texttt{SALib}}, \href{https://github.com/tensorflow/cleverhans}{\texttt{tensorflow/cleverhans}}, and \href{https://github.com/tensorflow/model-analysis}{\texttt{tensorflow/model-analysis}} \cite{rauber2017foolbox}, \cite{papernot2018cleverhans}, \cite{modeltracker}, \cite{papernot2018marauder}.
+\item \textbf{Model debugging}: \href{https://github.com/bethgelab/foolbox}{\texttt{foolbox}}, \href{https://github.com/SALib/SALib}{\texttt{SALib}}, \href{https://github.com/tensorflow/cleverhans}{\texttt{tensorflow/cleverhans}}, and \href{https://github.com/tensorflow/model-analysis}{\texttt{tensorflow/model-analysis}} \cite{rauber2017foolbox}, \cite{papernot2018cleverhans}, \cite{modeltracker}, \cite{papernot2018marauder}.
 
@@
 
-Interpretable models and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions.
+Interpretable models and post-hoc explanations can play an important role in increasing transparency into model mechanisms and predictions.
 As seen in Section \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain.
@@
 Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, providing persons affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank SR 11-7: Guidance on Model Risk Management}.} Among these, providing persons affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
 ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions can be difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for compliance, but also a failsafe procedure for those affected by ML decisions.
@@
 
-A significant body of research has emerged around exploring and fixing algorithmic discrimination \cite{friedler2019comparative}.
-Broadly, methodologies can be placed into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and more recently developed approaches that change the algorithms or input data themselves in order to mitigate disparities.
-Many approaches that have been developed in the last 5--7 years focus on altering the algorithm itself, preprocessing training data, or post-processing predictions in order to diminish class-control correlations or dependencies.
-Whether these more recent methods are suitable for a particular use case depends on the legal environment in which the model is deployed and on the use case itself.
-For comments on why these techniques could result in regulatory non-compliance in certain scenarios, see Appendix \ref{a_sec:comp}.
+A significant body of research has emerged around exploring and fixing ML discrimination \cite{friedler2019comparative}.
+Methods can be broadly placed into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and many approaches that have been developed in the last 5--7 years that alter the training algorithm, preprocess training data, or post-process predictions in order to diminish class-control correlations or dependencies.
+Whether these more recent methods are suitable for a particular use case depends on the legal environment where a model is deployed and on the use case itself.
+For comments on why some recent techniques could result in regulatory non-compliance in certain scenarios, see Appendix \ref{a_sec:comp}.
 
 Of the newer class of fairness enhancing interventions, within-algorithm discrimination mitigation techniques that do not use class information may be more likely to be acceptable in highly regulated settings today.
-These techniques often work by incorporating a loss function where more discriminatory paths or weights are penalized and will only be used by the model if improvements in model fit overcome the penalty.
-(The relative level of fit-to-discrimination penalty is usually determined via a hyperparameter choice.) Other mitigation strategies that only alter hyperparameters or algorithm choice are also likely to be acceptable.
-And feature selection techniques that have been used in traditional modeling (e.g., linear models and decision trees) are likely to continue to be accepted in regulatory environments.
-For further discussion of potential and utilized techniques that can mitigate DI in US financial services, see \citet{schmidt2019introduction}.
+These techniques often incorporate a loss function where more discriminatory paths or weights are penalized and only used by the model if improvements in fit overcome some penalty.
+(The relative level of fit-to-discrimination penalty is usually determined via hyperparameter.) Other mitigation strategies that only alter hyperparameters or algorithm choice are also likely to be acceptable.
+Traditional feature selection techniques (e.g., those used in linear models and decision trees) are also likely to continue to be accepted in regulatory environments.
+For further discussion of techniques that can mitigate DI in US financial services, see \citet{schmidt2019introduction}.
 
@@
 
-The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain data about many of the mortgage-related products that they either deny or originate on an annual basis.
+The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain loan-level data about many of the types of mortgage-related products on an annual basis.
 This information is first provided to the CFPB, which subsequently releases some of the data to the public.
@@
 	\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
-	\item Each $i$-th observation of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
+	\item Each $i$-th observed instance of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
 	$[x_0^{(i)}, x_1^{(i)}, \dots, x_{\textit{P}-1}^{(i)}]$, with corresponding $i$-th labels in $\mathbf{Y}, \mathbf{y}^{(i)}$, and corresponding predictions in $\mathbf{\hat{Y}}, \mathbf{\hat{y}}^{(i)}$.
 % = [y_0^{(i)}, y_1^{(i)}, \dots, y_{\textit{C}-1}^{(i)}]$.
-	\item $\mathbf{X}$ and $\mathbf{Y}$ consist of $N$ tuples of observations: $[(\mathbf{x}^{(0)},\mathbf{y}^{(0)}), (\mathbf{x}^{(1)},\mathbf{y}^{(1)}), \dots,(\mathbf{x}^{(N-1)},\mathbf{y}^{(N-1)})]$.
+	\item $\mathbf{X}$ and $\mathbf{Y}$ consist of $N$ tuples of observed instances: $[(\mathbf{x}^{(0)},\mathbf{y}^{(0)}), (\mathbf{x}^{(1)},\mathbf{y}^{(1)}), \dots,(\mathbf{x}^{(N-1)},\mathbf{y}^{(N-1)})]$.
 %\\ $\mathbf{x}^{(i)} \in \mathcal{X}$, $\mathbf{y}^{(i)} \in \mathcal{Y}$.
@@
 Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$) is considered.
-$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observations of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
+$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observed instances of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
 $\text{ICE}(x_j, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
@@
 \noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
-Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the observations in some set $\mathbf{X}$.
+Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the instances in some set $\mathbf{X}$.
 
@@
 	\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
-	\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
+	\caption{PD, ICE for 10 instances across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
 	\label{fig:sim_mgbm_glob_pdp_ice}
@@
 	\includegraphics[width=15cm]{img/sim_mgbm_loc.png}
-	\caption{Tree SHAP values for three observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.}
+	\caption{Tree SHAP values for three instances across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the simulated test data.}
 	\label{fig:sim_mgbm_loc}
@@
 	\includegraphics[width=15cm]{img/sim_xnn_loc.png}
-	\caption{Deep SHAP values for three observations across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
+	\caption{Deep SHAP values for three instances across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
 	\label{fig:sim_xnn_loc}
@@
 Tables \ref{tab:sim_disc}\textbf{a} and \ref{tab:sim_disc}\textbf{b} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, false negative rates (FNRs), their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\textsuperscript{11}
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\textsuperscript{11}
 

### 2020-02-06 article-information-2019.tex
more editing of Section 2, quick pass over appendices A, C, D, E, F, G

@@
 In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
-In this manuscript, the signal generating function and input features are modified in several ways.
+In this manuscript, the signal-generating function and input features are modified in several ways.
 Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings.
@@
 \noindent where $\mu_0$ is a global bias for $K$ individually specified ANN subnetworks, $n_k$, with weights $\gamma_k$.
-The inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest projection layer of $g^\text{XNN}$.
+The inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest, i.e., \textit{projection}, layer of $g^\text{XNN}$.
 
@@
 For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_{j,k}$ weights in the bottom projection layer.
-Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into the structure $g^\text{XNN}$ (also described in Equation \ref{eq:xnn}).
+Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into the structure of $g^\text{XNN}$ (also described in Equation \ref{eq:xnn}).
 \ref{fig:sim_xnn_glob_ridge}\textbf{a} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
@@
 
-The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain loan-level data about many of the types of mortgage-related products on an annual basis.
+The US HMDA law, originally enacted in 1975, requires many financial institutions that originate mortgage products to provide certain loan-level data about many types of mortgage-related products on an annual basis.
 This information is first provided to the CFPB, which subsequently releases some of the data to the public.
@@
 
-It is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI).
-DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
-With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI, colloquially known as \textit{unintentional discrimination}, occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
+It is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT) and disparate impact (DI).
+DT occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
+With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
 In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.
@@
 For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Subection \ref{ssec:xnn_sim}.
-Both $g^\text{MGBM}$ and $g^\text{XNN}$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
+Both $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
 
@@
 $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE.
-Also, as the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
-Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is probably not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}.
-The one-dimensional monotonic constraints may hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data.
-This is an interesting result for responsible ML practitioners.
-In some noisy scenarios, monotonicity constraints can increase model interpretability without causing a drastic drop in model accuracy, even when known interactions and non-monotonic behavior exist in training data.
+As the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
+Also because PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is likely not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal-generating function in Equation \ref{eq:sim}.
+The one-dimensional monotonic constraints may hinder the modeling of non-monotonic interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to the main drivers, $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$, all being constrained in the same direction and able to weakly interact as needed.
 
@@
 
-\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x}^{(i)})$ predictions.
+\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x})$ predictions.
 
@@
 Tables \ref{tab:sim_disc}\textbf{a} and \ref{tab:sim_disc}\textbf{b} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}$ and  $g^{\text{XNN}}$, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\textsuperscript{11}
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and  $g^{\text{XNN}}(\mathbf{X})$, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\textsuperscript{11}
 
@@
 
-For $g^{\text{XNN}}$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}\textbf{b}.
+For $g^{\text{XNN}}(\mathbf{X})$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}\textbf{b}.
 Of note is that 12.0\% is not a meaningful difference without context.
@@
 SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
-(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}$ has less DI than $g^\text{XNN}$ (-0.206 vs.
+(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}(\mathbf{X})$ has less DI than $g^\text{XNN}(\mathbf{X})$ (-0.206 vs.
 -0.274), but both are close to Cohen’s small effect threshold of -0.20.
@@
 But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
-Because disparity analyses are concerned with how people are affected by the deployed model, it is essential that any confusion matrix-based measures of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.
+Because disparity analyses are concerned with how people are affected by the deployed model, it is essential that any confusion matrix-based measures of disparity be calculated on the in-production classification decisions, rather than on cutoffs that are not related to what those affected by the model will experience.
 
@@
 
-Great care must be taken in order to ensure that the appropriate fairness measures are chosen, because certain measures may not be appropriate for some use cases.
-Additionally, the effects of changing the model must be viewed holistically.
+Great care must be taken to ensure that the appropriate discrimination measures are employed for any given use case.
+Additionally, the effects of changing a model must be viewed holistically.
 For example, the mortgage data disparity analysis in Subsection \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
-
 Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.

### 2020-02-07 article-information-2019.tex
edits to appendix B

@@
 
-To facilitate descriptions of data and modeling, explanatory, and discrimination testing techniques, notation for input and output spaces, datasets, and models is defined.
+To facilitate descriptions of data, modeling, and other post-hoc techniques, notation for input and output spaces, datasets, and models is defined.
 
@@
 \begin{itemize}[leftmargin=*,labelsep=5.8mm]
-	\item The input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
+	\item An input dataset $\mathbf{X}$ is composed of observed instances of the set $\mathcal{X}$ with a corresponding dataset of labels $\mathbf{Y}$, observed instances of the set $\mathcal{Y}$. 
 	\item Each $i$-th observed instance of $\mathbf{X}$ is denoted as $\mathbf{x}^{(i)} = $  
@@
 	\item $g$ generates learned output responses on the input dataset $g(\mathbf{X}) = \mathbf{\hat{Y}}$, and on the general input space $g(\mathcal{X}) = \mathcal{\hat{Y}}$.
-	\item The model to be explained and tested for discrimination is denoted as $g$.
+	\item A model to be explained or tested for discrimination is denoted as $g$.
 \end{itemize}
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in the in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits, $T_b(\mathbf{x}^{(i)};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, and a regularization term that penalizes complexity in the current tree, $\Omega(T_b)$.
-For the $b$-th iteration, the loss function, $\mathcal{L}_{b}$, can generally be defined as:
+\noindent For some $g^\text{MGBM}$ model (see Equation \ref{eq:gbm}), monotonic splitting rules, $\Theta^{\text{mono}}_b$, are selected in a greedy, additive fashion by minimizing a regularized loss function, $\mathcal{L}$,  that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits applied to some instance $\mathbf{x}$, $T_b(\mathbf{x};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, in addition to a regularization term that penalizes complexity in the $b$-th tree, $\Omega(T_b)$.
+For the $b$-th iteration over $N$ instances, $\mathcal{L}_{b}$, can generally be defined as:
 
@@
 
-\noindent In addition to $\mathcal{L}_{b}$, $g^\text{MGBM}$ training is characterized by additional splitting rules and constraints on tree node weights.
-Each binary splitting rule, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left and right child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$.
+\noindent In addition to $\mathcal{L}$, $g^\text{MGBM}$ training is characterized by monotonic splitting rules and constraints on tree node weights.
+Each binary splitting rule in $T_b$, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left, $L$, and right, $R$, child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$.
 For terminal nodes, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$ can be direct numeric components of some $g^\text{MGBM}$ prediction.
-For two values of some feature $X_j$, $x^{\alpha}_j \le x^{\beta}_j$, $g^{\text{MGBM}}$ is positive monotonic with respect to some $X_j$ if $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
-The following rules and constraints ensure positive monotonicity in $\Theta_b$, where the prediction for each value results in $T_b(x^{\alpha}_j; \Theta_b) = w_\alpha$ and $T_b(x^{\beta}_j; \Theta_b) = w_\beta$.
+For two values, $x^{\alpha}_j$ and $x^{\beta}_j$, of some feature $X_j$, where $x^{\alpha}_j \le x^{\beta}_j$, $g^{\text{MGBM}}$ is positive monotonic with respect to $X_j$ if $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~\{x^{\alpha}_j \le x^{\beta}_j\} \in X_j$.
+The following rules and constraints ensure positive monotonicity in $\Theta_b^\text{mono}$:
 
@@
 
-\noindent Note that for any one $X_j$ and $T_b \in g^{\text{MGBM}}$ left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each $T_b$ output, with the application of a monotonic logit or softmax link function for classification problems.
+\noindent Note that for any one $X_j$ and subtree in $g^{\text{MGBM}}$, left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each full $T_b$ prediction, with the application of a monotonic logit or softmax link function for classification problems.
 Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0}) = w_{b,0}$.
-Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j)  \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$.
-For a negative monotonic constraint, i.e., $g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall ~x^{\alpha}_j \le x^{\beta}_j \in X_j$, left and right splitting rules and node weight constraints are switched.
+Together these additional splitting rules and node weight constraints ensure that $g^{\text{MGBM}}(x^{\alpha}_j) \le g^{\text{MGBM}}(x^{\beta}_j) ~\forall~ \{x^{\alpha}_j \le x^{\beta}_j\} \in X_j$.
+For a negative monotonic constraint, i.e., $g^{\text{MGBM}}(x^{\alpha}_j)  \ge g^{\text{MGBM}}(x^{\beta}_j) ~\forall~ \{x^{\alpha}_j \le x^{\beta}_j\} \in X_j$, left and right splitting rules and node weight constraints are switched.
 Also consider that MGBM models with independent monotonicity constraints between some $X_j$ and $\mathbf{y}$ likely restrict non-monotonic interactions between multiple $X_j$.
@@
 	\includegraphics[width=8cm]{img/xnn_arch.png}
-	\caption{A diagram of an XNN with three metalayers: the bottom combination layer with  $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, the middle metalayer with with $K$ hidden and separate $n_k$ ridge functions, and the output combination layer that generates $g^{\text{XNN}}$ predictions.
+	\caption{A general diagram of an XNN with three metalayers: the bottom combination layer with  $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, the middle metalayer with with $K$ hidden and separate $n_k$ ridge functions, and the output combination layer that generates $g^{\text{XNN}}(\mathbf{X})$ predictions.
 Figure adapted from \citet{wf_xnn}.}
@@
 \begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-	\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, which should learn higher magnitude weights for important $X_j$, is known as the \textit{projection layer}.
+	\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units (see Equation \ref{eq:xnn}), which should learn higher magnitude weights for each important input, $X_j$, is known as the \textit{projection layer}.
 It is fully connected to each input $X_j$.
@@
 	\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
-Each $n_k$ is a neural network, which can be parameterized to suit a given modeling task.
-To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer hidden unit, $\sum_j\beta_{k,j}x_j$.
+Each $n_k$ is a neural network itself, which can be parameterized to suit a given modeling task.
+To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer $\sum_j\beta_{k,j}x_j$ hidden unit.
 Each $n_k$ can contain several bias terms.
-	\item The output meta-layer, called the \textit{combination layer}, is another linear unit comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
-Again, subnetwork output is restricted to 1-dimension for interpretation and visualization purposes.
+	\item The output meta-layer, called the \textit{combination layer}, is an output neuron comprised of a global bias term, $\mu_0$, and the $K$ weighted 1-dimensional outputs of each subnetwork, $\gamma_kn_k(\sum_j\beta_{k,j}x_j)$.
+Again, each $n_k$ subnetwork output into the combination layer is restricted to 1-dimension for interpretation and visualization purposes.
 \end{enumerate}
@@
 
-Following \citet{esl} a single feature $X_j \in \mathbf{X}$ and its complement set $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$ (where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$) is considered.
-$\text{PD}(X_j, g)$ for a given feature $X_j$ is estimated as the average output of the learned function $g(\mathbf{X})$ when all the observed instances of $X_j$ are set to a constant $x \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
-$\text{ICE}(x_j, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
-PD and ICE curves are usually plotted over some set of constants $x \in \mathcal{X}$, as displayed in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
+Following \citet{esl} a single input feature, $X_j \in \mathbf{X}$, and its complement set, $\mathbf{X}_{\mathcal{P} \setminus \{j\}} \in \mathbf{X}$, where $X_j \cup \mathbf{X}_{\mathcal{P} \setminus \{j\}} = \mathbf{X}$ is considered.
+$\text{PD}(X_j, g)$ for a given $X_j$ is the estimated average output of the learned function $g(\mathbf{X})$ when all the observed instances of $X_j$ are set to a constant $x^\gamma \in \mathcal{X}$ and $\mathbf{X}_{\mathcal{P} \setminus \{j\}}$ is left unchanged.
+$\text{ICE}(x_j, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x^\gamma \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
+PD and ICE curves are usually plotted over some set of constants drawn from $\mathcal{X}$, as displayed in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
@@
 \begin{aligned}
-g(\mathbf{x}) = \phi_0 + \sum_{j=0}^{j=\mathcal{P} - 1} \phi_j \mathbf{z}_j
+g(\mathbf{x}) = \phi_0 + \sum_{j=0}^{j=\mathcal{P} - 1} \phi_j z_j
 \end{aligned}
@@
 \noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
-Each $\phi_j$ is the local feature contribution value associated with $x_j$, and $\phi_0$ is the average of $g(\mathbf{X})$.
-Each $\phi_j$ is a weighted combination of model scores, $g_x(\mathbf{x})$, with $x_j$, $g_x(S \cup \{j\})$, and the model scores without $x_j$, $g_x(S)$, for every subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
+Each Shapley value, $\phi_j$, is the local feature contribution value associated with $x_j$, and $\phi_0$ is the average of $g(\mathbf{X})$.
+Each $\phi_j$ is a weighted combination of model predictions with $x_j$, $g_x(S \cup \{j\})$, and the model predictions without $x_j$, $g_x(S)$, for every possible subset of features $S$ not including $j$, $S \subseteq \mathcal{P} \setminus \{j\}$, where $g_x$ incorporates the mapping between $\mathbf{x}$ and the binary vector $\mathbf{z}$.
 
@@
 
-\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
+\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j \in \mathbf{x}$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
 Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the instances in some set $\mathbf{X}$.

### 2020-02-10 article-information-2019.tex
submitted, revised version

@@
 % Authors, for the paper (add full first names)
-\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,3,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger}$}
+\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,3,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger, *}$}
 
@@
 For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and with global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD), and also with straightforward group fairness measures.
+The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD), along with straightforward group fairness measures.
 By combining interpretable models, post-hoc explanations, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
@@
 This manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form interpretable models, post-hoc explanation, and DI and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervisory and Regulatory (SR) Letter 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and DI and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervisory and Regulatory (SR) Letter 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
 
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discussed discrimination testing methods (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
+Discrimination testing methods discussed (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
 All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).
@@
 XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
-XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}M), i.e., GAMs which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
+XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}Ms), which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
 XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
@@
 The discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
-See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
+See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings, and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
 statistical significance for discrimination measures.
@@
 \noindent where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
-ME is a favored DI measure used by the the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
+ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
 
@@
 \noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome and the proportion of the control class that receives a favorable outcome.
-Statistically significant AIR values below 0.8 can be considered \textit{prima facia} evidence of discrimination.
+Statistically significant AIR values below 0.8 can be considered \textit{prima facie} evidence of discrimination.
 An additional long-standing and pertinent measure of DI is SMD.
@@
 Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort} to showcase the example workflow.
-$g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec:c_v_uc_sim} the constrained variants of both model architectures do not show large differences in model fit with respect to unconstrained variants.
+$g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec:c_v_uc_sim}, the constrained variants of both model architectures do not show large differences in model fit with respect to unconstrained variants.
 Assuming that in high-stakes applications small fit differences on static test data do not outweigh the need for enhanced model debugging facilitated by high interpretability, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
@@
 Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for ICE and on average with PD.
-Because PD curves generally follow the patterns of the ICE curves for both features, it's also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
+Because PD curves generally follow the patterns of the ICE curves for both features, it is also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
 Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the training data (e.g., LTV ratio and the no introductory rate period flag, see Figure \ref{fig:mort_xnn_glob_ridge}).
@@
 PD and ICE are displayed with a histogram to highlight any sparse regions in an input feature's domain.
-Because most ML models will always issue a prediction on any instance with a correct schema, it's crucial to consider whether a given model learned enough about an instance to make an accurate prediction.
+Because most ML models will always issue a prediction on any instance with a correct schema, it is crucial to consider whether a given model learned enough about an instance to make an accurate prediction.
 Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
@@
 Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns established by ICE curves.
-While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$, and lower, exhibit different prediction behavior with respect to property value.
-Some individuals at these lower percentiles display monotonically decreasing prediction behavior while others appear to show fluctuating prediction behavior.
+While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$ and lower exhibit different prediction behavior with respect to property value.
+Some individuals at these lower percentiles display monotonically decreasing prediction behavior, while others appear to show fluctuating prediction behavior.
 Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn.
@@
 
-\noindent The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
+\noindent The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values at the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
 Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to accurately and consistently summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
@@
 Of course, there is more to the responsible practice of ML than interpretable models, post-hoc explanation, and discrimination testing, even from a technology perspective, and Section \ref{sec:disc} also points out numerous additional references and open source Python software assets that are available to researchers and practitioners today to increase human trust and understanding in ML systems.
-While the complex and messy problems of racism, sexism, privacy violations, and cyber crime can probably never be solved by technology alone, this work (and many others) illustrate numerous ways for ML practitioners to mitigate such risks.
+While the complex and messy problems of racism, sexism, privacy violations, and cyber crime can probably never be solved by technology alone, this work and many others illustrate numerous ways for ML practitioners to mitigate such risks.
 
@@
 Wells Fargo is a customer of, and investor in, H2O.ai and a client of BLDS, LLC.
-However, communications regarding XNN between Wells Fargo and the authors have been extremely limited prior to and during the drafting of this manuscript.
+However, communications regarding XNN between Wells Fargo and Patrick Hall at H2O.ai have been extremely limited prior to and during the drafting of this manuscript.
 Moreover, Wells Fargo exerted absolutely no editorial control over the text or results herein.}
@@
 
-A finding of \textit{practical significance} means that discovered disparity is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facia} evidence of illegal discrimination.
+A finding of \textit{practical significance} means that discovered disparity is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facie} evidence of illegal discrimination.
 Practical significance acknowledges that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
@@
 Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
-One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facia} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facie} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
 ``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.

### 2020-02-26 article-information-2019.tex
edits after mdpi edits ...

@@
+% CC by 4.0
+
 % TODO:
@@
 %=================================================================
-\documentclass[information,article,submit,moreauthors,pdftex]{definitions/mdpi}
-
+%%%\documentclass[information,article,submit,moreauthors,pdftex]{definitions/mdpi} 
+\documentclass[information,article,accept,moreauthors,pdftex]{Definitions/mdpi}
+\usepackage{soul}
 %=================================================================
@@
 \articlenumber{5}
-\pubyear{2019}
-\copyrightyear{2019}
+\pubyear{2020}
+\copyrightyear{2020}
 %\externaleditor{Academic Editor: name}
-\history{Received: date; Accepted: date; Published: date}
+\history{Received: 21 December 2019; Accepted: 25 February 2020; Published: date}
 %\updates{yes} % If there is an update available, un-comment this line
@@
 % Full title of the paper (Capitalized)
-\Title{A Responsible Machine Learning Workflow\\\normalsize{With Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}
+\Title{A Responsible Machine Learning Workflow\\{with Focus on Interpretable Models, Post-hoc Explanation, and Discrimination Testing}}%Attention AE/ME.
+The following layout issues have not been checked by the English Editing Department and must be carefully verified by the AE/Layout Department: All callout issues, bold usage of callouts, and references to callouts in the text.
+Correct callout usage in figures.
+Figure and Table layout issues.
+Footnote formatting and Glossaries have not been checked.
+En dash usage for negative values, en dash usage to indicate relationships, en dash usage to indicate bonds (especially in chemistry).
+The English Editing Department is not responsible for correct italic usage for genes, proteins and technical terminology.
+This responsibility belongs to the authors.
+The following are also not checked: spacing between numbers and units of measurement, ratios, en dashes for ranges, date and time formats, punctuation in equation lines, and less than/more than spacing (< >).
+Finally, capitalization and layout of titles/headings must be properly checked as well as ensuring 'Eq.' and 'Fig.' are properly spelled out, as these are layout issues.
 
 % Authors, for the paper (add full first names)
-\Author{Navdeep Gill $^{1, \dagger}$, Patrick Hall $^{1,3,\dagger,*}$, Kim Montgomery $^{1,\dagger}$, and Nicholas Schmidt $^{2, \dagger, *}$}
+\Author{\hl{Navdeep Gill} $^{1, \dagger}$, Patrick Hall $^{1,2,}$*$^{,\dagger}$, Kim Montgomery $^{1,\dagger}$ and Nicholas Schmidt $^{3,}$*$^{,\dagger}$} %%%Author order is not the same as on the system, please confirm which one is correct.
+%Please carefully check the accuracy of names and affiliations.
 
 % Authors, for metadata in PDF
-\AuthorNames{Navdeep Gill, Patrick Hall, Kim Montgomery, and Nicholas Schmidt}
+\AuthorNames{Navdeep Gill, Patrick Hall, Kim Montgomery and Nicholas Schmidt}
 
@@
 \address{%
-$^{1}$ \quad H2O.ai\\
-$^{2}$ \quad BLDS, LLC\\
-$^{3}$ \quad The George Washington University}
+$^{1}$ \quad \hl{H2O.ai}, 2307 Leghorn St., Mountain View, CA, 94043, USA; navdeep.gill@h2o.ai (N.G.); kim.montgomery@h2o.ai (K.M.)\\%%%Please provide the department, university,zip code, city, country.
+$^{2}$ \quad \hl{The George Washington University}, Department of Decision Sciences, 2201 G Street NW,
+Washington, D.C., 20052, USA\\%%%Please provide the department, zip code, city, country.
+$^{3}$ \quad \hl{BLDS, LLC}, 1608 Walnut St., Suite 1108, Philadelphia, PA, 19103, USA}%%%Please provide the department, university,zip code, city, country.
 
 % Contact information of the corresponding author
-\corres{Correspondence: phall@h2o.ai; nschmidt@bldsllc.com}
+\corres{Correspondence: phall@h2o.ai (P.H.); nschmidt@bldsllc.com (N.S.)}
 
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{This manuscript outlines a viable approach for training and evaluating machine learning (ML) systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
-The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines (MGBMs) and explainable neural networks (XNNs), a deep learning architecture well-suited for structured data, are assessed on simulated data and publicly available mortgage data.
-For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence (PD) and individual conditional expectation (ICE) and with global and local Shapley feature importance.
-The constrained model predictions are also tested for disparate impact (DI) and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD), along with straightforward group fairness measures.
-By combining interpretable models, post-hoc explanations, and discrimination testing with accessible software tools, this text aims to provide a template workflow for important ML applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
+\abstract{This manuscript outlines a viable approach for training and evaluating machine learning systems for high-stakes, human-centered, or regulated applications using common Python programming tools.
+The accuracy and intrinsic interpretability of two types of constrained models, monotonic gradient boosting machines and explainable neural networks, a deep learning architecture well-suited for structured data, are assessed on simulated data and publicly available mortgage data.
+For maximum transparency and the potential generation of personalized adverse action notices, the constrained models are analyzed using post-hoc explanation techniques including plots of partial dependence and individual conditional expectation and with global and local Shapley feature importance.
+The constrained model predictions are also tested for disparate impact and other types of discrimination using measures with long-standing legal precedents, adverse impact ratio, marginal effect, and standardized mean difference, along with straightforward group fairness measures.
+By combining interpretable models, post-hoc explanations, and discrimination testing with accessible software tools, this text aims to provide a template workflow for machine learning applications that require high accuracy and interpretability and that mitigate risks of discrimination.}
 
 % Keywords
-\keyword{Deep Learning; Disparate Impact; Explanation; Fairness; Gradient Boosting Machine; Interpretable; Machine Learning; Neural Network; Python}
+\keyword{deep learning; disparate impact; explanation; fairness; gradient boosting machine; interpretable; machine learning; neural network; Python}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{section}{-1}
+\setcounter{section}{0}
 
@@
 
-Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds.\footnote{See: \href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}{Responsible Artificial Intelligence}, \href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, \href{https://www.pwc.com/gx/en/issues/data-and-analytics/artificial-intelligence/what-is-responsible-ai.html}{PwC's Responsible AI}, \href{https://ai.google/responsibilities/responsible-ai-practices/}{Responsible AI Practices}.} Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear.
-ML models, a common application of AI, can present serious risks.
-ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{See: \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{\textit{When a Computer Program Keeps You in Jail}}.} ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}, and ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}, \cite{shokri2019privacy}.
-This manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}, \ref{sec:res}, and \ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and DI and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems.\footnote{This text and associated software are not, and should not be construed as, legal advice or requirements for regulatory compliance.}\textsuperscript{,}\footnote{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervisory and Regulatory (SR) Letter 11-7, and the European Union (EU) General Data Protection Regulation (GDPR) Article 22 \cite{ff_interpretability}.}
+Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that~perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds ({\hl{See}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+%is the text color necessary? please confirm if need remove the link of the text,
+%% we have removed the link of text, and keep web link
+: 
+% THESE HYPERLINKS ARE NECESSARY - PH
+\href{https://ec.europa.eu/jrc/communities/sites/jrccties/files/03_dignum_v.pdf}
+{Responsible Artificial Intelligence}, 
+\href{https://www.accenture.com/_acnmedia/pdf-92/accenture-afs-responsible-ai.pdf}
+{\textit{Responsible AI: A Framework for Building Trust in Your AI Solutions}}, 
+\href{https://www.pwc.com/gx/en/issues/data-and-analytics/artificial-intelligence/what-is-responsible-ai.html}
+{PwC's Responsible AI}, 
+\href{https://ai.google/responsibilities/responsible-ai-practices/}
+{Responsible AI Practices}).
+Although responsible AI is today a somewhat broad and amorphous notion, at least one aspect is becoming clear.
+Machine learning (ML) models, a common application of AI, can present serious risks.
+ML~models can be inaccurate and unappealable %INCORRECT - PH -- Please confirm that the intended meaning has been retained.
+black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop} ({\hl{See}% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+: 
+\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}
+{\textit{When a Computer Program Keeps You in Jail}}).
+ML models can perpetuate and exacerbate discrimination \cite{feldman2015certifying,dwork2012fairness,gender_shades}, and ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml,model_stealing,membership_inference,shokri2019privacy}.
+This~manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena.
+However, Sections \ref{sec:m_and_m}--\ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and disparate impact (DI) and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems ({\hl{This text and associated software are not, and should not be construed as, legal advice or regulatory compliance requirements}}), ({\hl{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervision and Regulation (SR) Letter 11-7, the European Union (EU) General Data Protection Regulation (GDPR) Article} 22, or other laws and regulations \cite{ff_interpretability}.})%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
 
 Section \ref{sec:m_and_m} describes methods and materials, including training datasets, interpretable and constrained models, post-hoc explanations, tests for DI and other social discrimination, and public and open source software resources associated with this text.
-In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and post-hoc explanation and discrimination testing results are also presented for interpretable models.
+In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models, and~post-hoc explanation and discrimination testing results are also presented for interpretable models.
 Of course, an even wider array of tools and techniques are likely helpful to fully minimize discrimination, inaccuracy, privacy, and security risks associated with ML models.
@@
 
-The simulated data (see Subsection \ref{ssec:sim_data}) is based on the well-known Friedman datasets.
-Its known feature importance and augmented discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree}, \cite{friedman1991multivariate}.
-The mortgage data (see Subsection \ref{ssec:mort_data}) is sourced from the Home Mortgage Disclosure Act (HMDA) database and is a fairly realistic data source for demonstrating the template workflow.\footnote{See: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}.} To provide a sense of fit differences, performance of more interpretable constrained ML models and less interpretable unconstrained ML models is compared on simulated data and collected mortgage data.
-Because the unconstrained ML models, gradient boosting machines (GBMs, e.g.
-\cite{friedman2001greedy}, \cite{esl}) and artificial neural networks (ANNs, e.g.
-\cite{recht2011hogwild}, \cite{hinton2012improving}, \cite{sutskever2013importance}, \cite{zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present the unmitigated risks discussed above, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop}, \cite{fair_washing}, \cite{scaffolding}.
-Here, MGBMs\footnote{As implemented in \href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}{\texttt{XGBoost}} or \href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}{\texttt{h2o}}.} (See Subsection \ref{ssec:mgbm}) and XNNs (see Subsection \ref{ssec:xnn}, \cite{wf_xnn} \cite{yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
+The simulated data (see Section~\ref{ssec:sim_data}) are based on the well-known Friedman datasets.
+Its~known feature importance and augmented discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree,friedman1991multivariate}.
+The~mortgage data (see Section~\ref{ssec:mort_data}) are sourced from the Home Mortgage Disclosure Act (HMDA) database, a fairly realistic data source for demonstrating the template workflow \cite{hmda} ({\hl{See:}
+\href{https://www.consumerfinance.gov/data-research/hmda/}
+{Mortgage data (HMDA})}). %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+To provide a sense of fit differences, performance is compared on simulated data and collected mortgage data between the more interpretable constrained ML models and the less interpretable unconstrained ML models.
+Because the unconstrained ML models, gradient boosting machines (GBMs, e.g., \cite{friedman2001greedy,esl}) and artificial neural networks (ANNs, e.g., \cite{recht2011hogwild,hinton2012improving,sutskever2013importance,zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present the unmitigated risks discussed above, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop,fair_washing,scaffolding}.
+Here, monotonic gradient boosting machines ({\hl{as implemented in}
+%\href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}
+{\texttt{\hl{XGBoost}}} or 
+%\href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}
+{\texttt{\hl{h2o}}}.}) %% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. % we have removed the link of text, and keep web link
+(MGBMs, see Section~\ref{ssec:mgbm}) and explainable neural networks (XNNs, see Section~\ref{ssec:xnn} \cite{,wf_xnn,yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
 MGBM and XNN interpretable model architectures are selected for the example workflow because they are straightforward variants of popular unconstrained ML models.
-If practitioners are working with GBM and ANN models, it should be relatively uncomplicated to also evaluate the constrained versions of these models.
+If~practitioners are working with GBM and ANN models, it should be relatively uncomplicated to also evaluate the constrained versions of these models.
 
@@
 Due to their post-hoc nature, they can often be shoe-horned into existing ML workflows and pipelines.
-Presented explanation techniques include PD and ICE (see Subsection \ref{ssec:pd_ice}) and Shapley values (see Subsection \ref{ssec:shap}) \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}.
+Presented explanation techniques include partial dependence (PD) and individual conditional expectation (ICE) (see Section~\ref{ssec:pd_ice}) and Shapley values (see Section~\ref{ssec:shap}) \cite{esl,ice_plots,shapley,tree_shap}.
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discrimination testing methods discussed (see Subsection \ref{ssec:di}) include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
-All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Subsection \ref{ssec:soft}).
+Discrimination testing methods discussed (see Section~\ref{ssec:di}) include adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD)  \cite{feldman2015certifying,cohen1988statistical,cohen1992power}.
+({\hl{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures} (1978) \textsection 1607.4.}).
+%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+ Accuracy and other confusion matrix measures are also reported by demographic segment \cite{zafar2017fairness}.
+All outlined materials and methods are implemented in open source Python code, and are made available on GitHub (see Section~\ref{ssec:soft}).
 
@@
 
-Simulated data is created based on a signal-generating function, $f$, applied to input data, $\mathbf{X}$, first proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
-
+Simulated data are created based on a signal-generating function, $f$, applied to input data, $\mathbf{X}$, first~proposed in \citet{friedman1979tree} and extended in \citet{friedman1991multivariate}:
 \begin{equation}
@@
 \end{equation}
-
-\noindent where each $X_{\text{Friedman},j}$ is a random uniform feature in $[0,1]$.
+where each $X_{\text{Friedman},j}$ is a random uniform feature in $[0,1]$.
 In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies.
@@
 Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing.
-The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 instances.
-The simulated data is then split into a training and test set, with 80,000 and 20,000 instances, respectively.
-Within the training set, a 5-fold cross-validation indicator is used for training all models.
-For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.
+The simulated data are generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 instances.
+The simulated data are then split into a training and test set, with 80,000 and 20,000 instances, respectively.
+Within the training set, a~five-fold cross-validation indicator is used for training all models.
+For an exact specification of the simulated data, see the software resources referenced in Section~\ref{ssec:soft}.
 
@@
 A selection of features is used to predict whether a loan is \textit{high-priced}, i.e., the annual percentage rate (APR) charged was 150 basis points (1.5\%) or more above a survey-based estimate of other similar loans offered around the time of the given loan.
-After data cleaning and preprocessing to encode categorical features and create missing markers, the mortgage data contains ten input features and the binary outcome, \textit{high-priced}.
-The data is split into a training set with 160,338 loans and a marker for 5-fold cross-validation and a test set containing 39,662 loans.
+After data cleaning and preprocessing to encode categorical features and create missing markers, the mortgage data contain ten input features and the binary outcome, \textit{high-priced}.
+The data are split into a training set with 160,338 loans and a marker for 5-fold cross-validation and a test set containing 39,662 loans.
 While lenders would almost certainly use more information than the selected features to determine whether to offer and originate a high-priced loan, the selected input features (loan to value (LTV) ratio, debt to income (DTI) ratio, property value, loan amount, introductory interest rate, customer income, etc.) are likely to be some of the most influential factors that a lender would consider.
@@
 An MGBM remains an additive combination of $B$ trees trained by gradient boosting, $T_b$, and each tree learns a set of splitting rules that respect monotonicity constraints,  $\Theta^\text{mono}_b$.
-For some instance, $\mathbf{x}$, a trained MGBM model, $g^{\text{MGBM}}$, takes the form:
+For an instance, $\mathbf{x}$, a trained MGBM model, $g^{\text{MGBM}}$, takes the form:
 
@@
 
-\noindent As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x};\Theta^\text{mono}_b)$, in a numeric loss function (e.g., squared loss, Huber loss), and a regularization term that penalizes complexity in the current tree.
-See Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
+ As in unconstrained GBM, $\Theta^{\text{mono}}_b$ is selected in a greedy, additive fashion by minimizing a regularized loss function that considers known target labels, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, and the $b$-th tree splits applied to $\mathbf{x}$, $T_b(\mathbf{x};\Theta^\text{mono}_b)$, in a numeric loss function (e.g., squared loss, Huber loss), and a regularization term that penalizes complexity in the current tree.
+See~Appendices \ref{a_ssec:not} and \ref{a_ssec:mgbm} for details pertaining to MGBM training.
 
@@
 One on the simulated data and one on the mortgage data.
-In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and 5-fold cross-validation and test partitions are used for model assessment.
-For exact parameterization of the two $g^\text{MGBM}$ models, see the software resources referenced in Subsection \ref{ssec:soft}.
+In both cases, positive and negative monotonic constraints for each $X_j$ are selected using domain knowledge, random grid search is used to determine other hyperparameters, and five-fold cross-validation and test partitions are used for model assessment.
+For exact parameterization of the two $g^\text{MGBM}$ models, see the software resources referenced in Section~\ref{ssec:soft}.
 
@@
 XNNs are an alternative formulation of additive index models in which the ridge functions are neural networks \cite{wf_xnn}.
-XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}Ms), which consider main effects and a small number of 2-way interactions and may also incorporate boosting into their training \cite{esl}, \cite{ga2m}.
+XNNs also bear a strong resemblance to generalized additive models (GAMs) and so-called explainable boosting machines (EBMs or GA\textsuperscript{2}Ms), which consider main effects and a small number of two-way interactions and may also incorporate boosting into their training \cite{esl,ga2m}.
 XNNs enable users to tailor interpretable neural network architectures to a given prediction problem and to visualize model behavior by plotting ridge functions.
 A trained XNN function, $g^{\text{XNN}}$, applied to some instance , $\mathbf{x}$, is defined as:
-
 \begin{equation}
@@
 \end{equation}
-
-\noindent where $\mu_0$ is a global bias for $K$ individually specified ANN subnetworks, $n_k$, with weights $\gamma_k$.
-The inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest, i.e., \textit{projection}, layer of $g^\text{XNN}$.
+where $\mu_0$ is a global bias for $K$ individually specified ANN subnetworks, $n_k$, with weights $\gamma_k$.
+The~inputs to each $n_k$ are themselves a linear combination of the $J$ modeling inputs and their associated $\beta_{k,j}$ coefficients in the deepest, i.e., {projection}, layer of $g^\text{XNN}$.
 
 Two $g^\text{XNN}$ models are trained by mini-batch stochastic gradient descent (SGD) on the simulated data and mortgage data.
-Each $g^\text{XNN}$ is assessed in 5 training folds and in a test data partition.
-$L_1$ regularization is applied to network weights to induce a sparse and interpretable model, where each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
-The $g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size.
-Be aware that $g^{XNN}$ model architectures may require manual and judicious feature selection due to long training times.
-For more details regarding $g^\text{XNN}$ training, see the software resources in Subsection \ref{ssec:soft} and Appendices  \ref{a_ssec:not} and \ref{a_ssec:xnn}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{One-dimensional Partial Dependence and Individual Conditional Expectation}\label{ssec:pd_ice}
+Each $g^\text{XNN}$ is assessed in five training folds and in a test data partition.
+$L_1$ regularization is applied to network weights to induce a sparse and interpretable model, where~each $n_k$ and corresponding $\gamma_k$ are ideally associated with an important $X_j$ or combination thereof.
+$g^\text{XNN}$ models appear highly sensitive to weight initialization and batch size.
+Be aware that $g^{XNN}$ architectures may require manual and judicious feature selection due to long training times.
+For more details regarding $g^\text{XNN}$ training, see the software resources in Section~\ref{ssec:soft} and Appendices  \ref{a_ssec:not} and \ref{a_ssec:xnn}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{One-Dimensional Partial Dependence and Individual Conditional Expectation}\label{ssec:pd_ice}
 
@@
 PD and ICE can be overlaid in the same plot to create a holistic global and local portrait of the predictions for some $g$ and $X_j$ \cite{ice_plots}.
-When $\text{PD}(X_j, g)$ and $\text{ICE}(x_j,  g)$ curves diverge, such plots can also be indicative of modeled interactions in $g$ or expose flaws in PD estimation, e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots}, \cite{ale_plot}.
-For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Subsection \ref{ssec:soft} and Appendices \ref{a_ssec:not} and \ref{a_ssec:pd_ice}.
+When $\text{PD}(X_j, g)$ and $\text{ICE}(x_j,  g)$ curves diverge, such plots can also be indicative of modeled interactions in $g$ or expose flaws in PD estimation, e.g., inaccuracy in the presence of strong interactions and correlations \cite{ice_plots,ale_plot}.
+For details regarding the calculation of one-dimensional PD and ICE, see the software resources in Section~\ref{ssec:soft} and Appendices~\ref{a_ssec:not} and \ref{a_ssec:pd_ice}.
 
@@
 
-Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley}, \cite{shapley1988shapley}.
-Shapley explanations are the only known locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to the model's prediction, $g(\mathbf{x})$, for any instance $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some instance of $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley}, \cite{tree_shap}.
+Shapley explanations are a class of additive, locally accurate feature contribution measures with long-standing theoretical support \cite{shapley,shapley1988shapley}.
+Shapley explanations are the only known locally accurate and globally consistent feature contribution values, meaning that Shapley explanation values for input features always sum to the model's prediction, $g(\mathbf{x})$, for any instance $\mathbf{x}$, and that Shapley explanation values should not decrease in magnitude for some instance of $x_j$ when $g$ is changed such that $x_j$ truly makes a stronger contribution to $g(\mathbf{x})$ \cite{shapley,tree_shap}.
 Shapley values can be estimated in different ways, many of which are intractable for datasets with large numbers of input features.
-Tree SHAP (SHapley Additive exPlanations) is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
-Tree SHAP has been implemented in popular gradient boosting libraries such as \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}{\texttt{h2o}}, \href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}{\texttt{LightGBM}}, and \href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}{\texttt{XGBoost}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
+Tree~Shapley Additive Explanations (SHAP) is a specific implementation of Shapley explanations that relies on traversing internal decision tree structures to efficiently estimate the contribution of each $x_j$ for some $g(\mathbf{x})$ \cite{tree_shap}.
+Tree SHAP has been implemented in popular gradient boosting libraries such as
+%\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}
+{\texttt{\hl{h2o}}}, 
+%\href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}
+{\texttt{\hl{LightGBM}}}, and 
+%\href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}
+{\texttt{\hl{XGBoost}}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}.
-Deep SHAP is implemented in the \href{https://github.com/slundberg/shap}{\texttt{shap}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
-For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.
+Deep SHAP is implemented in the
+%\href{https://github.com/slundberg/shap}
+{\texttt{\hl{shap}}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
+For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.%% we have removed the link of text, and keep web link
 
@@
 
-Because many current technical discussions of fairness in ML appear inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
-The discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
-See Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings, and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
-statistical significance for discrimination measures.
-One such common measure of DI used in US litigation and regulatory settings is ME.
-ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.
-
+Because many current technical discussions of fairness in ML appear inconclusive ({See:} 
+\href{https://www.youtube.com/watch?v=jIXIuYdnyyk}
+{Tutorial: 21 Fairness Definitions and Their Politics}.)%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
+The~discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
+(See~Appendix \ref{a_sec:types} for a brief discussion regarding different types of discrimination in US legal and regulatory settings, and Appendix \ref{a_sec:prac_v_stat} for remarks on practical vs.
+statistical significance for discrimination measures.) One such common measure of DI used in US litigation and regulatory settings is ME.
+ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome:
 \begin{equation}
@@
 \end{equation}
-
-\noindent where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
-ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important DI measure is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.
+ where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
+ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.
+({See:
+\href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}
+{Supervisory Highlights, Issue 9, Fall 2015}.) %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+ Another important DI measure is AIR, more~commonly known as a {relative risk ratio} in settings outside of regulatory compliance.
 
@@
 \end{equation}
-
-\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome and the proportion of the control class that receives a favorable outcome.
-Statistically significant AIR values below 0.8 can be considered \textit{prima facie} evidence of discrimination.
+ AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome and the proportion of the control class that receives a favorable outcome.
+Statistically significant AIR values below 0.8 can be considered {prima facie} evidence of discrimination.
 An additional long-standing and pertinent measure of DI is SMD.
-SMD is often used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending.
-It originates from work on statistical power, and is more formally known as \textit{Cohen’s d}.
-SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.} Cohen defined values of this measure to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
+SMD is often used to assess disparities in continuous features, such~as income differences in employment analyses, or interest rate differences in lending.
+It originates from work on statistical power, and is more formally known as {Cohen’s d}.
+SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.
+({\hl{There are several measures of the standard deviation of the score that are typically used}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+: \textbf{\hl{1}.} the standard deviation of the population, irrespective of protected class status, \textbf{\hl{2}.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{\hl{3}.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.}) Cohen defined values of this measure to have {small}, {medium}, and {large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
+%is bold necessary?
 
@@
 
-\noindent The numerator in the SMD is roughly equivalent to ME but adds the standard deviation divisor as a standardizing factor.
+ The numerator in the SMD is roughly equivalent to ME but adds the standard deviation divisor as a standardizing factor.
 Because of this standardization factor, SMD allows for a comparison across different types of outcomes, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.
-In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail in Appendix \ref{a_sec:prac_v_stat}.
-Finally, confusion matrix measures in demographic groups, such as accuracy, false positive rate (FPR), false negative rate (FNR), and their ratios, are also considered as measures of DI in Subsection \ref{ssec:dis_mort} and Appendix \ref{a_ssec:dis_sim}.
+In this, one may apply definitions in \citet{cohen1988statistical} of {small}, {medium}, and~{large} effect sizes, which represent a measure of {practical significance}, which is described in detail in Appendix \ref{a_sec:prac_v_stat}.
+Finally, confusion matrix measures in demographic groups, such as accuracy, false~positive rate (FPR), false negative rate (FNR), and their ratios, are also considered as measures of DI in Section~\ref{ssec:dis_mort} and Appendix \ref{a_ssec:dis_sim}.
 
@@
 
-Python code to reproduce discussed results is available at: \texttt{\url{https://github.com/h2oai/article-information-2019}}.
-The primary Python packages employed are: \href{https://numpy.org/}{\texttt{numpy}} and \href{https://pandas.pydata.org/}{\texttt{pandas}} for data manipulation, \href{https://github.com/h2oai/h2o-3}{\texttt{h2o}}, \href{https://keras.io/}{\texttt{keras}}, \href{https://github.com/slundberg/shap}{\texttt{shap}}, and \href{https://www.tensorflow.org/}{\texttt{tensorflow}} for modeling, explanation, and discrimination testing, and typically \href{https://matplotlib.org/}{\texttt{matplotlib}} for plotting.
+Python code to reproduce discussed results is available at: {\url{https://github.com/h2oai/article-information-2019}}. The primary Python packages employed are:  
+%\href{https://numpy.org/}
+{\texttt{\hl{numpy}}} 1.14.5 and 
+%\href{https://pandas.pydata.org/}
+{\texttt{\hl{pandas}}} 0.22.0 for data manipulation, 
+%\href{https://github.com/h2oai/h2o-3}
+{\texttt{\hl{h2o}}} 3.26.0.9, 
+%\href{https://keras.io/}
+{\texttt{\hl{Keras}}} 2.3.1, 
+%\href{https://github.com/slundberg/shap}
+{\texttt{\hl{shap}}} 0.31.0, and 
+%\href{https://www.tensorflow.org/}
+{\texttt{\hl{tensorflow}}} 1.14.0 for modeling, explanation, and discrimination testing, and typically 
+%\href{https://matplotlib.org/}
+{\texttt{\hl{matplotlib}}} 2.2.2 for plotting.
+
+% we have removed the link of text, and keep web link
 
@@
 
-Fit comparisons between unconstrained and constrained models and XNN interpretability results are discussed in Subsections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
+Fit comparisons between unconstrained and constrained models and XNN interpretability results are discussed in Sections \ref{ssec:c_v_uc_sim} and \ref{ssec:xnn_sim}.
 As model training and assessment on the simulated data is a rough validation exercise meant to showcase expected results on data with known characteristics, and given that most of the techniques in the proposed workflow are already used widely or have been validated elsewhere, reporting of simulated data results in the main text will focus mostly on fit measures and the more novel $g^\text{XNN}$ interpretability results.
@@
 Table \ref{tab:sim_acc} presents a variety of fit measures for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data.
-$g^\text{XNN}$ exhibits the best performance, but the models exhibit only a fairly small range of fit results.
-Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
-For the displayed measures, $g^\text{MGBM}$ performs $\sim$2\% worse on average than $g^\text{GBM}$.
-$g^\text{XNN}$ performs $\sim$0.5\% better on average than $g^\text{XNN}$, and $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across all fit measures except specificity.
-Fit measures that require a probability cutoff are taken at the best F1 threshold for each model.
-
-\begin{table}[H]
+$g^\text{XNN}$ exhibited the best performance, but the models exhibited only a fairly small range of fit results.
+Interpretability and explainability benefits of the constrained models appeared to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all.
+For the displayed measures, $g^\text{MGBM}$ performed $\sim$2\% worse on average than $g^\text{GBM}$.
+$g^\text{XNN}$ performed $\sim$0.5\% better on average than $g^\text{XNN}$, and $g^\text{XNN}$ actually showed slightly better fit than $g^\text{ANN}$ across all fit measures except specificity.
+Fit measures that required a probability cutoff were taken at the best F1 threshold for each model.
+
+\begin{table}[H]%% we have removed the link of text
 	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.
-Arrows indicate the direction of improvement for each measure.}
+Arrows indicate the direction of improvement for each measure and the best result in each column is displayed in bold font.}
 	\centering
@@
 		\textbf{Model} & 
-		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ &
-		\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
+		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}
+		{\textbf{\hl{Accuracy}}}~$\uparrow$ & 
+		%\href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}
+		{\textbf{\hl{AUC}}}~$\uparrow$ & 
+		%\href{https://en.wikipedia.org/wiki/F1_score}
+		{\textbf{\hl{F1}}}~$\uparrow$ & 
+		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}
+		{\textbf{\hl{Logloss}}}~$\downarrow$ & 
+		%\href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}
+		{\textbf{\hl{MCC}}}~$\uparrow$ & 
+		%\href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}
+		{\textbf{\hl{RMSE}}}~$\downarrow$ & 
+		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
+		{\textbf{\hl{Sensitivity}}}~$\uparrow$ & 
+		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
+		{\textbf{\hl{Specificity}}}~$\uparrow$ \\
 		\midrule
-		$g^\text{GBM}$		& 0.757 	& 0.847 	& 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & 0.657 \\
+		$g^\text{GBM}$		& 0.757 & 0.847 & 0.779 & 0.486 & 0.525 & 0.400 & 0.858 & \textbf{0.657} \\
 		$g^\text{MGBM}$		& 0.744 & 0.842 & 0.771 & 0.502 & 0.504 & 0.407 & 0.864 & 0.625 \\
-		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & 0.657 \\
-		$g^\text{XNN}$		& 0.758	& 0.851	& 0.781 & 0.479 & 0.528 & 0.397 & 0.867 & 0.648 \\
+		$g^\text{ANN}$		& 0.757	& 0.850	& 0.779 & 0.480 & 0.525 & 0.398 & 0.858 & \textbf{0.657} \\
+		$g^\text{XNN}$		& \textbf{0.758}	& \textbf{0.851}	& \textbf{0.781} & \textbf{0.479} & \textbf{0.528} & \textbf{0.397} & \textbf{0.867} & 0.648 \\
 		\bottomrule
@@
 
-For $g^\text{XNN}$, inherent interpretability manifests as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_{j,k}$ weights in the bottom projection layer.
-Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into the structure of $g^\text{XNN}$ (also described in Equation \ref{eq:xnn}).
-\ref{fig:sim_xnn_glob_ridge}\textbf{a} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ are associated with large magnitude weights.
-The $n_k$ subnetwork ridge functions appear in \ref{fig:sim_xnn_glob_ridge}\textbf{b} as simplistic but distinctive functional forms.
-Color-coding between \ref{fig:sim_xnn_glob_ridge}\textbf{a} and \ref{fig:sim_xnn_glob_ridge}\textbf{b} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.
+For $g^\text{XNN}$, inherent interpretability manifested as plots of sparse $\gamma_k$ output layer weights, $n_k$ subnetwork ridge functions, and sparse $\beta_{j,k}$ weights in the bottom projection layer.
+Figure \ref{fig:sim_xnn_glob_ridge} provides detailed insights into the structure of $g^\text{XNN}$ (also described in Equation (\ref{eq:xnn})).
+Figure \ref{fig:sim_xnn_glob_ridge}{a} displays the sparse $\gamma_k$ weights of the output layer, where only $n_k$ subnetworks with $k \in \{1, 4, 7, 8, 9\}$ were associated with large magnitude weights.
+The $n_k$ subnetwork ridge functions appear in Figure \ref{fig:sim_xnn_glob_ridge}{b} as simplistic but distinctive functional forms.
+Color-coding between Figure \ref{fig:sim_xnn_glob_ridge}{a},{b} visually reinforces the direct feed-forward relationship between the $n_k$ subnetworks and the $\gamma_k$ weights of the output layer.% we have removed the link of text
 
 \begin{figure}[H]
-	\hskip-35pt\includegraphics[width=19cm]{img/sim_xnn_glob_ridge.png}
+\centering
+	%\includegraphics[width=15cm]{img/sim_xnn_glob_ridge.png}
+	\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
 	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ subnetwork ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the simulated data.}
@@
 
-\noindent $n_k$ subnetworks are plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between \ref{fig:sim_xnn_glob_ridge}\textbf{b} and \ref{fig:sim_xnn_glob_ridge}\textbf{c} link the $\beta_{j,k}$ weights to their $n_k$ subnetworks.
-Most of the heavily utilized $n_k$ subnetworks have sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
-In particular, subnetwork $n_1$ appears to be almost solely a function of $X_{\text{Friedman}, 3}$ and appears to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
-Subnetworks $n_7$, $n_8$, and $n_9$ appear to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and these subnetworks, especially $n_7$ and $n_8$, appear to display some noticeable sinusoidal characteristics.
-Subnetwork $n_4$ seems to be a linear combination of all the original input $X_j$ features, but does weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
-As a whole, Figure \ref{fig:sim_xnn_glob_ridge}\textbf{a}, \textbf{b}, and \textbf{c} exhibit evidence that $g^\text{XNN}$ has learned about the signal-generating function in Equation \ref{eq:sim} and the displayed information should help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
+  $n_k$ subnetworks were plotted across the output values of their associated $\sum_j\beta_{k,j}x_j$ projection layer hidden units, and color-coding between Figure \ref{fig:sim_xnn_glob_ridge}{b},c link the $\beta_{j,k}$ weights to their $n_k$ subnetworks.
+Most of the heavily utilized $n_k$ subnetworks had sparse weights in their $\sum_j\beta_{k,j}x_j$ projection layer hidden units.
+In particular, subnetwork $n_1$ appeared to be almost solely a function of $X_{\text{Friedman}, 3}$ and appeared to exhibit the expected quadratic behavior for $X_{\text{Friedman}, 3}$.
+Subnetworks $n_7$, $n_8$, and $n_9$ appeaed to be most associated with the globally important $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ features, likely betraying the effort required for $g^{\text{XNN}}$ to model the nonlinear $\text{sin}()$ function of the $X_{\text{Friedman}, 1}$ and $X_{\text{Friedman}, 2}$ product, and these subnetworks, especially $n_7$ and $n_8$, appeared to display some noticeable sinusoidal characteristics.
+Subnetwork $n_4$ seemed to be a linear combination of all the original input $X_j$ features, but did weigh the linear $X_{\text{Friedman},4}$ and $X_{\text{Friedman},5}$ terms roughly in the correct two-to-one ratio.
+As~a whole, Figure \ref{fig:sim_xnn_glob_ridge}a--c exhibited evidence that $g^\text{XNN}$ learned about the signal-generating function in Equation (\ref{eq:sim}) and the displayed information should help practitioners understand which original input $X_j$ features were weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$ output.
 See Appendix \ref{a_ssec:xnn} for additional details regarding general XNN architecture.
@@
 
-Results for the mortgage data are presented in Subsections \ref{ssec:c_v_uc_mort} -- \ref{ssec:dis_mort} to showcase the example workflow.
-$g^\text{ANN}$ and $g^\text{XNN}$ outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Subsection \ref{ssec:c_v_uc_sim}, the constrained variants of both model architectures do not show large differences in model fit with respect to unconstrained variants.
-Assuming that in high-stakes applications small fit differences on static test data do not outweigh the need for enhanced model debugging facilitated by high interpretability, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
+Results for the mortgage data are presented in Sections \ref{ssec:c_v_uc_mort}---\ref{ssec:dis_mort} to showcase the example workflow.
+$g^\text{ANN}$ and $g^\text{XNN}$ outperformed $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data, but as in Section~\ref{ssec:c_v_uc_sim}, the constrained variants of both model architectures did not show large differences in model fit with respect to unconstrained variants.
+Assuming that in high-stakes applications small fit differences on static test data did not outweigh the need for enhanced model debugging facilitated by high interpretability, only $g^\text{MGBM}$ and $g^\text{XNN}$ interpretability, post-hoc explainability, and discrimination testing results are presented.
 
@@
 
-Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data for most of the fit measures.
-This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g., numerical encoding of categorical features and missing values.
+Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperformed $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data for most of the fit measures.
+This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and TensorFlow, e.g., numerical encoding of categorical features and missing values.
 This preprocessing appears to hamstring some of the tree-based models' inherent capabilities.
-$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced area under the curve (AUC) values of $\sim$0.81 (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
-
-\begin{table}[htb]
+$g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced receiver operating characteristic area under the curve (AUC) values of $\sim$0.81 (not shown, but available in resources discussed in Section~\ref{ssec:soft}).
+
+\begin{table}[H]%% we have removed the link of text
 	\caption{Fit measures for $g^\text{GBM}(\mathbf{X})$, $g^\text{MGBM}(\mathbf{X})$, $g^\text{ANN}(\mathbf{X})$, and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.
-Arrows indicate the direction of improvement for each measure.}
+Arrows indicate the direction of improvement for each measure and the best result in each column is displayed in bold font.}
 \centering
@@
 	\textbf{Model} & 	
-	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}{\textbf{Accuracy}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}{\textbf{AUC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/F1_score}{\textbf{F1}}~$\uparrow$ &
-	\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}{\textbf{Logloss}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}{\textbf{MCC}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}{\textbf{RMSE}}~$\downarrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Sensitivity}}~$\uparrow$ & \href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}{\textbf{Specificity}}~$\uparrow$ \\
+	%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}
+	{\textbf{\hl{Accuracy}}}~$\uparrow$ & 
+	%\href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}
+	{\textbf{\hl{AUC}}}~$\uparrow$ & 
+	%\href{https://en.wikipedia.org/wiki/F1_score}
+	{\textbf{\hl{F1}}}~$\uparrow$ & 
+	%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}
+	{\textbf{\hl{Logloss}}}~$\downarrow$ & 
+	%\href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}
+	{\textbf{\hl{MCC}}}~$\uparrow$ & 
+	%\href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}
+	{\textbf{\hl{RMSE}}}~$\downarrow$ & 
+	%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
+	{\textbf{\hl{Sensitivity}}}~$\uparrow$ & 
+	%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
+	{\textbf{\hl{Specificity}}}~$\uparrow$ \\
 	\midrule	
 $g^\text{GBM}$		& 0.795	& 0.828	& 0.376 & 0.252 & 0.314 & 0.276 & 0.634 & 0.813 \\
-$g^\text{MGBM}$	& 0.765	& 0.814	& 0.362 & 0.259 & 0.305 & 0.278 & 0.684 & 0.773 \\
-$g^\text{ANN}$		& 0.865	& 0.871	& 0.474 & 0.231 & 0.418 & 0.262 & 0.624 & 0.891 \\
-$g^\text{XNN}$		& 0.869	& 0.868	& 0.468 & 0.233 & 0.409 & 0.263 & 0.594 & 0.898 \\
+$g^\text{MGBM}$		& 0.765	& 0.814	& 0.362 & 0.259 & 0.305 & 0.278 & \textbf{0.684} & 0.773 \\
+$g^\text{ANN}$		& 0.865	& \textbf{0.871} & \textbf{0.474} & \textbf{0.231} & \textbf{0.418} & \textbf{0.262} & 0.624 & 0.891 \\
+$g^\text{XNN}$		& \textbf{0.869} & 0.868 & 0.468 & 0.233 & 0.409 & 0.263 & 0.594 & \textbf{0.898} \\
 \bottomrule
@@
 
-\noindent Regardless of the fit differences between the two families of models, the difference between the constrained and unconstrained variants within the two types of models is small for the GBMs and smaller for the ANNs, $\sim$3.5\% and $\sim$1\% worse fit respectively, averaged across the measures in Table \ref{tab:mort_acc}.
+ Regardless of the fit differences between the two families of models, the difference between the constrained and unconstrained variants within the two types of models is small for the GBMs and smaller for the ANNs, $\sim$3.5\% and $\sim$1\% worse fit respectively, averaged across the measures in Table \ref{tab:mort_acc}.
 
@@
 
-For $g^\text{MGBM}(\mathbf{X})$, intrinsic interpretability is evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis is used to create global and local feature importance.
+For $g^\text{MGBM}(\mathbf{X})$, intrinsic interpretability was evaluated with PD and ICE plots of mostly monotonic prediction behavior for several important $X_j$, and post-hoc Shapley explanation analysis was used to create global and local feature importance.
 Global Shapley feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data is reported in Figure \ref{fig:mort_mgbm_glob}.
-$g^\text{MGBM}$ places high importance on LTV ratio, perhaps too high, and also weighs DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
+$g^\text{MGBM}$ placed high importance on LTV ratio, perhaps too high, and also weighed DTI ratio, property value, loan amount, and introductory rate period heavily in many of its predictions.
 Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the mortgage test data in the $g^\text{MGBM}(\mathbf{X})$ margin space.
-The potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the de-emphasis of the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}(\mathbf{X})$ as compared to $g^\text{XNN}(\mathbf{X})$.
-
-\begin{figure}[htb]
-\centering
-\captionsetup{width=17cm}
-\includegraphics[width=10cm]{img/mort_mgbm_glob.png}
-\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
+The~potential over-emphasis of LTV ratio, and the de-emphasis of income, likely an important feature from a business perspective, and the de-emphasis of the encoded no introductory rate period flag feature may also contribute to the decreased performance of $g^\text{MGBM}(\mathbf{X})$ as compared to $g^\text{XNN}(\mathbf{X})$.
+
+\begin{figure}[H]
+\centering
+%\captionsetup{width=17cm}
+%\includegraphics[width=15cm]{img/mort_mgbm_glob.png}
+\includegraphics[width=12cm]{img/mort_mgbm_glob.png}
+\caption{Global mean absolute Tree Shapley Additive Explanations (SHAP) feature importance for  $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data.}
 \label{fig:mort_mgbm_glob}
@@
 Domain knowledge was used to positively constrain DTI ratio and LTV ratio and to negatively constrain income and the loan term flag under $g^\text{MGBM}$.
-The monotonicity constraints for DTI ratio and LTV ratio are confirmed for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data in Figure \ref{fig:mort_mgbm_glob_pdp_ice}.
-Both DTI ratio and LTV ratio display positive monotonic behavior at all selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for ICE and on average with PD.
-Because PD curves generally follow the patterns of the ICE curves for both features, it is also likely that no strong interactions are at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
-Of course, the monotonicity constraints themselves can dampen the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they do exist in the training data (e.g., LTV ratio and the no introductory rate period flag, see Figure \ref{fig:mort_xnn_glob_ridge}).
-This rigidity could also play a role in the performance differences between $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ in the mortgage data not observed for the simulated data, wherein strong interactions appear to be between features with the same monotonicity constraints (e.g., $X_{\text{Friedman},1}$ and $X_{\text{Friedman},2}$, see Figure \ref{fig:sim_xnn_glob_ridge}).
+The monotonicity constraints for DTI ratio and LTV ratio were confirmed for $g^\text{MGBM}(\mathbf{X})$ on the mortgage test data in Figure \ref{fig:mort_mgbm_glob_pdp_ice}.
+Both DTI ratio and LTV ratio displayed positive monotonic behavior at all selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for ICE and on average with PD.
+Because PD curves generally followed the patterns of the ICE curves for both features, it is also likely that no strong interactions were at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
+Of course, the monotonicity constraints themselves could have dampened the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they did exist in the training data (e.g., LTV ratio and the no introductory rate period flag, see\hl{ Figure} \ref{fig:mort_xnn_glob_ridge}).
+This rigidity could also have played a role in the performance differences between $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ in the mortgage data not observed for the simulated data, wherein strong interactions appeared to be between features with the same monotonicity constraints (e.g., $X_{\text{Friedman},1}$ and $X_{\text{Friedman},2}$, see Figure \ref{fig:sim_xnn_glob_ridge}).%Figures  should be cited in sequential numerical order.
+Please confirm if it's invalid citation, if no, please revise the order.
 
@@
 Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.
-DTI ratio and LTV ratio do appear to have sparse regions in their univariate distributions.
+The DTI ratio and LTV ratio do appear to have had sparse regions in their univariate distributions.
 The monotonicity constraints likely play to the advantage of $g^\text{MGBM}$ in this regard, as $g^\text{MGBM}(\mathbf{X})$ appears to carry reasonable predictions learned from dense domains into the sparse domains of both features.
 
-\begin{figure}[htb]
-\hskip-10pt\includegraphics[width=16.5cm]{img/mort_mgbm_glob_pdp_ice.png}
-\caption{PD, ICE for 10 individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
+\begin{figure}[H]
+\centering
+ \includegraphics[width=15cm]{img/mort_mgbm_glob_pdp_ice.png}
+\caption{PD, ICE for ten individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the mortgage test data.}
 \label{fig:mort_mgbm_glob_pdp_ice}
@@
 Figure \ref{fig:mort_mgbm_glob_pdp_ice} also displays PD and ICE for the unconstrained feature property value.
-Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns established by ICE curves.
-While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$ and lower exhibit different prediction behavior with respect to property value.
-Some individuals at these lower percentiles display monotonically decreasing prediction behavior, while others appear to show fluctuating prediction behavior.
-Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn.
-For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions.
-However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and appear in a sparse region of property value.
-This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
-Fluctuations in ICE could also be caused by overfitting or by leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
-
-In Figure \ref{fig:mort_mgbm_loc}, local Tree SHAP values are displayed for selected individuals at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in the mortgage test data.
+Unlike the DTI ratio and LTV ratio, PD for property value did not always follow the patterns established by ICE curves.
+While PD showed monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$ and lower exhibited different prediction behavior with respect to property value.
+Some individuals at these lower percentiles displayed monotonically decreasing prediction behavior, while others appeared to show fluctuating prediction behavior.
+Property value was strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn.
+For the most part, reasonable predictions did appear to be carried from more densely populated regions to more sparsely populated regions.
+However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles were visible, and appeared in a sparse region of property value.
+This divergence of PD and ICE could be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Section~\ref{ssec:soft}).
+Fluctuations in ICE could also have been caused by overfitting or by leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.
+
+In Figure \ref{fig:mort_mgbm_loc}, local Tree SHAP values are displayed for selected individuals at the 10{th}, 50{th}, and 90{th} percentiles of $g^{\text{MGBM}}(\mathbf{X})$ in the mortgage test data.
 Each Shapley value in Figure \ref{fig:mort_mgbm_loc} represents the difference in $g^\text{MGBM}(\mathbf{x})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with this instance of some input feature $x_j$ \cite{molnar}.
@@
 
-\begin{figure}[htb]
-\centering
-\includegraphics[width=16cm]{img/mort_mgbm_loc.png}
+\begin{figure}[H]
+\centering
+\includegraphics[width=15cm]{img/mort_mgbm_loc.png}
 \caption{Tree SHAP values for three individuals across selected percentiles of $g^\text{MGBM}(\mathbf{X})$ for the mortgage test data.}
@@
 
-\noindent The selected individuals show an expected progression of mostly negative Shapley values at the 10\textsuperscript{th} percentile, a mixture of positive and negative Shapley values at the 50\textsuperscript{th} percentile, mostly positive Shapley values at the 90\textsuperscript{th} percentile, and with globally important features driving most local model decisions.
+  The selected individuals showed an expected progression of mostly negative Shapley values at the 10{th} percentile, a mixture of positive and negative Shapley values at the 50{th} percentile, mostly positive Shapley values at the 90{th} percentile, and with globally important features driving most local model decisions.
 Deeper significance for Figure \ref{fig:mort_mgbm_loc} lies in the ability of Tree SHAP to accurately and consistently summarize any single $g^\text{MGBM}(\mathbf{x})$ prediction in this manner, which is generally important for enabling logical appeal or override of ML-based decisions, and is specifically important in the context of lending, where applicable regulations often require lenders to provide consumer-specific reasons for denying credit to an individual.
@@
 Like $g^\text{MGBM}$, $g^\text{XNN}$ puts little emphasis on the other flag features.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=10cm]{img/mort_xnn_glob.png}
+Unlike $g^\text{MGBM}$, $g^\text{XNN}$ assigned higher importance to property value, loan amount, and income, and lower importance on LTV ratio and DTI ratio.
+
+\begin{figure}[H]
+\centering
+%\includegraphics[width=15cm]{img/mort_xnn_glob.png}
+\includegraphics[width=12cm]{img/mort_xnn_glob.png}
 \caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
@@
 
-\noindent As compared to $g^\text{MGBM}$, $g^\text{XNN}$ assigns higher importance to property value, loan amount, and income, and lower importance on LTV ratio and DTI ratio.
-
 The capability of $g^\text{XNN}$ to model nonlinear phenomenon and high-degree interactions, and to do so in an interpretable manner, is on display in Figure \ref{fig:mort_xnn_glob_ridge}.
-\ref{fig:mort_xnn_glob_ridge}\textbf{a} presents the sparse $\gamma_k$ weights of the $g^\text{XNN}$ output layer in which the $n_k$ subnetworks with $k \in \{0,1,2,3,5,8,9\}$ have large magnitude weights and $n_k$ subnetworks, $k \in \{4,6,7\}$, have small or near-zero magnitude weights.
-Distinctive ridge functions that feed into those large magnitude $\gamma_k$ weights are highlighted in \ref{fig:mort_xnn_glob_ridge}\textbf{b} and color-coded to pair with their corresponding $\gamma_k$ weight.
-As in the Subsection \ref{ssec:xnn_sim}, $n_k$ ridge function plots vary with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in \ref{fig:mort_xnn_glob_ridge}\textbf{c}.
-In both the simulated and mortgage data, $n_k$ ridge functions appear to be elementary functional forms that the output layer learns to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
-\ref{fig:mort_xnn_glob_ridge}\textbf{c} displays the sparse $\beta_{j,k}$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function.
-For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.
+Figure \ref{fig:mort_xnn_glob_ridge}{a} presents the sparse $\gamma_k$ weights of the $g^\text{XNN}$ output layer in which the $n_k$ subnetworks with $k \in \{0,1,2,3,5,8,9\}$ had large magnitude weights and $n_k$ subnetworks, $k \in \{4,6,7\}$, had small or near-zero magnitude weights.
+Distinctive ridge functions that fed into those large magnitude $\gamma_k$ weights are highlighted in Figure \ref{fig:mort_xnn_glob_ridge}{b} and color-coded to pair with their corresponding $\gamma_k$ weight.
+As in the Section~\ref{ssec:xnn_sim}, $n_k$ ridge function plots varied with the output of the corresponding projection layer $\sum_j\beta_{k,j}x_j$ hidden unit, with weights displayed in matching colors in Figure \ref{fig:mort_xnn_glob_ridge}{c}.
+In both the simulated and mortgage data, $n_k$ ridge functions appeared to be elementary functional forms that the output layer learned to combine to generate accurate predictions, reminiscent of basis functions for the modeled space.
+Figure \ref{fig:mort_xnn_glob_ridge}{c} displays the sparse $\beta_{j,k}$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that were associated with each $n_k$ subnetwork ridge function.
+For instance, subnetwork $n_3$ was influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ was nearly completely dominated by the weight for income.
 See Appendix \ref{a_ssec:xnn} for details regarding general XNN architecture.
 
-\begin{figure}[htb]
-\hskip-35pt\includegraphics[width=19cm]{img/mort_xnn_glob_ridge.png}
+\begin{figure}[H]
+\includegraphics[width=17cm]{img/mort_xnn_glob_ridge.png}
 \caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the mortgage data.}
@@
 
-To compliment the global interpretability of $g^\text{XNN}$, Figure \ref{fig:mort_xnn_loc} displays local Shapley values for selected individuals, estimated from the projection layer using Deep SHAP in the $g^\text{XNN}$ probability space.
+To complement the global interpretability of $g^\text{XNN}$, Figure \ref{fig:mort_xnn_loc} displays local Shapley values for selected individuals, estimated from the projection layer using Deep SHAP in the $g^\text{XNN}$ probability space.
 Similar to Tree SHAP, local Deep SHAP values should sum to $g^\text{XNN}(\mathbf{x})$.
-While the Shapley values appear to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc}, their true value is their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions.
-
-\begin{figure}[htb]
-\centering
-\includegraphics[width=16cm]{img/mort_xnn_loc.png}
+While the Shapley values appeared to follow the roughly increasing pattern established in Figures \ref{fig:sim_mgbm_loc}, \ref{fig:sim_xnn_loc}, and \ref{fig:mort_mgbm_loc}, their true value was their ability to be calculated for any $g^\text{XNN}(\mathbf{x})$ prediction, as a means to summarize model reasoning and allow for appeal and override of specific ML-based decisions.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=15cm]{img/mort_xnn_loc.png}
 \caption{Deep SHAP values for three individuals across selected percentiles of $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
@@
 
-Tables \ref{tab:mort_disc}\textbf{a} and \ref{tab:mort_disc}\textbf{b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
-As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and $g^{\text{XNN}}(\mathbf{X})$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome).\footnote{See Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  are predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score is favorable.
+Table \ref{tab:mort_disc}{a,b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
+As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and $g^{\text{XNN}}(\mathbf{X})$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome)({\hl{See Appendix} \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.).
+%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  were predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score was favorable.
 Thus, one might consider FPR ratios as a measure of the class-control disparities.
-FPR ratios are higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs.
-2.10) in Table \ref{tab:mort_disc}\textbf{b}, but overall FPRs are lower for blacks under $g^{\text{XNN}}$ (0.295 vs.
-0.315) in Table \ref{tab:mort_disc}\textbf{a}.
+FPR ratios were higher under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$ (2.45 vs.
+2.10) in Table \ref{tab:mort_disc}{b}, but overall FPRs were lower for blacks under $g^{\text{XNN}}$ (0.295 vs.
+0.315) in Table \ref{tab:mort_disc}{a}.
 This is the same pattern seen in the simulated data results in Appendix \ref{a_ssec:dis_sim}, leading to the question of whether a fairness goal should not only consider class-control relative rates, but also intra-class improvements in the chosen fairness measure.
-Similar results are found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}(\mathbf{X})$ and 1.21 for $g^{\text{XNN}}(\mathbf{X})$.
-
-\begin{table}[htb]
-	\hskip-10pt\begin{subtable}{0.45\textwidth}
-	\captionsetup{width=7cm}
-	\caption{Group size, accuracy, and FPR for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.} 
-	\centering
-	\resizebox{\columnwidth}{!}{%
+Similar results were found for the female-male comparison, but the relative rates are less stark: 1.15 for $g^{\text{MGBM}}(\mathbf{X})$ and 1.21 for $g^{\text{XNN}}(\mathbf{X})$.
+
+%\begin{table}[H]
+%	\hskip-10pt\begin{subtable}{0.45\textwidth}
+%	\captionsetup{width=7cm}
+%	\caption{Group size, accuracy, and FPR for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.} 
+%	\centering
+%	\resizebox{\columnwidth}{!}{%
+%	\begin{tabular}{ccccc}
+%		\toprule
+%		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy}$\uparrow$ & \textbf{FPR}$\downarrow$ \\
+%		\midrule
+%		Black & 2,608 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} & \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} \\
+%		\midrule			 
+%		White & 28,361 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} & \begin{tabular}{@{}c@{}} 0.150 \\ 0.120\end{tabular}\\
+%		\midrule
+%		Female & 8,301 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} & \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}\\
+%		\midrule
+%		Male & 13,166 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\ 0.847\end{tabular} & \begin{tabular}{@{}c@{}} 0.182 \\ 0.131\end{tabular}\\
+%		\bottomrule
+%		\label{tab:mort_n_acc}
+%	\end{tabular}%
+%	}
+%	\end{subtable}
+%	\hspace{10pt}
+%	\begin{subtable}{0.45\textwidth}
+%	\vspace{-14pt}
+%	\captionsetup{width=10cm}
+%	\caption{%
+%		\leftskip=45pt
+%		AIR, ME, SMD, and FPR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
+%	\centering
+%	\scalebox{0.95}{
+%	\begin{tabular}{ccccccc}
+%		\toprule
+%		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}$\uparrow$ & \textbf{ME}$\downarrow$ & \textbf{SMD}$\downarrow$ & \begin{tabular}[x]{@{}c@{}}\textbf{FPR}\\\textbf{Ratio}\end{tabular}$\downarrow$\\
+%		\midrule
+%		$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}	& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular} & \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular} & \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular} & \begin{tabular}{@{}c@{}} 2.10 \\ 1.15\end{tabular} \\
+%		\midrule 
+%		$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular} & \begin{tabular}{@{}c@{}} White \\ Male\end{tabular} & \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular} & \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular} & \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular} & \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular} \\
+%		\bottomrule
+%		\label{tab:mort_disc_detail}
+%	\end{tabular}
+%	}
+%	\end{subtable}
+%\captionsetup{width=10cm}
+%\caption{Discrimination measures for the mortgage test data. Arrows indicate the direction of improvement for each measure.}
+%\label{tab:mort_disc}
+%\end{table}
+
+\begin{table}[H]
+\centering
+
+\caption{Discrimination measures for the mortgage test data. Arrows indicate the direction of improvement for each measure.}
+\label{tab:mort_disc}
 \begin{tabular}{ccccc}
 		\toprule
+		\multicolumn{5}{l}{\begin{tabular}{@{}c@{}} (\textbf{a}) Group size, accuracy, and FPR for $g^\text{MGBM}(\mathbf{X})$\\ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data. \end{tabular}}\\
+		\midrule
 		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy}$\uparrow$ & \textbf{FPR}$\downarrow$ \\
 		\midrule
-		Black & 2,608 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} & \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} \\
+		Black & 2608 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} & \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} \\
 		\midrule			 
@@
 		\midrule
-		Female & 8,301 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} & \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}\\
+		Female & 8301 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} & \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}\\
 		\midrule
@@
 		\label{tab:mort_n_acc}
-	\end{tabular}%
-	}
-	\end{subtable}
-	\hspace{10pt}
-	\begin{subtable}{0.45\textwidth}
-	\vspace{-14pt}
-	\captionsetup{width=10cm}
-	\caption{%
-		\leftskip=45pt
-		AIR, ME, SMD, and FPR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the mortgage test data.}
-	\centering
-	\scalebox{0.95}{
+	\end{tabular}\\
 \begin{tabular}{ccccccc}
 		\toprule
+		 \multicolumn{7}{l}{\begin{tabular}{@{}c@{}} (\textbf{b}) AIR, ME, SMD, and FPR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$\\ on the mortgage test data. \end{tabular}}   \\
+		\midrule
 		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}$\uparrow$ & \textbf{ME}$\downarrow$ & \textbf{SMD}$\downarrow$ & \begin{tabular}[x]{@{}c@{}}\textbf{FPR}\\\textbf{Ratio}\end{tabular}$\downarrow$\\
@@
 	\end{tabular}
-	}
-	\end{subtable}
-\captionsetup{width=10cm}
-\caption{Discrimination measures for the mortgage test data. Arrows indicate the direction of improvement for each measure.}
-\label{tab:mort_disc}
 \end{table}
 
-Both ME and AIR show higher disparities for blacks under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$.
+Both ME and AIR showed higher disparities for blacks under $g^{\text{XNN}}$ than $g^{\text{MGBM}}$.
 Blacks receive high-priced loans 21.4\% more frequently using $g^{\text{XNN}}$ vs.
 18.3\% for $g^{\text{MGBM}}$.
-Both $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ show AIRs that are statistically significantly below parity (not shown, but available in resources discussed in Subsection \ref{ssec:soft}), and which are also below the EEOC’s 0.80 threshold.
-This would typically indicate need for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Subsection \ref{ssec:disc_rem}.
-On the other hand, women improve under $g^{\text{XNN}}$ vs.
+Both $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ showed AIRs that were statistically significantly below parity (not shown, but available in resources discussed in Section~\ref{ssec:soft}), and which were also below the EEOC’s 0.80 threshold.
+This would typically indicate need for further review to determine the cause and validity of these disparities, and a few relevant remediation techniques for such discovered discrimination are discussed in Section~\ref{ssec:disc_rem}.
+On the other hand, women improved under $g^{\text{XNN}}$ vs.
 $g^{\text{MGBM}}$ (MEs of 3.6\% vs.
@@
 0.948).
-The AIRs, while statistically significantly below parity, are well above the EEOC’s threshold of 0.80.
+The AIRs, while statistically significantly below parity, were well above the EEOC’s threshold of 0.80.
 In most situations, the values of these measures alone would not likely flag a model for further review.
-Black SMDs for $g^{\text{XNN}}(\mathbf{X})$ and $g^{\text{MGBM}}(\mathbf{X})$ are similar: 0.621 and 0.628, respectively.
-These exceed Cohen’s guidelines of 0.5 for a medium effect size and would likely trigger further review.
-Female SMDs are well below Cohen’s definition of small effect size: 0.105 and 0.084 for $g^{\text{XNN}}(\mathbf{X})$ and $g^{\text{MGBM}}(\mathbf{X})$, respectively.
+Black SMDs for $g^{\text{XNN}}(\mathbf{X})$ and $g^{\text{MGBM}}(\mathbf{X})$ were similar: 0.621 and 0.628, respectively.
+These exceeded Cohen’s guidelines of 0.5 for a medium effect size and would likely trigger further review.
+Female SMDs were well below Cohen’s definition of small effect size: 0.105 and 0.084 for $g^{\text{XNN}}(\mathbf{X})$ and $g^{\text{MGBM}}(\mathbf{X})$, respectively.
 Similar to results for female AIR, these values alone are unlikely to prompt further review.
@@
 \section{Discussion}\label{sec:disc}
+\vspace{-6pt}
 
@@
 
-Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, and deployment meant to decrease discrimination, inaccuracy, privacy, and security risks for high-stakes, human-centered, or regulated ML applications.\footnote{See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure \ref{fig:blueprint}.} While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
+Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, and deployment meant to decrease discrimination, inaccuracy, and privacy and security risks for high-stakes, human-centered, or regulated ML applications (See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure~\ref{fig:blueprint}.}).
+While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
 
@@
 	\centering
-	\includegraphics[width=12cm]{img/blueprint.png}
+	\includegraphics[width=15cm]{img/blueprint.png}
 	\caption{An example responsible ML workflow in which interpretable models, post-hoc explanations, discrimination testing and remediation techniques, among several other processes, can create an understandable and trustworthy ML system for high-stakes, human-centered, or regulated applications.}
@@
 
-Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University and EBMs, or GA\textsuperscript{2}Ms, pioneered by researchers at Microsoft and Cornell.
+Any discussion of interpretable ML models would be incomplete without references to the seminal work of the Rudin group at Duke University  and EBMs, or GA\textsuperscript{2}Ms, pioneered by researchers at Microsoft and Cornell \cite{ga2m,osdt,this_looks_like_that}.
 In keeping with a major theme of this manuscript, models from these leading researchers and several other kinds of interpretable ML models are now available as open source Python packages.
-Among several types of currently available interpretable models, practitioners can now use Python to evaluate EBM in the \href{https://github.com/interpretml/interpret}{\texttt{interpret}} package, optimal sparse decision trees, GAMs in the \href{https://github.com/dswah/pyGAM}{\texttt{pyGAM}} package, a variant of Friedman's RuleFit  in the \href{https://github.com/scikit-learn-contrib/skope-rules}{\texttt{skope-rules}} package, monotonic calibrated interpolated lookup tables in \href{https://github.com/tensorflow/lattice}{\texttt{tensorflow/lattice}}, and \textit{this looks like that} interpretable deep learning \cite{osdt}, \cite{rulefit}, \cite{lattice}, \cite{this_looks_like_that}.\footnote{See: \href{https://github.com/xiyanghu/OSDT}{Optimal sparse decision trees}.}\textsuperscript{,}\footnote{See: \href{https://github.com/cfchen-duke/ProtoPNet}{\textit{This looks like that} interpretable deep learning}.} Additional, relevant references and Python functionality include:
-
-\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-
-\item \textbf{Exploratory data analysis (EDA)}: \texttt{H2OAggregatorEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}{\texttt{h2o}} \cite{wilkinson2018visualizing}.
-
-\item \textbf{Sparse feature extraction}: \texttt{H2OGeneralizedLowRankEstimator} in \href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogeneralizedlowrankestimator}{\texttt{h2o}} \cite{udell2016generalized}.
-
-\item \textbf{Preprocessing \& models for privacy}: \href{https://github.com/IBM/differential-privacy-library}{\texttt{diffprivlib}}, \href{https://github.com/tensorflow/privacy/}{\texttt{tensorflow/privacy}} \cite{diffprivlib}, \cite{ji2014differential}, \cite{pate}, \cite{abadi2016deep}.
-
-\item \textbf{Causal inference \& probabilistic programming}: \href{https://github.com/microsoft/dowhy}{dowhy}, \href{https://docs.pymc.io/}{PyMC3} \cite{book_of_why}.
-
-\item \textbf{Post-hoc explanation}: structured data explanations with \href{https://github.com/SeldonIO/alibi}{\texttt{alibi}} and \href{https://github.com/SauceCat/PDPbox}{\texttt{PDPbox}}, image classification explanations with \href{https://github.com/marcoancona/DeepExplain}{\texttt{DeepExplain}}, and natural language explanations with \href{https://github.com/allenai/allennlp}{\texttt{allennlp}} \cite{wachter2017counterfactual}, \cite{grad_attr}, \cite{wallace2019allennlp}.
-
-\item \textbf{Discrimination testing}: \href{https://github.com/dssg/aequitas}{\texttt{aequitas}}, \href{https://github.com/LASER-UMASS/Themis}{\texttt{Themis}}.
-
-\item \textbf{Discrimination remediation}: Reweighing, adversarial de-biasing, learning fair representations, and reject option classification with \href{http://aif360.mybluemix.net/}{\texttt{AIF360}} \cite{kamiran2012data}, \cite{zhang2018mitigating}, \cite{lfr}, \cite{kamiran2012decision}.
-
-\item \textbf{Model debugging}: \href{https://github.com/bethgelab/foolbox}{\texttt{foolbox}}, \href{https://github.com/SALib/SALib}{\texttt{SALib}}, \href{https://github.com/tensorflow/cleverhans}{\texttt{tensorflow/cleverhans}}, and \href{https://github.com/tensorflow/model-analysis}{\texttt{tensorflow/model-analysis}} \cite{rauber2017foolbox}, \cite{papernot2018cleverhans}, \cite{modeltracker}, \cite{papernot2018marauder}.
-
-\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.,  \href{https://github.com/openai/gpt-2/blob/master/model_card.md}{GPT-2 model card}, \href{https://modelcards.withgoogle.com/object-detection}{Object Detection model card}.
+Among several types of currently available interpretable models, practitioners can now use Python to evaluate EBM in the  % we have removed the link of text
+%\href{https://github.com/interpretml/interpret}
+{\texttt{\hl{interpret}}} package, optimal sparse decision trees, GAMs in the 
+%\href{https://github.com/dswah/pyGAM}
+{\texttt{\hl{pyGAM}}} package, a variant of Friedman's RuleFit  in the 
+%\href{https://github.com/scikit-learn-contrib/skope-rules}
+{\texttt{\hl{skope-rules}}} package, monotonic calibrated interpolated lookup tables in 
+%\href{https://github.com/tensorflow/lattice}
+{\texttt{\hl{tensorflow/lattice}}}, and \textit{this looks like that} interpretable deep learning \cite{osdt,rulefit,lattice,this_looks_like_that} ({See: 
+\href{https://github.com/xiyanghu/OSDT}{optimal sparse decision trees}; see: 
+\href{https://github.com/cfchen-duke/ProtoPNet}{this looks like that}.). Additional, relevant references and Python functionality include:
+
+\begin{itemize}[leftmargin=*,labelsep=5.8mm]%is bold necessary?
+
+\item \textbf{\hl{Exploratory data analysis (EDA)}}: \texttt{H2OAggregatorEstimator} in 
+%\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}
+{\texttt{\hl{h2o}}} \cite{wilkinson2018visualizing}.
+
+\item \textbf{Sparse feature extraction}: \texttt{H2OGeneralizedLowRankEstimator} in 
+%\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogeneralizedlowrankestimator}
+{\texttt{\hl{h2o}}} \cite{udell2016generalized}.
+
+\item \textbf{Preprocessing and models for privacy}: 
+%\href{https://github.com/IBM/differential-privacy-library}
+{\texttt{diffprivlib}}, 
+%\href{https://github.com/tensorflow/privacy/}
+{\texttt{\hl{tensorflow/privacy}}} \cite{diffprivlib,ji2014differential,pate,abadi2016deep}.
+
+\item \textbf{Causal inference and probabilistic programming}: 
+%\href{https://github.com/microsoft/dowhy}
+{\hl{dowhy}}, 
+%\href{https://docs.pymc.io/}
+{PyMC3} \cite{book_of_why}.
+
+\item \textbf{Post-hoc explanation}: structured data explanations with 
+%\href{https://github.com/SeldonIO/alibi}
+{\texttt{\hl{alibi}}} and 
+%\href{https://github.com/SauceCat/PDPbox}
+{\texttt{\hl{PDPbox}}}, image classification explanations with 
+%\href{https://github.com/marcoancona/DeepExplain}
+{\texttt{\hl{DeepExplain}}}, and natural language explanations with 
+%\href{https://github.com/allenai/allennlp}
+{\texttt{\hl{allennlp}}} \cite{wachter2017counterfactual,grad_attr,wallace2019allennlp}.
+
+\item \textbf{Discrimination testing}: 
+%\href{https://github.com/dssg/aequitas}
+{\texttt{aequitas}}, 
+%\href{https://github.com/LASER-UMASS/Themis}
+{\texttt{Themis}}.
+
+\item \textbf{Discrimination remediation}: Reweighing, adversarial de-biasing, learning fair representations, and reject option classification with 
+%\href{http://aif360.mybluemix.net/}
+{\texttt{\hl{AIF}360}} \cite{kamiran2012data,zhang2018mitigating,lfr,kamiran2012decision}.
+
+\item \textbf{Model debugging}: 
+%\href{https://github.com/bethgelab/foolbox}
+{\texttt{\hl{foolbox}}}, 
+%\href{https://github.com/SALib/SALib}
+{\texttt{\hl{SALib}}}, 
+%\href{https://github.com/tensorflow/cleverhans}
+{\texttt{\hl{tensorflow/cleverhans}}}, and 
+%\href{https://github.com/tensorflow/model-analysis}
+{\texttt{\hl{tensorflow/model- analysis}}} \cite{rauber2017foolbox,papernot2018cleverhans,modeltracker,papernot2018marauder}.
+
+\item \textbf{Model documentation}: models cards \cite{model_cards}, e.g.,  
+\href{https://github.com/openai/gpt-2/blob/master/model_card.md}{GPT-2 model card}, 
+\href{https://modelcards.withgoogle.com/object-detection}{Object Detection model card}.
 
@@
 
-\noindent See \href{https://github.com/jphall663/awesome-machine-learning-interpretability}{Awesome Machine Learning Interpretability} for a longer, community-curated metalist of related software packages and resources.
+See 
+\href{https://github.com/jphall663/awesome-machine-learning-interpretability}
+{Awesome Machine Learning Interpretability} for a longer, community-curated metalist of related software packages and resources.
 
@@
 Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}.
-Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank SR 11-7: Guidance on Model Risk Management}.} Among these, providing persons affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
-ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions can be difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for compliance, but also a failsafe procedure for those affected by ML decisions.
+Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes (See:
+\href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}
+{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.
+{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.}).
+Transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations ({\hl{E.g.:}  % we have removed the link of text
+%\href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}
+{\hl{US Federal Reserve Bank SR 11-7: Guidance on Model Risk Management}}.}).%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+ Among these, providing persons affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
+ML models are often wrong ({``\hl{All models are wrong, but some are useful.''---George Box, Statistician (1919--2013)}}) %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+ and appealing black-box decisions can be difficult.
+For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for compliance, but also a failsafe procedure for those affected by ML decisions.
+%is this necessary?
 
@@
 A significant body of research has emerged around exploring and fixing ML discrimination \cite{friedler2019comparative}.
-Methods can be broadly placed into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and many approaches that have been developed in the last 5--7 years that alter the training algorithm, preprocess training data, or post-process predictions in order to diminish class-control correlations or dependencies.
+Methods can be broadly placed into two groups: more traditional methods that mitigate discrimination by searching across possible algorithmic and feature specifications, and many approaches that have been developed in the last 5--7 years that alter the training algorithm, preprocess training data, or~post-process predictions in order to diminish class-control correlations or dependencies.
 Whether these more recent methods are suitable for a particular use case depends on the legal environment where a model is deployed and on the use case itself.
@@
 
-The often black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the privacy harms and security vulnerabilities inherent in ML are each serious and difficult problems on their own.
-However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+The often black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or~the privacy harms and security vulnerabilities inherent in ML are each serious and difficult problems on their own.
+However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the {fairwashing} or {scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy,fair_washing,scaffolding} (See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}
+ {\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.} While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
-However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks, and that explanations can mask ML discrimination.
-Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory.
+This~is sometimes known as the {transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation}.
+{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.}).
+Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and~discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
+However, they should also consider that explanations can ease model stealing, data~extraction, and membership inference attacks, and that explanations can mask ML discrimination.
+Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and~specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory.
 Accuracy, discrimination, and security characteristics of a system can change over time as well.
-Simply testing for these problems at training time, as presented in Section \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems.
-Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
+Simply testing for these problems at training time, as presented in Section \ref{sec:res}, is~not adequate for high-stakes, human-centered, or regulated ML systems.
+Accuracy, discrimination, and~security should be monitored in real-time and over time, as long as a model is deployed.
+% we have removed the link of text
 
@@
 
+\vspace{+6pt}
+%\AuthorNames{Navdeep Gill, Patrick Hall, Kim Montgomery and Nicholas Schmidt}
 %%%%%%%%
-\authorcontributions{NG, data cleaning, GBM and MGBM assessment and results; PH, primary author; KM, ANN and XNN implementation, assessment, and results; NS, secondary author, data simulation and collection, and discrimination testing.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+\authorcontributions{N.G., data cleaning; GBM and MGBM, assessment and results; P.H., primary author; K.M., ANN, and XNN, implementation, assessment, and results; N.S., secondary author, data simulation and collection, and discrimination testing.
+All authors have read and agreed to the published version of the manuscript.}%For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
 The following statements should be used ``conceptualization, X.X.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{XNN was first made public by the corporate model validation team at Wells Fargo bank.
+\conflictsofinterest{\hl{The authors declare no conflict of interest}. %%Please confirm.
+XNN was first made public by the corporate model validation team at Wells Fargo bank.
 Wells Fargo is a customer of, and investor in, H2O.ai and a client of BLDS, LLC.
@@
 %% optional
-\abbreviations{The following abbreviations are used in this text: 
-AI -- artificial intelligence, % -------------------------------------------------------------------------- x
-AIR - adverse impact ratio, % ----------------------------------------------------------------------------- x
-ALE - accumulated local effect, % ------------------------------------------------------------------------- x
-ANN -- artificial neural network, % ----------------------------------------------------------------------- x
-APR -- annual percentage rate, % -------------------------------------------------------------------------- x
-AUC -- area under the curve, % ---------------------------------------------------------------------------- x
-CFPB -- Consumer Financial Protection Bureau, % ----------------------------------------------------------- x
-DI -- disparate impact, % --------------------------------------------------------------------------------- x
-DT -- disparate treatment, % ------------------------------------------------------------------------------ x
-DTI -- debt to income, % ---------------------------------------------------------------------------------- x
-EBM or GA\textsuperscript{2}M -- explainable boosting machine, i.e.
-variants GAMs that consider two-way interactions and may incorporate boosting into training, % ------------------------------------------------ x
-EEOC -- Equal Employment Opportunity Commission, % -------------------------------------------------------- x
-ECOA - Equal Credit Opportunity Act, % -------------------------------------------------------------------- x
-EDA -- exploratory data analysis, % ----------------------------------------------------------------------- x
-EU -- European Union, % ----------------------------------------------------------------------------------- x
-FCRA -- Fair Credit Reporting Act, % ---------------------------------------------------------------------- x
-FNR -- false negative rate, % ----------------------------------------------------------------------------- x
-FPR -- false positive rate, % ----------------------------------------------------------------------------- x
-GAM -- generalized additive model, % ---------------------------------------------------------------------- x
-GBM -- gradient boosting machine, % ----------------------------------------------------------------------- x
-GDPR - General Data Protection Regulation, % -------------------------------------------------------------- x
-HMDA -- Home Mortgage Disclosure Act, % ------------------------------------------------------------------- x
-ICE -- individual conditional expectation, % -------------------------------------------------------------- x
-LTV -- loan to value, % ----------------------------------------------------------------------------------- x
-MCC -- Matthews correlation coefficient, % ---------------------------------------------------------------- x
-ME -- marginal effect, % ---------------------------------------------------------------------------------- x
-MGBM -- monotonic gradient boosting machine, % ------------------------------------------------------------ x
-ML -- machine learning, % --------------------------------------------------------------------------------- x
-PD -- partial dependence, % ------------------------------------------------------------------------------- x
-RMSE -- root mean square error, % ------------------------------------------------------------------------- x
-SGD -- stochastic gradient descent, % --------------------------------------------------------------------- x
-SHAP -- SHapley Additive exPlanation, % ------------------------------------------------------------------- x
-SMD - standardized mean difference, % --------------------------------------------------------------------- x
-SR -- supervision and regulation, % ----------------------------------------------------------------------- x
-US -- United States, % ------------------------------------------------------------------------------------ x
-XNN -- explainable neural network.} % --------------------------------------------------------------------- x
+%\abbreviations{The following abbreviations are used in this text}
+
+\abbreviations{The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+AI & artificial intelligence\\ 
+AIR & adverse impact ratio\\ 
+ALE & accumulated local effect\\ 
+ANN & artificial neural network\\ % ----------------------------------------------------------------------- x
+APR & annual percentage rate\\ % -------------------------------------------------------------------------- x
+AUC & area under the curve\\ % ---------------------------------------------------------------------------- x
+CFPB & Consumer Financial Protection Bureau\\% ----------------------------------------------------------- x
+DI & disparate impact\\ % --------------------------------------------------------------------------------- x
+DT & disparate treatment\\% ------------------------------------------------------------------------------ x
+DTI & debt to income\\ % ---------------------------------------------------------------------------------- x
+\multirow{2}{*}{EBM or GA$\textsuperscript{2}$M} & explainable boosting machine, i.e., variants GAMs that consider two-way interactions and \\ &may incorporate boosting into training\\%
+EEOC & Equal Employment Opportunity Commission\\%
+ECOA & Equal Credit Opportunity Act\\% -------------------------------------------------------------------- x
+EDA & exploratory data analysis\\ % ----------------------------------------------------------------------- x
+EU & European Union\\ % ----------------------------------------------------------------------------------- x
+FCRA & Fair Credit Reporting Act\\ % ---------------------------------------------------------------------- x
+FNR & false negative rate\\ % ----------------------------------------------------------------------------- x
+FPR & false positive rate\\ % ----------------------------------------------------------------------------- x
+GAM & generalized additive model\\ % ---------------------------------------------------------------------- x
+GBM & gradient boosting machine\\ % ----------------------------------------------------------------------- x
+GDPR & General Data Protection Regulation\\% -------------------------------------------------------------- x
+%MGBM & monotonic gradient boosting machine\\
+%ML & machine learning\\ % --------------------------------------------------------------------------------- x
+%PD & partial dependence\\ % ------------------------------------------------------------------------------- x
+%RMSE & root mean square error\\ % ------------------------------------------------------------------------- x
+%SGD & stochastic gradient descent\\% --------------------------------------------------------------------- x
+%SHAP & Shapley Additive Explanation\\ % ------------------------------------------------------------------- x
+%SMD & standardized mean difference\\ % --------------------------------------------------------------------- x
+%SR & supervision and regulation\\ % ----------------------------------------------------------------------- x
+%US & United States\\ 
+%XNN & explainable neural network 
+\end{tabular}
+
+\noindent 
+\begin{tabular}{@{}ll}
+HMDA & Home Mortgage Disclosure Act\\% ------------------------------------------------------------------- x
+ICE & individual conditional expectation\\% -------------------------------------------------------------- x
+LTV & loan to value\\ % ----------------------------------------------------------------------------------- x
+MCC & Matthews correlation coefficient\\% ---------------------------------------------------------------- x
+ME & marginal effect\\ % ---------------------------------------------------------------------------------- x
+MGBM & monotonic gradient boosting machine\\
+ML & machine learning\\ % --------------------------------------------------------------------------------- x
+PD & partial dependence\\ % ------------------------------------------------------------------------------- x
+RMSE & root mean square error\\ % ------------------------------------------------------------------------- x
+SGD & stochastic gradient descent\\% --------------------------------------------------------------------- x
+SHAP & Shapley Additive Explanation\\ % ------------------------------------------------------------------- x
+SMD & standardized mean difference\\ % --------------------------------------------------------------------- x
+SR & supervision and regulation\\ % ----------------------------------------------------------------------- x
+US & United States\\ 
+XNN & explainable neural network 
+\end{tabular}}
 
@@
 This information is first provided to the CFPB, which subsequently releases some of the data to the public.
-Regulators often use HMDA data to, ``...show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory.''\textsuperscript{5} In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
+Regulators often use HMDA data to, ``...
+show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory'' (see: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}).
+In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
 The publicly available data that the CFPB releases includes information such as the lender, the type of loan, loan amount, LTV ratio, DTI ratio, and other important financial descriptors.
@@
 Because the data includes information on these protected class characteristics, certain measures that can be indicative of discrimination in lending can be calculated directly using the HDMA data.
-Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections \ref{sec:m_and_m} and \ref{sec:res}.
+Ultimately, the HMDA data represent the most comprehensive source of data on highly-regulated mortgage lending that is publicly available, which makes it an ideal dataset to use for the types of analyses set forth in Sections~\ref{sec:m_and_m}~and~\ref{sec:res}.
 
 \section{Selected Algorithmic Details}\label{a_sec:algo}
+\vspace{-6pt}
 
@@
 	\item Input features come from the set $\mathcal{X}$ contained in a \textit{P}-dimensional input space, $\mathcal{X} \subset \mathbb{R}^P$.
-An arbitrary, potentially unobserved, or future instance of $\mathcal{X}$ is denoted $\mathbf{x}$, $\mathbf{x} \in \mathcal{X}$.
+An~arbitrary, potentially unobserved, or future instance of $\mathcal{X}$ is denoted $\mathbf{x}$, $\mathbf{x} \in \mathcal{X}$.
 	\item Labels corresponding to instances of $\mathcal{X}$ come from the set $\mathcal{Y}$.
@@
 
-\noindent For some $g^\text{MGBM}$ model (see Equation \ref{eq:gbm}), monotonic splitting rules, $\Theta^{\text{mono}}_b$, are selected in a greedy, additive fashion by minimizing a regularized loss function, $\mathcal{L}$,  that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits applied to some instance $\mathbf{x}$, $T_b(\mathbf{x};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, in addition to a regularization term that penalizes complexity in the $b$-th tree, $\Omega(T_b)$.
+ For some $g^\text{MGBM}$ model (see Equation (\ref{eq:gbm})), monotonic splitting rules, $\Theta^{\text{mono}}_b$, are selected in a greedy, additive fashion by minimizing a regularized loss function, $\mathcal{L}$,  that considers known target labels, $\mathbf{y}$, the predictions of all subsequently trained trees in $g^{\text{MGBM}}$, $g^{\text{MGBM}}_{b-1}(\mathbf{X})$, and the $b$-th tree splits applied to some instance $\mathbf{x}$, $T_b(\mathbf{x};\Theta^\text{mono}_b)$, in a numeric error function (e.g., squared error, Huber error), $l$, in addition to a regularization term that penalizes complexity in the $b$-th tree, $\Omega(T_b)$.
 For the $b$-th iteration over $N$ instances, $\mathcal{L}_{b}$, can generally be defined as:
@@
 
-\noindent In addition to $\mathcal{L}$, $g^\text{MGBM}$ training is characterized by monotonic splitting rules and constraints on tree node weights.
+ In addition to $\mathcal{L}$, $g^\text{MGBM}$ training is characterized by monotonic splitting rules and constraints on tree node weights.
 Each binary splitting rule in $T_b$, $\theta_{b,j,k} \in \Theta_b$, is associated with a feature, $X_j$, is the $k$-th split associated with $X_j$ in $T_b$, and results in left, $L$, and right, $R$, child nodes with a numeric weights, $\{w_{b,j,k,L}, w_{b,j,k,R}\}$.
@@
 
-\noindent Note that for any one $X_j$ and subtree in $g^{\text{MGBM}}$, left subtrees will alway produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each full $T_b$ prediction, with the application of a monotonic logit or softmax link function for classification problems.
+ Note that for any one $X_j$ and subtree in $g^{\text{MGBM}}$, left subtrees will always produce lower predictions than right subtrees, and that any $g^{\text{MGBM}}(\mathbf{x})$ is an addition of each full $T_b$ prediction, with the application of a monotonic logit or softmax link function for classification problems.
 Moreover, each tree's root node corresponds to some constant node weight that by definition obeys monotonicity constraints, $ T(x^{\alpha}_j; \theta_{b,0}) = T(x^{\beta}_j; \theta_{b,0}) = w_{b,0}$.
@@
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=8cm]{img/xnn_arch.png}
-	\caption{A general diagram of an XNN with three metalayers: the bottom combination layer with  $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, the middle metalayer with with $K$ hidden and separate $n_k$ ridge functions, and the output combination layer that generates $g^{\text{XNN}}(\mathbf{X})$ predictions.
-Figure adapted from \citet{wf_xnn}.}
-	\label{fig:xnn_arch}
-\end{figure}
-
-\noindent $g^{\text{XNN}}$ is comprised of 3 meta-layers:
+ $g^{\text{XNN}}$ is comprised of 3 meta-layers:
 
 \begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-	\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units (see Equation \ref{eq:xnn}), which should learn higher magnitude weights for each important input, $X_j$, is known as the \textit{projection layer}.
+	\item The first and deepest meta-layer, composed of $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units (see Equation (\ref{eq:xnn})), which should learn higher magnitude weights for each important input, $X_j$, is known as the \textit{projection layer}.
 It is fully connected to each input $X_j$.
@@
 	\item The second meta-layer contains $K$ hidden and separate $n_k$ ridge functions, or \textit{subnetworks}.
-Each $n_k$ is a neural network itself, which can be parameterized to suit a given modeling task.
+Each $n_k$ is a neural network itself, which can be parametrized to suit a given modeling task.
 To facilitate direct interpretation and visualization, the input to each subnetwork is the 1-dimensional output of its associated projection layer $\sum_j\beta_{k,j}x_j$ hidden unit.
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=12cm]{img/xnn_arch.png}
+	\caption{{A general diagram} of an XNN with three metalayers: the bottom combination layer with  $K$ linear $\sum_j\beta_{k,j}x_j$ hidden units, the middle metalayer with with $K$ hidden and separate $n_k$ ridge functions, and the output combination layer that generates $g^{\text{XNN}}(\mathbf{X})$ predictions.
+Figure adapted from \citet{wf_xnn}.}
+	\label{fig:xnn_arch}%Figure A1 is not mentioned in the maunscript
+\end{figure}
+
 \subsection{One-dimensional Partial Dependence and Individual Conditional Expectation Details}\label{a_ssec:pd_ice}
@@
 $\text{ICE}(x_j, g)$ for a given instance $\mathbf{x}$ and feature $x_j$ is estimated as the output of $g(\mathbf{x})$ when $x_j$ is set to a constant $x^\gamma \in \mathcal{X}$ and all other features $\mathbf{x} \in \mathbf{X}_{\mathcal{P} \setminus \{j\}}$ are left untouched.
-PD and ICE curves are usually plotted over some set of constants drawn from $\mathcal{X}$, as displayed in Subsection \ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
+PD and ICE curves are usually plotted over some set of constants drawn from $\mathcal{X}$, as displayed in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 Due to known problems for PD in the presence of strong correlation and interactions, PD should not be used alone.
-PD should be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots}, \cite{ale_plot}.
+PD should be paired with ICE or be replaced with accumulated local effect (ALE) plots \cite{ice_plots,ale_plot}.
 
@@
 
-\noindent In Equation \ref{eq:shap_additive}, $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
+ In Equation (\ref{eq:shap_additive}), $\mathbf{z} \in \{0,1\}^\mathcal{P}$ is a binary representation of $\mathbf{x}$ where 0 indicates missingness.
 Each Shapley value, $\phi_j$, is the local feature contribution value associated with $x_j$, and $\phi_0$ is the average of $g(\mathbf{X})$.
@@
 
-\noindent Local, per-instance explanations using Shapley values tend to involve ranking $x_j \in \mathbf{x}$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say 5.
+ Local, per-instance explanations using Shapley values tend to involve ranking $x_j \in \mathbf{x}$ by $\phi_j$ values or delineating a set of the $X_j$ names associated with the $k$-largest $\phi_j$ values for some $\mathbf{x}$, where $k$ is some small positive integer, say five.
 Global explanations are typically the absolute mean of the $\phi_j$ associated with a given $X_j$ across all of the instances in some set $\mathbf{X}$.
@@
 DT occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.
-With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
+With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.
+DI occurs when some element of a decisioning process includes a {facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.
 In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.
@@
 
-A finding of \textit{practical significance} means that discovered disparity is not only statistically significant, but also passes beyond a chosen threshold that would constitute \textit{prima facie} evidence of illegal discrimination.
+A finding of \textit{practical significance} means that discovered disparity is not only statistically significant, but also passes beyond a chosen threshold that would constitute {prima facie} evidence of illegal discrimination.
 Practical significance acknowledges that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not truly meaningful.
@@
 Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
-One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes \textit{prima facie} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
-``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.
+One exception, however, is in employment discrimination analyses, where~the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes {prima facie} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination ({Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).}).
+It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.
 
@@
 The first guidance in US courts occurred in a case involving discrimination in jury selection, \textit{Castaneda vs.
-Partida}.\footnote{\textit{Castaneda vs.
-Partida}, 430 US 482 - Supreme Court (1977)}   Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts vs.
-United States.}\footnote{\textit{Hazelwood School Dist.
-vs.
-United States}, 433 US 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
+Partida} (430 US 482 - Supreme Court (1977)).
+Here, the US Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts vs.
+United States} (433 US 299 (1977)).
+Out of this, a 5\% two-sided test (z = 1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.
 
@@
 
-As seen in Subsection \ref{ssec:c_v_uc_sim}, little or no trade-off is required in terms of model to fit to use the constrained models.
-Hence, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Appendices \ref{a_ssec:int_ex_sim} - \ref{a_ssec:dis_sim}.
+As seen in Section~\ref{ssec:c_v_uc_sim}, little or no trade-off is required in terms of model to fit to use the constrained models.
+Hence, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Appendices \ref{a_ssec:int_ex_sim}  and  \ref{a_ssec:dis_sim}.
 Intrinsic interpretability for $g^\text{MGBM}$ is evaluated with PD and ICE, and post-hoc explainability is highlighted via global and local Shapley explanations.
 For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Subection \ref{ssec:xnn_sim}.
-Both $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
+Both~$g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[H]
+\centering
 	\captionsetup{width=16cm}
 	\centering
-	\includegraphics[width=10cm]{img/sim_mgbm_glob.png}
+	\includegraphics[width=12cm]{img/sim_mgbm_glob.png}
 	\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}(\mathbf{X})$ on the simulated test data.}
@@
 
-\noindent As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.
+As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[H]
 	\centering
 	\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
-	\caption{PD, ICE for 10 instances across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
+	\caption{PD, ICE for ten instances across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
 	\label{fig:sim_mgbm_glob_pdp_ice}
@@
 As the PD curves generally follow the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$.
-Also because PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is likely not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal-generating function in Equation \ref{eq:sim}.
+Also because PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is likely not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal-generating function in Equation (\ref{eq:sim}).
 The one-dimensional monotonic constraints may hinder the modeling of non-monotonic interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to the main drivers, $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$, all being constrained in the same direction and able to weakly interact as needed.
 
-Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
-
-\begin{figure}[htb]
+Local Shapley values for records at the 10{th}, 50{th}, and 90{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}.
+
+\begin{figure}[H]
 	\centering
@@
 
-\noindent The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result.
-For the lower prediction at the 10\textsuperscript{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative.
-At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive.
+The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result.
+For the lower prediction at the 10{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative.
+At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive.
 In each case, large local contributions generally follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.
@@
 
-\begin{figure}[htb]
-	\centering
-	\includegraphics[width=10cm]{img/sim_xnn_glob.png}
-	\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
+\begin{figure}[H]
+\centering
+	\captionsetup{width=16cm}
+	\centering
+	\includegraphics[width=12cm]{img/sim_xnn_glob.png}
+	\caption{Global mean absolute Deep SHAP feature importance for $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
 	\label{fig:sim_xnn_glob}
@@
 
-\noindent Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features.
+Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features.
 The consistency between the feature rankings of $g^\text{MGBM}$ and $g^\text{XNN}$ is somewhat striking, given their different hypothesis families and architectures.
@@
 
-\begin{figure}[htb]
+\begin{figure}[H]
 	\centering
@@
 
-\noindent As expected, Deep SHAP values generally increase from the 10\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90\textsuperscript{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x})$ predictions.
+ As expected, Deep SHAP values generally increase from the 10{th} percentile of $g^\text{XNN}(\mathbf{X})$ to the 90{th} percentile of $g^\text{XNN}(\mathbf{X})$, with primarily important global drivers of model behavior contributing to the selected local $g^\text{XNN}(\mathbf{x})$ predictions.
 
@@
 
-Tables \ref{tab:sim_disc}\textbf{a} and \ref{tab:sim_disc}\textbf{b} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
-Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and  $g^{\text{XNN}}(\mathbf{X})$, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\textsuperscript{11}
+Tables \ref{tab:sim_disc}{a,b} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.
+Several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and  $g^{\text{XNN}}(\mathbf{X})$, FNRs, their ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).
 
 Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might consider the relative FNRs as a measure of the class-control disparities.
-Table \ref{tab:sim_disc}\textbf{b} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
+Table \ref{tab:sim_disc}{b} shows that protected group 1 has higher relative FNRs under $g^{\text{XNN}}$ (1.13 vs.
 1.06).
-However, in Table \ref{tab:sim_disc}\textbf{a}  the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
+However, in Table \ref{tab:sim_disc}{a}  the overall FNRs were lower for  $g^{\text{XNN}}$ (0.357 vs.
 0.401).
@@
 
-\begin{table}[htb]
-	\hskip-10pt
-	\begin{subtable}{0.45\textwidth}
-		\captionsetup{width=7cm}
-		\caption{Group size, accuracy, and FNR for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.} 
-		\centering
-		\resizebox{\columnwidth}{!}{%
+%\begin{table}[H]
+%	\hskip-10pt
+%	\begin{subtable}{0.45\textwidth}
+%		\captionsetup{width=7cm}
+%		\caption{Group size, accuracy, and FNR for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.} 
+%		\centering
+%		\resizebox{\columnwidth}{!}{%
+%			\begin{tabular}{ccccc}
+%				\toprule
+%		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy}$\uparrow$ & \textbf{FNR}$\downarrow$ \\
+%				\midrule
+%				Protected 1	& 3,057 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} & \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
+%				\midrule			 
+%				Control 1 & 16,943 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.739 \\ 0.756\end{tabular} & \begin{tabular}{@{}c@{}} 0.378 \\ 0.314\end{tabular} \\
+%				\midrule
+%				Protected 2	& 9,916 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular} & \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
+%				\midrule
+%				Control 2 & 10,084 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.729 \\ 0.756\end{tabular} & \begin{tabular}{@{}c@{}} 0.420 \\ 0.332\end{tabular} \\
+%				\bottomrule
+%				\label{tab:sim_n_acc}
+%			\end{tabular}%
+%		}
+%	\end{subtable}
+%	\hspace{5pt}
+%	\begin{subtable}{0.45\textwidth}
+%		\vspace{-11pt}
+%		\captionsetup{width=10cm}
+%		\caption{%
+%			\leftskip=45pt
+%			AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
+%		\centering
+%		\scalebox{0.95}{
+%			\begin{tabular}{ccccccc}
+%				\toprule
+%		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}$\uparrow$ & \textbf{ME}$\downarrow$ & \textbf{SMD}$\downarrow$ & \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}$\downarrow$\\
+%				\midrule
+%				$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular} & \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
+%				\midrule 
+%				$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular} & \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
+%				\bottomrule
+%				\label{tab:sim_disc_detail}
+%			\end{tabular}
+%		}
+%	\end{subtable}
+%	\captionsetup{width=10cm}
+%	\caption{Discrimination measures for the simulated test data. Arrows indicate the direction of improvement for each measure.}
+%	\label{tab:sim_disc}
+%\end{table}
+
+\begin{table}[H]
+\centering
+\caption{Discrimination measures for the simulated test data. Arrows indicate the direction of improvement for each measure.}
+	\label{tab:sim_disc}
 \begin{tabular}{ccccc}
 \toprule
+		\multicolumn{5}{c}{\begin{tabular}{@{}c@{}} (\textbf{a}) Group size, accuracy, and FNR for $g^\text{MGBM}(\mathbf{X})$\\ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.
+\end{tabular}}    \\
+		\midrule
+
 		\textbf{Class} & $\mathbf{N}$ & \textbf{Model} & \textbf{Accuracy}$\uparrow$ & \textbf{FNR}$\downarrow$ \\
 				\midrule
-				Protected 1	& 3,057 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} & \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
+				Protected 1	& 3057 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.770 \\ 0.771\end{tabular} & \begin{tabular}{@{}c@{}} 0.401 \\ 0.357\end{tabular} \\
 				\midrule			 
@@
 				\midrule
-				Protected 2	& 9,916 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular} & \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
+				Protected 2	& 9916 & \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular} & \begin{tabular}{@{}c@{}} 0.758 \\ 0.762\end{tabular} & \begin{tabular}{@{}c@{}} 0.331 \\ 0.302\end{tabular} \\
 				\midrule
@@
 				\label{tab:sim_n_acc}
-			\end{tabular}%
-		}
-	\end{subtable}
-	\hspace{5pt}
-	\begin{subtable}{0.45\textwidth}
-		\vspace{-11pt}
-		\captionsetup{width=10cm}
-		\caption{%
-			\leftskip=45pt
-			AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ on the simulated test data.}
-		\centering
-		\scalebox{0.95}{
+			\end{tabular}\\
+
 		\begin{tabular}{ccccccc}
 				\toprule
+		\multicolumn{7}{c}{\begin{tabular}{@{}c@{}} (\textbf{b}) AIR, ME, SMD, and FNR ratio for $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$\\ on the simulated test data. \end{tabular}}   \\
+		\midrule
 		\textbf{Model} &  \begin{tabular}[x]{@{}c@{}}\textbf{Protected}\\\textbf{Class}\end{tabular} & \begin{tabular}[x]{@{}c@{}}\textbf{Control}\\\textbf{Class}\end{tabular} & \textbf{AIR}$\uparrow$ & \textbf{ME}$\downarrow$ & \textbf{SMD}$\downarrow$ & \begin{tabular}[x]{@{}c@{}}\textbf{FNR}\\\textbf{Ratio}\end{tabular}$\downarrow$\\
 				\midrule
-				$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular} & \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
+				$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular} & \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} $-$0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
 				\midrule 
-				$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular} & \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
+				$g^\text{XNN}$ & \begin{tabular}{@{}c@{}} 1 \\ 2\end{tabular} & \begin{tabular}{@{}c@{}} 1 \\  2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.727 \\ 0.976\end{tabular}	& \begin{tabular}{@{}c@{}} 12.0\%\\ 1.0\%\end{tabular}	& \begin{tabular}{@{}c@{}} $-$0.274 \\ 0.001\end{tabular}	& \begin{tabular}{@{}c@{}} 1.13 \\ 0.907\end{tabular} \\
 				\bottomrule
@@
 			\end{tabular}
-		}
-	\end{subtable}
-	\captionsetup{width=10cm}
-	\caption{Discrimination measures for the simulated test data. Arrows indicate the direction of improvement for each measure.}
-	\label{tab:sim_disc}
 \end{table}
 
-For $g^{\text{XNN}}(\mathbf{X})$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}\textbf{b}.
+For $g^{\text{XNN}}(\mathbf{X})$, 12.0\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}{b}.
 Of note is that 12.0\% is not a meaningful difference without context.
 If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 12.0\% could represent an extremely large difference and would require remediation.
-But if they represent substantially different populations, then 12.0\% could represent a reasonable deviation from parity.
+However, if they represent substantially different populations, then 12.0\% could represent a reasonable deviation from parity.
 As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 12.0\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population.
 Protected group 1’s AIR under $g^\text{XNN}$ is 0.727, below the EEOC 4/5ths rule threshold.
-It is also highly statistically significant (not shown, but available in resources discussed in Subsection \ref{ssec:soft}).
+It is also highly statistically significant (not shown, but available in resources discussed in Section~\ref{ssec:soft}).
 Together these would indicate that there may be evidence of illegal DI.
@@
 However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.
-Some pertinent remediation strategies for discovered discrimination are discussed in Subsection \ref{ssec:disc_rem}.
+Some pertinent remediation strategies for discovered discrimination are discussed in Section~\ref{ssec:disc_rem}.
 
 SMD is used here to measure $g^\text{MGBM}(\mathbf{X})$ and $g^\text{MGBM}(\mathbf{X})$ probabilities prior to being transformed into classifications.
-(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}(\mathbf{X})$ has less DI than $g^\text{XNN}(\mathbf{X})$ (-0.206 vs.
--0.274), but both are close to Cohen’s small effect threshold of -0.20.
+(This measurement would be particularly relevant if the probabilities are used in combination with other models to determine an outcome.)  The results show that $g^\text{MGBM}(\mathbf{X})$ has less DI than $g^\text{XNN}(\mathbf{X})$ ($-$0.206 vs.
+$-$0.274), but both are close to Cohen’s small effect threshold of $-$0.20.
 Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.
@@
 For purposes of evaluating the quality of the model, she may review confusion matrix statistics (accuracy, recall, precision, etc.) using cutoffs based on the maximum F1 score.
-But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
+However, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.
 Because disparity analyses are concerned with how people are affected by the deployed model, it is essential that any confusion matrix-based measures of disparity be calculated on the in-production classification decisions, rather than on cutoffs that are not related to what those affected by the model will experience.
@@
 Additionally, the effects of changing a model must be viewed holistically.
-For example, the mortgage data disparity analysis in Subsection \ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
+For example, the mortgage data disparity analysis in Section~\ref{ssec:dis_mort} shows that if one were to choose $g^\text{MGBM}$ over $g^\text{XNN}$ because $g^\text{MGBM}$ has a lower FPR ratio for blacks, it would ultimately lead to a higher FPR for blacks overall, which may represent doing more harm than good.
 Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
-A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.''\footnote{In the pay discrimination case, \textit{Bazemore vs.
-Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.} A model developed without inclusion of class status (or proxies thereof) considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will have the same model output (though there may be some small or random differences in outcomes due to computational issues).
+A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.'' ({\hl{In the pay discrimination case,} \textit{Bazemore vs.
+Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.}) % footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+A model developed without inclusion of class status (or proxies thereof) considers similarly situated people the same on the dimensions included in the model: people who have the same feature values will have the same model output (though there may be some small or random differences in outcomes due to computational issues).
 Obviously, the inclusion of protected class status will change model output by class.
-With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.\footnote{In a reverse discrimination case, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.}  Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
+With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal ({\hl{In a reverse discrimination case}, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.}) Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
 
@@
 %=====================================
-\externalbibliography{yes}
-\bibliography{article-information-2019}
+%%%\externalbibliography{yes}
+%%%\bibliography{article-information-2019}
+\begin{thebibliography}{999}
+\providecommand{\natexlab}[1]{#1}
+
+\bibitem[Rudin(2018)]{please_stop}
+Rudin, C.
+\newblock Please {S}top {E}xplaining {B}lack {B}ox {M}odels for {H}igh {S}takes
+  {D}ecisions and {U}se {I}nterpretable {M}odels {I}nstead.
+\newblock {\em arXiv} {\bf 2018}, arXiv:1811.10154.
+\newblock  Available online: \url{https://arxiv.org/pdf/1811.10154.pdf} (accessed on 26 February 2020).
+
+\bibitem[Feldman {et~al.}(2015)Feldman, Friedler, Moeller, Scheidegger, and
+  Venkatasubramanian]{feldman2015certifying}
+Feldman, M.; Friedler, S.A.; Moeller, J.; Scheidegger, C.; Venkatasubramanian,
+  S.
+\newblock Certifying and {R}emoving {D}isparate {I}mpact.
+\newblock In {Proceedings of the 21{st} ACM SIGKDD International
+  Conference on Knowledge Discovery and Data Mining},  Sydney, Australia, 10--13 August 2015; pp. 259--268.
+\newblock Available online: \url{https://arxiv.org/pdf/1412.3756.pdf} (accessed on 26 February 2020).
+
+\bibitem[Dwork {et~al.}(2012)Dwork, Hardt, Pitassi, Reingold, and
+  Zemel]{dwork2012fairness}
+Dwork, C.; Hardt, M.; Pitassi, T.; Reingold, O.; Zemel, R.
+\newblock Fairness {T}hrough {A}wareness.
+\newblock  In {Proceedings of the 3rd Innovations in Theoretical Computer Science
+  Conference}, Cambridge, MA, USA, 8--10 January 2012. pp. 214--226. 
+\newblock Available online: \url{https://arxiv.org/pdf/1104.3913.pdf} (accessed on 26 February 2020).
+
+\bibitem[Buolamwini and Gebru(2018)]{gender_shades}
+Buolamwini, J.; Gebru, T.
+\newblock Gender {S}hades: {I}ntersectional {A}ccuracy {D}isparities in
+  {C}ommercial {G}ender {C}lassification.
+\newblock In Proceedings of the  Conference on Fairness, Accountability and {Transparency}, New York, NY, USA, 23--24 February 2018. pp. 77--91.
+\newblock Available online: 
+  \url{http://proceedings.mlr.press/v81/buolamwini18a/buolamwini18a.pdf} (accessed on 26 February 2020).
+
+\bibitem[Barreno {et~al.}(2010)Barreno, Nelson, Joseph, and
+  Tygar]{security_of_ml}
+Barreno, M.; Nelson, B.; Joseph, A.D.; Tygar, J.
+\newblock The {S}ecurity of {M}achine {L}earning.
+\newblock {\em Mach. Learn.} {\bf 2010}, {\em 81},~121--148.
+\newblock Available online: 
+  \url{http://people.ischool.berkeley.edu/~tygar/papers/SML/sec_mach_learn_journal.pdf} (accessed on 26 February 2020).
+
+\bibitem[Tram{\`e}r {et~al.}(2016)Tram{\`e}r, Zhang, Juels, Reiter, and
+  Ristenpart]{model_stealing}
+Tram{\`e}r, F.; Zhang, F.; Juels, A.; Reiter, M.K.; Ristenpart, T.
+\newblock Stealing {M}achine {L}earning {M}odels via {P}rediction {A}{P}{I}s.
+\newblock In Proceedings of the 25th USENIX Security Symposium, Austin, TX, USA, 10--12 August 2016. pp. 601--618.
+%%%\newblock Available online: 
+%%%  \url{https://www.usenix.org/system/files/conference/usenixsecurity16/sec16_paper_tramer.pdf} ((accessed on 26 February 2020)).
+
+\bibitem[Shokri {et~al.}(2017)Shokri, Stronati, Song, and
+  Shmatikov]{membership_inference}
+Shokri, R.; Stronati, M.; Song, C.; Shmatikov, V.
+\newblock Membership {I}nference {A}ttacks {A}gainst {M}achine {L}earning
+  {M}odels.
+\newblock  In Proceedings of the 2017 IEEE Symposium on Security and Privacy (SP), San Jose, CA, USA, 25 May 2017, pp.
+  3--18.
+\newblock Available online: \url{https://arxiv.org/pdf/1610.05820.pdf} (accessed on 26 February 2020).
+
+\bibitem[Shokri {et~al.}(2019)Shokri, Strobel, and Zick]{shokri2019privacy}
+Shokri, R.; Strobel, M.; Zick, Y.
+\newblock Privacy {R}isks of {E}xplaining {M}achine {L}earning {M}odels.
+\newblock {\em arXiv} {\bf 2019}, arXiv:1907.00164.
+\newblock Available online: \url{https://arxiv.org/pdf/1907.00164.pdf} (accessed on 26 February 2020).
+
+\bibitem[Williams {et~al.}(2017)Williams et~al.]{ff_interpretability}
+Williams, M.
+\newblock {\em Interpretability}; \hl{Fast Forward Labs}: Brooklyn, NY, USA.  2017.   %Please add the location of the publisher 
+\newblock Available online: 
+  \url{https://www.cloudera.com/products/fast-forward-labs-research.html} (accessed on 26 February 2020).
+
+\bibitem[Friedman(1979)]{friedman1979tree}
+Friedman, J.H.
+\newblock A {T}ree-structured {A}pproach to {N}onparametric {M}ultiple
+  {R}egression. In {\em Smoothing Techniques for Curve Estimation}; Springer: {Berlin/Heidelberg, Germany,} 
+  1979; pp. 5--22.
+%%%\newblock Available online: 
+%%%  \url{http://inspirehep.net/record/140963/files/slac-pub-2336.pdf} (\hl{accessed~on}).
+
+\bibitem[Friedman {et~al.}(1991)Friedman et~al.]{friedman1991multivariate}
+Friedman, J.H.
+\newblock Multivariate {A}daptive {R}egression {S}plines.
+\newblock {\em  Ann. Stat.} {\bf 1991}, {\em 19},~1--67.
+%%%\newblock Available online: 
+%%%  \url{https://projecteuclid.org/download/pdf_1/euclid.aos/1176347963} (\hl{accessed~on}).
+
+\bibitem[hmd()]{hmda}
+Mortgage Data ({H}{M}{D}{A}).
+\newblock Available online:  \url{https://www.consumerfinance.gov/data-research/hmda/} ({accessed~on 24 February 2020}).
+%\newblock Accessed: 2020-02-24.
+
+\bibitem[Friedman(2001)]{friedman2001greedy}
+Friedman, J.H.
+\newblock Greedy {F}unction {A}pproximation: a {G}radient {B}oosting {M}achine.
+\newblock {\em  Ann. Stat.} {\bf 2001}, 1189--1232.
+%%%\newblock Available online: 
+%%%  \url{https://projecteuclid.org/download/pdf_1/euclid.aos/1013203451} (\hl{accessed~on}).
+
+\bibitem[Friedman {et~al.}(2001)Friedman, Hastie, and Tibshirani]{esl}
+Friedman, J.H.; Hastie, T.; Tibshirani, R.
+\newblock {\em  {The Elements of Statistical Learning}}; Springer: Berlin/Heidelberg, Germany.
+\newblock Available online: 
+  \url{https://web.stanford.edu/~hastie/ElemStatLearn/printings/ESLII\_print12.pdf} (accessed on 26 February 2020).
+
+\bibitem[Recht {et~al.}(2011)Recht, Re, Wright, and Niu]{recht2011hogwild}
+Recht, B.; Re, C.; Wright, S.; Niu, F.
+\newblock {H}{O}{G}{W}{I}{L}{D}: {A} {L}ock-free {A}pproach to {P}arallelizing
+  {S}tochastic {G}radient {D}escent.
+%%%\newblock In \emph{Advances in Neural Information Processing Systems (NIPS)}; \hl{2011}; pp.
+%%%  693--701.   %Please add the publisher and location , or add %please add accessed on date.
+\newblock Available online: 
+  \url{https://papers.nips.cc/paper/4390-hogwild-a-lock-free-approach-to-parallelizing-stochastic-gradient-descent.pdf}  (accessed on 26 February 2020).
+
+\bibitem[Hinton {et~al.}(2012)Hinton, Srivastava, Krizhevsky, Sutskever, and
+  Salakhutdinov]{hinton2012improving}
+Hinton, G.E.; Srivastava, N.; Krizhevsky, A.; Sutskever, I.; Salakhutdinov,
+  R.R.
+\newblock Improving {N}eural {N}etworks by {P}reventing {C}o-adaptation of
+  {F}eature {D}etectors.
+\newblock {\em arXiv} {\bf 2012}, arXiv:1207.0580.
+\newblock Available online: \url{https://arxiv.org/pdf/1207.0580.pdf}  (accessed on 26 February 2020).
+
+\bibitem[Sutskever {et~al.}(2013)Sutskever, Martens, Dahl, and
+  Hinton]{sutskever2013importance}
+Sutskever, I.; Martens, J.; Dahl, G.; Hinton, G.
+\newblock On the {I}mportance of {I}nitialization and {M}omentum in {D}eep
+  {L}earning.
+\newblock  In Proceedings of the International Conference on Machine Learning, Atlanta, GA, USA, 16--21 June 2013, pp. 1139--1147. 
+\newblock Available online: \url{http://proceedings.mlr.press/v28/sutskever13.pdf} (accessed on 26 February 2020).
+
+\bibitem[Zeiler(2012)]{zeiler2012adadelta}
+Zeiler, M.D.
+\newblock {A}{D}{A}{D}{E}{L}{T}{A}: {A}n {A}daptive {L}earning {R}ate {M}ethod.
+\newblock {\em arXiv} {\bf 2012}, arXiv:1212.5701.
+\newblock Available online: \url{https://arxiv.org/pdf/1212.5701.pdf} (accessed on 26 February 2020).
+
+\bibitem[A{\"\i}vodji {et~al.}(2019)A{\"\i}vodji, Arai, Fortineau, Gambs,
+  Hara, and Tapp]{fair_washing}
+A{\"\i}vodji, U.; Arai, H.; Fortineau, O.; Gambs, S.; Hara, S.; Tapp, A.
+\newblock Fairwashing: {T}he {R}isk of {R}ationalization.
+\newblock {\em arXiv } {\bf 2019}, arXiv:1901.09749.
+\newblock Available online: \url{https://arxiv.org/pdf/1901.09749.pdf} (accessed on 26 February 2020).
+
+\bibitem[Slack {et~al.}(2019)Slack, Hilgard, Jia, Singh, and
+  Lakkaraju]{scaffolding}
+Slack, D.; Hilgard, S.; Jia, E.; Singh, S.; Lakkaraju, H.
+\newblock {F}ooling {L}{I}{M}{E} and {S}{H}{A}{P}: {A}dversarial {A}ttacks on
+  {P}ost-hoc {E}xplanation {M}ethods.
+\newblock {\em arXiv} {\bf 2019}, arXiv:1911.02508.
+\newblock Available online: \url{https://arxiv.org/pdf/1911.02508.pdf} (accessed on 26 February 2020).
+
+\bibitem[Vaughan {et~al.}(2018)Vaughan, Sudjianto, Brahimi, Chen, and
+  Nair]{wf_xnn}
+Vaughan, J.; Sudjianto, A.; Brahimi, E.; Chen, J.; Nair, V.N.
+\newblock Explainable {N}eural {N}etworks {B}ased on {A}dditive {I}ndex
+  {M}odels.
+\newblock {\em arXiv} {\bf 2018}, arXiv:1806.01933.
+\newblock Available online: \url{https://arxiv.org/pdf/1806.01933.pdf} (accessed on 26 February 2020).
+
+\bibitem[Yang {et~al.}(2019)Yang, Zhang, and Sudjianto]{yang2019enhancing}
+Yang, Z.; Zhang, A.; Sudjianto, A.
+\newblock Enhancing {E}xplainability of {N}eural {N}etworks {T}hrough
+  {A}rchitecture {C}onstraints.
+\newblock {\em arXiv} {\bf 2019}, arXiv:1901.03838.
+\newblock Available online: \url{https://arxiv.org/pdf/1901.03838.pdf} (accessed on 26 February 2020).
+
+\bibitem[Goldstein {et~al.}(2015)Goldstein, Kapelner, Bleich, and
+  Pitkin]{ice_plots}
+Goldstein, A.; Kapelner, A.; Bleich, J.; Pitkin, E.
+\newblock Peeking {I}nside the {B}lack {B}ox: {V}isualizing {S}tatistical
+  {L}earning with {P}lots of {I}ndividual {C}onditional {E}xpectation.
+\newblock {\em J. Comput. Graph. Stat.} {\bf 2015},
+  {\em 24}.
+\newblock Available online: \url{https://arxiv.org/pdf/1309.6392.pdf} (accessed on 26 February 2020).
+
+\bibitem[Lundberg and Lee(2017)]{shapley}
+Lundberg, S.M.; Lee, S.I.
+\newblock A {U}nified {A}pproach to {I}nterpreting {M}odel {P}redictions. In
+  {\em Advances in Neural Information Processing Systems (NIPS)}; Guyon, I.,
+  Luxburg, U.V., Bengio, S., Wallach, H., Fergus, R., Vishwanathan, S.,
+  Garnett, R., Eds.; Curran Associates, Inc.: Red Hook, NY, USA, 2017; pp. 4765--4774.  
+\newblock Available online: 
+  \url{http://papers.nips.cc/paper/7062-a-unified-approach-to-interpreting-model-predictions.pdf} (accessed on 26 February 2020).
+
+\bibitem[Lundberg {et~al.}(2017)Lundberg, Erion, and Lee]{tree_shap}
+Lundberg, S.M.; Erion, G.G.; Lee, S.I.
+\newblock Consistent {I}ndividualized {F}eature {A}ttribution for {T}ree
+  {E}nsembles. In Proceedings of the 2017 ICML Workshop on Human
+  Interpretability in Machine Learning (WHI 2017), Sydney, Australia, 10 August 2017, pp. 15--21. 
+\newblock Available online: \url{https://openreview.net/pdf?id=ByTKSo-m-} (accessed on 26 February 2020).
+
+\bibitem[Cohen(1988)]{cohen1988statistical}
+Cohen, J.
+\newblock {\em {Statistical {P}ower {A}nalysis for the {B}ehavioral
+  {S}ciences}}; Lawrence Erlbaum Associates: Mahwah, NJ, USA, 1988. 
+\newblock Available online: 
+  \url{http://www.utstat.toronto.edu/~brunner/oldclass/378f16/readings/CohenPower.pdf}  (accessed on 26 February 2020).
+
+\bibitem[Cohen(1992)]{cohen1992power}
+Cohen, J.
+\newblock A {P}ower {P}rimer.
+\newblock {\em Psychol. Bull.} {\bf 1992}, {\em 112},~155.
+\newblock Available online: \url{https://www.ime.usp.br/~abe/lista/pdfn45sGokvRe.pdf} (accessed on 26 February 2020).
+
+\bibitem[Zafar {et~al.}(2017)Zafar, Valera, Gomez~Rodriguez, and
+  Gummadi]{zafar2017fairness}
+Zafar, M.B.; Valera, I.; Gomez~Rodriguez, M.; Gummadi, K.P.
+\newblock Fairness {B}eyond {D}isparate {T}reatment \& {D}isparate {I}mpact:
+  {L}earning {C}lassification {W}ithout {D}isparate {M}istreatment.
+\newblock  In Proceedings of the  26th International Conference on World Wide Web, Perth, Australia, 3--7 April  2017, pp.
+  1171--1180. 
+\newblock Available online: \url{https://arxiv.org/pdf/1610.08452.pdf} (accessed on 26 February 2020).
+
+\bibitem[Lou {et~al.}(2013)Lou, Caruana, Gehrke, and Hooker]{ga2m}
+Lou, Y.; Caruana, R.; Gehrke, J.; Hooker, G.
+\newblock Accurate {I}ntelligible {M}odels with {P}airwise {I}nteractions.
+\newblock In Proceedings of the  19th ACM SIGKDD International Conference on
+  Knowledge Discovery and Data Mining,  Chicago, IL, USA, 11--14 August 2013, pp. 623--631. 
+\newblock Available online: 
+  \url{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.352.7682&rep=rep1&type=pdf} (accessed on 26 February 2020).
+
+\bibitem[Apley(2016)]{ale_plot}
+Apley, D.W.
+\newblock Visualizing the {E}ffects of {P}redictor {V}ariables in {B}lack {B}ox
+  {S}upervised {L}earning {M}odels.
+\newblock {\em arXiv} {\bf 2016}, arXiv:1612.08468.
+\newblock Available online: \url{https://arxiv.org/pdf/1612.08468.pdf} (accessed on 26 February 2020).
+
+\bibitem[Shapley {et~al.}(1988)Shapley, Roth, et~al.]{shapley1988shapley}
+Shapley, L.S.; Roth, A.E.
+\newblock {\em {The Shapley value: {E}ssays in Honor of {L}loyd {S}.
+  {S}hapley}}; Cambridge University Press: Cambridge, UK, 1988. 
+\newblock Available online: \url{http://www.library.fa.ru/files/Roth2.pdf} (accessed on 26 February 2020).
+
+\bibitem[Hall(2019)]{art_and_sci}
+Hall, P.
+\newblock On the {A}rt and {S}cience of {M}achine {L}earning {E}xplanations.
+%%%\newblock  KDD '19 XAI Workshop.  2019.
+\newblock Available online: \url{https://arxiv.org/pdf/1810.02909.pdf} (accessed on 26 February 2020).
+
+\bibitem[Molnar(2018)]{molnar}
+Molnar, C.
+\newblock {\em {\textit{{I}nterpretable {M}achine {L}earning}}}.
+%%%  christophm.github.io.  2018.
+\newblock Available online: \url{https://christophm.github.io/interpretable-ml-book/} (accessed on 26 February 2020).
+
+\bibitem[Hu {et~al.}(2019)Hu, Rudin, and Seltzer]{osdt}
+Hu, X.; Rudin, C.; Seltzer, M.
+\newblock Optimal {S}parse {D}ecision {T}rees.
+\newblock {\em arXiv} {\bf 2019}, arXiv:1904.12847.
+\newblock Available online: \url{https://arxiv.org/pdf/1904.12847.pdf} (accessed on 26 February 2020).
+
+\bibitem[Chen {et~al.}(2019)Chen, Li, Barnett, Su, and
+  Rudin]{this_looks_like_that}
+Chen, C.; Li, O.; Barnett, A.; Su, J.; Rudin, C.
+\newblock {T}his {L}ooks {L}ike {T}hat: {D}eep {L}earning for {I}nterpretable
+  {I}mage {R}ecognition.
+\newblock  In Proceedings of Neural Information Processing Systems {(NeurIPS)}, Vancouver, BC, Canada, 8--14 December 2019. 
+\newblock Available online: \url{https://arxiv.org/pdf/1806.10574.pdf} (accessed on 26 February 2020).
+
+\bibitem[Friedman {et~al.}(2008)Friedman, Popescu, et~al.]{rulefit}
+Friedman, J.H.; Popescu, B.E.
+\newblock Predictive {L}earning {V}ia {R}ule {E}nsembles.
+\newblock {\em  Ann. Appl. Stat.} {\bf 2008}, {\em 2},~916--954.
+\newblock Available online: 
+  \url{https://projecteuclid.org/download/pdfview_1/euclid.aoas/1223908046} (accessed on 26 February 2020).
+
+\bibitem[Gupta {et~al.}(2016)Gupta, Cotter, Pfeifer, Voevodski, Canini,
+  Mangylov, Moczydlowski, and Van~Esbroeck]{lattice}
+Gupta, M.; Cotter, A.; Pfeifer, J.; Voevodski, K.; Canini, K.; Mangylov, A.;
+  Moczydlowski, W.; Van~Esbroeck, A.
+\newblock Monotonic {C}alibrated {I}nterpolated {L}ookup {T}ables.
+\newblock {\em  J. Mach. Learn. Res.} {\bf 2016}, {\em
+  17},~3790--3836.
+\newblock Available online: \url{http://www.jmlr.org/papers/volume17/15-243/15-243.pdf} (accessed on 26 February 2020).
+
+\bibitem[Wilkinson(2018)]{wilkinson2018visualizing}
+Wilkinson, L.
+\newblock Visualizing Big Data Outliers through Distributed Aggregation.
+\newblock {\em IEEE Trans. Vis. Comput. Graph.} {\bf
+  2018}.
+\newblock Available online: 
+  \url{https://www.cs.uic.edu/~wilkinson/Publications/outliers.pdf} (accessed on 26 February 2020).
+
+\bibitem[Udell {et~al.}(2016)Udell, Horn, Zadeh, Boyd,
+  et~al.]{udell2016generalized}
+Udell, M.; Horn, C.; Zadeh, R.; Boyd, S.
+\newblock Generalized {L}ow {R}ank {M}odels.
+\newblock {\em Found. Trends{\textregistered} Mach. Learn.}
+  {\bf 2016}, {\em 9},~1--118.
+\newblock Available online: \url{https://www.nowpublishers.com/article/Details/MAL-055} (accessed on 26 February 2020).
+
+\bibitem[Holohan {et~al.}(2019)Holohan, Braghin, Mac~Aonghusa, and
+  Levacher]{diffprivlib}
+Holohan, N.; Braghin, S.; Mac~Aonghusa, P.; Levacher, K.
+\newblock {D}iffprivlib: {T}he {I}{B}{M} {D}ifferential {P}rivacy {L}ibrary.
+\newblock {\em arXiv} {\bf 2019}, arXiv:1907.02444.
+\newblock Available online: \url{https://arxiv.org/pdf/1907.02444.pdf} (accessed on 26 February 2020).
+
+\bibitem[Ji {et~al.}(2014)Ji, Lipton, and Elkan]{ji2014differential}
+Ji, Z.; Lipton, Z.C.; Elkan, C.
+\newblock Differential {P}rivacy and {M}achine {L}earning: {A} {S}urvey and
+  {R}eview.
+\newblock {\em arXiv} {\bf 2014}, arXiv:1412.7584.
+\newblock Available online: \url{https://arxiv.org/pdf/1412.7584.pdf} (accessed on 26 February 2020).
+
+\bibitem[Papernot {et~al.}(2018)Papernot, Song, Mironov, Raghunathan,
+  Talwar, and Erlingsson]{pate}
+Papernot, N.; Song, S.; Mironov, I.; Raghunathan, A.; Talwar, K.; Erlingsson,
+  {\'U}.
+\newblock Scalable {P}rivate {L}earning with {P}{A}{T}{E}.
+\newblock {\em arXiv} {\bf 2018}, arXiv:1802.08908.
+\newblock Available online: \url{https://arxiv.org/pdf/1802.08908.pdf} (accessed on 26 February 2020).
+
+\bibitem[Abadi {et~al.}(2016)Abadi, Chu, Goodfellow, McMahan, Mironov,
+  Talwar, and Zhang]{abadi2016deep}
+Abadi, M.; Chu, A.; Goodfellow, I.; McMahan, H.B.; Mironov, I.; Talwar, K.;
+  Zhang, L.
+\newblock Deep {L}earning with {D}ifferential {P}rivacy.
+\newblock  In Proceedings of the 2016 ACM SIGSAC Conference on Computer and
+  Communications Security, Vienna, Austria, 24--28 October 2016, pp. 308--318. 
+\newblock Available online: \url{https://arxiv.org/pdf/1607.00133.pdf} (accessed on 26 February 2020).
+
+\bibitem[Pearl and Mackenzie(2018)]{book_of_why}
+Pearl, J.; Mackenzie, D.
+\newblock {\em The {B}ook of {W}hy: {T}he {N}ew {S}cience of {C}ause and
+  {E}ffect}.
+%%%{Basic Books}:  2018.
+\newblock Available online: 
+  \url{http://cdar.berkeley.edu/wp-content/uploads/2017/04/Lisa-Goldberg-reviews-The-Book-of-Why.pdf} (accessed on 26 February 2020).
+
+\bibitem[Wachter {et~al.}(2017)Wachter, Mittelstadt, and
+  Russell]{wachter2017counterfactual}
+Wachter, S.; Mittelstadt, B.; Russell, C.
+\newblock Counterfactual {E}xplanations without {O}pening the {B}lack {B}ox:
+  {A}utomated {D}ecisions and the {G}{P}{D}{R}.
+\newblock {\em Harv. JL  Tech.} {\bf 2017}, {\em 31},~841.
+\newblock Available online: \url{https://arxiv.org/pdf/1711.00399.pdf} (accessed on 26 February 2020).
+
+\bibitem[Ancona {et~al.}(2018)Ancona, Ceolini, Oztireli, and
+  Gross]{grad_attr}
+Ancona, M.; Ceolini, E.; Oztireli, C.; Gross, M.
+\newblock Towards {B}etter {U}nderstanding of {G}radient-based {A}ttribution
+  {M}ethods for {D}eep {N}eural {N}etworks.
+\newblock  In Proceedings of the 6th International Conference on Learning Representations (ICLR
+  2018), Vancouver, BC, Canada, 30 April--3 May 2018. 
+\newblock Available online: 
+  \url{https://www.research-collection.ethz.ch/bitstream/handle/20.500.11850/249929/Flow_ICLR_2018.pdf} (accessed on 26 February 2020).
+
+\bibitem[Wallace {et~al.}(2019)Wallace, Tuyls, Wang, Subramanian, Gardner,
+  and Singh]{wallace2019allennlp}
+Wallace, E.; Tuyls, J.; Wang, J.; Subramanian, S.; Gardner, M.; Singh, S.
+\newblock AllenNLP Interpret: A Framework for Explaining Predictions of NLP
+  Models.
+\newblock {\em arXiv} {\bf 2019}, arXiv:1909.09251.
+\newblock Available online: \url{https://arxiv.org/pdf/1909.09251.pdf} (accessed on 26 February 2020).
+
+\bibitem[Kamiran and Calders(2012)]{kamiran2012data}
+Kamiran, F.; Calders, T.
+\newblock Data {P}reprocessing {T}echniques for {C}lassification {W}ithout
+  {D}iscrimination.
+\newblock {\em Knowl. Inf. Syst.} {\bf 2012}, {\em 33},~1--33.
+\newblock Available online: \url{https://bit.ly/2lH95lQ} (accessed on 26 February 2020).
+
+\bibitem[Zhang {et~al.}(2018)Zhang, Lemoine, and
+  Mitchell]{zhang2018mitigating}
+Zhang, B.H.; Lemoine, B.; Mitchell, M.
+\newblock Mitigating {U}nwanted {B}iases with {A}dversarial {L}earning.
+\newblock  In Proceedings of the 2018 AAAI/ACM Conference on AI, Ethics, and
+  Society, New Orleans, LA, USA, 2--3 February 2018, pp. 335--340.
+\newblock Available online: \url{https://arxiv.org/pdf/1801.07593.pdf} (accessed on 26 February 2020).
+\bibitem[Zemel {et~al.}(2013)Zemel, Wu, Swersky, Pitassi, and Dwork]{lfr}
+Zemel, R.; Wu, Y.; Swersky, K.; Pitassi, T.; Dwork, C.
+\newblock Learning {F}air {R}epresentations.
+\newblock In Proceedings of the  International Conference on Machine Learning, Atlanta, GA, USA, 16--21 June 2013, pp.
+325--333.
+%Please add the location and day month year for the conference , or add %please add accessed on date.
+\newblock Available online: \url{http://proceedings.mlr.press/v28/zemel13.pdf} (accessed on 26 February 2020).
+
+\bibitem[Kamiran {et~al.}(2012)Kamiran, Karim, and
+  Zhang]{kamiran2012decision}
+Kamiran, F.; Karim, A.; Zhang, X.
+\newblock Decision {T}heory for {D}iscrimination-aware {C}lassification.
+\newblock  In Proceedings of the 2012 IEEE 12th International Conference on Data Mining, Brussels, Belgium, 10 December 2012, pp. 924--929.
+\newblock Available online: 
+  \url{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.722.3030&rep=rep1&type=pdf} (accessed on 26 February 2020).
+
+\bibitem[Rauber {et~al.}(2017)Rauber, Brendel, and
+  Bethge]{rauber2017foolbox}
+Rauber, J.; Brendel, W.; Bethge, M.
+\newblock {F}oolbox: {A} {P}ython {T}oolbox to {B}enchmark the {R}obustness of
+  {M}achine {L}earning {M}odels.
+\newblock {\em arXiv} {\bf 2017}, arXiv:1707.04131.
+\newblock Available online: \url{https://arxiv.org/pdf/1707.04131.pdf} (accessed on 26 February 2020).
+
+\bibitem[Papernot {et~al.}(2018)Papernot, Faghri, Carlini, Goodfellow,
+  Feinman, Kurakin, Xie, Sharma, Brown, Roy, Matyasko, Behzadan, Hambardzumyan,
+  Zhang, Juang, Li, Sheatsley, Garg, Uesato, Gierke, Dong, Berthelot,
+  Hendricks, Rauber, and Long]{papernot2018cleverhans}
+Papernot, N.; Faghri, F.; Carlini, N.; Goodfellow, I.; Feinman, R.; Kurakin,
+  A.; Xie, C.; Sharma, Y.; Brown, T.; Roy, A.; et al.
+\newblock {T}echnical {R}eport on the {C}leverHans v2.1.0 {A}dversarial
+  {E}xamples {L}ibrary.
+\newblock {\em arXiv} {\bf 2018}, arXiv:1610.00768.
+\newblock Available online: \url{https://arxiv.org/pdf/1610.00768.pdf} (accessed on 26 February 2020).
+
+\bibitem[Amershi {et~al.}(2015)Amershi, Chickering, Drucker, Lee, Simard,
+  and Suh]{modeltracker}
+Amershi, S.; Chickering, M.; Drucker, S.M.; Lee, B.; Simard, P.; Suh, J.
+\newblock Modeltracker: {R}edesigning {P}erformance {A}nalysis {T}ools for
+  {M}achine {L}earning.
+\newblock  In Proceedings of the 33rd Annual ACM Conference on Human Factors in
+  Computing Systems, Seoul, Korea, 18--23 April 2015, pp. 337--346. %% Please add the location and day month year for the conference. please add ACM location , or add %please add accessed on date.
+\newblock Available online: 
+  \url{https://www.microsoft.com/en-us/research/wp-content/uploads/2016/02/amershi.CHI2015.ModelTracker.pdf} (accessed on 26 February 2020).
+
+\bibitem[Papernot(2018)]{papernot2018marauder}
+Papernot, N.
+\newblock A {M}arauder's {M}ap of {S}ecurity and {P}rivacy in {M}achine
+  {L}earning: {A}n overview of current and future research directions for
+  making machine learning secure and private.
+\newblock  In Proceedings of the 11th ACM Workshop on Artificial Intelligence and
+  Security, Toronto, ON, Canada, 19 October 2018. %%% Please add the location and day month year for the conference. please add ACM location , or add %please add accessed on date.
+\newblock Available online: \url{https://arxiv.org/pdf/1811.01134.pdf} (accessed on 26 February 2020).
+
+\bibitem[Mitchell {et~al.}(2019)Mitchell, Wu, Zaldivar, Barnes, Vasserman,
+  Hutchinson, Spitzer, Raji, and Gebru]{model_cards}
+Mitchell, M.; Wu, S.; Zaldivar, A.; Barnes, P.; Vasserman, L.; Hutchinson, B.;
+  Spitzer, E.; Raji, I.D.; Gebru, T.
+\newblock Model {C}ards for {M}odel {R}eporting.
+\newblock  In Proceedings of the Conference on Fairness, Accountability, and
+  Transparency, Atlanta, GA, USA, 29--31 January 2019. pp. 220--229. %% Please add the location and day month year for the conference. please add ACM location , or add %please add accessed on date.
+\newblock Available online: \url{https://arxiv.org/pdf/1810.03993.pdf} (accessed on 26 February 2020).
+
+\bibitem[Bracke {et~al.}(2019)Bracke, Datta, Jung, and
+  Sen]{bracke2019machine}
+Bracke, P.; Datta, A.; Jung, C.; Sen, S.
+\newblock Machine {L}earning {E}xplainability in {F}inance: {A}n {A}pplication
+  to {D}efault {R}isk {A}nalysis {\bf 2019}.
+\newblock Available online: 
+  \url{https://www.bankofengland.co.uk/-/media/boe/files/working-paper/2019/machine-learning-explainability-in-finance-an-application-to-default-risk-analysis.pdf} (accessed on 26 February 2020).
+
+\bibitem[Friedler {et~al.}(2019)Friedler, Scheidegger, Venkatasubramanian,
+  Choudhary, Hamilton, and Roth]{friedler2019comparative}
+Friedler, S.A.; Scheidegger, C.; Venkatasubramanian, S.; Choudhary, S.;
+  Hamilton, E.P.; Roth, D.
+\newblock A {C}omparative {S}tudy of {F}airness-enhancing {I}nterventions in
+  {M}achine {L}earning.
+\newblock  In Proceedings of the Conference on Fairness, Accountability, and
+  Transparency, Atlanta, GA, USA, 29--31 January 2019; pp. 329--338. 
+\newblock Available online: \url{https://arxiv.org/pdf/1802.04422.pdf} (accessed on 26 February 2020).
+
+\bibitem[Schmidt and Stephens(2019)]{schmidt2019introduction}
+Schmidt, N.; Stephens, B.
+\newblock An {I}ntroduction to {A}rtificial {I}ntelligence and {S}olutions to
+  the {P}roblems of {A}lgorithmic {D}iscrimination.
+\newblock {\em Conf. Consum. Financ. Law Q. Rep.} {\bf 2019},
+  {\em 73},~130--144.
+\newblock Available online: \url{https://arxiv.org/pdf/1911.05755.pdf} (accessed on 26 February 2020).
+
+\bibitem[Hoare(1981)]{hoare19811980}
+Hoare, C.A.R.
+\newblock The 1980 {A}{C}{M} {T}uring {A}ward {L}ecture.
+\newblock {\em Communications} {\bf 1981}.
+\newblock Available online: \url{http://www.cs.fsu.edu/~engelen/courses/COP4610/hoare.pdf}  (accessed on 26 February 2020).
+
+\end{thebibliography}
 

### 2020-02-26 definitions/journalnames.tex
edits after mdpi edits ...

\DeclareOption{acoustics}{	\gdef\@journal{acoustics}	\gdef\@journalshort{Acoustics}	\gdef\@journalfull{Acoustics}	\gdef\@doiabbr{acoustics}	\gdef\@ISSN{2624-599X}	}
\DeclareOption{actuators}{	\gdef\@journal{actuators}	\gdef\@journalshort{Actuators}	\gdef\@journalfull{Actuators}	\gdef\@doiabbr{act}	\gdef\@ISSN{2076-0825}	}
\DeclareOption{addictions}{	\gdef\@journal{addictions}	\gdef\@journalshort{Addictions}	\gdef\@journalfull{Addictions}	\gdef\@doiabbr{}	\gdef\@ISSN{0006-0006}	}
\DeclareOption{admsci}{	\gdef\@journal{admsci}	\gdef\@journalshort{Adm. Sci.}	\gdef\@journalfull{Administrative Sciences}	\gdef\@doiabbr{admsci}	\gdef\@ISSN{2076-3387}	}
\DeclareOption{aerospace}{	\gdef\@journal{aerospace}	\gdef\@journalshort{Aerospace}	\gdef\@journalfull{Aerospace}	\gdef\@doiabbr{aerospace}	\gdef\@ISSN{2226-4310}	}
\DeclareOption{agriculture}{	\gdef\@journal{agriculture}	\gdef\@journalshort{Agriculture}	\gdef\@journalfull{Agriculture}	\gdef\@doiabbr{agriculture}	\gdef\@ISSN{2077-0472}	}
\DeclareOption{agriengineering}{	\gdef\@journal{agriengineering}	\gdef\@journalshort{AgriEngineering}	\gdef\@journalfull{AgriEngineering}	\gdef\@doiabbr{agriengineering}	\gdef\@ISSN{2624-7402}	}
\DeclareOption{agronomy}{	\gdef\@journal{agronomy}	\gdef\@journalshort{Agronomy}	\gdef\@journalfull{Agronomy}	\gdef\@doiabbr{agronomy}	\gdef\@ISSN{2073-4395}	}
\DeclareOption{algorithms}{	\gdef\@journal{algorithms}	\gdef\@journalshort{Algorithms}	\gdef\@journalfull{Algorithms}	\gdef\@doiabbr{a}	\gdef\@ISSN{1999-4893}	}
\DeclareOption{animals}{	\gdef\@journal{animals}	\gdef\@journalshort{Animals}	\gdef\@journalfull{Animals}	\gdef\@doiabbr{ani}	\gdef\@ISSN{2076-2615}	}
\DeclareOption{antibiotics}{	\gdef\@journal{antibiotics}	\gdef\@journalshort{Antibiotics}	\gdef\@journalfull{Antibiotics}	\gdef\@doiabbr{antibiotics}	\gdef\@ISSN{2079-6382}	}
\DeclareOption{antibodies}{	\gdef\@journal{antibodies}	\gdef\@journalshort{Antibodies}	\gdef\@journalfull{Antibodies}	\gdef\@doiabbr{antib}	\gdef\@ISSN{2073-4468}	}
\DeclareOption{antioxidants}{	\gdef\@journal{antioxidants}	\gdef\@journalshort{Antioxidants}	\gdef\@journalfull{Antioxidants}	\gdef\@doiabbr{antiox}	\gdef\@ISSN{2076-3921}	}
\DeclareOption{applsci}{	\gdef\@journal{applsci}	\gdef\@journalshort{Appl. Sci.}	\gdef\@journalfull{Applied Sciences}	\gdef\@doiabbr{app}	\gdef\@ISSN{2076-3417}	}
\DeclareOption{arts}{	\gdef\@journal{arts}	\gdef\@journalshort{Arts}	\gdef\@journalfull{Arts}	\gdef\@doiabbr{arts}	\gdef\@ISSN{2076-0752}	}
\DeclareOption{asc}{	\gdef\@journal{asc}	\gdef\@journalshort{Autom. Syst. Control}	\gdef\@journalfull{Automatic Systems and Control}	\gdef\@doiabbr{}	\gdef\@ISSN{}	}
\DeclareOption{asi}{	\gdef\@journal{asi}	\gdef\@journalshort{Appl. Syst. Innov.}	\gdef\@journalfull{Applied System Innovation}	\gdef\@doiabbr{asi}	\gdef\@ISSN{2571-5577}	}
\DeclareOption{atmosphere}{	\gdef\@journal{atmosphere}	\gdef\@journalshort{Atmosphere}	\gdef\@journalfull{Atmosphere}	\gdef\@doiabbr{atmos}	\gdef\@ISSN{2073-4433}	}
\DeclareOption{atoms}{	\gdef\@journal{atoms}	\gdef\@journalshort{Atoms}	\gdef\@journalfull{Atoms}	\gdef\@doiabbr{atoms}	\gdef\@ISSN{2218-2004}	}
\DeclareOption{axioms}{	\gdef\@journal{axioms}	\gdef\@journalshort{Axioms}	\gdef\@journalfull{Axioms}	\gdef\@doiabbr{axioms}	\gdef\@ISSN{2075-1680}	}
\DeclareOption{batteries}{	\gdef\@journal{batteries}	\gdef\@journalshort{Batteries}	\gdef\@journalfull{Batteries}	\gdef\@doiabbr{batteries}	\gdef\@ISSN{2313-0105}	}
\DeclareOption{bdcc}{	\gdef\@journal{bdcc}	\gdef\@journalshort{Big Data Cogn. Comput.}	\gdef\@journalfull{Big Data and Cognitive Computing}	\gdef\@doiabbr{bdcc}	\gdef\@ISSN{2504-2289}	}
\DeclareOption{behavsci}{ 	\gdef\@journal{behavsci} 	\gdef\@journalshort{Behav. Sci.}	\gdef\@journalfull{Behavioral Sciences}	\gdef\@doiabbr{bs}	\gdef\@ISSN{2076-328X}	}
\DeclareOption{beverages}{	\gdef\@journal{beverages}	\gdef\@journalshort{Beverages}	\gdef\@journalfull{Beverages}	\gdef\@doiabbr{beverages}	\gdef\@ISSN{2306-5710}	}
\DeclareOption{bioengineering}{	\gdef\@journal{bioengineering}	\gdef\@journalshort{Bioengineering}	\gdef\@journalfull{Bioengineering}	\gdef\@doiabbr{bioengineering}	\gdef\@ISSN{2306-5354}	}
\DeclareOption{biology}{	\gdef\@journal{biology}	\gdef\@journalshort{Biology}	\gdef\@journalfull{Biology}	\gdef\@doiabbr{biology}	\gdef\@ISSN{2079-7737}	}
\DeclareOption{biomedicines}{	\gdef\@journal{biomedicines}	\gdef\@journalshort{Biomedicines}	\gdef\@journalfull{Biomedicines}	\gdef\@doiabbr{biomedicines}	\gdef\@ISSN{2227-9059}	}
\DeclareOption{biomimetics}{	\gdef\@journal{biomimetics}	\gdef\@journalshort{Biomimetics}	\gdef\@journalfull{Biomimetics}	\gdef\@doiabbr{biomimetics}	\gdef\@ISSN{2313-7673}	}
\DeclareOption{biomolecules}{	\gdef\@journal{biomolecules}	\gdef\@journalshort{Biomolecules}	\gdef\@journalfull{Biomolecules}	\gdef\@doiabbr{biom}	\gdef\@ISSN{2218-273X}	}
\DeclareOption{biosensors}{	\gdef\@journal{biosensors}	\gdef\@journalshort{Biosensors}	\gdef\@journalfull{Biosensors}	\gdef\@doiabbr{bios}	\gdef\@ISSN{2079-6374}	}
\DeclareOption{brainsci}{ 	\gdef\@journal{brainsci} 	\gdef\@journalshort{Brain Sci.}	\gdef\@journalfull{Brain Sciences}	\gdef\@doiabbr{brainsci}	\gdef\@ISSN{2076-3425}	}
\DeclareOption{buildings}{	\gdef\@journal{buildings}	\gdef\@journalshort{Buildings}	\gdef\@journalfull{Buildings}	\gdef\@doiabbr{buildings}	\gdef\@ISSN{2075-5309}	}
\DeclareOption{cancers}{	\gdef\@journal{cancers}	\gdef\@journalshort{Cancers}	\gdef\@journalfull{Cancers}	\gdef\@doiabbr{cancers}	\gdef\@ISSN{2072-6694}	}
\DeclareOption{carbon}{ 	\gdef\@journal{carbon} 	\gdef\@journalshort{C}	\gdef\@journalfull{C}	\gdef\@doiabbr{c}	\gdef\@ISSN{2311-5629}	}
\DeclareOption{catalysts}{	\gdef\@journal{catalysts}	\gdef\@journalshort{Catalysts}	\gdef\@journalfull{Catalysts}	\gdef\@doiabbr{catal}	\gdef\@ISSN{2073-4344}	}
\DeclareOption{cells}{	\gdef\@journal{cells}	\gdef\@journalshort{Cells}	\gdef\@journalfull{Cells}	\gdef\@doiabbr{cells}	\gdef\@ISSN{2073-4409}	}
\DeclareOption{ceramics}{	\gdef\@journal{ceramics}	\gdef\@journalshort{Ceramics}	\gdef\@journalfull{Ceramics}	\gdef\@doiabbr{ceramics}	\gdef\@ISSN{2571-6131}	}
\DeclareOption{challenges}{	\gdef\@journal{challenges}	\gdef\@journalshort{Challenges}	\gdef\@journalfull{Challenges}	\gdef\@doiabbr{challe}	\gdef\@ISSN{2078-1547}	}
\DeclareOption{chemengineering}{	\gdef\@journal{chemengineering}	\gdef\@journalshort{ChemEngineering}	\gdef\@journalfull{ChemEngineering}	\gdef\@doiabbr{chemengineering}	\gdef\@ISSN{2305-7084}	}
\DeclareOption{chemistry}{	\gdef\@journal{chemistry}	\gdef\@journalshort{Chemistry}	\gdef\@journalfull{Chemistry}	\gdef\@doiabbr{chemistry}	\gdef\@ISSN{2624-8549}	}
\DeclareOption{chemosensors}{	\gdef\@journal{chemosensors}	\gdef\@journalshort{Chemosensors}	\gdef\@journalfull{Chemosensors}	\gdef\@doiabbr{chemosensors}	\gdef\@ISSN{2227-9040}	}
\DeclareOption{children}{	\gdef\@journal{children}	\gdef\@journalshort{Children}	\gdef\@journalfull{Children}	\gdef\@doiabbr{children}	\gdef\@ISSN{2227-9067}	}
\DeclareOption{cleantechnol}{	\gdef\@journal{cleantechnol}	\gdef\@journalshort{Clean Technol.}	\gdef\@journalfull{Clean Technologies}	\gdef\@doiabbr{cleantechnol}	\gdef\@ISSN{2571-8797}	}
\DeclareOption{climate}{	\gdef\@journal{climate}	\gdef\@journalshort{Climate}	\gdef\@journalfull{Climate}	\gdef\@doiabbr{cli}	\gdef\@ISSN{2225-1154}	}
\DeclareOption{clockssleep}{	\gdef\@journal{clockssleep}	\gdef\@journalshort{Clocks\&Sleep}	\gdef\@journalfull{Clocks \& Sleep}	\gdef\@doiabbr{clockssleep}	\gdef\@ISSN{2624-5175}	}
\DeclareOption{cmd}{	\gdef\@journal{cmd}	\gdef\@journalshort{Corros. Mater. Degrad.}	\gdef\@journalfull{Corrosion and Materials Degradation}	\gdef\@doiabbr{cmd}	\gdef\@ISSN{2624-5558}	}
\DeclareOption{coatings}{	\gdef\@journal{coatings}	\gdef\@journalshort{Coatings}	\gdef\@journalfull{Coatings}	\gdef\@doiabbr{coatings}	\gdef\@ISSN{2079-6412}	}
\DeclareOption{colloids}{	\gdef\@journal{colloids}	\gdef\@journalshort{Colloids Interfaces}	\gdef\@journalfull{Colloids Interfaces}	\gdef\@doiabbr{colloids}	\gdef\@ISSN{2504-5377}	}
\DeclareOption{computation}{	\gdef\@journal{computation}	\gdef\@journalshort{Computation}	\gdef\@journalfull{Computation}	\gdef\@doiabbr{computation}	\gdef\@ISSN{2079-3197}	}
\DeclareOption{computers}{	\gdef\@journal{computers}	\gdef\@journalshort{Computers}	\gdef\@journalfull{Computers}	\gdef\@doiabbr{computers}	\gdef\@ISSN{2073-431X}	}
\DeclareOption{condensedmatter}{	\gdef\@journal{condensedmatter}	\gdef\@journalshort{Condens. Matter}	\gdef\@journalfull{Condensed Matter}	\gdef\@doiabbr{condmat}	\gdef\@ISSN{2410-3896}	}
\DeclareOption{cosmetics}{	\gdef\@journal{cosmetics}	\gdef\@journalshort{Cosmetics}	\gdef\@journalfull{Cosmetics}	\gdef\@doiabbr{cosmetics}	\gdef\@ISSN{2079-9284}	}
\DeclareOption{cryptography}{	\gdef\@journal{cryptography}	\gdef\@journalshort{Cryptography}	\gdef\@journalfull{Cryptography}	\gdef\@doiabbr{cryptography}	\gdef\@ISSN{2410-387X}	}
\DeclareOption{crystals}{	\gdef\@journal{crystals}	\gdef\@journalshort{Crystals}	\gdef\@journalfull{Crystals}	\gdef\@doiabbr{cryst}	\gdef\@ISSN{2073-4352}	}
\DeclareOption{dairy}{	\gdef\@journal{dairy}	\gdef\@journalshort{Dairy}	\gdef\@journalfull{Dairy}	\gdef\@doiabbr{dairy}	\gdef\@ISSN{2624-862X}	}
\DeclareOption{data}{	\gdef\@journal{data}	\gdef\@journalshort{Data}	\gdef\@journalfull{Data}	\gdef\@doiabbr{data}	\gdef\@ISSN{2306-5729}	}
\DeclareOption{dentistry}{	\gdef\@journal{dentistry}	\gdef\@journalshort{Dent. J.}	\gdef\@journalfull{Dentistry Journal}	\gdef\@doiabbr{dj}	\gdef\@ISSN{2304-6767}	}
\DeclareOption{designs}{ 	\gdef\@journal{designs} 	\gdef\@journalshort{Designs}	\gdef\@journalfull{Designs}	\gdef\@doiabbr{designs}	\gdef\@ISSN{2411-9660}	}
\DeclareOption{diagnostics}{	\gdef\@journal{diagnostics}	\gdef\@journalshort{Diagnostics}	\gdef\@journalfull{Diagnostics}	\gdef\@doiabbr{diagnostics}	\gdef\@ISSN{2075-4418}	}
\DeclareOption{diseases}{	\gdef\@journal{diseases}	\gdef\@journalshort{Diseases}	\gdef\@journalfull{Diseases}	\gdef\@doiabbr{diseases}	\gdef\@ISSN{2079-9721}	}
\DeclareOption{diversity}{	\gdef\@journal{diversity}	\gdef\@journalshort{Diversity}	\gdef\@journalfull{Diversity}	\gdef\@doiabbr{d}	\gdef\@ISSN{1424-2818}	}
\DeclareOption{drones}{	\gdef\@journal{drones}	\gdef\@journalshort{Drones}	\gdef\@journalfull{Drones}	\gdef\@doiabbr{drones}	\gdef\@ISSN{2504-446X}	}
\DeclareOption{econometrics}{	\gdef\@journal{econometrics}	\gdef\@journalshort{Econometrics}	\gdef\@journalfull{Econometrics}	\gdef\@doiabbr{econometrics}	\gdef\@ISSN{2225-1146}	}
\DeclareOption{economies}{	\gdef\@journal{economies}	\gdef\@journalshort{Economies}	\gdef\@journalfull{Economies}	\gdef\@doiabbr{economies}	\gdef\@ISSN{2227-7099}	}
\DeclareOption{education}{	\gdef\@journal{education}	\gdef\@journalshort{Educ. Sci.}	\gdef\@journalfull{Education Sciences}	\gdef\@doiabbr{educsci}	\gdef\@ISSN{2227-7102}	}
\DeclareOption{ejihpe}{	\gdef\@journal{ejihpe}	\gdef\@journalshort{Eur. J. Investig. Health Psychol. Educ.}	\gdef\@journalfull{European Journal of Investigation in Health, Psychology and Education}	\gdef\@doiabbr{ejihpe}	\gdef\@ISSN{2254-9625}	}
\DeclareOption{electrochem}{	\gdef\@journal{electrochem}	\gdef\@journalshort{Electrochem}	\gdef\@journalfull{Electrochem}	\gdef\@doiabbr{electrochem}	\gdef\@ISSN{}	}
\DeclareOption{electronics}{	\gdef\@journal{electronics}	\gdef\@journalshort{Electronics}	\gdef\@journalfull{Electronics}	\gdef\@doiabbr{electronics}	\gdef\@ISSN{2079-9292}	}
\DeclareOption{energies}{	\gdef\@journal{energies}	\gdef\@journalshort{Energies}	\gdef\@journalfull{Energies}	\gdef\@doiabbr{en}	\gdef\@ISSN{1996-1073}	}
\DeclareOption{entropy}{	\gdef\@journal{entropy}	\gdef\@journalshort{Entropy}	\gdef\@journalfull{Entropy}	\gdef\@doiabbr{e}	\gdef\@ISSN{1099-4300}	}
\DeclareOption{environments}{	\gdef\@journal{environments}	\gdef\@journalshort{Environments}	\gdef\@journalfull{Environments}	\gdef\@doiabbr{environments}	\gdef\@ISSN{2076-3298}	}
\DeclareOption{epigenomes}{	\gdef\@journal{epigenomes}	\gdef\@journalshort{Epigenomes}	\gdef\@journalfull{Epigenomes}	\gdef\@doiabbr{epigenomes}	\gdef\@ISSN{2075-4655}	}
\DeclareOption{est}{	\gdef\@journal{est}	\gdef\@journalshort{Electrochem. Sci. Technol.}	\gdef\@journalfull{Electrochemical Science and Technology}	\gdef\@doiabbr{}	\gdef\@ISSN{}	}
\DeclareOption{fermentation}{	\gdef\@journal{fermentation}	\gdef\@journalshort{Fermentation}	\gdef\@journalfull{Fermentation}	\gdef\@doiabbr{fermentation}	\gdef\@ISSN{2311-5637}	}
\DeclareOption{fibers}{	\gdef\@journal{fibers}	\gdef\@journalshort{Fibers}	\gdef\@journalfull{Fibers}	\gdef\@doiabbr{fib}	\gdef\@ISSN{2079-6439}	}
\DeclareOption{fire}{	\gdef\@journal{fire}	\gdef\@journalshort{Fire}	\gdef\@journalfull{Fire}	\gdef\@doiabbr{fire}	\gdef\@ISSN{2571-6255}	}
\DeclareOption{fishes}{	\gdef\@journal{fishes}	\gdef\@journalshort{Fishes}	\gdef\@journalfull{Fishes}	\gdef\@doiabbr{fishes}	\gdef\@ISSN{2410-3888}	}
\DeclareOption{fluids}{	\gdef\@journal{fluids}	\gdef\@journalshort{Fluids}	\gdef\@journalfull{Fluids}	\gdef\@doiabbr{fluids}	\gdef\@ISSN{2311-5521}	}
\DeclareOption{foods}{	\gdef\@journal{foods}	\gdef\@journalshort{Foods}	\gdef\@journalfull{Foods}	\gdef\@doiabbr{foods}	\gdef\@ISSN{2304-8158}	}
\DeclareOption{forecasting}{	\gdef\@journal{forecasting}	\gdef\@journalshort{Forecasting}	\gdef\@journalfull{Forecasting}	\gdef\@doiabbr{forecast}	\gdef\@ISSN{2571-9394}	}
\DeclareOption{forests}{	\gdef\@journal{forests}	\gdef\@journalshort{Forests}	\gdef\@journalfull{Forests}	\gdef\@doiabbr{f}	\gdef\@ISSN{1999-4907}	}
\DeclareOption{fractalfract}{	\gdef\@journal{fractalfract}	\gdef\@journalshort{Fractal Fract.}	\gdef\@journalfull{Fractal and Fractional}	\gdef\@doiabbr{fractalfract}	\gdef\@ISSN{2504-3110}	}
\DeclareOption{futureinternet}{	\gdef\@journal{futureinternet}	\gdef\@journalshort{Future Internet}	\gdef\@journalfull{Future Internet}	\gdef\@doiabbr{fi}	\gdef\@ISSN{1999-5903}	}
\DeclareOption{futurephys}{	\gdef\@journal{futurephys}	\gdef\@journalshort{Future Phys.}	\gdef\@journalfull{Future Physics}	\gdef\@doiabbr{futurephys}	\gdef\@ISSN{2624-6503}	}
\DeclareOption{galaxies}{	\gdef\@journal{galaxies}	\gdef\@journalshort{Galaxies}	\gdef\@journalfull{Galaxies}	\gdef\@doiabbr{galaxies}	\gdef\@ISSN{2075-4434}	}
\DeclareOption{games}{	\gdef\@journal{games}	\gdef\@journalshort{Games}	\gdef\@journalfull{Games}	\gdef\@doiabbr{g}	\gdef\@ISSN{2073-4336}	}
\DeclareOption{gastrointestdisord}{	\gdef\@journal{gastrointestdisord}	\gdef\@journalshort{Gastrointest. Disord.}	\gdef\@journalfull{Gastrointestinal Disorders}	\gdef\@doiabbr{gidisord}	\gdef\@ISSN{2624-5647}	}
\DeclareOption{gels}{	\gdef\@journal{gels}	\gdef\@journalshort{Gels}	\gdef\@journalfull{Gels}	\gdef\@doiabbr{gels}	\gdef\@ISSN{2310-2861}	}
\DeclareOption{genealogy}{	\gdef\@journal{genealogy}	\gdef\@journalshort{Genealogy}	\gdef\@journalfull{Genealogy}	\gdef\@doiabbr{genealogy}	\gdef\@ISSN{2313-5778}	}
\DeclareOption{genes}{	\gdef\@journal{genes}	\gdef\@journalshort{Genes}	\gdef\@journalfull{Genes}	\gdef\@doiabbr{genes}	\gdef\@ISSN{2073-4425}	}
\DeclareOption{geohazards}{	\gdef\@journal{geohazards}	\gdef\@journalshort{GeoHazards}	\gdef\@journalfull{GeoHazards}	\gdef\@doiabbr{geohazards}	\gdef\@ISSN{2624-795X}	}
\DeclareOption{geosciences}{	\gdef\@journal{geosciences}	\gdef\@journalshort{Geosciences}	\gdef\@journalfull{Geosciences}	\gdef\@doiabbr{geosciences}	\gdef\@ISSN{2076-3263}	}
\DeclareOption{geriatrics}{	\gdef\@journal{geriatrics}	\gdef\@journalshort{Geriatrics}	\gdef\@journalfull{Geriatrics}	\gdef\@doiabbr{geriatrics}	\gdef\@ISSN{2308-3417}	}
\DeclareOption{hazardousmatters}{	\gdef\@journal{hazardousmatters}	\gdef\@journalshort{Hazard. Matters}	\gdef\@journalfull{Hazardous Matters}	\gdef\@doiabbr{}	\gdef\@ISSN{0014-0014}	}
\DeclareOption{healthcare}{	\gdef\@journal{healthcare}	\gdef\@journalshort{Healthcare}	\gdef\@journalfull{Healthcare}	\gdef\@doiabbr{healthcare}	\gdef\@ISSN{2227-9032}	}
\DeclareOption{heritage}{	\gdef\@journal{heritage}	\gdef\@journalshort{Heritage}	\gdef\@journalfull{Heritage}	\gdef\@doiabbr{heritage}	\gdef\@ISSN{2571-9408}	}
\DeclareOption{highthroughput}{	\gdef\@journal{highthroughput}	\gdef\@journalshort{High-Throughput}	\gdef\@journalfull{High-Throughput}	\gdef\@doiabbr{ht}	\gdef\@ISSN{2571-5135}	}
\DeclareOption{horticulturae}{	\gdef\@journal{horticulturae}	\gdef\@journalshort{Horticulturae}	\gdef\@journalfull{Horticulturae}	\gdef\@doiabbr{horticulturae}	\gdef\@ISSN{2311-7524}	}
\DeclareOption{humanities}{	\gdef\@journal{humanities}	\gdef\@journalshort{Humanities}	\gdef\@journalfull{Humanities}	\gdef\@doiabbr{h}	\gdef\@ISSN{2076-0787}	}
\DeclareOption{hydrology}{	\gdef\@journal{hydrology}	\gdef\@journalshort{Hydrology}	\gdef\@journalfull{Hydrology}	\gdef\@doiabbr{hydrology}	\gdef\@ISSN{2306-5338}	}
\DeclareOption{ijerph}{	\gdef\@journal{ijerph}	\gdef\@journalshort{Int. J. Environ. Res. Public Health}	\gdef\@journalfull{International Journal of Environmental Research and Public Health}	\gdef\@doiabbr{ijerph}	\gdef\@ISSN{1660-4601}	}
\DeclareOption{ijfs}{	\gdef\@journal{ijfs}	\gdef\@journalshort{Int. J. Financial Stud.}	\gdef\@journalfull{International Journal of Financial Studies}	\gdef\@doiabbr{ijfs}	\gdef\@ISSN{2227-7072}	}
\DeclareOption{ijgi}{	\gdef\@journal{ijgi}	\gdef\@journalshort{ISPRS Int. J. Geo-Inf.}	\gdef\@journalfull{ISPRS International Journal of Geo-Information}	\gdef\@doiabbr{ijgi}	\gdef\@ISSN{2220-9964}	}
\DeclareOption{ijms}{	\gdef\@journal{ijms}	\gdef\@journalshort{Int. J. Mol. Sci.}	\gdef\@journalfull{International Journal of Molecular Sciences}	\gdef\@doiabbr{ijms}	\gdef\@ISSN{1422-0067}	}
\DeclareOption{ijtpp}{	\gdef\@journal{ijtpp}	\gdef\@journalshort{Int. J. Turbomach. Propuls. Power}	\gdef\@journalfull{International Journal of Turbomachinery, Propulsion and Power}	\gdef\@doiabbr{ijtpp}	\gdef\@ISSN{2504-186X}	}
\DeclareOption{informatics}{	\gdef\@journal{informatics}	\gdef\@journalshort{Informatics}	\gdef\@journalfull{Informatics}	\gdef\@doiabbr{informatics}	\gdef\@ISSN{2227-9709}	}
\DeclareOption{information}{	\gdef\@journal{information}	\gdef\@journalshort{Information}	\gdef\@journalfull{Information}	\gdef\@doiabbr{info}	\gdef\@ISSN{2078-2489}	}
\DeclareOption{infrastructures}{	\gdef\@journal{infrastructures}	\gdef\@journalshort{Infrastructures}	\gdef\@journalfull{Infrastructures}	\gdef\@doiabbr{infrastructures}	\gdef\@ISSN{2412-3811}	}
\DeclareOption{inorganics}{	\gdef\@journal{inorganics}	\gdef\@journalshort{Inorganics}	\gdef\@journalfull{Inorganics}	\gdef\@doiabbr{inorganics}	\gdef\@ISSN{2304-6740}	}
\DeclareOption{insects}{	\gdef\@journal{insects}	\gdef\@journalshort{Insects}	\gdef\@journalfull{Insects}	\gdef\@doiabbr{insects}	\gdef\@ISSN{2075-4450}	}
\DeclareOption{instruments}{	\gdef\@journal{instruments}	\gdef\@journalshort{Instruments}	\gdef\@journalfull{Instruments}	\gdef\@doiabbr{instruments}	\gdef\@ISSN{2410-390X}	}
\DeclareOption{inventions}{	\gdef\@journal{inventions}	\gdef\@journalshort{Inventions}	\gdef\@journalfull{Inventions}	\gdef\@doiabbr{inventions}	\gdef\@ISSN{2411-5134}	}
\DeclareOption{iot}{	\gdef\@journal{iot}	\gdef\@journalshort{IoT}	\gdef\@journalfull{IoT}	\gdef\@doiabbr{iot}	\gdef\@ISSN{2624-831X}	}
\DeclareOption{j}{	\gdef\@journal{j}	\gdef\@journalshort{J}	\gdef\@journalfull{J}	\gdef\@doiabbr{j}	\gdef\@ISSN{2571-8800}	}
\DeclareOption{jcdd}{	\gdef\@journal{jcdd}	\gdef\@journalshort{J. Cardiovasc. Dev. Dis.}	\gdef\@journalfull{Journal of Cardiovascular Development and Disease}	\gdef\@doiabbr{jcdd}	\gdef\@ISSN{2308-3425}	}
\DeclareOption{jcm}{	\gdef\@journal{jcm}	\gdef\@journalshort{J. Clin. Med.}	\gdef\@journalfull{Journal of Clinical Medicine}	\gdef\@doiabbr{jcm}	\gdef\@ISSN{2077-0383}	}
\DeclareOption{jcp}{	\gdef\@journal{jcp}	\gdef\@journalshort{J. Cybersecur. Priv.}	\gdef\@journalfull{Journal of Cybersecurity and Privacy}	\gdef\@doiabbr{jcp}	\gdef\@ISSN{2624-800X}	}
\DeclareOption{jcs}{	\gdef\@journal{jcs}	\gdef\@journalshort{J. Compos. Sci.}	\gdef\@journalfull{Journal of Composites Science}	\gdef\@doiabbr{jcs}	\gdef\@ISSN{2504-477X}	}
\DeclareOption{jdb}{	\gdef\@journal{jdb}	\gdef\@journalshort{J. Dev. Biol.}	\gdef\@journalfull{Journal of Developmental Biology}	\gdef\@doiabbr{jdb}	\gdef\@ISSN{2221-3759}	}
\DeclareOption{jfb}{	\gdef\@journal{jfb}	\gdef\@journalshort{J. Funct. Biomater.}   	\gdef\@journalfull{Journal of Functional Biomaterials}   	\gdef\@doiabbr{jfb}	\gdef\@ISSN{2079-4983}	}
\DeclareOption{jfmk}{	\gdef\@journal{jfmk}	\gdef\@journalshort{J. Funct. Morphol. Kinesiol.}   	\gdef\@journalfull{Journal of Functional Morphology and Kinesiology}   	\gdef\@doiabbr{jfmk}	\gdef\@ISSN{2411-5142}	}
\DeclareOption{jimaging}{	\gdef\@journal{jimaging}	\gdef\@journalshort{J. Imaging}	\gdef\@journalfull{Journal of Imaging}	\gdef\@doiabbr{jimaging}	\gdef\@ISSN{2313-433X}	}
\DeclareOption{jintelligence}{	\gdef\@journal{jintelligence}	\gdef\@journalshort{J. Intell.}	\gdef\@journalfull{Journal of Intelligence}	\gdef\@doiabbr{jintelligence}	\gdef\@ISSN{2079-3200}	}
\DeclareOption{jlpea}{	\gdef\@journal{jlpea}	\gdef\@journalshort{J. Low Power Electron. Appl.}	\gdef\@journalfull{Journal of Low Power Electronics and Applications}	\gdef\@doiabbr{jlpea}	\gdef\@ISSN{2079-9268}	}
\DeclareOption{jmmp}{	\gdef\@journal{jmmp}	\gdef\@journalshort{J. Manuf. Mater. Process.}	\gdef\@journalfull{Journal of Manufacturing and Materials Processing}	\gdef\@doiabbr{jmmp}	\gdef\@ISSN{2504-4494}	}
\DeclareOption{jmse}{	\gdef\@journal{jmse}	\gdef\@journalshort{J. Mar. Sci. Eng.}	\gdef\@journalfull{Journal of Marine Science and Engineering}	\gdef\@doiabbr{jmse}	\gdef\@ISSN{2077-1312}	}
\DeclareOption{jnt}{	\gdef\@journal{jnt}	\gdef\@journalshort{J. Nanotheranostics}	\gdef\@journalfull{Journal of Nanotheranostics}	\gdef\@doiabbr{jnt}	\gdef\@ISSN{2624-845X}	}
\DeclareOption{jof}{	\gdef\@journal{jof}	\gdef\@journalshort{J. Fungi}	\gdef\@journalfull{Journal of Fungi}	\gdef\@doiabbr{jof}	\gdef\@ISSN{2309-608X}	}
\DeclareOption{joitmc}{	\gdef\@journal{joitmc}	\gdef\@journalshort{J. Open Innov. Technol. Mark. Complex.}	\gdef\@journalfull{Journal of Open Innovation: Technology, Market, and Complexity}	\gdef\@doiabbr{joitmc}	\gdef\@ISSN{2199-8531}	}
\DeclareOption{jpm}{	\gdef\@journal{jpm}	\gdef\@journalshort{J. Pers. Med.}	\gdef\@journalfull{Journal of Personalized Medicine}	\gdef\@doiabbr{jpm}	\gdef\@ISSN{2075-4426}	}
\DeclareOption{jrfm}{	\gdef\@journal{jrfm}	\gdef\@journalshort{J. Risk Financial Manag.}	\gdef\@journalfull{Journal of Risk and Financial Management}	\gdef\@doiabbr{jrfm}	\gdef\@ISSN{1911-8074}	}
\DeclareOption{jsan}{	\gdef\@journal{jsan}	\gdef\@journalshort{J. Sens. Actuator Netw.}	\gdef\@journalfull{Journal of Sensor and Actuator Networks}	\gdef\@doiabbr{jsan}	\gdef\@ISSN{2224-2708}	}
\DeclareOption{land}{	\gdef\@journal{land}	\gdef\@journalshort{Land}	\gdef\@journalfull{Land}	\gdef\@doiabbr{land}	\gdef\@ISSN{2073-445X}	}
\DeclareOption{languages}{	\gdef\@journal{languages}	\gdef\@journalshort{Languages}	\gdef\@journalfull{Languages}	\gdef\@doiabbr{languages}	\gdef\@ISSN{2226-471X}	}
\DeclareOption{laws}{	\gdef\@journal{laws}	\gdef\@journalshort{Laws}	\gdef\@journalfull{Laws}	\gdef\@doiabbr{laws}	\gdef\@ISSN{2075-471X}	}
\DeclareOption{life}{	\gdef\@journal{life}	\gdef\@journalshort{Life}	\gdef\@journalfull{Life}	\gdef\@doiabbr{life}	\gdef\@ISSN{2075-1729}	}
\DeclareOption{literature}{	\gdef\@journal{literature}	\gdef\@journalshort{Literature}	\gdef\@journalfull{Literature}	\gdef\@doiabbr{}	\gdef\@ISSN{2410-9789}	}
\DeclareOption{logistics}{	\gdef\@journal{logistics}	\gdef\@journalshort{Logistics}	\gdef\@journalfull{Logistics}	\gdef\@doiabbr{logistics}	\gdef\@ISSN{2305-6290}	}
\DeclareOption{lubricants}{	\gdef\@journal{lubricants}	\gdef\@journalshort{Lubricants}	\gdef\@journalfull{Lubricants}	\gdef\@doiabbr{lubricants}	\gdef\@ISSN{2075-4442}	}
\DeclareOption{machines}{	\gdef\@journal{machines}	\gdef\@journalshort{Machines}	\gdef\@journalfull{Machines}	\gdef\@doiabbr{machines}	\gdef\@ISSN{2075-1702}	}
\DeclareOption{magnetochemistry}{	\gdef\@journal{magnetochemistry}	\gdef\@journalshort{Magnetochemistry}	\gdef\@journalfull{Magnetochemistry}	\gdef\@doiabbr{magnetochemistry}	\gdef\@ISSN{2312-7481}	}
\DeclareOption{make}{	\gdef\@journal{make}	\gdef\@journalshort{Mach. Learn. Knowl. Extr.}	\gdef\@journalfull{Machine Learning and Knowledge Extraction}	\gdef\@doiabbr{make}	\gdef\@ISSN{2504-4990}	}
\DeclareOption{marinedrugs}{	\gdef\@journal{marinedrugs}	\gdef\@journalshort{Mar. Drugs}	\gdef\@journalfull{Marine Drugs}	\gdef\@doiabbr{md}	\gdef\@ISSN{1660-3397}	}
\DeclareOption{materials}{	\gdef\@journal{materials}	\gdef\@journalshort{Materials}	\gdef\@journalfull{Materials}	\gdef\@doiabbr{ma}	\gdef\@ISSN{1996-1944}	}
\DeclareOption{mathematics}{	\gdef\@journal{mathematics}	\gdef\@journalshort{Mathematics}	\gdef\@journalfull{Mathematics}	\gdef\@doiabbr{math}	\gdef\@ISSN{2227-7390}	}
\DeclareOption{mca}{	\gdef\@journal{mca}	\gdef\@journalshort{Math. Comput. Appl.}	\gdef\@journalfull{Mathematical and Computational Applications}	\gdef\@doiabbr{mca}	\gdef\@ISSN{2297-8747}	}
\DeclareOption{medicina}{	\gdef\@journal{medicina}	\gdef\@journalshort{Medicina}	\gdef\@journalfull{Medicina}	\gdef\@doiabbr{medicina}	\gdef\@ISSN{1010-660X}	}
\DeclareOption{medicines}{	\gdef\@journal{medicines}	\gdef\@journalshort{Medicines}	\gdef\@journalfull{Medicines}	\gdef\@doiabbr{medicines}	\gdef\@ISSN{2305-6320}	}
\DeclareOption{medsci}{	\gdef\@journal{medsci}	\gdef\@journalshort{Med. Sci.}	\gdef\@journalfull{Medical Sciences}	\gdef\@doiabbr{medsci}	\gdef\@ISSN{2076-3271}	}
\DeclareOption{membranes}{	\gdef\@journal{membranes}	\gdef\@journalshort{Membranes}	\gdef\@journalfull{Membranes}	\gdef\@doiabbr{membranes}	\gdef\@ISSN{2077-0375}	}
\DeclareOption{metabolites}{	\gdef\@journal{metabolites}	\gdef\@journalshort{Metabolites}	\gdef\@journalfull{Metabolites}	\gdef\@doiabbr{metabo}	\gdef\@ISSN{2218-1989}	}
\DeclareOption{metals}{	\gdef\@journal{metals}	\gdef\@journalshort{Metals}	\gdef\@journalfull{Metals}	\gdef\@doiabbr{met}	\gdef\@ISSN{2075-4701}	}
\DeclareOption{microarrays}{	\gdef\@journal{microarrays}	\gdef\@journalshort{Microarrays}	\gdef\@journalfull{Microarrays}	\gdef\@doiabbr{}	\gdef\@ISSN{2076-3905}	}
\DeclareOption{micromachines}{	\gdef\@journal{micromachines}	\gdef\@journalshort{Micromachines}	\gdef\@journalfull{Micromachines}	\gdef\@doiabbr{mi}	\gdef\@ISSN{2072-666X}	}
\DeclareOption{microorganisms}{	\gdef\@journal{microorganisms}	\gdef\@journalshort{Microorganisms}	\gdef\@journalfull{Microorganisms}	\gdef\@doiabbr{microorganisms}	\gdef\@ISSN{2076-2607}	}
\DeclareOption{minerals}{	\gdef\@journal{minerals}	\gdef\@journalshort{Minerals}	\gdef\@journalfull{Minerals}	\gdef\@doiabbr{min}	\gdef\@ISSN{2075-163X}	}
\DeclareOption{modelling}{	\gdef\@journal{modelling}	\gdef\@journalshort{Modelling}	\gdef\@journalfull{Modelling}	\gdef\@doiabbr{}	\gdef\@ISSN{0012-0012}	}
\DeclareOption{molbank}{	\gdef\@journal{molbank}	\gdef\@journalshort{Molbank}	\gdef\@journalfull{Molbank}	\gdef\@doiabbr{M}	\gdef\@ISSN{1422-8599}	}
\DeclareOption{molecules}{	\gdef\@journal{molecules}	\gdef\@journalshort{Molecules}	\gdef\@journalfull{Molecules}	\gdef\@doiabbr{molecules}	\gdef\@ISSN{1420-3049}	}
\DeclareOption{mps}{	\gdef\@journal{mps}	\gdef\@journalshort{Methods Protoc.}	\gdef\@journalfull{Methods and Protocols}	\gdef\@doiabbr{mps}	\gdef\@ISSN{2409-9279}	}
\DeclareOption{mti}{	\gdef\@journal{mti}	\gdef\@journalshort{Multimodal Technol. Interact.}	\gdef\@journalfull{Multimodal Technologies and Interaction}	\gdef\@doiabbr{mti}	\gdef\@ISSN{2414-4088}	}
\DeclareOption{nanomaterials}{	\gdef\@journal{nanomaterials}	\gdef\@journalshort{Nanomaterials}	\gdef\@journalfull{Nanomaterials}	\gdef\@doiabbr{nano}	\gdef\@ISSN{2079-4991}	}
\DeclareOption{ncrna}{	\gdef\@journal{ncrna}	\gdef\@journalshort{Non-coding RNA}	\gdef\@journalfull{Non-coding RNA}	\gdef\@doiabbr{ncrna}	\gdef\@ISSN{2311-553X}	}
\DeclareOption{ijns}{	\gdef\@journal{ijns}	\gdef\@journalshort{Int. J. Neonatal Screen.}	\gdef\@journalfull{International Journal of Neonatal Screening}	\gdef\@doiabbr{ijns}	\gdef\@ISSN{2409-515X}	}
\DeclareOption{neuroglia}{	\gdef\@journal{neuroglia}	\gdef\@journalshort{Neuroglia}	\gdef\@journalfull{Neuroglia}	\gdef\@doiabbr{neuroglia}	\gdef\@ISSN{2571-6980}	}
\DeclareOption{nitrogen}{	\gdef\@journal{nitrogen}	\gdef\@journalshort{Nitrogen}	\gdef\@journalfull{Nitrogen}	\gdef\@doiabbr{nitrogen}	\gdef\@ISSN{2504-3129}	}
\DeclareOption{notspecified}{	\gdef\@journal{notspecified}	\gdef\@journalshort{Journal Not Specified}	\gdef\@journalfull{Journal Not Specified}	\gdef\@doiabbr{}	\gdef\@ISSN{}	}
\DeclareOption{nutrients}{	\gdef\@journal{nutrients}	\gdef\@journalshort{Nutrients}	\gdef\@journalfull{Nutrients}	\gdef\@doiabbr{nu}	\gdef\@ISSN{2072-6643}	}
\DeclareOption{ohbm}{	\gdef\@journal{ohbm}	\gdef\@journalshort{J. Otorhinolaryngol. Hear. Balance Med.}	\gdef\@journalfull{Journal of Otorhinolaryngology, Hearing and Balance Medicine}	\gdef\@doiabbr{ohbm}	\gdef\@ISSN{2504-463X}	}
\DeclareOption{optics}{	\gdef\@journal{optics}	\gdef\@journalshort{Optics}	\gdef\@journalfull{Optics}	\gdef\@doiabbr{opt}	\gdef\@ISSN{}	}
\DeclareOption{particles}{	\gdef\@journal{particles}	\gdef\@journalshort{Particles}	\gdef\@journalfull{Particles}	\gdef\@doiabbr{particles}	\gdef\@ISSN{2571-712X}	}
\DeclareOption{pathogens}{	\gdef\@journal{pathogens}	\gdef\@journalshort{Pathogens}	\gdef\@journalfull{Pathogens}	\gdef\@doiabbr{pathogens}	\gdef\@ISSN{2076-0817}	}
\DeclareOption{pharmaceuticals}{	\gdef\@journal{pharmaceuticals}	\gdef\@journalshort{Pharmaceuticals}	\gdef\@journalfull{Pharmaceuticals}	\gdef\@doiabbr{ph}	\gdef\@ISSN{1424-8247}	}
\DeclareOption{pharmaceutics}{	\gdef\@journal{pharmaceutics}	\gdef\@journalshort{Pharmaceutics}	\gdef\@journalfull{Pharmaceutics}	\gdef\@doiabbr{pharmaceutics}	\gdef\@ISSN{1999-4923}	}
\DeclareOption{pharmacy}{	\gdef\@journal{pharmacy}	\gdef\@journalshort{Pharmacy}	\gdef\@journalfull{Pharmacy}	\gdef\@doiabbr{pharmacy}	\gdef\@ISSN{2226-4787}	}
\DeclareOption{philosophies}{	\gdef\@journal{philosophies}	\gdef\@journalshort{Philosophies}	\gdef\@journalfull{Philosophies}	\gdef\@doiabbr{philosophies}	\gdef\@ISSN{2409-9287}	}
\DeclareOption{photonics}{	\gdef\@journal{photonics}	\gdef\@journalshort{Photonics}	\gdef\@journalfull{Photonics}	\gdef\@doiabbr{photonics}	\gdef\@ISSN{2304-6732}	}
\DeclareOption{physics}{	\gdef\@journal{physics}	\gdef\@journalshort{Physics}	\gdef\@journalfull{Physics}	\gdef\@doiabbr{physics}	\gdef\@ISSN{2624-8174}	}
\DeclareOption{plants}{	\gdef\@journal{plants}	\gdef\@journalshort{Plants}	\gdef\@journalfull{Plants}	\gdef\@doiabbr{plants}	\gdef\@ISSN{2223-7747}	}
\DeclareOption{plasma}{	\gdef\@journal{plasma}	\gdef\@journalshort{Plasma}	\gdef\@journalfull{Plasma}	\gdef\@doiabbr{plasma}	\gdef\@ISSN{2571-6182}	}
\DeclareOption{polymers}{	\gdef\@journal{polymers}	\gdef\@journalshort{Polymers}	\gdef\@journalfull{Polymers}	\gdef\@doiabbr{polym}	\gdef\@ISSN{2073-4360}	}
\DeclareOption{polysaccharides}{	\gdef\@journal{polysaccharides}	\gdef\@journalshort{Polysaccharides}	\gdef\@journalfull{Polysaccharides}	\gdef\@doiabbr{}	\gdef\@ISSN{}	}
\DeclareOption{preprints}{ 	\gdef\@journal{preprints} 	\gdef\@journalshort{Preprints}	\gdef\@journalfull{Preprints}	\gdef\@doiabbr{}	\gdef\@ISSN{}	}
\DeclareOption{proceedings}{	\gdef\@journal{proceedings}	\gdef\@journalshort{Proceedings}	\gdef\@journalfull{Proceedings}	\gdef\@doiabbr{proceedings}	\gdef\@ISSN{2504-3900}	}
\DeclareOption{processes}{	\gdef\@journal{processes}	\gdef\@journalshort{Processes}	\gdef\@journalfull{Processes}	\gdef\@doiabbr{pr}	\gdef\@ISSN{2227-9717}	}
\DeclareOption{prosthesis}{	\gdef\@journal{prosthesis}	\gdef\@journalshort{Prosthesis}	\gdef\@journalfull{Prosthesis}	\gdef\@doiabbr{prosthesis}	\gdef\@ISSN{}	}
\DeclareOption{proteomes}{	\gdef\@journal{proteomes}	\gdef\@journalshort{Proteomes}	\gdef\@journalfull{Proteomes}	\gdef\@doiabbr{proteomes}	\gdef\@ISSN{2227-7382}	}
\DeclareOption{psych}{	\gdef\@journal{psych}	\gdef\@journalshort{Psych}	\gdef\@journalfull{Psych}	\gdef\@doiabbr{psych}	\gdef\@ISSN{2624-8611}	}
\DeclareOption{publications}{	\gdef\@journal{publications}	\gdef\@journalshort{Publications}	\gdef\@journalfull{Publications}	\gdef\@doiabbr{publications}	\gdef\@ISSN{2304-6775}	}
\DeclareOption{quantumrep}{	\gdef\@journal{quantumrep}	\gdef\@journalshort{Quantum Rep.}	\gdef\@journalfull{Quantum Reports}	\gdef\@doiabbr{quantum}	\gdef\@ISSN{2624-960X}	}
\DeclareOption{quaternary}{	\gdef\@journal{quaternary}	\gdef\@journalshort{Quaternary}	\gdef\@journalfull{Quaternary}	\gdef\@doiabbr{quat}	\gdef\@ISSN{2571-550X}	}
\DeclareOption{qubs}{	\gdef\@journal{qubs}	\gdef\@journalshort{Quantum Beam Sci.}	\gdef\@journalfull{Quantum Beam Science}	\gdef\@doiabbr{qubs}	\gdef\@ISSN{2412-382X}	}
\DeclareOption{reactions}{	\gdef\@journal{reactions}	\gdef\@journalshort{Reactions}	\gdef\@journalfull{Reactions}	\gdef\@doiabbr{reactions}	\gdef\@ISSN{2624-781X}	}
\DeclareOption{recycling}{	\gdef\@journal{recycling}	\gdef\@journalshort{Recycling}	\gdef\@journalfull{Recycling}	\gdef\@doiabbr{recycling}	\gdef\@ISSN{2313-4321}	}
\DeclareOption{religions}{	\gdef\@journal{religions}	\gdef\@journalshort{Religions}	\gdef\@journalfull{Religions}	\gdef\@doiabbr{rel}	\gdef\@ISSN{2077-1444}	}
\DeclareOption{remotesensing}{	\gdef\@journal{remotesensing}	\gdef\@journalshort{Remote Sens.}	\gdef\@journalfull{Remote Sensing}	\gdef\@doiabbr{rs}	\gdef\@ISSN{2072-4292}	}
\DeclareOption{reports}{	\gdef\@journal{reports}	\gdef\@journalshort{Reports}	\gdef\@journalfull{Reports}	\gdef\@doiabbr{reports}	\gdef\@ISSN{2571-841X}	}
\DeclareOption{resources}{	\gdef\@journal{resources}	\gdef\@journalshort{Resources}	\gdef\@journalfull{Resources}	\gdef\@doiabbr{resources}	\gdef\@ISSN{2079-9276}	}
\DeclareOption{risks}{	\gdef\@journal{risks}	\gdef\@journalshort{Risks}	\gdef\@journalfull{Risks}	\gdef\@doiabbr{risks}	\gdef\@ISSN{2227-9091}	}
\DeclareOption{robotics}{	\gdef\@journal{robotics}	\gdef\@journalshort{Robotics}	\gdef\@journalfull{Robotics}	\gdef\@doiabbr{robotics}	\gdef\@ISSN{2218-6581}	}
\DeclareOption{safety}{	\gdef\@journal{safety}	\gdef\@journalshort{Safety}	\gdef\@journalfull{Safety}	\gdef\@doiabbr{safety}	\gdef\@ISSN{2313-576X}	}
\DeclareOption{sci}{	\gdef\@journal{sci}	\gdef\@journalshort{Sci}	\gdef\@journalfull{Sci}	\gdef\@doiabbr{sci}	\gdef\@ISSN{2413-4155}	}
\DeclareOption{scipharm}{	\gdef\@journal{scipharm}	\gdef\@journalshort{Sci. Pharm.}	\gdef\@journalfull{Scientia Pharmaceutica}	\gdef\@doiabbr{scipharm}	\gdef\@ISSN{2218-0532}	}
\DeclareOption{sensors}{	\gdef\@journal{sensors}	\gdef\@journalshort{Sensors}	\gdef\@journalfull{Sensors}	\gdef\@doiabbr{s}	\gdef\@ISSN{1424-8220}	}
\DeclareOption{separations}{	\gdef\@journal{separations}	\gdef\@journalshort{Separations}	\gdef\@journalfull{Separations}	\gdef\@doiabbr{separations}	\gdef\@ISSN{2297-8739}	}
\DeclareOption{sexes}{	\gdef\@journal{sexes}	\gdef\@journalshort{Sexes}	\gdef\@journalfull{Sexes}	\gdef\@doiabbr{}	\gdef\@ISSN{2411-5118}	}
\DeclareOption{signals}{	\gdef\@journal{signals}	\gdef\@journalshort{Signals}	\gdef\@journalfull{Signals}	\gdef\@doiabbr{signals}	\gdef\@ISSN{2624-6120}	}
\DeclareOption{sinusitis}{	\gdef\@journal{sinusitis}	\gdef\@journalshort{Sinusitis}	\gdef\@journalfull{Sinusitis}	\gdef\@doiabbr{sinusitis}	\gdef\@ISSN{2309-107X}	}
\DeclareOption{smartcities}{	\gdef\@journal{smartcities}	\gdef\@journalshort{Smart Cities}	\gdef\@journalfull{Smart Cities}	\gdef\@doiabbr{smartcities}	\gdef\@ISSN{2624-6511}	}
\DeclareOption{sna}{	\gdef\@journal{sna}	\gdef\@journalshort{Sinusitis Asthma}	\gdef\@journalfull{Sinusitis and Asthma}	\gdef\@doiabbr{sna}	\gdef\@ISSN{2624-7003}	}
\DeclareOption{societies}{	\gdef\@journal{societies}	\gdef\@journalshort{Societies}	\gdef\@journalfull{Societies}	\gdef\@doiabbr{soc}	\gdef\@ISSN{2075-4698}	}
\DeclareOption{socsci}{	\gdef\@journal{socsci}	\gdef\@journalshort{Soc. Sci.}	\gdef\@journalfull{Social Sciences}	\gdef\@doiabbr{socsci}	\gdef\@ISSN{2076-0760}	}
\DeclareOption{soilsystems}{	\gdef\@journal{soilsystems}	\gdef\@journalshort{Soil Syst.}	\gdef\@journalfull{Soil Systems}	\gdef\@doiabbr{soilsystems}	\gdef\@ISSN{2571-8789}	}
\DeclareOption{sports}{	\gdef\@journal{sports}	\gdef\@journalshort{Sports}	\gdef\@journalfull{Sports}	\gdef\@doiabbr{sports}	\gdef\@ISSN{2075-4663}	}
\DeclareOption{standards}{	\gdef\@journal{standards}	\gdef\@journalshort{Standards}	\gdef\@journalfull{Standards}	\gdef\@doiabbr{}	\gdef\@ISSN{2305-6703}	}
\DeclareOption{stats}{	\gdef\@journal{stats}	\gdef\@journalshort{Stats}	\gdef\@journalfull{Stats}	\gdef\@doiabbr{stats}	\gdef\@ISSN{2571-905X}	}
\DeclareOption{surfaces}{	\gdef\@journal{surfaces}	\gdef\@journalshort{Surfaces}	\gdef\@journalfull{Surfaces}	\gdef\@doiabbr{surfaces}	\gdef\@ISSN{2571-9637}	}
\DeclareOption{surgeries}{	\gdef\@journal{surgeries}	\gdef\@journalshort{Surgeries}	\gdef\@journalfull{Surgeries}	\gdef\@doiabbr{}	\gdef\@ISSN{2017-2017}	}
\DeclareOption{sustainability}{	\gdef\@journal{sustainability}	\gdef\@journalshort{Sustainability}	\gdef\@journalfull{Sustainability}	\gdef\@doiabbr{su}	\gdef\@ISSN{2071-1050}	}
\DeclareOption{symmetry}{	\gdef\@journal{symmetry}	\gdef\@journalshort{Symmetry}	\gdef\@journalfull{Symmetry}	\gdef\@doiabbr{sym}	\gdef\@ISSN{2073-8994}	}
\DeclareOption{systems}{	\gdef\@journal{systems}	\gdef\@journalshort{Systems}	\gdef\@journalfull{Systems}	\gdef\@doiabbr{systems}	\gdef\@ISSN{2079-8954}	}
\DeclareOption{technologies}{	\gdef\@journal{technologies}	\gdef\@journalshort{Technologies}	\gdef\@journalfull{Technologies}	\gdef\@doiabbr{technologies}	\gdef\@ISSN{2227-7080}	}
\DeclareOption{test}{	\gdef\@journal{test}	\gdef\@journalshort{Test}	\gdef\@journalfull{Test}	\gdef\@doiabbr{}	\gdef\@ISSN{}	}
\DeclareOption{toxics}{	\gdef\@journal{toxics}	\gdef\@journalshort{Toxics}	\gdef\@journalfull{Toxics}	\gdef\@doiabbr{toxics}	\gdef\@ISSN{2305-6304}	}
\DeclareOption{toxins}{	\gdef\@journal{toxins}	\gdef\@journalshort{Toxins}	\gdef\@journalfull{Toxins}	\gdef\@doiabbr{toxins}	\gdef\@ISSN{2072-6651}	}
\DeclareOption{tropicalmed}{	\gdef\@journal{tropicalmed}	\gdef\@journalshort{Trop. Med. Infect. Dis.}	\gdef\@journalfull{Tropical Medicine and Infectious Disease}	\gdef\@doiabbr{tropicalmed}	\gdef\@ISSN{2414-6366}	}
\DeclareOption{universe}{	\gdef\@journal{universe}	\gdef\@journalshort{Universe}	\gdef\@journalfull{Universe}	\gdef\@doiabbr{universe}	\gdef\@ISSN{2218-1997}	}
\DeclareOption{urbansci}{	\gdef\@journal{urbansci}	\gdef\@journalshort{Urban Sci.}	\gdef\@journalfull{Urban Science}	\gdef\@doiabbr{urbansci}	\gdef\@ISSN{2413-8851}	}
\DeclareOption{vaccines}{	\gdef\@journal{vaccines}	\gdef\@journalshort{Vaccines}	\gdef\@journalfull{Vaccines}	\gdef\@doiabbr{vaccines}	\gdef\@ISSN{2076-393X}	}
\DeclareOption{vehicles}{	\gdef\@journal{vehicles}	\gdef\@journalshort{Vehicles}	\gdef\@journalfull{Vehicles}	\gdef\@doiabbr{vehicles}	\gdef\@ISSN{2624-8921}	}
\DeclareOption{vetsci}{	\gdef\@journal{vetsci}	\gdef\@journalshort{Vet. Sci.}	\gdef\@journalfull{Veterinary Sciences}	\gdef\@doiabbr{vetsci}	\gdef\@ISSN{2306-7381}	}
\DeclareOption{vibration}{	\gdef\@journal{vibration}	\gdef\@journalshort{Vibration}	\gdef\@journalfull{Vibration}	\gdef\@doiabbr{vibration}	\gdef\@ISSN{2571-631X}	}
\DeclareOption{viruses}{	\gdef\@journal{viruses}	\gdef\@journalshort{Viruses}	\gdef\@journalfull{Viruses}	\gdef\@doiabbr{v}	\gdef\@ISSN{1999-4915}	}
\DeclareOption{vision}{	\gdef\@journal{vision}	\gdef\@journalshort{Vision}	\gdef\@journalfull{Vision}	\gdef\@doiabbr{vision}	\gdef\@ISSN{2411-5150}	}
\DeclareOption{water}{	\gdef\@journal{water}	\gdef\@journalshort{Water}	\gdef\@journalfull{Water}	\gdef\@doiabbr{w}	\gdef\@ISSN{2073-4441}	}
\DeclareOption{wem}{	\gdef\@journal{wem}	\gdef\@journalshort{Wildl. Ecol. Manag.}	\gdef\@journalfull{Wildlife Ecology and Management}	\gdef\@doiabbr{}	\gdef\@ISSN{1234-4321}	}
\DeclareOption{wevj}{	\gdef\@journal{wevj}	\gdef\@journalshort{World Electric Vehicle Journal}	\gdef\@journalfull{World Electric Vehicle Journal}	\gdef\@doiabbr{wevj}	\gdef\@ISSN{2032-6653}	}

### 2020-02-28 article-information-2019.tex
final, final, final edits

@@
 % Authors, for the paper (add full first names)
-\Author{\hl{Navdeep Gill} $^{1, \dagger}$, Patrick Hall $^{1,2,}$*$^{,\dagger}$, Kim Montgomery $^{1,\dagger}$ and Nicholas Schmidt $^{3,}$*$^{,\dagger}$} %%%Author order is not the same as on the system, please confirm which one is correct.
+\Author{{Navdeep Gill} $^{1, \dagger}$, Patrick Hall $^{1,2,}$*$^{,\dagger}$, Kim Montgomery $^{1,\dagger}$ and Nicholas Schmidt $^{3,}$*$^{,\dagger}$} %%%Author order is not the same as on the system, please confirm which one is correct.
 %Please carefully check the accuracy of names and affiliations.
@@
 \address{%
-$^{1}$ \quad \hl{H2O.ai}, 2307 Leghorn St., Mountain View, CA, 94043, USA; navdeep.gill@h2o.ai (N.G.); kim.montgomery@h2o.ai (K.M.)\\%%%Please provide the department, university,zip code, city, country.
-$^{2}$ \quad \hl{The George Washington University}, Department of Decision Sciences, 2201 G Street NW,
-Washington, D.C., 20052, USA\\%%%Please provide the department, zip code, city, country.
-$^{3}$ \quad \hl{BLDS, LLC}, 1608 Walnut St., Suite 1108, Philadelphia, PA, 19103, USA}%%%Please provide the department, university,zip code, city, country.
+$^{1}$ \quad {H2O.ai}, Mountain View 94043, CA, USA; navdeep.gill@h2o.ai (N.G.); kim.montgomery@h2o.ai (K.M.)\\
+$^{2}$ \quad {Department of Decision Sciences, The George Washington University}, Washington 20052, DC,  USA\\%%%Please provide the department, zip code, city, country.
+$^{3}$ \quad {BLDS, LLC}, Philadelphia 19103, PA, USA}
 
@@
 
-Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that~perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds ({\hl{See}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
-%is the text color necessary? please confirm if need remove the link of the text,
-%% we have removed the link of text, and keep web link
-: 
+Responsible artificial intelligence (AI) has been variously conceptualized as AI-based products or projects that use transparent technical mechanisms, that create appealable decisions or outcomes, that~perform reliably and in a trustworthy manner over time, that exhibit minimal social discrimination, and that are designed by humans with diverse experiences, both in terms of demographics and professional backgrounds (e.g.,
 % THESE HYPERLINKS ARE NECESSARY - PH
@@
 ML~models can be inaccurate and unappealable %INCORRECT - PH -- Please confirm that the intended meaning has been retained.
-black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop} ({\hl{See}% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
-: 
+black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop} (e.g.,
 \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}
@@
 This~manuscript makes no claim that these interdependent issues of ML opaqueness, discrimination, privacy harms, and security vulnerabilities have been resolved, even as singular entities, and much less as complex intersectional phenomena.
-However, Sections \ref{sec:m_and_m}--\ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and disparate impact (DI) and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems ({\hl{This text and associated software are not, and should not be construed as, legal advice or regulatory compliance requirements}}), ({\hl{In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervision and Regulation (SR) Letter 11-7, the European Union (EU) General Data Protection Regulation (GDPR) Article} 22, or other laws and regulations \cite{ff_interpretability}.})%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+However, Sections \ref{sec:m_and_m}--\ref{sec:disc} do propose some specific technical countermeasures, mostly in the form of interpretable models, post-hoc explanation, and disparate impact (DI) and discrimination testing, that responsible practitioners can use to address a subset of these vexing problems.
+(In the United States (US), interpretable models, explanations, DI testing, and the model documentation they enable may also be required under the Civil Rights Acts of 1964 and 1991, the Americans with Disabilities Act, the Genetic Information Nondiscrimination Act, the Health Insurance Portability and Accountability Act, the Equal Credit Opportunity Act (ECOA), the Fair Credit Reporting Act (FCRA), the Fair Housing Act, Federal Reserve Supervision and Regulation (SR) Letter 11-7, the European Union (EU) General Data Protection Regulation (GDPR) Article 22, or other laws and regulations \cite{ff_interpretability}.
+Note that this text and associated software are not, and should not be construed as, legal advice or regulatory compliance requirements.)
 
@@
 Its~known feature importance and augmented discrimination characteristics are used to gauge the validity of interpretable modeling, post-hoc explanation, and discrimination testing techniques \cite{friedman1979tree,friedman1991multivariate}.
-The~mortgage data (see Section~\ref{ssec:mort_data}) are sourced from the Home Mortgage Disclosure Act (HMDA) database, a fairly realistic data source for demonstrating the template workflow \cite{hmda} ({\hl{See:}
+The~mortgage data (see Section~\ref{ssec:mort_data}) are sourced from the Home Mortgage Disclosure Act (HMDA) database, a fairly realistic data source for demonstrating the template workflow \cite{hmda} (see
 \href{https://www.consumerfinance.gov/data-research/hmda/}
-{Mortgage data (HMDA})}). %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+{Mortgage data (HMDA)}). %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
 To provide a sense of fit differences, performance is compared on simulated data and collected mortgage data between the more interpretable constrained ML models and the less interpretable unconstrained ML models.
 Because the unconstrained ML models, gradient boosting machines (GBMs, e.g., \cite{friedman2001greedy,esl}) and artificial neural networks (ANNs, e.g., \cite{recht2011hogwild,hinton2012improving,sutskever2013importance,zeiler2012adadelta}), do not exhibit convincing accuracy benefits on the simulated or mortgage data and can also present the unmitigated risks discussed above, further explanation and discrimination analyses are applied only to the constrained, interpretable ML models \cite{please_stop,fair_washing,scaffolding}.
-Here, monotonic gradient boosting machines ({\hl{as implemented in}
+Here, monotonic gradient boosting machines
+(MGBMs, as implemented in 
 %\href{https://xgboost.readthedocs.io/en/latest/tutorials/monotonic.html}
-{\texttt{\hl{XGBoost}}} or 
+\texttt{XGBoost} or 
 %\href{https://github.com/h2oai/h2o-3/blob/master/h2o-py/demos/H2O_tutorial_gbm_monotonicity.ipynb}
-{\texttt{\hl{h2o}}}.}) %% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. % we have removed the link of text, and keep web link
-(MGBMs, see Section~\ref{ssec:mgbm}) and explainable neural networks (XNNs, see Section~\ref{ssec:xnn} \cite{,wf_xnn,yang2019enhancing}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
+\texttt{h2o}, see Section~\ref{ssec:mgbm}) 
+and explainable neural networks (XNNs, e.g., \cite{,wf_xnn,yang2019enhancing}, see Section~\ref{ssec:xnn}) will serve as those more interpretable models for subsequent explanatory and discrimination analyses.
 MGBM and XNN interpretable model architectures are selected for the example workflow because they are straightforward variants of popular unconstrained ML models.
@@
 PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models.
-Discrimination testing methods discussed (see Section~\ref{ssec:di}) include adverse impact ratio (AIR), marginal effect (ME), and standardized mean difference (SMD)  \cite{feldman2015certifying,cohen1988statistical,cohen1992power}.
-({\hl{See: Part 1607 - Uniform Guidelines on Employee Selection Procedures} (1978) \textsection 1607.4.}).
+Discrimination testing methods discussed (see Section~\ref{ssec:di}) include adverse impact ratio
+(AIR, see Part 1607---Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4), 
+marginal effect (ME), and standardized mean difference (SMD)  \cite{feldman2015certifying,cohen1988statistical,cohen1992power}.
 %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
@@
 %\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogradientboostingestimator}
-{\texttt{\hl{h2o}}}, 
+{\texttt{{h2o}}}, 
 %\href{https://lightgbm.readthedocs.io/en/latest/Python-Intro.html}
-{\texttt{\hl{LightGBM}}}, and 
+{\texttt{{LightGBM}}}, and 
 %\href{https://xgboost.readthedocs.io/en/latest/python/python_intro.html}
-{\texttt{\hl{XGBoost}}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
+{\texttt{{XGBoost}}}, and Tree SHAP is used to calculate accurate and consistent global and local feature importance for MGBM models in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 Deep SHAP is an approximate Shapley value technique that creates SHAP values for ANNs \cite{shapley}.
@@
 %\href{https://github.com/slundberg/shap}
-{\texttt{\hl{shap}}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
+{\texttt{{shap}}} package and is used to generate SHAP values for the two $g^{XNN}$ models discussed in Section~\ref{ssec:int_ex_mort} and Appendix \ref{a_ssec:int_ex_sim}.
 For more information pertaining to the calculation of Shapley values, see Appendices \ref{a_ssec:not} and \ref{a_ssec:shap}.%% we have removed the link of text, and keep web link
@@
 
-Because many current technical discussions of fairness in ML appear inconclusive ({See:} 
-\href{https://www.youtube.com/watch?v=jIXIuYdnyyk}
-{Tutorial: 21 Fairness Definitions and Their Politics}.)%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
-, this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
+Because many current technical discussions of fairness in ML appear inconclusive (e.g., 
+\href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}), this text will draw on regulatory and legal standards that have been used for years in regulated, high-stakes employment and financial decisions.
 The~discussed measures are also representative of fair lending analyses and pair well with the mortgage data.
@@
  where $\hat{\mathbf{y}}$ are the model decisions, $X_p$ and $X_c$ represent binary markers created from some demographic attribute, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and $\text{Pr}(\cdot)$ is the operator for conditional probability.
-ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market.
-({See:
+ME is a favored DI measure used by the US Consumer Financial Protection Bureau (CFPB), the primary agency charged with regulating fair lending laws at the largest US lending institutions and for various other participants in the consumer financial market (see
 \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}
-{Supervisory Highlights, Issue 9, Fall 2015}.) %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
- Another important DI measure is AIR, more~commonly known as a {relative risk ratio} in settings outside of regulatory compliance.
+{\textit{Supervisory Highlights}, Issue 9, Fall 2015}).
+ Another important DI measure is AIR, more~commonly known as a relative risk ratio in settings outside of regulatory compliance.
 
@@
 SMD is equal to the difference in the average protected class outcome, $\bar{\hat{\mathbf{y}}}_p$, minus the control class outcome, $\bar{\hat{\mathbf{y}}}_c$, divided by a measure of the standard deviation of the population, $\sigma_{\hat{\mathbf{y}}}$.
-({\hl{There are several measures of the standard deviation of the score that are typically used}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
-: \textbf{\hl{1}.} the standard deviation of the population, irrespective of protected class status, \textbf{\hl{2}.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{\hl{3}.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.}) Cohen defined values of this measure to have {small}, {medium}, and {large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
+({{There are several measures of the standard deviation of the score that are typically used}%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
+: \textbf{{1}.} the standard deviation of the population, irrespective of protected class status, \textbf{{2}.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{{3}.} a pooled standard deviation, using the standard deviations for each of the two groups with weights.}) Cohen defined values of this measure to have {small}, {medium}, and {large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.
 %is bold necessary?
@@
 %\href{https://numpy.org/}
-{\texttt{\hl{numpy}}} 1.14.5 and 
+{\texttt{{numpy}}} 1.14.5 and 
 %\href{https://pandas.pydata.org/}
-{\texttt{\hl{pandas}}} 0.22.0 for data manipulation, 
+{\texttt{{pandas}}} 0.22.0 for data manipulation, 
 %\href{https://github.com/h2oai/h2o-3}
-{\texttt{\hl{h2o}}} 3.26.0.9, 
+{\texttt{{h2o}}} 3.26.0.9, 
 %\href{https://keras.io/}
-{\texttt{\hl{Keras}}} 2.3.1, 
+{\texttt{{Keras}}} 2.3.1, 
 %\href{https://github.com/slundberg/shap}
-{\texttt{\hl{shap}}} 0.31.0, and 
+{\texttt{{shap}}} 0.31.0, and 
 %\href{https://www.tensorflow.org/}
-{\texttt{\hl{tensorflow}}} 1.14.0 for modeling, explanation, and discrimination testing, and typically 
+{\texttt{{tensorflow}}} 1.14.0 for modeling, explanation, and discrimination testing, and typically 
 %\href{https://matplotlib.org/}
-{\texttt{\hl{matplotlib}}} 2.2.2 for plotting.
+{\texttt{{matplotlib}}} 2.2.2 for plotting.
 
@@
 		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}
-		{\textbf{\hl{Accuracy}}}~$\uparrow$ & 
+		{\textbf{{Accuracy}}}~$\uparrow$ & 
 		%\href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}
-		{\textbf{\hl{AUC}}}~$\uparrow$ & 
+		{\textbf{{AUC}}}~$\uparrow$ & 
 		%\href{https://en.wikipedia.org/wiki/F1_score}
-		{\textbf{\hl{F1}}}~$\uparrow$ & 
+		{\textbf{{F1}}}~$\uparrow$ & 
 		%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}
-		{\textbf{\hl{Logloss}}}~$\downarrow$ & 
+		{\textbf{{Logloss}}}~$\downarrow$ & 
 		%\href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}
-		{\textbf{\hl{MCC}}}~$\uparrow$ & 
+		{\textbf{{MCC}}}~$\uparrow$ & 
 		%\href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}
-		{\textbf{\hl{RMSE}}}~$\downarrow$ & 
+		{\textbf{{RMSE}}}~$\downarrow$ & 
 		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
-		{\textbf{\hl{Sensitivity}}}~$\uparrow$ & 
+		{\textbf{{Sensitivity}}}~$\uparrow$ & 
 		%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
-		{\textbf{\hl{Specificity}}}~$\uparrow$ \\
+		{\textbf{{Specificity}}}~$\uparrow$ \\
 		\midrule
@@
 
+% start a new page
+\newpage
+% change it to landscape
+\paperwidth=\pdfpageheight
+\paperheight=\pdfpagewidth
+\pdfpageheight=\paperheight
+\pdfpagewidth=\paperwidth
+\newgeometry{layoutwidth=297mm,layoutheight=210 mm, left=2.7cm,right=2.7cm,top=1.8cm,bottom=1.5cm, includehead,includefoot}
+\fancyheadoffset[LO,RE]{0cm}
+\fancyheadoffset[RO,LE]{0cm}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[H]
@@
 	%\includegraphics[width=15cm]{img/sim_xnn_glob_ridge.png}
-	\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
+	\includegraphics[width=23cm]{img/sim_xnn_glob_ridge.png}
 	\caption{Output layer $\gamma_k$ weights, corresponding $n_k$ subnetwork ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the simulated data.}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% change everything back
+\newpage
+\restoregeometry
+\paperwidth=\pdfpageheight
+\paperheight=\pdfpagewidth
+\pdfpageheight=\paperheight
+\pdfpagewidth=\paperwidth
+\headwidth=\textwidth
 
@@
 	%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#accuracy-score}
-	{\textbf{\hl{Accuracy}}}~$\uparrow$ & 
+	{\textbf{{Accuracy}}}~$\uparrow$ & 
 	%\href{https://en.wikipedia.org/wiki/Receiver_operating_characteristic#Area_under_the_curve}
-	{\textbf{\hl{AUC}}}~$\uparrow$ & 
+	{\textbf{{AUC}}}~$\uparrow$ & 
 	%\href{https://en.wikipedia.org/wiki/F1_score}
-	{\textbf{\hl{F1}}}~$\uparrow$ & 
+	{\textbf{{F1}}}~$\uparrow$ & 
 	%\href{https://scikit-learn.org/stable/modules/model_evaluation.html#log-loss}
-	{\textbf{\hl{Logloss}}}~$\downarrow$ & 
+	{\textbf{{Logloss}}}~$\downarrow$ & 
 	%\href{https://en.wikipedia.org/wiki/Matthews_correlation_coefficient}
-	{\textbf{\hl{MCC}}}~$\uparrow$ & 
+	{\textbf{{MCC}}}~$\uparrow$ & 
 	%\href{https://en.wikipedia.org/wiki/Root-mean-square_deviation}
-	{\textbf{\hl{RMSE}}}~$\downarrow$ & 
+	{\textbf{{RMSE}}}~$\downarrow$ & 
 	%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
-	{\textbf{\hl{Sensitivity}}}~$\uparrow$ & 
+	{\textbf{{Sensitivity}}}~$\uparrow$ & 
 	%\href{https://en.wikipedia.org/wiki/Sensitivity_and_specificity}
-	{\textbf{\hl{Specificity}}}~$\uparrow$ \\
+	{\textbf{{Specificity}}}~$\uparrow$ \\
 	\midrule	
@@
 Because PD curves generally followed the patterns of the ICE curves for both features, it is also likely that no strong interactions were at play for DTI ratio and LTV ratio under $g^\text{MGBM}$.
-Of course, the monotonicity constraints themselves could have dampened the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they did exist in the training data (e.g., LTV ratio and the no introductory rate period flag, see\hl{ Figure} \ref{fig:mort_xnn_glob_ridge}).
+Of course, the monotonicity constraints themselves could have dampened the effects of non-monotonic interactions under $g^\text{MGBM}$, even if they did exist in the training data (e.g., LTV ratio and the no introductory rate period flag, see {Figure} 6).
 This rigidity could also have played a role in the performance differences between $g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ in the mortgage data not observed for the simulated data, wherein strong interactions appeared to be between features with the same monotonicity constraints (e.g., $X_{\text{Friedman},1}$ and $X_{\text{Friedman},2}$, see Figure \ref{fig:sim_xnn_glob_ridge}).%Figures  should be cited in sequential numerical order.
@@
 
+% start a new page
+\newpage
+% change it to landscape
+\paperwidth=\pdfpageheight
+\paperheight=\pdfpagewidth
+\pdfpageheight=\paperheight
+\pdfpagewidth=\paperwidth
+\newgeometry{layoutwidth=297mm,layoutheight=210 mm, left=2.7cm,right=2.7cm,top=1.8cm,bottom=1.5cm, includehead,includefoot}
+\fancyheadoffset[LO,RE]{0cm}
+\fancyheadoffset[RO,LE]{0cm}
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[H]
-\includegraphics[width=17cm]{img/mort_xnn_glob_ridge.png}
+\includegraphics[width=23cm]{img/mort_xnn_glob_ridge.png}
 \caption{Output layer $\gamma_k$ weights, corresponding $n_k$ ridge functions, and associated projection layer $\beta_{k,j}$ weights for $g^\text{XNN}$ on the mortgage data.}
@@
 Table \ref{tab:mort_disc}{a,b} show the results of the discrimination tests using the mortgage data for two sets of class-control groups: blacks as compared to whites, and females as compared to males.
-As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and $g^{\text{XNN}}(\mathbf{X})$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome)({\hl{See Appendix} \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection.).
-%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
-} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  were predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score was favorable.
+As with the simulated data in Table \ref{tab:sim_disc}, several measures of disparities are shown, with the SMDs calculated using the probabilities from $g^{\text{MGBM}}(\mathbf{X})$ and $g^{\text{XNN}}(\mathbf{X})$, and the accuracy, FPRs, and  FPR ratios, MEs, and AIRs calculated using a binary outcome based on a cutoff of 0.20 (anyone with probabilities of 0.2 or less receives the favorable outcome; see Appendix \ref{a_sec:cut} for comments pertaining to discrimination testing and cutoff selection).
+Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$  were predicting the likelihood of receiving a high-priced loan, $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assume that a lower score was favorable.
 Thus, one might consider FPR ratios as a measure of the class-control disparities.
@@
 
-Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, and deployment meant to decrease discrimination, inaccuracy, and privacy and security risks for high-stakes, human-centered, or regulated ML applications (See: \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure~\ref{fig:blueprint}.}).
-While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
+Figure \ref{fig:blueprint} displays a holistic approach to ML model training, assessment, and deployment meant to decrease discrimination, inaccuracy, and privacy and security risks for high-stakes, human-centered, or regulated ML applications.
+(See \href{https://github.com/jphall663/hc_ml}{Toward Responsible Machine Learning} for details regarding Figure~\ref{fig:blueprint}.) While all the methods mentioned in Figure \ref{fig:blueprint} play an important role in increasing human trust and understanding of ML, a few pertinent references and Python resources are highlighted below as further reading to augment this this text's focus on certain interpretable models, post-hoc explanation, and discrimination testing techniques.
 
@@
 %\href{https://github.com/interpretml/interpret}
-{\texttt{\hl{interpret}}} package, optimal sparse decision trees, GAMs in the 
+{\texttt{{interpret}}} package, optimal sparse decision trees, GAMs in the 
 %\href{https://github.com/dswah/pyGAM}
-{\texttt{\hl{pyGAM}}} package, a variant of Friedman's RuleFit  in the 
+{\texttt{{pyGAM}}} package, a variant of Friedman's RuleFit  in the 
 %\href{https://github.com/scikit-learn-contrib/skope-rules}
-{\texttt{\hl{skope-rules}}} package, monotonic calibrated interpolated lookup tables in 
+{\texttt{{skope-rules}}} package, monotonic calibrated interpolated lookup tables in 
 %\href{https://github.com/tensorflow/lattice}
-{\texttt{\hl{tensorflow/lattice}}}, and \textit{this looks like that} interpretable deep learning \cite{osdt,rulefit,lattice,this_looks_like_that} ({See: 
-\href{https://github.com/xiyanghu/OSDT}{optimal sparse decision trees}; see: 
-\href{https://github.com/cfchen-duke/ProtoPNet}{this looks like that}.). Additional, relevant references and Python functionality include:
+{\texttt{{tensorflow/lattice}}}, and \textit{this looks like that} interpretable deep learning \cite{osdt,rulefit,lattice,this_looks_like_that} (see 
+\href{https://github.com/xiyanghu/OSDT}{Optimal Sparse Decision Trees},
+\href{https://github.com/cfchen-duke/ProtoPNet}{ProtoPNet (\textit{this looks like that})}). Additional, relevant references and Python functionality include:
 
@@
 
-\item \textbf{\hl{Exploratory data analysis (EDA)}}: \texttt{H2OAggregatorEstimator} in 
+\item \textbf{{Exploratory data analysis (EDA)}}: \texttt{H2OAggregatorEstimator} in 
 %\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2oaggregatorestimator}
-{\texttt{\hl{h2o}}} \cite{wilkinson2018visualizing}.
+{\texttt{{h2o}}} \cite{wilkinson2018visualizing}.
 
@@
 %\href{http://docs.h2o.ai/h2o/latest-stable/h2o-py/docs/modeling.html#h2ogeneralizedlowrankestimator}
-{\texttt{\hl{h2o}}} \cite{udell2016generalized}.
+{\texttt{{h2o}}} \cite{udell2016generalized}.
 
@@
 %\href{https://github.com/tensorflow/privacy/}
-{\texttt{\hl{tensorflow/privacy}}} \cite{diffprivlib,ji2014differential,pate,abadi2016deep}.
+{\texttt{{tensorflow/privacy}}} \cite{diffprivlib,ji2014differential,pate,abadi2016deep}.
 
@@
 %\href{https://github.com/microsoft/dowhy}
-{\hl{dowhy}}, 
+{{dowhy}}, 
 %\href{https://docs.pymc.io/}
@@
 %\href{https://github.com/SeldonIO/alibi}
-{\texttt{\hl{alibi}}} and 
+{\texttt{{alibi}}} and 
 %\href{https://github.com/SauceCat/PDPbox}
-{\texttt{\hl{PDPbox}}}, image classification explanations with 
+{\texttt{{PDPbox}}}, image classification explanations with 
 %\href{https://github.com/marcoancona/DeepExplain}
-{\texttt{\hl{DeepExplain}}}, and natural language explanations with 
+{\texttt{{DeepExplain}}}, and natural language explanations with 
 %\href{https://github.com/allenai/allennlp}
-{\texttt{\hl{allennlp}}} \cite{wachter2017counterfactual,grad_attr,wallace2019allennlp}.
+{\texttt{{allennlp}}} \cite{wachter2017counterfactual,grad_attr,wallace2019allennlp}.
 
@@
 %\href{http://aif360.mybluemix.net/}
-{\texttt{\hl{AIF}360}} \cite{kamiran2012data,zhang2018mitigating,lfr,kamiran2012decision}.
+{\texttt{{AIF}360}} \cite{kamiran2012data,zhang2018mitigating,lfr,kamiran2012decision}.
 
@@
 %\href{https://github.com/bethgelab/foolbox}
-{\texttt{\hl{foolbox}}}, 
+{\texttt{{foolbox}}}, 
 %\href{https://github.com/SALib/SALib}
-{\texttt{\hl{SALib}}}, 
+{\texttt{{SALib}}}, 
 %\href{https://github.com/tensorflow/cleverhans}
-{\texttt{\hl{tensorflow/cleverhans}}}, and 
+{\texttt{{tensorflow/cleverhans}}}, and 
 %\href{https://github.com/tensorflow/model-analysis}
-{\texttt{\hl{tensorflow/model- analysis}}} \cite{rauber2017foolbox,papernot2018cleverhans,modeltracker,papernot2018marauder}.
+{\texttt{{tensorflow/model- analysis}}} \cite{rauber2017foolbox,papernot2018cleverhans,modeltracker,papernot2018marauder}.
 
@@
 Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}.
-Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes (See:
+Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.
+(For examples uses in financial services, see
 \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}
-{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.
-{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
-Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.}).
-Transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations ({\hl{E.g.:}  % we have removed the link of text
-%\href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}
-{\hl{US Federal Reserve Bank SR 11-7: Guidance on Model Risk Management}}.}).%%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this. 
+{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}.
+Also note that many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}.
+Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings, such as credit lending \cite{bracke2019machine}.) In general, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some of which may also be required by applicable regulations (e.g., US Federal Reserve Bank SR 11-7: Guidance on Model Risk Management).
  Among these, providing persons affected by a model with the opportunity to appeal ML-based decisions may deserve the most attention.
-ML models are often wrong ({``\hl{All models are wrong, but some are useful.''---George Box, Statistician (1919--2013)}}) %%% footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
- and appealing black-box decisions can be difficult.
+ML models are often wrong (``All models are wrong, but some are useful.''---George Box, Statistician (1919--2013))
+ and appealing black-box decisions can be difficult (e.g.,
+ \href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}
+ {\textit{When a Computer Program Keeps You in Jail}}).
 For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for compliance, but also a failsafe procedure for those affected by ML decisions.
-%is this necessary?
 
@@
 The often black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or~the privacy harms and security vulnerabilities inherent in ML are each serious and difficult problems on their own.
-However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the {fairwashing} or {scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy,fair_washing,scaffolding} (See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}
- {\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.} While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
+However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g., the {fairwashing} or {scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy,fair_washing,scaffolding} (e.g., \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}
+ {\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}).
+(While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}.
 High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}.
-This~is sometimes known as the {transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation}.
-{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.}).
-Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and~discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
+This~is sometimes known as the {transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation (see \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}).) Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and~discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML.
 However, they should also consider that explanations can ease model stealing, data~extraction, and membership inference attacks, and that explanations can mask ML discrimination.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{\hl{The authors declare no conflict of interest}. %%Please confirm.
+\conflictsofinterest{{The authors declare no conflict of interest}. %%Please confirm.
 XNN was first made public by the corporate model validation team at Wells Fargo bank.
@@
 Regulators often use HMDA data to, ``...
-show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory'' (see: \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}).
+show whether lenders are serving the housing needs of their communities; they give public officials information that helps them make decisions and policies; and they shed light on lending patterns that could be discriminatory'' (see \href{https://www.consumerfinance.gov/data-research/hmda/}{Mortgage data (HMDA)}).
 In addition to regulatory use, public advocacy groups use these data for similar purposes, and the lenders themselves use the data to benchmark their community outreach relative to their peers.
@@
 Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.
-One exception, however, is in employment discrimination analyses, where~the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes {prima facie} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination ({Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
-``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).}).
-It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.
+One exception, however, is in employment discrimination analyses, where~the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes {prima facie} evidence of discrimination, which the model user must rebut in order for the DI not to be considered illegal discrimination.
+(Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.
+``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).) It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for measures other than the AIR.
 
@@
 Intrinsic interpretability for $g^\text{MGBM}$ is evaluated with PD and ICE, and post-hoc explainability is highlighted via global and local Shapley explanations.
-For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Subection \ref{ssec:xnn_sim}.
+For $g^\text{XNN}$, Shapley explanation techniques are also used to generate global and local feature importance to augment interpretability results exhibited in Section \ref{ssec:xnn_sim}.
 Both~$g^\text{MGBM}(\mathbf{X})$ and $g^\text{XNN}(\mathbf{X})$ are evaluated for discrimination using AIR, ME, SMD, and other measures.
@@
 Furthermore, using some recently developed discrimination mitigation methods may lead to non-compliance with anti-discrimination laws and regulations.
-A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.'' ({\hl{In the pay discrimination case,} \textit{Bazemore vs.
+A fundamental maxim of US anti-discrimination law is that (to slightly paraphrase), ``similarly situated people should be treated similarly.'' ({{In the pay discrimination case,} \textit{Bazemore vs.
 Friday}, 478 US 385 (1986), the US Supreme Court found that,``Each week's paycheck that delivers less to a black than to a similarly situated white is a wrong actionable ...''  Beyond the obvious conceptual meaning, what specifically constitutes \textit{similarly situated} is controversial and its interpretation differs by circuit.}) % footnote is not allowed in our jounral, we have moved them in the main text, please confirm if you are OK with this.
@@
 Obviously, the inclusion of protected class status will change model output by class.
-With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal ({\hl{In a reverse discrimination case}, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.}) Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
+With possible rare exceptions, this is likely to be viewed with legal and regulatory skepticism today, even if including class status is done with fairness as the goal.
+(In a reverse discrimination case, \textit{Ricci v Desafano}, 557 US 557 (2009), the court found that any consideration of race which is not justified by correcting for past proven discrimination is illegal and, moreover, a lack of fairness is not necessarily evidence of illegal discrimination.) Preprocessing and post-processing techniques may be similarly problematic, because industries that must provide explanations to those who receive unfavorable treatment (e.g., adverse action notices in US financial services) may have to incorporate the class adjustments into their explanations as well.
 
@@
 \newblock In Proceedings of the 25th USENIX Security Symposium, Austin, TX, USA, 10--12 August 2016. pp. 601--618.
-%%%\newblock Available online: 
-%%%  \url{https://www.usenix.org/system/files/conference/usenixsecurity16/sec16_paper_tramer.pdf} ((accessed on 26 February 2020)).
+\newblock Available online: 
+\url{https://www.usenix.org/system/files/conference/usenixsecurity16/sec16_paper_tramer.pdf} (accessed on 26 February 2020).
 
@@
 
-\bibitem[Williams {et~al.}(2017)Williams et~al.]{ff_interpretability}
+\bibitem[Williams {et~al.}(2017) Williams et~al.]{ff_interpretability}
 Williams, M.
-\newblock {\em Interpretability}; \hl{Fast Forward Labs}: Brooklyn, NY, USA.  2017.   %Please add the location of the publisher 
+\newblock {\em Interpretability}; {Fast Forward Labs}: Brooklyn, NY, USA, 2017.   %Please add the location of the publisher 
 \newblock Available online: 
@@
   1979; pp. 5--22.
-%%%\newblock Available online: 
-%%%  \url{http://inspirehep.net/record/140963/files/slac-pub-2336.pdf} (\hl{accessed~on}).
+\newblock Available online: 
+\url{http://inspirehep.net/record/140963/files/slac-pub-2336.pdf} (accessed on 28 February 2020).
 
@@
 \newblock {\em  Ann. Stat.} {\bf 1991}, {\em 19},~1--67.
-%%%\newblock Available online: 
-%%%  \url{https://projecteuclid.org/download/pdf_1/euclid.aos/1176347963} (\hl{accessed~on}).
+\newblock Available online: 
+\url{https://projecteuclid.org/download/pdf_1/euclid.aos/1176347963} (accessed on 28 February 2020).
 
@@
 \newblock {\em  Ann. Stat.} {\bf 2001}, 1189--1232.
-%%%\newblock Available online: 
-%%%  \url{https://projecteuclid.org/download/pdf_1/euclid.aos/1013203451} (\hl{accessed~on}).
+\newblock Available online: 
+\url{https://statweb.stanford.edu/~jhf/ftp/trebst.pdf} (accessed on 28 February 2020).
 
